### 2013-06-17 gapoly.Rmd
1st commit

GA Poly
========================================================

```{r, fig.width=12}
library(genalg)

x1 <- seq(0, 4 * pi, length.out = 201)
x2 <- seq(0, 4 * pi, length.out = 201)

# a latent variable, with a trick expression and some noise
y <- 2 * sin(x1) + cos(2 * x2) + 0.1 * x1 * x2 + 0.1 * rnorm(201)

# Caution: there are **ONLY** 201 observations (and not 201*201)
observations <- data.frame(y = y, x1 = x1, x2 = x2)

##############

make.component <- function(bits, n.vars) {
  vars <- paste0("x",1:n.vars)
  elements <- split(bits, ceiling(seq_along(bits)/(length(bits)/n.vars)))
  str <- ""
  for(i in 1:length(elements)) {
    power <- length(elements[[i]][elements[[i]]==1])
    if (power == 1)
      if (str=="")
        str = vars[i]
      else  
        str = paste0(str, "*", vars[i])
    if (power > 1)
      if (str=="")
        str = paste0(vars[i], "^", power)
      else  
        str = paste0(str, "*", vars[i], "^", power)
  }
  str
}

make.formula <- function(bits, n.vars, n.compts) {
  components <- split(bits, ceiling(seq_along(bits)/(length(bits)/n.compts)))
  str <- ""
  for(i in 1:length(components)) {
    comp <- make.component(components[[i]], n.vars)
    if (comp!="")
      if (str=="")
        str = paste0("I(",comp,")")
    else
        str = paste0(str," + I(", comp ,")")
  }
  str
}

############################

max.degree <- 3
n.vars     <- 3  # x1, x2,... 
n.compts   <- 4  # poly = compt1 + compt2 + ...

set.seed(101)
bits <- sample(c(0,1), max.degree*n.vars*n.compts, replace=TRUE)
paste0(bits, collapse = "") # just to show as string

##############################

x1 <- seq(0, 8 * pi, length.out = 201)/pi
x2 <- sin(x1)^2
x3 <- abs(cos(x2))
y  <- x1*x2 - x1*x3 + x3^3 + rnorm(201,0,.15) # a latent variable with some noise
df <- data.frame(y=y,x1=x1,x2=x2,x3=x3) # make data frame

evalFuncFactory <- function(df, n.vars, n.compts) {

  function(bits) {
    formula <- paste0("y ~ ", make.formula(bits, n.vars, n.compts))
    model <- lm(formula, data=df)
    return( sqrt(mean(residuals(model)^2)) )
  }
}

#############################

GAmodel <- rbga.bin(size = max.degree*n.vars*n.compts, 
                    popSize = 100, 
                    iters = 100, 
                    mutationChance = 0.01, 
                    elitism = TRUE, 
                    evalFunc = evalFuncFactory(df, n.vars, n.compts))

cat(summary.rbga(GAmodel))

best.solution <- GAmodel$population[1,]
best.formula <- paste0("y ~ ", make.formula(best.solution, n.vars, n.compts))
model <- lm(best.formula, data=df)
coef(model)
plot(df$x1,df$y,type="l")
title(best.formula)
y1<-predict(model,df[,-1])
lines(x1,y1,col="red") # approx
```

### 2013-06-17 gapoly.Rmd
code organization

@@
 
-```{r, fig.width=12}
-library(genalg)
+GA package use based on [http://www.r-bloggers.com/genetic-algorithms-a-simple-r-example/](http://www.r-bloggers.com/genetic-algorithms-a-simple-r-example/)
 
-x1 <- seq(0, 4 * pi, length.out = 201)
-x2 <- seq(0, 4 * pi, length.out = 201)
+<hr>
 
-# a latent variable, with a trick expression and some noise
-y <- 2 * sin(x1) + cos(2 * x2) + 0.1 * x1 * x2 + 0.1 * rnorm(201)
+Herein we code a polynomial based on variables $x_1, x_2 \ldots x_n$ into a binary vector (ie, made of 0s and 1s).
 
-# Caution: there are **ONLY** 201 observations (and not 201*201)
-observations <- data.frame(y = y, x1 = x1, x2 = x2)
+To achieve this we need to know three values in advance:
++ number of different variables
++ maximum number of different components (being a polynomial a sum of components)
++ maximum degree for any given variable
 
-##############
+```{r}
+n.vars     <- 3  # x1, x2,... 
+max.compts <- 4  # poly = compt1 + compt2 + ...
+max.degree <- 3
+```
 
+The chromosome coding will as follow:
++ the chromosome is split into `max.compts` sets of bits of equal size (a component)
++ each component is split into `n.vars` sets of `max.degree` size each (a variable)
++ for each variable, the number of 1s gives the variable's degree (this means that a given polynomial may have more than one possible representation)
+
+Eg (for the values in the previous code snippet):
+
+Consider polynomial $x_1^2 \times x_3 + x_3^3 + x_1 \times x_2$
+
+One possible coding would be (the semicolons separate components, the commas seperate variables):
+
+$$011,000,001;000,000,111;001,001,000;000,000,000$$
+
+The next functions transform the binary vector into a polynomial formula that `lm()` understands:
+
+```{r}
 make.component <- function(bits, n.vars) {
@@
 
-make.formula <- function(bits, n.vars, n.compts) {
-  components <- split(bits, ceiling(seq_along(bits)/(length(bits)/n.compts)))
+make.formula <- function(bits, n.vars, max.compts) {
+  components <- split(bits, ceiling(seq_along(bits)/(length(bits)/max.compts)))
   str <- ""
@@
 }
+```
 
-############################
+Let's check the previous eg, $x_1^2 \times x_3 + x_3^3 + x_1 \times x_2$:
 
-max.degree <- 3
-n.vars     <- 3  # x1, x2,... 
-n.compts   <- 4  # poly = compt1 + compt2 + ...
+```{r}
+code.eg <- c(0,1,1,0,0,0,0,0,1,0,0,0,0,0,0,1,1,1,0,0,1,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0)
+paste0(code.eg, collapse = "") # just to show as string
+make.formula(code.eg, n.vars, max.compts)
+```
 
+Let's create a dataframe for testing:
+
+```{r, fig.width=12}
 set.seed(101)
-bits <- sample(c(0,1), max.degree*n.vars*n.compts, replace=TRUE)
-paste0(bits, collapse = "") # just to show as string
-
-##############################
-
 x1 <- seq(0, 8 * pi, length.out = 201)/pi
@@
 df <- data.frame(y=y,x1=x1,x2=x2,x3=x3) # make data frame
+```
 
-evalFuncFactory <- function(df, n.vars, n.compts) {
+This binary coding is useful to use the binary GA rbga.bin() from package `genalg` (see below):
 
-  function(bits) {
-    formula <- paste0("y ~ ", make.formula(bits, n.vars, n.compts))
+```{r}
+library(genalg)
+
+# This is a factory for the evaluation function that the GA function
+# needs to execute.
+
+# The factory is a closure that includes the dataframe, the number of
+# variable and the max number of componentes, all required information
+# to perform the polynomial regression and then compute the residuals' rsme 
+# which will be the fitness of the respective chromosome
+evalFuncFactory <- function(df, n.vars, max.compts) {
+
+  function(chromosome) {
+    formula <- paste0("y ~ ", make.formula(chromosome, n.vars, max.compts))
     model <- lm(formula, data=df)
-    return( sqrt(mean(residuals(model)^2)) )
+    return( sqrt(mean(residuals(model)^2)) ) #rbga.bim tries to minimize
   }
@@
 #############################
-
-GAmodel <- rbga.bin(size = max.degree*n.vars*n.compts, 
+# Unleash the chromosomes of war!
+GAmodel <- rbga.bin(size = max.degree*n.vars*max.compts, 
                     popSize = 100, 
@@
                     elitism = TRUE, 
-                    evalFunc = evalFuncFactory(df, n.vars, n.compts))
+                    evalFunc = evalFuncFactory(df, n.vars, max.compts))
 
 cat(summary.rbga(GAmodel))
-
 best.solution <- GAmodel$population[1,]
-best.formula <- paste0("y ~ ", make.formula(best.solution, n.vars, n.compts))
+best.formula <- paste0("y ~ ", make.formula(best.solution, n.vars, max.compts))
 model <- lm(best.formula, data=df)

### 2013-06-17 gapoly.Rmd
changed dataframe eg

@@
 
-Eg (for the values in the previous code snippet):
+Eg: consider polynomial $x_1^2 \times x_3 + x_3^3 + x_1 \times x_2$
 
-Consider polynomial $x_1^2 \times x_3 + x_3^3 + x_1 \times x_2$
-
-One possible coding would be (the semicolons separate components, the commas seperate variables):
+One possible coding -- for the values in the previous code snippet -- would be:
 
 $$011,000,001;000,000,111;001,001,000;000,000,000$$
+
+(the semicolons separate components, the commas seperate variables)
 
@@
 
-Let's create a dataframe for testing:
+Let's create a dataframe for testing. Notice that the data frame columns name must be called $y$ for the output value, and $x_i$ for the input values:
 
 ```{r, fig.width=12}
-set.seed(101)
+set.seed(102)
 x1 <- seq(0, 8 * pi, length.out = 201)/pi
 x2 <- sin(x1)^2
-x3 <- abs(cos(x2))
+x3 <- abs(cos(x1))
 y  <- x1*x2 - x1*x3 + x3^3 + rnorm(201,0,.15) # a latent variable with some noise
@@
 
-This binary coding is useful to use the binary GA rbga.bin() from package `genalg` (see below):
+This binary coding is useful to use the binary GA `rbga.bin()` from package `genalg` (see below):
 
@@
     model <- lm(formula, data=df)
-    return( sqrt(mean(residuals(model)^2)) ) #rbga.bim tries to minimize
+    # rbga.bin() minimizes, so the rmse will be smaller as the interations advance
+    return( sqrt(mean(residuals(model)^2)) ) 
   }
@@
 
-#############################
+#################################
 # Unleash the chromosomes of war!
+#################################
 GAmodel <- rbga.bin(size = max.degree*n.vars*max.compts, 
-                    popSize = 100, 
+                    popSize = 150, 
                     iters = 100, 

### 2013-06-28 PaperDraft.Rmd
The first steps towards a paper.

Learning Polynomial Regression with Genetic Algorithms
======================================================

Francisco Coelho |  | João Neto
:---------------:|-----|:---------:
[`fc@di.uevora.pt`](mailto://fc@di.uevora.pt)| |[`jpn@di.fc.ul.pt`](mailto://jpn@di.fc.ul.pt)

Introduction
------------

_intro text_

### State of the Art

Em [1] vimos que...

Section 1
---------

References
----------

1. Hofwing, Magnus, Niclas Strömberg, and Martin Tapankov. "Optimal Polynomial Regression Models by using a Genetic Algorithm." Proceedings of the Second International Conference on Soft ComputingTechnology in Civil, Structural and Environmental Engineering Conference,(Crete, Greece), 2011009. 2011.

1. Wu, Chih-Hung, Gwo-Hshiung Tzeng, and Rong-Ho Lin. "A Novel hybrid genetic algorithm for kernel function and parameter optimization in support vector regression." Expert Systems with Applications 36.3 (2009): 4725-4735.

1. Gálvez, Akemi, Andrés Iglesias, and Jaime Puig-Pey. "Iterative two-step genetic-algorithm-based method for efficient polynomial B-spline surface reconstruction." Information Sciences 182.1 (2012): 56-76.

1. Ghai, Dhruva, Saraju P. Mohanty, and Garima Thakral. "Fast Analog Design Optimization using Regression based Modeling and Genetic Algorithm: A Nano-CMOS VCO Case Study." Proceedings of the 14th IEEE International Symposium on Quality Electronic Design (ISQED). 2013.

1. Rezania, Mohammad, Akbar A. Javadi, and Orazio Giustolisi. "Evaluation of liquefaction potential based on CPT results using evolutionary polynomial regression." Computers and Geotechnics 37.1 (2010): 82-92.

1. Zain, Azlan Mohd, Habibollah Haron, and Safian Sharif. "Genetic algorithm and simulated annealing to estimate optimal process parameters of the abrasive waterjet machining." Engineering with computers 27.3 (2011): 251-259.

1. Garg, A., and K. Tai. "Comparison of regression analysis, artificial neural network and genetic programming in handling the multicollinearity problem." Modelling, Identification & Control (ICMIC), 2012 Proceedings of International Conference on. IEEE, 2012.

### 2013-07-02 PaperDraft.Rmd
paper: initial introduction text.

@@
 
-_intro text_
+<!--Com notáveis excepções, a maioria das técnicas de _machine learning_ assenta em modelos lineares.
+Este pressuposto tem a seu favor inúmeras vantagens que abrangem, por exemplo, uma reduzida complexidade computacional e um forte enquadramento teórico.
+Embora por vezes os modelos lineares sejam adequados, restam muitos problemas não-lineares sem aproximações lineares aceitáveis, como pode ser constatado pelo número de artigos que descrevem aplicações de modelos não-lineares a problemas de engenharia, economia, linguística, produção animal, _etc_.
+
+Em complemento, os algoritmos genéticos proporcionam um esquema de optimização fácil de conceptualizar e de aplicação muito geral.
+Sendo necessário parameterizar um modelo não-linear, os algoritmos genéticos proporcionam um meio simples e frequentemente eficaz de encontrar aproximações de parâmetros localmente óptimos.
+
+Os polinómios proporcionam um dos casos mais simples de modelos não-lineares.
+Além disso as funções polinomiais constituem uma das áreas mais aprofundadas da matemática.
+Esta característica de limiar torna o estudo dos modelos polinomiais especialmente atractiva.
+-->
+With notable exceptions, most machine learning techniques are based on linear models.
+This assumption has many advantages in its favor spanning, for example, reduced computational complexity and strong theoretical framework.
+Though sometimes linear models are adequate, many nonlinear problems without acceptable linear approximations are left, as can be seen by the number of articles that describe applications of nonlinear models in engineering, economics, linguistics, animal production, _etc_.
+
+In addition, genetic algorithms provide an optimization scheme easy to conceptualize and with very broad application.
+If a nonlinear model requires parameterization, genetic algorithms provide a simple and often effective approach to search for local optimal parameters.
+
+Polynomials provide one of the simplest cases to nonlinear models.
+This threshold feature makes the study of polynomial models especially attractive when moving from linear to nonlinear models.
+Also the polynomial functions are one of the deepest studied subjects in mathematics.
 
@@
 
-Em [1] vimos que...
+1. _EXISTEM MUITAS APLICAÇÕES_
+2. _DESCRIÇÃO ESTRUTURAL (ÁRVORE DE PARSING) VS. COEFICIENTES DOS MONÓMIOS_
+3. _PRINCIPAIS PROBLEMAS E ACTUAIS PROPOSTAS DE RESOLUÇÃO_
 
@@
 ---------
+
+_texto para a section 1_
 

### 2013-07-02 PaperDraft.Rmd
paper: initial introduction text, 2

@@
 With notable exceptions, most machine learning techniques are based on linear models.
-This assumption has many advantages in its favor spanning, for example, reduced computational complexity and strong theoretical framework.
+This assumption has many advantages spanning, for example, reduced computational complexity and strong theoretical framework.
 Though sometimes linear models are adequate, many nonlinear problems without acceptable linear approximations are left, as can be seen by the number of articles that describe applications of nonlinear models in engineering, economics, linguistics, animal production, _etc_.
@@
 Also the polynomial functions are one of the deepest studied subjects in mathematics.
+
+In this setting, the use of genetic algorithms to search for optimal parameters of polynomial models is unavoidable.
+Indeed, many applied works can be found under this keywords, with special incidence on engineering and biology, where often the required models do not admit acceptable linear approximations.
+
+The specific nature of polynomials allows quite different representations of individual instances. For example, $$$p(x) = x^2 - 1 = (x-1)(x+1)$$$ can be described either by:
+
+* a list of coefficients: `[-1, 0, 1]`;
+* the product of two lower degree polynomials;
+* a parse tree: `sum(square(x),-1)`;
 

### 2013-07-02 gapoly.Rmd
included markdowns for 2 dataset tests

@@
 max.compts <- 4  # poly = compt1 + compt2 + ...
-max.degree <- 3
+max.degree <- 3  # meaning exponents up to 2^max.degree - 1
 ```
@@
 
-Eg: consider polynomial $x_1^2 \times x_3 + x_3^3 + x_1 \times x_2$
+Eg: consider polynomial $x_1^3 \times x_3 + x_3^7 + x_1 \times x_2$
 
@@
 ```{r}
+bin2dec <- function(x) 
+    sum(2^(which(rev(unlist(strsplit(as.character(x), "")) == 1))-1))
+
 make.component <- function(bits, n.vars) {
@@
   for(i in 1:length(elements)) {
-    power <- length(elements[[i]][elements[[i]]==1])
+    # power <- length(elements[[i]][elements[[i]]==1]) previous coding
+    power <- bin2dec(elements[[i]])
     if (power == 1)
@@
 
-Let's check the previous eg, $x_1^2 \times x_3 + x_3^3 + x_1 \times x_2$:
+Let's check the previous eg, $x_1^3 \times x_3 + x_3^7 + x_1 \times x_2$:
 

### 2013-07-02 tests_Ailerons.Rmd
included markdowns for 2 dataset tests

```{r, message = FALSE, tidy = FALSE, echo = F}
## knitr configuration: http://yihui.name/knitr/options#chunk_options
opts_chunk$set(comment = "", error= TRUE, warning = FALSE, message = FALSE,
               tidy = FALSE, cache = TRUE, echo = T,
               fig.width = 6, fig.height = 6)
```
Testing Ailerons
========================================================

```{r}
source("gapolyfuncs.R")

set.seed(101)

# Read Data

filename <- "Ailerons/ailerons"  # where is the data (under dataset folder)
filepath <- paste0("datasets/",filename)
my.data <- read.csv(filepath)

# Preprocess Data

my.data <- na.omit(my.data) # the dataset has some NAs
my.data <- my.data[,c(1:25,27:41)] # x26 is almost all zeros

names(my.data) <- c(paste0("x",1:(ncol(my.data)-1)),"y")

sapply(my.data, class)
my.data$x1 <- as.numeric(my.data$x1) # x1 was a factor, make it a numeric value
head(my.data,8)

# Execute ML methods

source("gapolyfuncs.R")
report <- make.report(my.data, population=200, iterations=200)

# See Results

coef(report$ga.model)
report$ga.error

boxplot(report$ga.error, 
        report$lm.error,
        report$svm.error,
        report$rpart.error,
        report$rf.error,
        report$citree.error,
        names=c("GA Poly", "Lin.Reg.", "SVM", "RPART", "Rnd.For.", "CI Tree"),
        ylab="RMSE")

df <- data.frame(ga.poly = report$ga.error, 
                 lin.reg = report$lm.error,
                 svm = report$svm.error,
                 rpart = report$rpart.error,
                 rf = report$rf.error,
                 ci.tree = report$citree.error)

save(report, file="Ailerons.Rda") # recover with load("Ailerons.Rda")

write.table(df, "Ailerons.txt")

# print(xtable(res.df), include.rownames = FALSE, floating = FALSE, type = "latex")
```

Tabela para Ailerons dataset
------

### 2013-07-02 tests_test.Rmd
included markdowns for 2 dataset tests

--- 2013-07-02 tests_Ailerons.Rmd
@@
 ```
-Testing Ailerons
+Testing Test
+
 ========================================================
@@
 ```{r}
-source("gapolyfuncs.R")
-
 set.seed(101)
@@
 
-filename <- "Ailerons/ailerons"  # where is the data (under dataset folder)
-filepath <- paste0("datasets/",filename)
-my.data <- read.csv(filepath)
+size <- 50
+my.data <- data.frame(x1 <- rpois(size,4),
+                      x2 <- rnorm(size),
+                      y <- x1*x2 + x1)
 
 # Preprocess Data
-
-my.data <- na.omit(my.data) # the dataset has some NAs
-my.data <- my.data[,c(1:25,27:41)] # x26 is almost all zeros
 
@@
 sapply(my.data, class)
-my.data$x1 <- as.numeric(my.data$x1) # x1 was a factor, make it a numeric value
+my.data$x1 <- as.numeric(my.data$x1) # x1 becomes a numeric value
 head(my.data,8)
@@
 source("gapolyfuncs.R")
-report <- make.report(my.data, population=200, iterations=200)
+report <- make.report(my.data, population=20, iterations=20)
 
@@
 
-df <- data.frame(ga.poly = report$ga.error, 
-                 lin.reg = report$lm.error,
-                 svm = report$svm.error,
-                 rpart = report$rpart.error,
-                 rf = report$rf.error,
-                 ci.tree = report$citree.error)
-
-save(report, file="Ailerons.Rda") # recover with load("Ailerons.Rda")
-
-write.table(df, "Ailerons.txt")
-
 # print(xtable(res.df), include.rownames = FALSE, floating = FALSE, type = "latex")
 ```
-
-Tabela para Ailerons dataset
-------

### 2013-07-02 old/gapolyfuncs - times.R

@@ -1,0 +1,170 @@
+#########################################################################
+# initial values, can be overriden
+n.vars     <- 3  # x1, x2,... 
+max.compts <- 4  # poly = compt1 + compt2 + ...
+max.degree <- 3  # meaning exponents up to 2^max.degree - 1
+
+#########################################################################
+# Root mean square error
+rsme <- function(v1, v2=c(0)) {
+  sqrt(mean((v1-v2)^2))
+}
+
+#########################################################################
+# convert a binary vector into an integer
+bin2dec <- function(x) 
+  sum(2^(which(rev(unlist(strsplit(as.character(x), "")) == 1))-1))
+
+#########################################################################
+# The chromosome coding will as follow:
+# + the chromosome is split into `max.compts` sets of bits of equal size 
+#   (a component)
+# + each component is split into `n.vars` sets of `max.degree` size each 
+#   (a variable)
+# + for each variable, the number of 1s gives the variable's degree 
+#   (this means that a given polynomial may have more than one possible 
+#   representation)
+# 
+# Eg: consider polynomial $x_1^3 \times x_3 + x_3^7 + x_1 \times x_2$
+# 
+# One possible coding -- for the values in the previous code snippet -- would be:
+# 
+# $$011,000,001;000,000,111;001,001,000;000,000,000$$
+# 
+# (the semicolons separate components, the commas seperate variables)
+# 
+# The next functions transform the binary vector into a polynomial formula 
+# that `lm()` understands:
+
+make.component <- function(bits, n.vars) {
+  vars <- paste0("x",1:n.vars)
+  elements <- split(bits, ceiling(seq_along(bits)/(length(bits)/n.vars)))
+  str <- ""
+  for(i in 1:length(elements)) {
+    power <- bin2dec(elements[[i]])
+    if (power == 1)
+      if (str=="")
+        str = vars[i]
+    else  
+      str = paste0(str, "*", vars[i])
+    if (power > 1)
+      if (str=="")
+        str = paste0(vars[i], "^", power)
+    else  
+      str = paste0(str, "*", vars[i], "^", power)
+  }
+  str
+}
+
+make.formula <- function(bits, n.vars, max.compts) {
+  components <- split(bits, ceiling(seq_along(bits)/(length(bits)/max.compts)))
+  str <- ""
[…]

### 2013-07-03 PaperDraft.Rmd
Some advances in the paper structure

--- 2013-06-28 PaperDraft.Rmd
@@
 
+**Abstract**
+
+Many applications require models that have no acceptable linear approximation and many nonlinear models are defined by polynomials.
+The use of genetic algorithms to find the parameters of a polynomial model is decades old but still poses challenges due to the complexity of the search and different definitions of "optimal" solution.
+
+GIVE A SUMMARY DESCRIPTION AND RESULTS OF "OUR" APPROACH.
+
 Introduction
@@
 
-_intro text_
+With notable exceptions, most machine learning techniques are based on linear models.
+This assumption has many advantages including, for example, reduced computational complexity and strong theoretical framework.
+However many application scenarios pose nonlinear problems that have no acceptable linear approximations.
+GIVE EXAMPLES AND EXPLAIN WHY LINEAR MODELS FAIL IN THOSE EXAMPLES.
 
-### State of the Art
+Polynomials, one of the most studied subjects in mathematics, generalize linear models and are, perhaps, one of the simplest and most used nonlinear models.
+Applications of (nonlinear) polynomial models include colorimetric calibration [APPL1-2005], finding explicit formulae for turbulent pipe flows [APPL2-1999], computational linguistics [APPL8-2009] and more recently, analytical techniques for cultural heritage materials [APPL3-2010], liquid epoxy molding process [APPL4-2011], B-spline surface reconstruction [APPL5-2012], product design [APPL6-2012] or forecasting cyanotoxins presence in water reservoirs [APPL7-2013].
+Besides the huge range of quite different areas, the work for each one of these polynomial models used, somewhere, a genetic algorithm.
 
-Em [1] vimos que...
+Genetic algorithms (GA), are perhaps one of the most popular "buzzwords" of recent decades but with good reason since they outline an optimization scheme easy to conceptualize and with very broad application.
+If a nonlinear model requires parameterization, GAs provide a simple and often effective approach to search for localy optimal parameters.
+Research related to genetic algorithms abound and spans from the 1950's seminal work of Nils Aall Barricelli in the Institute for Advanced Study of Princeton to today's principal area of study of thousands of researchers, covered in hundreds of conferences, workshops and other meetings.
+Perhaps the key impulse to GAs come from John Holland and his book [BOOK-JOHNHOLLAND].
+One interesting "flavour" of genetic algorithms, termed "genetic programming" by John Koza, proposed the use of GAs to search the syntatic stucture of complex functions.
+This syntatic structure search is keen to the central ideas of deep learning, the subarea of machine learning actually producing the most advanced results.
+INSERT MORE REFERENCES OF RECENT GA RESEARCH AND DEEPLEARNING RESULTS
 
-Section 1
----------
+**END** STATE-OF-THE-ART
+
+SUMMARIZE OUR PROBLEM STATEMENT, EXPLAIN WHY IT IS IMPORTANT, OUR RESOLUTION, EVALUATION STRATEGY AND RESULTS.
+
+Polynomial Regression with Genetic Algorithms
+---------------------------------------------
+
+HOW DO WE DO THE
+* polynomial encoding
+* cost function
+* genetic operators
+* selection of other search parameters (_eg_ population size, maxiter, _etc_)
+
+### Polinomial Encoding
+
+_Describe our method to encode polynomial instances_
+
+### Cost Function
+
+_Define the cost function_
+
+Given a dataset $D=\left\lbrace x_1^{(i)}, \ldots, x_m^{(i)}, y^{(i)} \in \mathbb{R}^{m+1} : i = 1,\ldots, n \right\rbrace$ with $n$ observations of $m+1$ variables $X_1, \ldots, X_m, Y$.
+We want to find a polynomial that fits $y = h\left( x_1, \ldots, x_m\right)$ and for that purpose we apply the usual _root-mean-square error_ as a cost function of a candidate fitting polynomial $h_{\Theta}$ defined by parameters $\Theta$:
+$$
+J_{fit}\left( \Theta; D \right) = \frac{1}{n} \sum_{i=1}^n \left( y^{(i)} - h_{\Theta}\left(x_1^{(i)}, \ldots, x_m^{(i)}\right) \right)^2
+$$
+
+### Genetic Operators
+
+_Define the genetic operators_
+
+### Other Search Parameters
+
+_Explain the remaining search parameters_
+
+Results
+-------
+
+* Measured quantities
+  - error
+  - number of iterations to convergence
+  - memory usage
+  - F1, ROC, ?  
+* Selection of datasets and regression algorithms
+* Summary Figures and Numeric results
+
+Conclusion
+----------
 
@@
 ----------
-
-1.
-Hofwing, Magnus, Niclas Strömberg, and Martin Tapankov.
-"Optimal Polynomial Regression Models by using a Genetic Algorithm." Proceedings of the Second International Conference on Soft ComputingTechnology in Civil, Structural and Environmental Engineering Conference,(Crete, Greece), 2011009.
-2011.
-
-1.
-Wu, Chih-Hung, Gwo-Hshiung Tzeng, and Rong-Ho Lin.
-"A Novel hybrid genetic algorithm for kernel function and parameter optimization in support vector regression." Expert Systems with Applications 36.3 (2009): 4725-4735.
-
-1.
-Gálvez, Akemi, Andrés Iglesias, and Jaime Puig-Pey.
-"Iterative two-step genetic-algorithm-based method for efficient polynomial B-spline surface reconstruction." Information Sciences 182.1 (2012): 56-76.
 
@@
 IEEE, 2012.
+
+[GAPOLY1-2011] http://dx.doi.org/10.4203/ccp.97.39 "Optimal Polynomial Regression Models by using a Genetic Algorithm"
+
+[GAPOLY2-2008] http://citeseerx.ist.psu.edu/viewdoc/summary?doi=10.1.1.152.7415 "Optimal Sampling of Genetic Algorithms on Polynomial Regression"
+
+[GAPOLY3-2011] http://www.eejournal.ktu.lt/index.php/elt/article/view/460 "Polynomial Curve Fitting with Varying Real Powers"
+
+[GAPOLY4-2009] http://dx.doi.org/10.1016/j.eswa.2008.06.046 "A Novel hybrid genetic algorithm for kernel function and parameter optimization in support vector regression"
+
+[GAPOLY5-2006] http://dx.doi.org/10.1007/s00500-005-0008-8 "Genetic polynomial regression as input selection algorithm for non-linear identification"
+
+[APPL1-2005]: http://dx.doi.org/10.1109/WISP.2005.1531626 "Adaptive polynomial regression for colorimetric scanner calibration using genetic algorithms"
+
+[APPL2-1999]: http://www.environmental-expert.com/Files/5302/articles/5912/art-4.pdf "Method for the identification of explicit polynomial formulae for the friction in turbulent pipe flow"
+
+[APPL3-2010]: http://discovery.ucl.ac.uk/459862/1/459862_Csefalvayova_2010_Talanta_EPS.pdf "Use of genetic algorithms with multivariate regression for determination of gelatine in historic papers based on FT-IR and NIR spectral data"
+
+[APPL4-2011]: http://dx.doi.org/10.1109/TII.2010.2100130 "Modeling of a liquid epoxy molding process using a particle swarm optimization-based fuzzy regression approach"
+
+[APPL5-2012]: http://dx.doi.org/10.1016/j.ins.2010.09.031 "Iterative two-step genetic-algorithm-based method for efficient polynomial B-spline surface reconstruction"
+
+[APPL6-2012]: http://dx.doi.org/10.1007/978-3-642-27476-3_6 "Development of Product Design Models Using Fuzzy Regression Based Genetic Programming"
+
+[APPL7-2013]: http://dx.doi.org/10.1016/j.envres.2013.01.001 "Hybrid modelling based on support vector regression with genetic algorithms in forecasting the cyanotoxins presence in the Trasona reservoir (Northern Spain)"
+
+[APPL8-2009]: http://dx.doi.org/10.1007/s00500-008-0362-4 "Obtaining linguistic fuzzy rule-based regression models from imprecise data with multiobjective genetic algorithms"
+
+[BOOK-JOHNHOLLAND]: http://mitpress.mit.edu/books/adaptation-natural-and-artificial-systems "Adaptation in Natural and Artificial Systems"

### 2013-07-03 tests_Kinematics.Rmd
add Kinematics tests

```{r, message = FALSE, tidy = FALSE, echo = F}
## knitr configuration: http://yihui.name/knitr/options#chunk_options
opts_chunk$set(comment = "", error= TRUE, warning = FALSE, message = FALSE,
               tidy = FALSE, cache = TRUE, echo = T,
               fig.width = 6, fig.height = 6)
```
Testing Kinematics
========================================================

```{r}
set.seed(101)

# Read Data

filename <- "Kinematics/kin8nm.data"  # where is the data (under dataset folder)
filepath <- paste0("datasets/",filename)
my.data <- read.csv(filepath, header=FALSE)

# Preprocess Data

# my.data <- na.omit(my.data) # the dataset has some NAs

names(my.data) <- c(paste0("x",1:(ncol(my.data)-1)),"y")

sapply(my.data, class)
# my.data$x1 <- as.numeric(my.data$x1) # x1 was a factor, make it a numeric value
head(my.data,8)

# Execute ML methods

source("gapolyfuncs.R")
report <- make.report(my.data, population=300, iterations=300)

# See Results

save(report, file="Kinematics300.Rda") # recover with load("Kinematics.Rda")

df <- data.frame(ga.poly = report$ga.error,
                 lin.reg = report$lm.error,
                 svm = report$svm.error,
                 rpart = report$rpart.error,
                 rf = report$rf.error,
                 ci.tree = report$citree.error)

coef(report$ga.model)
report$ga.error

library(ggplot2)

df <- data.frame(GA.Poly = report$ga.error,
                 Lin.Reg = report$lm.error,
                 SVM = report$svm.error,
                 RPART = report$rpart.error,
                 Rnd.For = report$rf.error,
                 CI.Tree = report$citree.error)

write.table(df, "Kinematics300.txt")

size <- length(report$ga.error)
list.data <- list(error=c(report$ga.error, 
                          report$lm.error,
                          report$svm.error,
                          report$rpart.error,
                          report$rf.error,
                          report$citree.error), 
                  method=c(rep("GA.Poly",size),
                           rep("Lin.Reg.",size),
                           rep("SVM",size), 
                           rep("RPART",size),
                           rep("Rnd.For.",size),
                           rep("CI.Tree",size)))

df <- as.data.frame(list.data)

write.table(df, "")

qplot(method,error,data=df,geom=c("jitter","boxplot")) + 
  ylab("root-mean-square error") + 
  xlab("regression method") +
  ggtitle("Kinematics dataset")

# print(xtable(res.df), include.rownames = FALSE, floating = FALSE, type = "latex")
```

### 2013-07-05 tests_test.Rmd
added regularization + parameter testing

@@
 my.data <- data.frame(x1 <- rpois(size,4),
-                      x2 <- rnorm(size),
-                      y <- x1*x2 + x1)
+                      x2 <- rnorm(size,10),
+                      x3 <- rpois(size,20),
+                      x4 <- rnorm(size,4,1),                      
+                      y <- x4^2*x2 + x1^2*x3 + 5)
 
@@
 source("gapolyfuncs.R")
-report <- make.report(my.data, population=20, iterations=20)
+report <- make.report(my.data, population=50, iterations=25, mutation.rate=0.05)
 
@@
 
-boxplot(report$ga.error, 
+# boxplot(report$ga.error, 
+#         report$lm.error,
+#         report$svm.error,
+#         report$rpart.error,
+#         report$rf.error,
+#         report$citree.error,
+#         names=c("GA Poly", "Lin.Reg.", "SVM", "RPART", "Rnd.For.", "CI Tree"),
+#         ylab="RMSE")
+
+# df <- data.frame(GA.Poly = report$ga.error,
+#                  Lin.Reg = report$lm.error,
+#                  SVM = report$svm.error,
+#                  RPART = report$rpart.error,
+#                  Rnd.For = report$rf.error,
+#                  CI.Tree = report$citree.error)
+# 
+# write.table(df, "Articial_dataset.txt")
+
+size <- length(report$ga.error)
+list.data <- list(error=c(report$ga.error, 
                           report$lm.error,
@@
                           report$rf.error,
-        report$citree.error,
-        names=c("GA Poly", "Lin.Reg.", "SVM", "RPART", "Rnd.For.", "CI Tree"),
-        ylab="RMSE")
+                          report$citree.error), 
+                  method=c(rep("GA.Poly",size),
+                           rep("Lin.Reg.",size),
+                           rep("SVM",size), 
+                           rep("RPART",size),
+                           rep("Rnd.For.",size),
+                           rep("CI.Tree",size)))
+
+df <- as.data.frame(list.data)
+
+library(ggplot2)
+qplot(method,error,data=df,geom=c("jitter","boxplot")) + 
+  ylab("root-mean-square error") + 
+  xlab("regression method") +
+  ggtitle("Artificial dataset")
+
+##############################################
+
+# Lambda Test
+lambda.values <- seq(1,2,length.out=20)
+n.runs <- 1
+
+source("gapolyfuncs.R")
+errors <- test.lambda(my.data, lambda.values, runs=n.runs, population=10, iterations=25)
+
+# df <- as.data.frame(errors)
+# names(df) <- round(lambda.values,3)
+
+df <- data.frame(error = as.vector(matrix(errors,nrow=1)),
+                 lambda = as.factor(rep(round(lambda.values,3), each=n.runs)))
+
+write.table(df, "lambda.errors.txt")
+
+library(ggplot2)
+qplot(lambda,error,data=df,geom=c("jitter","boxplot")) + 
+  ylab("root-mean-square error") + 
+  xlab("lambda value") +
+  coord_cartesian(ylim = c(0, 50))  +
+  ggtitle("Artificial dataset")
+
+library(plyr)
+best.errors <- 
+  colwise(function(x) min(as.numeric(x)))(as.data.frame(errors)) # get min for each column
+
+plot(lambda.values, best.errors, type="l")
+##############################################
 

### 2013-07-05 gapolyfuncs.R

@@ -1,8 +1,8 @@
 #########################################################################
 # initial values, can be overriden
-n.vars     <- 3  # x1, x2,... 
-max.compts <- 4  # poly = compt1 + compt2 + ...
-max.degree <- 3  # meaning exponents up to 2^max.degree - 1
+n.vars      <- 3  # x1, x2,... 
+max.monoids <- 4  # poly = monoid1 + monoid2 + ...
+max.degree  <- 3  # meaning exponents up to 2^max.degree - 1
 
 #########################################################################
 # Root mean square error
@@ -17,57 +17,63 @@ bin2dec <- function(x)
 
 #########################################################################
 # The chromosome coding will as follow:
-# + the chromosome is split into `max.compts` sets of bits of equal size 
-#   (a component)
-# + each component is split into `n.vars` sets of `max.degree` size each 
+# + an initial segment detailing which monoids are active (the 1st is always active)
+# + the chromosome is split into `max.monoids` sets of bits of equal size 
+#   (a monoid)
+# + each monoid is split into `n.vars` sets of `max.degree` size each 
 #   (a variable)
-# + for each variable, the number of 1s gives the variable's degree 
-#   (this means that a given polynomial may have more than one possible 
-#   representation)
+# + for each variable, the bits give the binary description of the variable's degree 
 # 
 # Eg: consider polynomial $x_1^3 \times x_3 + x_3^7 + x_1 \times x_2$
 # 
 # One possible coding -- for the values in the previous code snippet -- would be:
 # 
-# $$011,000,001;000,000,111;001,001,000;000,000,000$$
+# $$ 1111 011,000,001;000,000,111;001,001,000;000,000,000$$
 # 
-# (the semicolons separate components, the commas seperate variables)
+# (the semicolons separate monoids, the commas separate variables)
 # 
-# The next functions transform the binary vector into a polynomial formula 
+# The next two functions transform the binary vector into a polynomial formula 
 # that `lm()` understands:
 
-make.component <- function(bits, n.vars) {
-  vars <- paste0("x",1:n.vars)
+make.monoid <- function(bits, n.vars, vars) {
   elements <- split(bits, ceiling(seq_along(bits)/(length(bits)/n.vars)))
-  str <- ""
+  monoid <- ""
   for(i in 1:length(elements)) {
     power <- bin2dec(elements[[i]])
     if (power == 1)
[…]

### 2013-07-05 old/gapolyfuncs - without regularization.R

@@ -1,0 +1,171 @@
+#########################################################################
+# initial values, can be overriden
+n.vars     <- 3  # x1, x2,... 
+max.compts <- 4  # poly = compt1 + compt2 + ...
+max.degree <- 3  # meaning exponents up to 2^max.degree - 1
+
+#########################################################################
+# Root mean square error
+rsme <- function(v1, v2=c(0)) {
+  sqrt(mean((v1-v2)^2))
+}
+
+#########################################################################
+# convert a binary vector into an integer
+bin2dec <- function(x) 
+  sum(2^(which(rev(unlist(strsplit(as.character(x), "")) == 1))-1))
+
+#########################################################################
+# The chromosome coding will as follow:
+# + the chromosome is split into `max.compts` sets of bits of equal size 
+#   (a component)
+# + each component is split into `n.vars` sets of `max.degree` size each 
+#   (a variable)
+# + for each variable, the number of 1s gives the variable's degree 
+#   (this means that a given polynomial may have more than one possible 
+#   representation)
+# 
+# Eg: consider polynomial $x_1^3 \times x_3 + x_3^7 + x_1 \times x_2$
+# 
+# One possible coding -- for the values in the previous code snippet -- would be:
+# 
+# $$011,000,001;000,000,111;001,001,000;000,000,000$$
+# 
+# (the semicolons separate components, the commas seperate variables)
+# 
+# The next functions transform the binary vector into a polynomial formula 
+# that `lm()` understands:
+
+make.component <- function(bits, n.vars) {
+  vars <- paste0("x",1:n.vars)
+  elements <- split(bits, ceiling(seq_along(bits)/(length(bits)/n.vars)))
+  str <- ""
+  for(i in 1:length(elements)) {
+    power <- bin2dec(elements[[i]])
+    if (power == 1)
+      if (str=="")
+        str = vars[i]
+    else  
+      str = paste0(str, "*", vars[i])
+    if (power > 1)
+      if (str=="")
+        str = paste0(vars[i], "^", power)
+    else  
+      str = paste0(str, "*", vars[i], "^", power)
+  }
+  str
+}
+
+make.formula <- function(bits, n.vars, max.compts) {
+  components <- split(bits, ceiling(seq_along(bits)/(length(bits)/max.compts)))
+  str <- ""
[…]

### 2013-07-05 tests_test.Rmd
added monitorFunction

@@
 source("gapolyfuncs.R")
-report <- make.report(my.data, population=50, iterations=25, mutation.rate=0.05)
+report <- make.report(my.data, population=50, iterations=25, mutation.rate=0.05, runs=1)
 
@@
 source("gapolyfuncs.R")
-errors <- test.lambda(my.data, lambda.values, runs=n.runs, population=10, iterations=25)
+errors <- test.lambda(my.data, lambda.values, runs=n.runs, population=10, iterations=10)
 

### 2013-07-05 gapolyfuncs.R

@@ -1,6 +1,5 @@
 #########################################################################
 # initial values, can be overriden
-n.vars      <- 3  # x1, x2,... 
 max.monoids <- 4  # poly = monoid1 + monoid2 + ...
 max.degree  <- 3  # meaning exponents up to 2^max.degree - 1
 
@@ -96,12 +95,30 @@ evalFuncFactory <- function(df, n.vars, max.monoids, lambda=1.05) {
   }
 }
 
+monitorEvalFactory <- function(n.vars, max.monoids) {
+
+  function(obj) {
+    minEval <- min(obj$evaluations)
+    filter  <- obj$evaluations == minEval
+    bestObjectCount = sum(rep(1, obj$popSize)[filter]);
+    # ok, deal with the situation that more than one object is best
+    if (bestObjectCount > 1) {
+      bestSolution = obj$population[filter,][1,];
+    } else {
+      bestSolution = obj$population[filter,];
+    }
+    print(bestSolution)
+    print(make.formula(bestSolution, n.vars, max.monoids))
+  }
+}
+
 #########################################################################
 # Returns a dataset with the rsme for each one of the tested ML methods
 # including GA Poly
 
 # pre: both sets must have columns named x1...xn and the last one is y (the output)
 make.report <- function(my.data, 
+                        n.vars,
                         population=100, 
                         iterations=25, 
                         runs=10, 
@@ -133,12 +150,13 @@ make.report <- function(my.data,
     train.set <- my.data[inTrain,]
     test.set  <- my.data[-inTrain,]
     
-    GAmodel <- rbga.bin(size = max.degree*n.vars*max.monoids, 
+    GAmodel <- rbga.bin(size = max.monoids + max.degree*n.vars*max.monoids, 
                         popSize = population, 
                         iters = iterations, 
                         mutationChance = mutation.rate, 
                         elitism = TRUE, 
-                        evalFunc = evalFuncFactory(train.set, n.vars,max.monoids))
+                        evalFunc = evalFuncFactory(train.set, n.vars, max.monoids),
+                        monitorFunc = monitorEvalFactory(n.vars, max.monoids))
       
     best.solution <- GAmodel$population[1,]
     best.formula <- paste0("y ~ ", make.formula(best.solution, n.vars, max.monoids))
@@ -212,7 +230,7 @@ test.lambda <- function(my.data, lambda.values,
       train.set <- my.data[inTrain,]
       test.set  <- my.data[-inTrain,]
       
[…]

### 2013-07-05 tests_test.Rmd
added follow.fitness()

@@
 ##############################################
+# Check fitness progress
 
+source("gapolyfuncs.R")
+follow.fitness(my.data,population=200,iterations=50)
+
+# output:
+
+progress <- c(78.3925,78.3925,71.30467,71.30467,60.09713,20.41959,15.00416,15.00416,15.00416,15.00416,15.00416,15.00416,12.31946,12.31946,12.31946,12.31946,2.971683,2.971683,2.971683,2.971683,6.569273e-14,6.569273e-14,6.500183e-14,6.500183e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14,3.573077e-14)
+
+df <- data.frame(iteration = 1:length(progress),
+                 fitness = as.numeric(progress))
+
+qplot(iteration, fitness, data=df) +
+  geom_line() +
+  ylim(0, max(df$fitness)) +
+  ggtitle("Artificial dataset (population 200)")
+
+##############################################
 # print(xtable(res.df), include.rownames = FALSE, floating = FALSE, type = "latex")

### 2013-07-05 gapolyfuncs.R

@@ -95,23 +95,6 @@ evalFuncFactory <- function(df, n.vars, max.monoids, lambda=1.05) {
   }
 }
 
-monitorEvalFactory <- function(n.vars, max.monoids) {
-
-  function(obj) {
-    minEval <- min(obj$evaluations)
-    filter  <- obj$evaluations == minEval
-    bestObjectCount = sum(rep(1, obj$popSize)[filter]);
-    # ok, deal with the situation that more than one object is best
-    if (bestObjectCount > 1) {
-      bestSolution = obj$population[filter,][1,];
-    } else {
-      bestSolution = obj$population[filter,];
-    }
-    print(bestSolution)
-    print(make.formula(bestSolution, n.vars, max.monoids))
-  }
-}
-
 #########################################################################
 # Returns a dataset with the rsme for each one of the tested ML methods
 # including GA Poly
@@ -155,8 +138,7 @@ make.report <- function(my.data,
                         iters = iterations, 
                         mutationChance = mutation.rate, 
                         elitism = TRUE, 
-                        evalFunc = evalFuncFactory(train.set, n.vars, max.monoids),
-                        monitorFunc = monitorEvalFactory(n.vars, max.monoids))
+                        evalFunc = evalFuncFactory(train.set, n.vars, max.monoids))
       
     best.solution <- GAmodel$population[1,]
     best.formula <- paste0("y ~ ", make.formula(best.solution, n.vars, max.monoids))
@@ -249,4 +231,58 @@ test.lambda <- function(my.data, lambda.values,
   }
   
   errors  
+}
+
+#########################################################################
+# Follow fitness progress of GA
+
+fitness.progress <- c(0)
+
+follow.fitness <- function(my.data,
+                           population=100, 
+                           iterations=20, 
+                           mutation.rate=0.05,
+                           verbose=TRUE) {
+  
+  library(genalg)
+  
+  train.p.size <- 0.7 # percentage of training set
+  n.vars       <- ncol(my.data)-1
+  
+  # make train & test set
+  inTrain   <- sample(1:nrow(my.data), train.p.size * nrow(my.data))
+  train.set <- my.data[inTrain,]
+  test.set  <- my.data[-inTrain,]
+  
+  GAmodel <- rbga.bin(size = max.monoids + max.degree*n.vars*max.monoids, 
+                      popSize = population, 
+                      iters = iterations, 
[…]

### 2013-07-05 tests_Kinematics.Rmd
added regularization + parameter testing
some Kinematic results

@@
 
+########################################################
+# Lambda Test
+lambda.values <- seq(0.5,0.999,length.out=20)
+n.runs <- 5
+
+source("gapolyfuncs.R")
+errors <- test.lambda(my.data, lambda.values, runs=n.runs, population=50, iterations=25)
+
+df <- data.frame(error = as.vector(matrix(errors,nrow=1)),
+                 lambda = as.factor(rep(round(lambda.values,3), each=n.runs)))
+
+write.table(df, "lambda.errors.kinematics.txt")
+
+library(ggplot2)
+qplot(lambda,error,data=df,geom=c("jitter","boxplot")) + 
+  ylab("root-mean-square error") + 
+  xlab("lambda value") +
+ # coord_cartesian(ylim = c(0, 50))  +
+  ggtitle("Kinematics dataset")
+
+########################################################
+
 # print(xtable(res.df), include.rownames = FALSE, floating = FALSE, type = "latex")

### 2013-07-05 tests_Kinematics.Rmd
some bugs removed

@@
 source("gapolyfuncs.R")
-report <- make.report(my.data, population=300, iterations=300)
+report <- make.report(my.data, population=200, iterations=100)
 
@@
 
-save(report, file="Kinematics300.Rda") # recover with load("Kinematics.Rda")
-
-df <- data.frame(ga.poly = report$ga.error,
-                 lin.reg = report$lm.error,
-                 svm = report$svm.error,
-                 rpart = report$rpart.error,
-                 rf = report$rf.error,
-                 ci.tree = report$citree.error)
+save(report, file="Kinematics200.Rda") # recover with load("Kinematics.Rda")
 
@@
 
-write.table(df, "lambda.errors.kinematics.txt")
+write.table(df, "lambda.errors.kinematics2.txt")
+
+df <- read.table("lambda.errors.kinematics2.txt", header=TRUE)
+
+df$lambda <- as.factor(df$lambda)
 
@@
   ggtitle("Kinematics dataset")
+##############################################
+# Check fitness progress
+
+source("gapolyfuncs.R")
+follow.fitness(my.data,population=300,iterations=50,mutation.rate=0.15,lambda=0.95)
 

### 2013-07-05 gapolyfuncs.R

@@ -83,7 +83,7 @@ make.formula <- function(bits, n.vars, max.monoids) {
 # variable and the max number of componentes, all required information
 # to perform the polynomial regression and then compute the residuals' rsme 
 # which will be the fitness of the respective chromosome
-evalFuncFactory <- function(df, n.vars, max.monoids, lambda=1.05) {
+evalFuncFactory <- function(df, n.vars, max.monoids, lambda=1) {
   
   function(chromosome) {
     formula <- paste0("y ~ ", make.formula(chromosome, n.vars, max.monoids))
@@ -105,7 +105,8 @@ make.report <- function(my.data,
                         population=100, 
                         iterations=25, 
                         runs=10, 
-                        mutation.rate=0.05) {
+                        mutation.rate=0.05,
+                        verbose=TRUE) {
   
   library(genalg)
   library(e1071)
@@ -142,7 +143,7 @@ make.report <- function(my.data,
       
     best.solution <- GAmodel$population[1,]
     best.formula <- paste0("y ~ ", make.formula(best.solution, n.vars, max.monoids))
-    ga.model <- lm(best.formula, data=my.data)
+    ga.model <- lm(best.formula, data=train.set)
     ga.pred  <- predict(ga.model, test.set[,-ncol(test.set)]) 
     ga.error[i] <- rsme(ga.pred,test.set[,ncol(test.set)])
     
@@ -171,6 +172,9 @@ make.report <- function(my.data,
     citree.model <- ctree(y ~ ., data = train.set)
     citree.pred  <- predict(citree.model, test.set[,-ncol(test.set)]) 
     citree.error[i] <- rsme(citree.pred,test.set[,ncol(test.set)])    
+    
+    if (verbose)
+      cat(i)
   }
   
   list(ga.error=ga.error,         # make the errors report into a list
@@ -221,7 +225,7 @@ test.lambda <- function(my.data, lambda.values,
       
       best.solution <- GAmodel$population[1,]
       best.formula <- paste0("y ~ ", make.formula(best.solution, n.vars, max.monoids))
-      ga.model <- lm(best.formula, data=my.data)
+      ga.model <- lm(best.formula, data=test.set)
       ga.pred  <- predict(ga.model, test.set[,-ncol(test.set)]) 
 
       errors[i,k] <- rsme(ga.pred,test.set[,ncol(test.set)])
@@ -242,6 +246,7 @@ follow.fitness <- function(my.data,
                            population=100, 
                            iterations=20, 
[…]

### 2013-07-06 PaperDraft.Rmd
progressing the paper draft

@@
 This assumption has many advantages including, for example, reduced computational complexity and strong theoretical framework.
-However many application scenarios pose nonlinear problems that have no acceptable linear approximations.
-GIVE EXAMPLES AND EXPLAIN WHY LINEAR MODELS FAIL IN THOSE EXAMPLES.
+However many application scenarios pose nonlinear problems without acceptable linear approximations.
+Nonlinear models usually occur in systems with feedback loops so common in ecology, cybernetics, robotics and may other areas.
 
@@
 This syntatic structure search is keen to the central ideas of deep learning, the subarea of machine learning actually producing the most advanced results.
-INSERT MORE REFERENCES OF RECENT GA RESEARCH AND DEEPLEARNING RESULTS
+INSERT MORE REFERENCES OF RECENT GA RESEARCH, STRUCTURE LEARNING AND DEEP-LEARNING RESULTS
 

### 2013-07-06 PaperDraft.Rmd
more advances in the paper.

--- 2013-07-03 PaperDraft.Rmd
@@
-Learning Polynomial Regression with Genetic Algorithms
-======================================================
+# Learning Polynomial Regression with Genetic Algorithms
 
@@
 The use of genetic algorithms to find the parameters of a polynomial model is decades old but still poses challenges due to the complexity of the search and different definitions of "optimal" solution.
+This work describes a general method based in genetic algorithms to find "empirical" polynomial regressions.
 
@@
 
-Introduction
-------------
+## Introduction
 
-With notable exceptions, most machine learning techniques are based on linear models.
+With notable exceptions (_e.g._ $k$-means), most machine learning techniques are based on linear models.
 This assumption has many advantages including, for example, reduced computational complexity and strong theoretical framework.
-However many application scenarios pose nonlinear problems that have no acceptable linear approximations.
-GIVE EXAMPLES AND EXPLAIN WHY LINEAR MODELS FAIL IN THOSE EXAMPLES.
+However nonlinearity in unavoidable in many application scenarions or, _e.g._ in systems with feedback loops so common in ecology, cybernetics, robotics and other areas.
 
 Polynomials, one of the most studied subjects in mathematics, generalize linear models and are, perhaps, one of the simplest and most used nonlinear models.
-Applications of (nonlinear) polynomial models include colorimetric calibration [APPL1-2005], finding explicit formulae for turbulent pipe flows [APPL2-1999], computational linguistics [APPL8-2009] and more recently, analytical techniques for cultural heritage materials [APPL3-2010], liquid epoxy molding process [APPL4-2011], B-spline surface reconstruction [APPL5-2012], product design [APPL6-2012] or forecasting cyanotoxins presence in water reservoirs [APPL7-2013].
+Applications of (nonlinear) polynomial models include colorimetric calibration ([APPL1-2005]), finding explicit formulae for turbulent pipe flows ([APPL2-1999]), computational linguistics ([APPL8-2009]) and more recently, analytical techniques for cultural heritage materials ([APPL3-2010]), liquid epoxy molding process ([APPL4-2011]), B-spline surface reconstruction ([APPL5-2012]), product design ([APPL6-2012]) or forecasting cyanotoxins presence in water reservoirs ([APPL7-2013]).
 Besides the huge range of quite different areas, the work for each one of these polynomial models used, somewhere, a genetic algorithm.
@@
 Genetic algorithms (GA), are perhaps one of the most popular "buzzwords" of recent decades but with good reason since they outline an optimization scheme easy to conceptualize and with very broad application.
-If a nonlinear model requires parameterization, GAs provide a simple and often effective approach to search for localy optimal parameters.
-Research related to genetic algorithms abound and spans from the 1950's seminal work of Nils Aall Barricelli in the Institute for Advanced Study of Princeton to today's principal area of study of thousands of researchers, covered in hundreds of conferences, workshops and other meetings.
-Perhaps the key impulse to GAs come from John Holland and his book [BOOK-JOHNHOLLAND].
-One interesting "flavour" of genetic algorithms, termed "genetic programming" by John Koza, proposed the use of GAs to search the syntatic stucture of complex functions.
-This syntatic structure search is keen to the central ideas of deep learning, the subarea of machine learning actually producing the most advanced results.
-INSERT MORE REFERENCES OF RECENT GA RESEARCH AND DEEPLEARNING RESULTS
+If a nonlinear model requires parameterization GAs provide a simple and often effective approach to search for localy optimal parameters.
+Research related to genetic algorithms abound and spans from the 1950's seminal work of Nils Aall Barricelli ([BARRICELLI-1962]) in the Institute for Advanced Study of Princeton to today's principal area of study of thousands of researchers, covered in hundreds of conferences, workshops and other meetings.
+Perhaps the key impulse to GAs come from John Holland and his book ([BOOK-JOHNHOLLAND]).
 
-**END** STATE-OF-THE-ART
+One interesting "flavour" of genetic algorithms, termed "genetic programming" by John Koza ([KOZA1-1992]), proposed the use of GAs to search the syntatic stucture of complex functions.
+This syntatic structure search is keen to the central ideas of deep learning ([DEEPLEARNING1-2012] and [DEEPLEARNING2-2009]), the subarea of machine learning actually producing the most promising results.
+It is also related to the work presented in this paper in the sense that, unlike linear models that have a simple structure ($y=\sum_i \beta_i x_i$) nonlinear (in particular polynomial) models pose an aditional "structure" search problem.
 
-SUMMARIZE OUR PROBLEM STATEMENT, EXPLAIN WHY IT IS IMPORTANT, OUR RESOLUTION, EVALUATION STRATEGY AND RESULTS.
+The idea of using GAs to find a polynomial regression is not new ([GAPOLY5-2006], [GAPOLY2-2008] and [GAPOLY4-2009]) but still generates original research ([GAPOLY1-2011] and [GAPOLY3-2011]).
+In line with that research this work describes a general method to find a polynomial regression of a given dataset.
+The optimal regression minimizes a cost function that accounts for both the _mean square error_ and a _regularization_ factor to avoid overfitting by penalizing polynomials that are "too complex".
 
-Polynomial Regression with Genetic Algorithms
----------------------------------------------
+A method that produces complex models "directly" from observed data has many uses.
+For example by a scientist to better understand the source of the data or by an autonomous agents adapting to the environment.
+MORE REASONS?
+
+[OUR RESOLUTION]  
+We represent a polynomial by a sum of monomials
+
+$$
+p\left( x_1, \ldots, x_n\right) = \sum_i \beta_i m_i
+$$
+where each monomial is $m_i = \prod_{j\in M_i} x_j^{\alpha_{ij}}$ the exponents are $\alpha_{ij}\in\mathbb{N}_0$ and the coeficients are $\beta_i \in \mathbb{R}$. For example 
+$$
+p\left( x_1, x_2, x_3 \right) = 2 x_1 + x_2 x_3 + \frac{1}{2} x_1^2 x_3
+$$
+has monomials $m_1 = x_1, m_2 = x_2 x_3$ and $m_3 = x_1^2 x_3$ coefficients $\beta_1 = 2, \beta_2 = 1$ and $\beta_3 = 1/2$ and exponents $\alpha_{1,1} = 1, \alpha_{2,2}=1, \alpha_{2,3} = 1, \alpha_{3,1} = 2, \alpha_{3,3} = 1$ and all other $\alpha_{ij} = 0$.
+
+This makes the problem of structure search very clear: except for the trivial cases, the number of possible monomials given $n$ variables and a maximum joint degree $d$ grows exponentialy with either $n$ or $d$.
+More importantly, the polynomial regression problem can be split into two subproblems:
+
+1. for a given set of monomials $q_1, \ldots, q_n$, find regression coefficients $\beta_1,\ldots,\beta_n$ that minimize an adequate cost function;
+2. find the fittest set of monomials;
+
+Our line of work is to solve the first problem with linear regression and the second with GAs.
+
+[EVALUATION STRATEGY]
+
+[RESULTS]
+
+## Polynomial Regression with Genetic Algorithms
 
 HOW DO WE DO THE
+
 * polynomial encoding
@@
 
-Results
--------
+## Results
 
@@
 
-Conclusion
-----------
+## Conclusion
 
-References
-----------
+## References
+
+**the references are defined "here" in the source but hidden in the resulting documents**
 
@@
 
-[GAPOLY1-2011] http://dx.doi.org/10.4203/ccp.97.39 "Optimal Polynomial Regression Models by using a Genetic Algorithm"
+[GAPOLY1-2011]: http://dx.doi.org/10.4203/ccp.97.39 "Optimal Polynomial Regression Models by using a Genetic Algorithm"
 
-[GAPOLY2-2008] http://citeseerx.ist.psu.edu/viewdoc/summary?doi=10.1.1.152.7415 "Optimal Sampling of Genetic Algorithms on Polynomial Regression"
+[GAPOLY2-2008]: http://citeseerx.ist.psu.edu/viewdoc/summary?doi=10.1.1.152.7415 "Optimal Sampling of Genetic Algorithms on Polynomial Regression"
 
-[GAPOLY3-2011] http://www.eejournal.ktu.lt/index.php/elt/article/view/460 "Polynomial Curve Fitting with Varying Real Powers"
+[GAPOLY3-2011]: http://www.eejournal.ktu.lt/index.php/elt/article/view/460 "Polynomial Curve Fitting with Varying Real Powers"
 
-[GAPOLY4-2009] http://dx.doi.org/10.1016/j.eswa.2008.06.046 "A Novel hybrid genetic algorithm for kernel function and parameter optimization in support vector regression"
+[GAPOLY4-2009]: http://dx.doi.org/10.1016/j.eswa.2008.06.046 "A Novel hybrid genetic algorithm for kernel function and parameter optimization in support vector regression"
 
-[GAPOLY5-2006] http://dx.doi.org/10.1007/s00500-005-0008-8 "Genetic polynomial regression as input selection algorithm for non-linear identification"
+[GAPOLY5-2006]: http://dx.doi.org/10.1007/s00500-005-0008-8 "Genetic polynomial regression as input selection algorithm for non-linear identification"
 
@@
 [BOOK-JOHNHOLLAND]: http://mitpress.mit.edu/books/adaptation-natural-and-artificial-systems "Adaptation in Natural and Artificial Systems"
+
+[DEEPLEARNING1-2012]: http://arxiv.org/abs/1206.5538 "Representation Learning: A Review and New Perspectives"
+
+[DEEPLEARNING2-2009]: http://www.iro.umontreal.ca/~lisa/publications2/index.php/publications/show/239 "Learning Deep Architectures for AI"
+
+[KOZA1-1992]: http://mitpress.mit.edu/books/genetic-programming "Genetic Programming: On the Programming of Computers by Means of Natural Selection, by John R. Koza, (1992)"
+
+[BARRICELLI-1962]: http://dx.doi.org/10.1007%2FBF01556771 "Numerical testing of evolution theories : Part I Theoretical introduction and basic tests"

### 2013-07-06 PaperDraft.Rmd
more advances in the paper -- Introduction is almost complete and entering section #2

@@
 
-With notable exceptions (_e.g._ $k$-means), most machine learning techniques are based on linear models.
+With notable exceptions (_e.g._ neural networks), most machine learning regression techniques are based on linear models.
 This assumption has many advantages including, for example, reduced computational complexity and strong theoretical framework.
 However nonlinearity in unavoidable in many application scenarions or, _e.g._ in systems with feedback loops so common in ecology, cybernetics, robotics and other areas.
+Nevertheless the variety and number of phenomena that can be adapted into a linear model is amazing.
 
-Polynomials, one of the most studied subjects in mathematics, generalize linear models and are, perhaps, one of the simplest and most used nonlinear models.
-Applications of (nonlinear) polynomial models include colorimetric calibration ([APPL1-2005]), finding explicit formulae for turbulent pipe flows ([APPL2-1999]), computational linguistics ([APPL8-2009]) and more recently, analytical techniques for cultural heritage materials ([APPL3-2010]), liquid epoxy molding process ([APPL4-2011]), B-spline surface reconstruction ([APPL5-2012]), product design ([APPL6-2012]) or forecasting cyanotoxins presence in water reservoirs ([APPL7-2013]).
+Polynomials, one of the most studied subjects in mathematics, generalize linear functions and define, perhaps, the simplest and most used nonlinear models.
+Applications of polynomial include colorimetric calibration ([APPL1-2005]), explicit formulae for turbulent pipe flows ([APPL2-1999]), computational linguistics ([APPL8-2009]) and more recently, analytical techniques for cultural heritage materials ([APPL3-2010]), liquid epoxy molding process ([APPL4-2011]), B-spline surface reconstruction ([APPL5-2012]), product design ([APPL6-2012]) or forecasting cyanotoxins presence in water reservoirs ([APPL7-2013]).
 Besides the huge range of quite different areas, the work for each one of these polynomial models used, somewhere, a genetic algorithm.
 
-Genetic algorithms (GA), are perhaps one of the most popular "buzzwords" of recent decades but with good reason since they outline an optimization scheme easy to conceptualize and with very broad application.
-If a nonlinear model requires parameterization GAs provide a simple and often effective approach to search for localy optimal parameters.
+Genetic algorithms (GA) where, arguably, one the most popular "hot" topics of research in the recent decades but with good reason since they outline an optimization scheme easy to conceptualize and with very broad application.
+If a nonlinear (or otherwise) model requires parameterization GAs provide a simple and often effective approach to search for localy optimal parameters.
 Research related to genetic algorithms abound and spans from the 1950's seminal work of Nils Aall Barricelli ([BARRICELLI-1962]) in the Institute for Advanced Study of Princeton to today's principal area of study of thousands of researchers, covered in hundreds of conferences, workshops and other meetings.
-Perhaps the key impulse to GAs come from John Holland and his book ([BOOK-JOHNHOLLAND]).
+Perhaps the key impulse to GAs come from John Holland's work and his book ([BOOK-JOHNHOLLAND]).
 
-One interesting "flavour" of genetic algorithms, termed "genetic programming" by John Koza ([KOZA1-1992]), proposed the use of GAs to search the syntatic stucture of complex functions.
-This syntatic structure search is keen to the central ideas of deep learning ([DEEPLEARNING1-2012] and [DEEPLEARNING2-2009]), the subarea of machine learning actually producing the most promising results.
+One interesting "flavour" of genetic algorithms, named _genetic programming_ by John Koza ([KOZA1-1992]), proposed the use of GAs to search the syntatic stucture of complex functions.
+This syntatic structure search is keen to the central ideas of deep learning ([DEEPLEARNING1-2012] and [DEEPLEARNING2-2009]), the subarea of machine learning actually producing the most promising results (e.g.
+[DLAPPL1-2013]).
 It is also related to the work presented in this paper in the sense that, unlike linear models that have a simple structure ($y=\sum_i \beta_i x_i$) nonlinear (in particular polynomial) models pose an aditional "structure" search problem.
@@
 
-A method that produces complex models "directly" from observed data has many uses.
+A method that produces adequate models "directly" from observed complex data has many uses.
 For example by a scientist to better understand the source of the data or by an autonomous agents adapting to the environment.
@@
 
-[OUR RESOLUTION]  
-We represent a polynomial by a sum of monomials
+RESULTS OUTLINE
 
+The remainder of this paper is organized as usual: the next section describes the details of our method and is followed by a presentation of some performance results.
+The last section draws some conclusions and points future research tasks.
+
+## Polynomial Regression with Genetic Algorithms
+
+The canonical representation of a polynomial is a sum
 $$
-p\left( x_1, \ldots, x_n\right) = \sum_i \beta_i m_i
+p\left( x_1, \ldots, x_n\right) = \sum_i \beta_i m_i.
 $$
-where each monomial is $m_i = \prod_{j\in M_i} x_j^{\alpha_{ij}}$ the exponents are $\alpha_{ij}\in\mathbb{N}_0$ and the coeficients are $\beta_i \in \mathbb{R}$. For example 
-$$
-p\left( x_1, x_2, x_3 \right) = 2 x_1 + x_2 x_3 + \frac{1}{2} x_1^2 x_3
-$$
-has monomials $m_1 = x_1, m_2 = x_2 x_3$ and $m_3 = x_1^2 x_3$ coefficients $\beta_1 = 2, \beta_2 = 1$ and $\beta_3 = 1/2$ and exponents $\alpha_{1,1} = 1, \alpha_{2,2}=1, \alpha_{2,3} = 1, \alpha_{3,1} = 2, \alpha_{3,3} = 1$ and all other $\alpha_{ij} = 0$.
+
+In this sum each $m_i$ is a monomial, $m_i = \prod_{j\in M_i} x_j^{\alpha_{ij}}$, the exponents are non-negative integers, $\alpha_{ij}\in\mathbb{N}_0$, and the coeficients are real valued, $\beta_i \in \mathbb{R}$.
+For example $p\left( x_1, x_2, x_3 \right) = 2 x_1 + x_2 x_3 + \frac{1}{2} x_1^2 x_3$ has monomials $m_1 = x_1, m_2 = x_2 x_3$ and $m_3 = x_1^2 x_3$ coefficients $\beta_1 = 2, \beta_2 = 1$ and $\beta_3 = 1/2$ and exponents $\alpha_{1,1} = 1, \alpha_{2,2}=1, \alpha_{2,3} = 1, \alpha_{3,1} = 2, \alpha_{3,3} = 1$ and all other $\alpha_{ij} = 0$.
+
+> **NOTA** Este parágrafo deu-me uma ideia: os $\alpha$ definem uma matriz em que as linhas são monómios e as colunas variáveis; "Multiplicando" esta matriz por um vector, obtemos o polinómio.
+Além disso, podemos continuar a definir a condição de regularização como antes, mas também podemos considerar condições *na* matriz (por exemplo, o número de entradas não nulas ser esparso ou a sua soma ser dominada por $\log nm$)...
 
 This makes the problem of structure search very clear: except for the trivial cases, the number of possible monomials given $n$ variables and a maximum joint degree $d$ grows exponentialy with either $n$ or $d$.
-More importantly, the polynomial regression problem can be split into two subproblems:
+But more importantly, the polynomial regression problem can be split into two subproblems:
 
@@
 [RESULTS]
-
-## Polynomial Regression with Genetic Algorithms
 
@@
 [BARRICELLI-1962]: http://dx.doi.org/10.1007%2FBF01556771 "Numerical testing of evolution theories : Part I Theoretical introduction and basic tests"
+
+[DLAPPL1-2013]: http://www.cs.toronto.edu/~dtarlow/TarSwerCharSutZem_ICML2013.pdf "Tarlow, Daniel and Sutskever, Ilya and Zemel, Richard S.
+Stochastic k-Neighborhood Selection for Supervised and Unsupervised Learning.
+2013.
+in Proceedings of the 30th International Conference on Machine Learning (ICML)."

### 2013-07-08 texts/PaperDraft.Rmd
Further writting in the second section.

--- 2013-07-06 PaperDraft.Rmd
@@
-# Learning Polynomial Regression with Genetic Algorithms
+# Exploring Polynomial Structure of Data with Genetic Algorithms
 
@@
 
+>Antes de submeter, temos de considerar [Polynomial regression na wikipedia](http://en.wikipedia.org/wiki/Polynomial_regression) e a [regressão não-linear em geral](https://www.google.pt/search?q=nonlinear+regression).
+>
+>Como este trabalho anda no meio de áreas muito trabalhadas, temos de ter muito cuidado com a questão da **investigação original**.
+
 **Abstract**
@@
 Many applications require models that have no acceptable linear approximation and many nonlinear models are defined by polynomials.
-The use of genetic algorithms to find the parameters of a polynomial model is decades old but still poses challenges due to the complexity of the search and different definitions of "optimal" solution.
+The use of genetic algorithms to find polynomial models is decades old but still poses challenges due to the complexity of the search and different definitions of "optimal" solution.
 This work describes a general method based in genetic algorithms to find "empirical" polynomial regressions.
@@
 This assumption has many advantages including, for example, reduced computational complexity and strong theoretical framework.
-However nonlinearity in unavoidable in many application scenarions or, _e.g._ in systems with feedback loops so common in ecology, cybernetics, robotics and other areas.
+However nonlinearity in unavoidable in many application scenarions _e.g._ phase transitions or systems with feedback loops so common in ecology, cybernetics, robotics and other areas.
 Nevertheless the variety and number of phenomena that can be adapted into a linear model is amazing.
@@
 $$
-p\left( x_1, \ldots, x_n\right) = \sum_i \beta_i m_i.
+p\left( x_1, \ldots, x_k\right) = \sum_i \beta_i m_i.
 $$
 
-In this sum each $m_i$ is a monomial, $m_i = \prod_{j\in M_i} x_j^{\alpha_{ij}}$, the exponents are non-negative integers, $\alpha_{ij}\in\mathbb{N}_0$, and the coeficients are real valued, $\beta_i \in \mathbb{R}$.
+In this sum each $m_i$ is a monomial, $m_i = \prod_{j\in A_i} x_j^{\alpha_{ij}}$, the exponents are non-negative integers, $\alpha_{ij}\in\mathbb{N}_0$, and the coeficients are real valued, $\beta_i \in \mathbb{R}$.
 For example $p\left( x_1, x_2, x_3 \right) = 2 x_1 + x_2 x_3 + \frac{1}{2} x_1^2 x_3$ has monomials $m_1 = x_1, m_2 = x_2 x_3$ and $m_3 = x_1^2 x_3$ coefficients $\beta_1 = 2, \beta_2 = 1$ and $\beta_3 = 1/2$ and exponents $\alpha_{1,1} = 1, \alpha_{2,2}=1, \alpha_{2,3} = 1, \alpha_{3,1} = 2, \alpha_{3,3} = 1$ and all other $\alpha_{ij} = 0$.
 
-> **NOTA** Este parágrafo deu-me uma ideia: os $\alpha$ definem uma matriz em que as linhas são monómios e as colunas variáveis; "Multiplicando" esta matriz por um vector, obtemos o polinómio.
+> **NOTA** Este parágrafo deu-me uma ideia: os $\alpha$ definem uma matriz $A$ em que as linhas são monómios e as colunas variáveis; "Multiplicando" esta matriz por um vector, obtemos o polinómio: $P=CA$.
 Além disso, podemos continuar a definir a condição de regularização como antes, mas também podemos considerar condições *na* matriz (por exemplo, o número de entradas não nulas ser esparso ou a sua soma ser dominada por $\log nm$)...
@@
 
-1. for a given set of monomials $q_1, \ldots, q_n$, find regression coefficients $\beta_1,\ldots,\beta_n$ that minimize an adequate cost function;
+1. for a given set of monomials $q_1, \ldots, q_r$, find regression coefficients $\theta_1,\ldots,\theta_r$ that minimize an adequate cost function;
 2. find the fittest set of monomials;
@@
 
-Given a dataset $D=\left\lbrace x_1^{(i)}, \ldots, x_m^{(i)}, y^{(i)} \in \mathbb{R}^{m+1} : i = 1,\ldots, n \right\rbrace$ with $n$ observations of $m+1$ variables $X_1, \ldots, X_m, Y$.
-We want to find a polynomial that fits $y = h\left( x_1, \ldots, x_m\right)$ and for that purpose we apply the usual _root-mean-square error_ as a cost function of a candidate fitting polynomial $h_{\Theta}$ defined by parameters $\Theta$:
+Given a dataset $D=\left\lbrace x_1^{(i)}, \ldots, x_k^{(i)}, y^{(i)} \in \mathbb{R}^{k+1} : i = 1,\ldots, n \right\rbrace$ with $n$ observations of $k+1$ variables $X_1, \ldots, X_k, Y$.
+To find a polynomial that fits the observations of the dependent variable $Y$ the usual procedure is to explore parameters $\Theta$ of a polynomial hypothesis $h_{\Theta}$ and select a combination that minimizes the _root-mean-square error_ in the dataset $D$:
 $$
 J_{fit}\left( \Theta; D \right) = \frac{1}{n} \sum_{i=1}^n \left( y^{(i)} - h_{\Theta}\left(x_1^{(i)}, \ldots, x_m^{(i)}\right) \right)^2
+$$
+
+Since it is always possible to fit exactly a polynomial to any given dataset, _overfitting_ the available data, a _regularization_ factor is needed to penalyze hypothesis too biased by the training data:
+
+$$
+J_{reg} = ?
 $$

### 2013-07-10 tests_Auto-Mpg.Rmd
added Auto-Mpg dataset

--- 2013-07-05 tests_Kinematics.Rmd
@@
 ```
-Testing Kinematics
+Testing Auto-Mpg
 ========================================================
@@
 
-filename <- "Kinematics/kin8nm.data"  # where is the data (under dataset folder)
+filename <- "Auto-Mpg/auto.data"  # where is the data (under dataset folder)
 filepath <- paste0("datasets/",filename)
@@
 sapply(my.data, class)
-# my.data$x1 <- as.numeric(my.data$x1) # x1 was a factor, make it a numeric value
+my.data$x3 <- as.numeric(my.data$x3) # x3 was a factor, make it a numeric value
 head(my.data,8)
@@
 source("gapolyfuncs.R")
-report <- make.report(my.data, population=200, iterations=100)
+report <- make.report(my.data, population=25, iterations=10)
 
@@
 
-save(report, file="Kinematics200.Rda") # recover with load("Kinematics.Rda")
+save(report, file="Auto-Mpg.Rda") # recover with load("Auto-Mpg.Rda")
 
@@
 
-write.table(df, "Kinematics300.txt")
+write.table(df, "Auto-Mpg.txt")
 
@@
   xlab("regression method") +
-  ggtitle("Kinematics dataset")
+  ggtitle("Auto-Mpg dataset")
 
@@
 
-write.table(df, "lambda.errors.kinematics2.txt")
+write.table(df, "lambda.errors.Auto-Mpg.txt")
 
-df <- read.table("lambda.errors.kinematics2.txt", header=TRUE)
+df <- read.table("lambda.errors.Auto-Mpg.txt", header=TRUE)
 
@@
  # coord_cartesian(ylim = c(0, 50))  +
-  ggtitle("Kinematics dataset")
+  ggtitle("Auto-Mpg dataset")
 ##############################################

### 2013-07-10 gapolyfuncs.R

@@ -285,10 +285,10 @@ monitorEvalFactory <- function(train.set, n.vars, max.monoids) {
       bestSolution = obj$population[filter,];
     }
     # we have the best solution (chromosome), let's eval it and keep it
-    eval <- evalFuncFactory(train.set, n.vars, max.monoids)
+    eval <- evalFuncFactory(train.set, n.vars, max.monoids) # get eval function
         
-    fitness.progress <- c(fitness.progress, eval(bestSolution)) #TODO: does not work (?)
-    print(make.formula(bestSolution, n.vars, max.monoids))
+    # fitness.progress <- c(fitness.progress, eval(bestSolution)) #TODO: does not work (?)
+    # print(make.formula(bestSolution, n.vars, max.monoids))
     print(eval(bestSolution))
   }
 }

### 2013-07-10 tests_Auto-Mpg.Rmd
minor stuff

@@
 source("gapolyfuncs.R")
-report <- make.report(my.data, population=25, iterations=10)
+report <- make.report(my.data, population=300, iterations=50)
 
@@
 
-save(report, file="Auto-Mpg.Rda") # recover with load("Auto-Mpg.Rda")
+save(report, file="Auto-Mpg300.Rda") # recover with load("Auto-Mpg300.Rda")
 
@@
 
-write.table(df, "Auto-Mpg.txt")
+write.table(df, "Auto-Mpg300.txt")
 

### 2013-07-10 tests_Housing.Rmd
added Housing and Cal-Housing datasets

--- 2013-07-10 tests_Auto-Mpg.Rmd
@@
 ```
-Testing Auto-Mpg
+Testing Housing
 ========================================================
@@
 
-filename <- "Auto-Mpg/auto.data"  # where is the data (under dataset folder)
+filename <- "Housing/housing.data"  # where is the data (under dataset folder)
 filepath <- paste0("datasets/",filename)
@@
 sapply(my.data, class)
-my.data$x3 <- as.numeric(my.data$x3) # x3 was a factor, make it a numeric value
+# my.data$x3 <- as.numeric(my.data$x3) # x3 was a factor, make it a numeric value
 head(my.data,8)
@@
 
-save(report, file="Auto-Mpg300.Rda") # recover with load("Auto-Mpg300.Rda")
+save(report, file="Housing300.Rda") # recover with load("Housing300.Rda")
 
@@
 
-write.table(df, "Auto-Mpg300.txt")
+write.table(df, "Housing300.txt")
 
@@
   xlab("regression method") +
-  ggtitle("Auto-Mpg dataset")
+  ggtitle("Housing dataset")
 
@@
 
-write.table(df, "lambda.errors.Auto-Mpg.txt")
+write.table(df, "lambda.errors.Housing.txt")
 
-df <- read.table("lambda.errors.Auto-Mpg.txt", header=TRUE)
+df <- read.table("lambda.errors.Housing.txt", header=TRUE)
 
@@
  # coord_cartesian(ylim = c(0, 50))  +
-  ggtitle("Auto-Mpg dataset")
+  ggtitle("Housing dataset")
 ##############################################

### 2013-07-10 gapolyfuncs.R

@@ -99,13 +99,13 @@ evalFuncFactory <- function(df, n.vars, max.monoids, lambda=1) {
 # Returns a dataset with the rsme for each one of the tested ML methods
 # including GA Poly
 
-# pre: both sets must have columns named x1...xn and the last one is y (the output)
+# pre: my.data must have columns named x1...xn and the last one is y (the output)
 make.report <- function(my.data, 
-                        n.vars,
-                        population=100, 
-                        iterations=25, 
-                        runs=10, 
-                        mutation.rate=0.05,
+                        population=100,     # GA population
+                        iterations=25,      # GA iterations
+                        runs=10,            # number of runs
+                        mutation.rate=0.05, # GA mutation rate
+                        lambda=1.0,         # regularization variable
                         verbose=TRUE) {
   
   library(genalg)
@@ -117,6 +117,7 @@ make.report <- function(my.data,
   train.p.size <- 0.7 # percentage of training set
   n.vars       <- ncol(my.data)-1
   
+  # vectors keeping the rsme errors for each regression method
   ga.error     <- rep(0,runs)
   lm.error     <- rep(0,runs)
   svm.error    <- rep(0,runs)
@@ -124,12 +125,12 @@ make.report <- function(my.data,
   rf.error     <- rep(0,runs)
   citree.error <- rep(0,runs)
   
-  best.ga.model <- NULL
-  best.ga.error <- 10e10
+  best.ga.model <- NULL  # the best GA model 
+  best.ga.error <- 10e10 # & its error
   
   for (i in 1:runs) {
     
-    # make train & test set
+    # make train & test set (ie, each run has a different train+test sets)
     inTrain   <- sample(1:nrow(my.data), train.p.size * nrow(my.data))
     train.set <- my.data[inTrain,]
     test.set  <- my.data[-inTrain,]
@@ -139,7 +140,7 @@ make.report <- function(my.data,
                         iters = iterations, 
                         mutationChance = mutation.rate, 
                         elitism = TRUE, 
-                        evalFunc = evalFuncFactory(train.set, n.vars, max.monoids))
+                        evalFunc = evalFuncFactory(train.set, n.vars, max.monoids, lambda))
       
     best.solution <- GAmodel$population[1,]
     best.formula <- paste0("y ~ ", make.formula(best.solution, n.vars, max.monoids))
[…]

### 2013-07-10 tests_Housing.Rmd
some more graphics

@@
 ```{r}
-set.seed(101)
+set.seed(202)
 
@@
 source("gapolyfuncs.R")
-report <- make.report(my.data, population=300, iterations=50)
+report <- make.report(my.data, population=300, iterations=100, lambda=0.975)
 
@@
 
-save(report, file="Housing300.Rda") # recover with load("Housing300.Rda")
+save(report, file="Housing300_lambda0.975.Rda") # recover with load("Housing300.Rda")
 
@@
 
-write.table(df, "Housing300.txt")
+write.table(df, "Housing300_lambda0.975.txt")
 

### 2013-07-10 tests_CalHousing.Rmd
processing Cal-Housing dataset

--- 2013-07-10 tests_Housing.Rmd
@@
 ```
-Testing Housing
+Testing California Housing
 ========================================================
@@
 
-filename <- "Housing/housing.data"  # where is the data (under dataset folder)
+filename <- "Cal-Housing/cal_housing"  # where is the data (under dataset folder)
 filepath <- paste0("datasets/",filename)
@@
 source("gapolyfuncs.R")
-report <- make.report(my.data, population=300, iterations=100, lambda=0.975)
+report <- make.report(my.data, population=250, iterations=50)
 
@@
 
-save(report, file="Housing300_lambda0.975.Rda") # recover with load("Housing300.Rda")
+save(report, file="Cal-Housing250.Rda") # recover with load("Cal-Housing250.Rda")
 
@@
 
-write.table(df, "Housing300_lambda0.975.txt")
+write.table(df, "Cal-Housing250.txt")
+# df <- read.table("Cal-Housing250.txt")
 
@@
 
-write.table(df, "")
-
 qplot(method,error,data=df,geom=c("jitter","boxplot")) + 
@@
   xlab("regression method") +
-  ggtitle("Housing dataset")
+  ggtitle("California Housing dataset")
 
@@
  # coord_cartesian(ylim = c(0, 50))  +
-  ggtitle("Housing dataset")
+  ggtitle("California Housing dataset")
 ##############################################

### 2013-07-10 tests_Mg.Rmd
Mg dataset

--- 2013-07-10 tests_CalHousing.Rmd
@@
 ```
-Testing California Housing
+Testing MG
 ========================================================
+
+Ref:
+
+Gary William Flake and Steve Lawrence. 
+Efficient SVM regression training with SMO. 
+Machine Learning, 46:271-290, 2002.
+
+# of data: 1,385
+# of features: 6
 
@@
 
-filename <- "Cal-Housing/cal_housing"  # where is the data (under dataset folder)
+filename <- "Mg/mg_scale"  # where is the data (under dataset folder)
 filepath <- paste0("datasets/",filename)
@@
 # my.data <- na.omit(my.data) # the dataset has some NAs
+my.data <- my.data[,c(2:7,1)] # place y as the last column
 
@@
 
-save(report, file="Cal-Housing250.Rda") # recover with load("Cal-Housing250.Rda")
+save(report, file="Mg250.Rda") # recover with load("Mg250.Rda")
 
@@
 
-write.table(df, "Cal-Housing250.txt")
-# df <- read.table("Cal-Housing250.txt")
+write.table(df, "Mg250.txt")
+# df <- read.table("Mg250.txt")
 
@@
   xlab("regression method") +
-  ggtitle("California Housing dataset")
+  ggtitle("Mg dataset")
 
@@
 
-write.table(df, "lambda.errors.Housing.txt")
+write.table(df, "lambda.errors.Mg.txt")
 
-df <- read.table("lambda.errors.Housing.txt", header=TRUE)
+df <- read.table("lambda.errors.Mg.txt", header=TRUE)
 
@@
  # coord_cartesian(ylim = c(0, 50))  +
-  ggtitle("California Housing dataset")
+  ggtitle("Mg250 dataset")
 ##############################################

### 2013-07-10 tests_Mg.Rmd
extra results

@@
 source("gapolyfuncs.R")
-report <- make.report(my.data, population=250, iterations=50)
+report <- make.report(my.data, population=250, iterations=50, lambda=0.975)
 
@@
 
-write.table(df, "Mg250.txt")
+write.table(df, "Mg250_lamda0.975.txt")
 # df <- read.table("Mg250.txt")
@@
   xlab("regression method") +
-  ggtitle("Mg dataset")
+  ggtitle("Mg dataset (lambda 0.975)")
 
@@
 # Lambda Test
-lambda.values <- seq(0.5,0.999,length.out=20)
+lambda.values <- seq(0.5,1.5,length.out=20)
 n.runs <- 5
@@
 source("gapolyfuncs.R")
-errors <- test.lambda(my.data, lambda.values, runs=n.runs, population=50, iterations=25)
+errors <- test.lambda(my.data, lambda.values, runs=n.runs, population=50, iterations=50)
 
 df <- data.frame(error = as.vector(matrix(errors,nrow=1)),
-                 lambda = as.factor(rep(round(lambda.values,3), each=n.runs)))
+                 lambda = as.factor(rep(round(lambda.values,2), each=n.runs)))
 
@@
 df <- read.table("lambda.errors.Mg.txt", header=TRUE)
-
 df$lambda <- as.factor(df$lambda)
@@
  # coord_cartesian(ylim = c(0, 50))  +
-  ggtitle("Mg250 dataset")
+  ggtitle("Mg dataset")
 ##############################################

### 2013-07-15 tests_Auto-Mpg.Rmd
beginning latex section of Results

@@
 my.data$x3 <- as.numeric(my.data$x3) # x3 was a factor, make it a numeric value
+my.data <- as.data.frame(scale(my.data)) # scale it
 head(my.data,8)
@@
 source("gapolyfuncs.R")
-report <- make.report(my.data, population=300, iterations=50)
+report <- make.report(my.data, population=300, iterations=50, lambda=0.8)
 
@@
 # Lambda Test
-lambda.values <- seq(0.5,0.999,length.out=20)
-n.runs <- 5
+lambda.values <- seq(0.7,1.3,by=0.05)
+n.runs <- 10
 
 source("gapolyfuncs.R")
-errors <- test.lambda(my.data, lambda.values, runs=n.runs, population=50, iterations=25)
+errors <- test.lambda(my.data, lambda.values, runs=n.runs, population=60, iterations=60)
 

### 2013-07-15 tests_Housing.Rmd
beginning latex section of Results

--- 2013-07-10 tests_Housing.Rmd
@@
 head(my.data,8)
+my.data <- as.data.frame(scale(my.data))
 
@@
 source("gapolyfuncs.R")
-report <- make.report(my.data, population=300, iterations=50)
+report <- make.report(my.data, population=300, iterations=100, lambda=0.8, runs=25)
 
@@
 
-save(report, file="Housing300.Rda") # recover with load("Housing300.Rda")
+save(report, file="Housing300_lambda0.8.Rda") # recover with load("Housing300.Rda")
 
@@
 
-write.table(df, "Housing300.txt")
+write.table(df, "Housing300_lambda0.8.txt")
 
@@
 df <- as.data.frame(list.data)
-
-write.table(df, "")
 

### 2013-07-15 tests_test.Rmd
beginning latex section of Results

@@
 source("gapolyfuncs.R")
-report <- make.report(my.data, population=50, iterations=25, mutation.rate=0.05, runs=1)
+report <- make.report(my.data, population=100, iterations=50, mutation.rate=0.05)
 
@@
 # Lambda Test
-lambda.values <- seq(1,2,length.out=20)
+lambda.values <- seq(0.5,2,length.out=20)
 n.runs <- 1

### 2013-07-15 texts/experiments.tex
beginning latex section of Results

\documentclass[a4paper, 12pt]{article}
\usepackage[isolatin]{inputenc}
\usepackage[normalem]{ulem}
\usepackage{graphicx}
\usepackage{cite}

\begin{document}

\subsection{Polynomial Encoding}

The chromosome coding will as follow:
\begin{enumerate}
\item an initial segment detailing which monomials are active (the 1st monomial is always active), these are represented in unary description
\item the chromosome is split into `max.monomials` sets of bits of equal size 
  (a monomial)
\item each monomial is split into `n.vars` sets of `max.degree` size each, i.e., a variable
\item for each variable, the remaining bits give the binary description of the variable degree, i.e., the maximum exponent is given by $2^max.degree-1$
\end{enumerate}

Eg: consider polynomial $x_1^3 x_3 + x_3^7 + x_1 x_2$ with max.monomials = 4, n.vars = 3 and max.degree=3. One possible encoding would be:

$$110 - 011,000,001~~;~~000,000,111~~;~~001,001,000~~;~~110,010,101$$

(for reading purposes the semicolons separate monomials, the commas separate variables)

The first 3 bits inform that the second and third monomials are active while the fourth is not (as said, the first monomial is always active). This last monomial does not enter neither in the polynomial regression nor in the fitness evaluation. However, it acts as a kind of junk DNA, becoming active when the third bit of the entire sequence flips from 0 to 1.

Let's interpret the first monomial description, $011,000,001$. It is divided by three since n.vars=3. The first triple $011$ is the binary description of the exponent of variable $x_1$ which is 3, so the first monomial includes $x_1^3$. The second triple, $000$, means that $x_2$ is not part of the monomial. The thrid triple $001$ says that variable $x_3$ has exponent $1$, so the first monomial consists of $x_1^3 x_3$. All the remaining sets of $9$ bits are interpreted the same way and we get the previous polynomial.

Notice that all binary descriptions give rise to valid polynomials. However, if the coding consists of entirely zeros, by convention, it describes polynomial $x_1$. This has to do with the execution of the polynomial regression that would fail if we interpret it as the zero polynomial. Anyway, for progressive larger  binary descriptions, the chances of getting this zero description decrease exponentially, so it does not impact in any meaningful way in the algorithm's performance.

Somewhat unexpectedly after some experiences it was found that lower values for $\lambda$ sometimes provide better, even if marginal, results. The next figure shows results for the Abalone dataset with five runs for each $\lambda$ (cf. Experimental Results section for more info on this dataset).

\begin{figure}[htb]
	\begin{center}
		\includegraphics[height=6cm,width=10cm,angle=0]
			{figures/Abalone_dataset_lambdas.eps}
		\caption{fitness progress for Abalone dataset}
	\end{center}
\end{figure}

The values for another dataset:

\begin{figure}[htb]
	\begin{center}
		\includegraphics[height=6cm,width=10cm,angle=0]
			{figures/Auto-Mpg_dataset_lambdas.eps}
		\caption{fitness progress for Abalone dataset}
	\end{center}
\end{figure}

The inflection point is around $\lambda = 0.8$. We'll use the regularisation parameter with this value.

\subsection{Regularization}

The polynomial fitness considered so far is based on the ability to predict the test set after the polynomial regression has found the appropriate coefficients $\theta_i$ for each one of the polynomial's monomials $m_i$.

This fitness function tends to prefer more complex polynomials, namely in the number of monomials which provides the regression algorithm for more fitting possibilities. One way to balance this is to provide a regularization term into the fitness function. Our proposal is to include a multiplicative factor into the fitness function proportional to the number of monomials. Thus, the fitness function becomes,

$$
J_{fit}\left( \Theta; D \right) = \lambda^{M} \sqrt{ \frac{1}{n} \sum_{i=1}^n \left( y^{(i)} - h_{\Theta}\left(x_1^{(i)}, \ldots, x_m^{(i)}\right) \right)^2 }
$$

where M is the number of monomials the polynomial has. A $\lambda$ greater than one punishes polynomials with more monomials.

\section{Experimental Results}

These results were found using R's programming language~\cite{R}.

To compare this paper's algorithm we applied the exact same train and test samples with other learning algorithms, namely: Linear Regression, Support Vector Machines~\cite{Meyer12}, Regression Trees~\cite{Therneau13}, Conditional Inference Trees~\cite{Hothorn06}~\cite{Strobl07}~\cite{Strobl08} and Random Forests~\cite{Liaw02}

In order to train and test the performance of GApoly it was used several datasets (see below). For each dataset, we selected 70\% for training purposes and the remaining observations to make the test set in order to compute the estimated rmse. To achieve more robust results, for each dataset there were ten runs, each one with different samples for the train and test sets. The results below are boxplots for the test set error predictions over these different runs.

\textbf{Artificial}: this is an artificial dataset with four numeric features, $x_1, \ldots x_4$, where $x_1,x_3$ are outcomes from Poisson random variables, and $x_2,x_4$ from Normal random variables. The dependent variable $y$ is given by expression $x_2x_4^2 + x_1^2x_3 + 5$. The dataset includes 50 observations.

\begin{figure}[htb]
	\begin{center}
		\includegraphics[height=6cm,width=10cm,angle=0]
			{figures/artificial_dataset1_lambda1.0.eps}
		\caption{results for Artificial dataset}
	\end{center}
\end{figure}

This dataset was used in order to verify if GApoly was able to find the polynomial relation, which the algorithm did. In this case, all runs used a population of 100 with 50 iterations for each run.

In the next datasets, the population was 250 with 100 iterations for each run

\textbf{Housing}: This data set concerns the task of predicting housing values in areas of Boston. There are 13 continuous attributes and the dependent variable is the median value of owner-occupied homes in \$1000's\footnote{This and the remaining regression datasets were selected from Luis Torgo's data repository, http://www.dcc.fc.up.pt/~ltorgo/Regression/DataSets.html. Most of these set originally come from UCI ML repository, http://archive.ics.uci.edu/ml/}. There are 506 observations.

\begin{figure}[htb]
	\begin{center}
		\includegraphics[height=6cm,width=10cm,angle=0]
			{figures/Housing_dataset_lambda0.975.eps}
		\caption{results for Boston Housing dataset}
	\end{center}
\end{figure}

\textbf{Abalone}: This dataset can be used to predict the age of a abalone shell using the given 8 numeric attributes concerning several physical measurements. There are 4177 observations.

\begin{figure}[htb]
	\begin{center}
		\includegraphics[height=6cm,width=10cm,angle=0]
			{figures/Abalone_dataset_lambda1.0.eps}
		\caption{results for Abalone dataset}
	\end{center}
\end{figure}

\textbf{Auto-MPG}: This dataset is used to predict fuel consumption in miles per gallon, based on two discrete and five continuous attributes. There are 398 observations. Here are the initial observations:

\begin{center}
\begin{tabular}{cccccccc}
x1 & x2 & x3 & x4 & x5 & x6 & x7 & y \\
8 & 307 & 17 & 3504 & 12.0 & 70 & 1 & 18 \\
8 & 350 & 35 & 3693 & 11.5 & 70 & 1 & 15 \\
8 & 318 & 29 & 3436 & 11.0 & 70 & 1 & 18 \\
8 & 304 & 29 & 3433 & 12.0 & 70 & 1 & 16 \\
8 & 302 & 24 & 3449 & 10.5 & 70 & 1 & 17 
\end{tabular}
\end{center}

\begin{figure}[htb]
	\begin{center}
		\includegraphics[height=6cm,width=10cm,angle=0]
			{figures/Auto-Mpg_dataset_lambda0.8.eps}
		\caption{results for Auto-MPG dataset}
	\end{center}
\end{figure}

% The best polynomial, with a rmse of 2.15 for the test set, was

% $$y = -3.4 10^{-6} x_4 x_6^2 + 1.05 10^{-8} x_6^5 + 33.7$$

\textbf{Kinematics}: This dataset is concerned with the realistic simulation of the forward kinematics of an 8 link robot arm. The task is to predict the distance of the end-effector from a target using 8 continuous attributes. There are 8192 observations.

\begin{figure}[htb]
	\begin{center}
		\includegraphics[height=6cm,width=10cm,angle=0]
			{figures/Kinematics_dataset_lambda1.0.eps}
		\caption{results for Kinematics dataset}
	\end{center}
\end{figure}

\subsection{Convergence speed}

The GA quickly proceeds in the first 50 to 100 generations to reasonable error rates. Then, it proceeds slower achieving best solutions with marginal error reduction. Since the entire learning process takes some time, in the current R implementation\footnote{the processing time can probably be speeded by one to two orders in magnitude if the process is implemented in a low level programming language like C++. However, speed optimisation was not the focus of this article}, placing a limit between 50 to 100 generations already achieves good results, relative to higher iteration values. The next figure shows a typical error evolution for the dataset Abalone given two different values for $\lambda$.

\begin{figure}[htb]
	\begin{center}
		\includegraphics[height=6cm,width=10cm,angle=0]
			{figures/Abalone_fitnessProgress.eps}
		\caption{fitness progress for Abalone dataset}
	\end{center}
\end{figure}

Somewhat unexpectedly after some experiences it was found that lower values for $\lambda$ tend to provide better, even if marginal, results. The next figure shows results for the artific

\subsection{Discussion}

The proposed method has competitive results comparing with some well-known regression methods. Only Random Forest outperforms GApoly systematically (it outperforms all the others) except in the Auto-MPG test and, of course, in the artificial dataset that uses a straightforward polynomial relation.

\section{Bibliography}

\bibliographystyle{plain}
\bibliography{mybib}

\end{document}

### 2013-07-15 tests_Housing.Rmd
more results into latex
scaling all datasets, using 25 runs, lambda 0.8

@@
 
-library(ggplot2)
-
 df <- data.frame(GA.Poly = report$ga.error,
@@
 
+library(ggplot2)
 size <- length(report$ga.error)
@@
   xlab("regression method") +
+  coord_cartesian(ylim = c(0.2, 0.75))  +
   ggtitle("Housing dataset")

### 2013-07-15 tests_Kinematics.Rmd
made some graphics
more results into latex
scaling all datasets, using 25 runs, lambda 0.8

@@
 # my.data$x1 <- as.numeric(my.data$x1) # x1 was a factor, make it a numeric value
+my.data <- as.data.frame(scale(my.data))
 head(my.data,8)
@@
 source("gapolyfuncs.R")
-report <- make.report(my.data, population=200, iterations=100)
+report <- make.report(my.data, population=300, iterations=100, runs=25, lambda=0.8)
 
@@
 
-save(report, file="Kinematics200.Rda") # recover with load("Kinematics.Rda")
+save(report, file="Kinematics300_lambda0.8_25runs.Rda") # recover with load("Kinematics.Rda")
 
@@
 
-write.table(df, "Kinematics300.txt")
+write.table(df, "Kinematics200.txt")
 

### 2013-07-15 texts/experiments.tex
more results into latex
scaling all datasets, using 25 runs, lambda 0.8

@@
 For each dataset, we selected 70\% for training purposes and the remaining observations to make the test set in order to compute the estimated rmse.
-To achieve more robust results, for each dataset there were ten runs, each one with different samples for the train and test sets.
+To achieve more robust results, for most datasets there were 25 runs, each one with different samples for the train and test sets.
+For the datasets with attribute values of different magnitudes, a preliminary scaling was executed.
 The results below are boxplots for the test set error predictions over these different runs.
@@
 		\includegraphics[height=6cm,width=10cm,angle=0]
-			{figures/Housing_dataset_lambda0.975.eps}
+			{figures/Housing_dataset_lambda0.8_25runs.eps}
 		\caption{results for Boston Housing dataset}
@@
 \end{figure}
+
+Just as an example of the model the GApoly algorithm outputs: in this dataset the best polynomial, with a rmse of 0.38 for the test set, was the following:
+
+$$y = -0.12 x_6 x_9^4 + 0.78 x_6 - 0.4 x_9^2 x_{13} + 0.17 x_{13}^2 - 0.044$$
+
+where the attributes mean:
+$x_6$, RM average number of rooms per dwelling;
+$x_9$, RAD index of accessibility to radial highways;
+$x_{13}$, Lower status of the population.
 
@@
 		\includegraphics[height=6cm,width=10cm,angle=0]
-			{figures/Abalone_dataset_lambda1.0.eps}
+			{figures/Abalone_dataset_lambda0.8_25runs.eps}
 		\caption{results for Abalone dataset}
@@
 \end{figure}
+
+% best polynomial, rsme 0.65
+% $$y = 0.52 s_1^2 x_5 - 0.13 x_2 x_8 -0.75 x_1^2 x_6 + 0.78 x_8 + 0.18 $$
 
@@
 
-\begin{center}
-\begin{tabular}{cccccccc}
-x1 & x2 & x3 & x4 & x5 & x6 & x7 & y \\
-8 & 307 & 17 & 3504 & 12.0 & 70 & 1 & 18 \\
-8 & 350 & 35 & 3693 & 11.5 & 70 & 1 & 15 \\
-8 & 318 & 29 & 3436 & 11.0 & 70 & 1 & 18 \\
-8 & 304 & 29 & 3433 & 12.0 & 70 & 1 & 16 \\
-8 & 302 & 24 & 3449 & 10.5 & 70 & 1 & 17 
-\end{tabular}
-\end{center}
-
 \begin{figure}[htb]
@@
 		\includegraphics[height=6cm,width=10cm,angle=0]
-			{figures/Auto-Mpg_dataset_lambda0.8.eps}
+			{figures/Auto-Mpg_dataset_lambda0.8_25runs.eps}
 		\caption{results for Auto-MPG dataset}
@@
 
-% The best polynomial, with a rmse of 2.15 for the test set, was
-
-% $$y = -3.4 10^{-6} x_4 x_6^2 + 1.05 10^{-8} x_6^5 + 33.7$$
+% best polynomial, rmse 0.32
+%$$ y = 0.4 x_6 + 0.02 x_5^3 x_6 - 0.85 x_4 + 0.23 x_4^2 - 0.26 $$
 

### 2013-07-16 texts/experiments.tex
added all data results on latex file

@@
 Eg: consider polynomial $x_1^3 x_3 + x_3^7 + x_1 x_2$ with max.monomials = 4, n.vars = 3 and max.degree=3. One possible encoding would be:
-
 $$110 - 011,000,001~~;~~000,000,111~~;~~001,001,000~~;~~110,010,101$$
@@
 
-Somewhat unexpectedly after some experiences it was found that lower values for $\lambda$ sometimes provide better, even if marginal, results.
-The next figure shows results for the Abalone dataset with five runs for each $\lambda$ (cf.
-Experimental Results section for more info on this dataset).
-
-\begin{figure}[htb]
-	\begin{center}
-		\includegraphics[height=6cm,width=10cm,angle=0]
-			{figures/Abalone_dataset_lambdas.eps}
-		\caption{fitness progress for Abalone dataset}
-	\end{center}
-\end{figure}
-
-The values for another dataset:
-
-\begin{figure}[htb]
-	\begin{center}
-		\includegraphics[height=6cm,width=10cm,angle=0]
-			{figures/Auto-Mpg_dataset_lambdas.eps}
-		\caption{fitness progress for Abalone dataset}
-	\end{center}
-\end{figure}
-
-The inflection point is around $\lambda = 0.8$. We'll use the regularisation parameter with this value.
-
 \subsection{Regularization}
@@
 where M is the number of monomials the polynomial has. A $\lambda$ greater than one punishes polynomials with more monomials.
+
+Somewhat unexpectedly after some experiences it was found that lower values for $\lambda$ sometimes provide better, even if marginal, results.
+Figure~\ref{Abalone_dataset_lambdas} shows results for the Abalone dataset with five runs for each $\lambda$ (cf.
+Experimental Results includes information about these datasets).
+
+\begin{figure}[htb]
+	\begin{center}
+		\includegraphics[height=6cm,width=10cm,angle=0]
+			{figures/Abalone_dataset_lambdas.eps}
+		\caption{fitness progress for Abalone dataset}
+		\label{Abalone_dataset_lambdas}
+	\end{center}
+\end{figure}
+
+Figure~\ref{Auto-Mpg_dataset_lambdas} presents regularization results for the Auto-MPG dataset:
+
+\begin{figure}[htb]
+	\begin{center}
+		\includegraphics[height=6cm,width=10cm,angle=0]
+			{figures/Auto-Mpg_dataset_lambdas.eps}
+		\caption{fitness progress for Abalone dataset}
+		\label{Auto-Mpg_dataset_lambdas}
+	\end{center}
+\end{figure}
+
+The typical inflection point is around $\lambda = 0.8$. We'll use the regularisation parameter with this value.
 
@@
 		\caption{results for Artificial dataset}
+		\label{artificial_dataset1_lambda1.0}
 	\end{center}
@@
 
-This dataset was used in order to verify if GApoly was able to find the polynomial relation, which the algorithm did. In this case, all runs used a population of 100 with 50 iterations for each run.
+This dataset was used in order to verify if GApoly was able to find the polynomial relation, which the algorithm did (cf.
+figure~\ref{artificial_dataset1_lambda1.0}).
+In this case, all runs used a population of 100 with 50 iterations for each run.
 
-In the next datasets, the population was 250 with 100 iterations for each run
+In the next datasets, the population has size 250 with 100 iterations for each run.
 
@@
 There are 506 observations.
+Figure~\ref{Housing_dataset_lambda0.8_25runs} presents the results for this dataset.
 
@@
 		\caption{results for Boston Housing dataset}
+		\label{Housing_dataset_lambda0.8_25runs}
 	\end{center}
@@
 
-\textbf{Abalone}: This dataset can be used to predict the age of a abalone shell using the given 8 numeric attributes concerning several physical measurements. There are 4177 observations.
+\textbf{Abalone}: This dataset can be used to predict the age of a abalone shell using the given 8 numeric attributes concerning several physical measurements.
+There are 4177 observations.
+Figure~\ref{Abalone_dataset_lambda0.8_25runs} presents the results for this dataset.
 
@@
 		\caption{results for Abalone dataset}
+		\label{Abalone_dataset_lambda0.8_25runs}
 	\end{center}
@@
 There are 398 observations.
-Here are the initial observations:
+Figure~\ref{Auto-Mpg_dataset_lambda0.8_25runs} presents the results for this dataset.
 
@@
 		\caption{results for Auto-MPG dataset}
+		\label{Auto-Mpg_dataset_lambda0.8_25runs}
 	\end{center}
@@
 There are 8192 observations.
+Figure~\ref{Kinematics300_lambda0.8_25runs} presents the results for this dataset.
 
@@
 		\includegraphics[height=6cm,width=10cm,angle=0]
-			{figures/Kinematics_dataset_lambda1.0.eps}
+			{figures/Kinematics300_lambda0.8_25runs.eps}
 		\caption{results for Kinematics dataset}
+		\label{Kinematics300_lambda0.8_25runs}
 	\end{center}
 \end{figure}
+
+% best polynomial, rmse 0.74
+% $$ y = 0.24 x_5 - 0.23 x_5 x_6 x_7 - 0.19 x_4 x_7 - 0.5 x_3 $$
 
@@
 Then, it proceeds slower achieving best solutions with marginal error reduction.
-Since the entire learning process takes some time, in the current R implementation\footnote{the processing time can probably be speeded by one to two orders in magnitude if the process is implemented in a low level programming language like C++.
+Since the entire learning process takes some time, in the current R implementation\footnote{The processing time can probably be speeded by one to two orders in magnitude if the process is implemented in a low level programming language like C++.
 However, speed optimisation was not the focus of this article}, placing a limit between 50 to 100 generations already achieves good results, relative to higher iteration values.
-The next figure shows a typical error evolution for the dataset Abalone given two different values for $\lambda$.
+Figure~\ref{Abalone_fitnessProgress} shows a typical error evolution for the dataset Abalone given two different values for $\lambda$.
 
@@
 		\caption{fitness progress for Abalone dataset}
+		\label{Abalone_fitnessProgress}
 	\end{center}
 \end{figure}
-
-Somewhat unexpectedly after some experiences it was found that lower values for $\lambda$ tend to provide better, even if marginal, results. The next figure shows results for the artific
 

### 2013-07-17 texts/paper-draft.tex
Paper integration; Experimentation with figures.

\documentclass[12pt]{article}
\usepackage[latin1]{inputenc}
\usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\geometry{a4paper} % or letter or a5paper or ... etc
% \geometry{landscape} % rotated page geometry
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{acronym}
\usepackage{pgfplots}
\pgfplotsset{compat=1.8}
\usepgfplotslibrary{statistics}

\usepackage{algpseudocode}
\usepackage{algorithm}

\title{Selection of Polynomial Features by Genetic Algorithms}
\author{Francisco Coelho \and João Neto}
\date{[{\sc draft} \today]} % delete this line to display the current date

\acrodef{GA}{Genetic algorithms}
\acrodef{rmse}{root-mean-square error}
\acrodef{GAPoly}[{\sc GAPoly}]{Genetic Algorithms for Polynomials}

%%% BEGIN DOCUMENT
\begin{document}

\maketitle
\begin{abstract}
Many applications require models that have no acceptable linear approximation and many nonlinear models are defined by polynomials. The use of genetic algorithms to find polynomial models is decades old but still poses challenges due to the complexity of the search and different definitions of optimal solution.

This paper describes a two-step method that uses genetic algorithms and linear regression to find empirical polynomial regressions. Experiments on common datasets show that, discounting the training computational effort, this method is quite competitive.
\end{abstract}

\section{Introduction}

With notable exceptions (\emph{e.g.} neural networks), most machine learning regression techniques are based on linear models. This assumption has many advantages including reduced computational complexity and strong theoretical framework. However nonlinearity is unavoidable in many application scenarios, specially those with phase transitions or feedback loops, so common in ecology, cybernetics, robotics and other areas.

Polynomials, one of the most studied subjects in mathematics, generalize linear functions and define, perhaps, the simplest and most used nonlinear models. Applications of polynomial include colorimetric calibration \cite{mendes2005adaptive}, explicit \emph{formulae} for turbulent pipe flows \cite{davidson1999method}, computational linguistics \cite{sanchez2009obtaining} and more recently, analytical techniques for cultural heritage materials \cite{csefalvayova2010use}, liquid epoxy molding process \cite{chan2011modeling}, B-spline surface reconstruction \cite{galvez2012iterative}, product design \cite{chan2012development} or forecasting cyanotoxins presence in water reservoirs \cite{garcia2013hybrid}. These examples not only illustrate the wide spectrum of applications but also the work in each one of these polynomial models uses, at some point, a genetic algorithm.

\ac{GA} where, arguably, one the hottest topics of research in the recent decades and with good reason since they outline an optimisation scheme easy to conceptualise and with very broad application. If a nonlinear (or otherwise) model requires parameterisation, \acp{GA} provide a simple and often effective approach to search for locally optimal parameters. Research related to genetic algorithms abound and spans from the 1950s seminal work of Nils Aall Barricelli \cite{barricelli1962numerical} in the Institute for Advanced Study of Princeton to today's principal area of study of thousands of researchers, covered in hundreds of conferences, workshops and other meetings. Perhaps the key impulse to \acp{GA} come from John Holland's work and his book ``Adaptation in Natural and Artificial Systems'' \cite{holland1975adaptation}.

One interesting ``flavour" of genetic algorithms, named \emph{genetic programming} by John Koza \cite{koza1992genetic}, proposed the use of \acp{GA} to search the syntatic stucture of complex functions. This syntatic structure search is keen to the central ideas of deep learning \cite{bengio2013representation, bengio2009learning}, the subarea of machine learning actually producing the most promising results (\emph{e.g.} \cite{tarlowstochastic}). It is also related to the work presented in this paper in the sense that, unlike linear models that have a simple structure ($y=\sum_i \beta_i x_i$) nonlinear (in particular polynomial) models pose an additional ``structure" search problem.

The idea of using \acp{GA} to find a polynomial regression is not new \cite{maertens2006genetic, yu2008optimal, wu2009novel} but still generates original research \cite{hofwing2011optimal, cetisli2011polynomial}. In line with that research this work describes a general method to find a polynomial regression of a given dataset. The optimal regression minimises a cost function that accounts for both the \ac{rmse} and a regularisation factor to avoid over-fitting.

%A method that produces \emph{adequate} models from observed complex data has many uses. For example by a scientist to better understand the source of the data or by an autonomous agents adapting to the environment.

It turns out that, discarding the computational cost of training, the polynomial regression method presented here, \ac{GAPoly}, provides a quite competitive regression method. Indeed, it is only systematically out-performed by random forests, an \emph{ensemble} method.

The remainder of this paper is organised as usual: the next section describes the details of our method and is followed by a presentation of some performance results. The last section draws some conclusions and points future research tasks.

\section{Polynomial Regression with Genetic Algorithms}

This section starts with a brief description of the polynomial representation used to support the encoding of polynomial instances. This encoding is required to define the populations the \acp{GA} will evolve. Then is defined the cost function, including regularisation for over-fitting.

A usual representation of polynomials is
$$
p\left( x_1, \ldots, x_k\right) = \sum_i \beta_i m_i
$$
where each $m_i$ is a monomial, $m_i = \prod_{j\in A_i} x_j^{\alpha_{ij}}$, the exponents are non-negative integers, $\alpha_{ij}\in\mathbb{N}_0$, and the coefficients are real valued, $\beta_i \in \mathbb{R}$. For example $p\left( x_1, x_2, x_3 \right) = 2 x_1 + x_2 x_3 + \frac{1}{2} x_1^2 x_3$ has monomials $m_1 = x_1, m_2 = x_2 x_3$ and $m_3 = x_1^2 x_3$ coefficients $\beta_1 = 2, \beta_2 = 1$ and $\beta_3 = 1/2$ and exponents $\alpha_{1,1} = 1, \alpha_{2,2}=1, \alpha_{2,3} = 1, \alpha_{3,1} = 2, \alpha_{3,3} = 1$ and all other $\alpha_{ij} = 0$.

By using the exponents alone we have a matrix that defines the monomial structure of the polynomial, $A = \left[\alpha_{ij}\right]$. For the example above
$$
A = \left[ 
\begin{array}{ccc}
1 & 0 & 0 \\
0 & 1 & 1 \\
2 & 0 & 1
\end{array}
\right] \sim 
\left[ 
\begin{array}{ccc}
x_1 \\
x_2 x_3 \\
x_1^2 x_3
\end{array}
\right]
$$
where each row defines a monomial and each column represents a variable. Changing the order of the rows doesn't change the polynomial whereas changing the order of the columns corresponds to changing the respective variables.
%Notice that if we interpret matrix $A$ as a single column of monomials and set $B = \left[ \begin{array}{ccc} 2 & 1 & \frac{1}{2} \end{array} \right ]$ then
%$$
%p\left( x_1, x_2, x_3 \right) = BA = \left[ \begin{array}{ccc} 2 & 1 & \frac{1}{2} \end{array} \right ]\left[ 
%\begin{array}{ccc}
%x_1 \\
%x_2 x_3 \\
%x_1^2 x_3
%\end{array}
%\right].
%$$

This representation of polynomials makes the problem of structure search very clear: except for the trivial cases, the number of possible monomials given $n$ variables and a maximum joint degree $d$ grows exponentialy with either $n$ or $d$. But more importantly, the polynomial regression problem can be naturally split into two subproblems:

\begin{enumerate}
\item for a given set of monomials $q_1, \ldots, q_r$, find regression coefficients $\theta_1,\ldots,\theta_r$ that minimize an adequate cost function;
\item find the fittest set of monomials;
\end{enumerate}

Algorithm \ref{alg:gapoly} solves the first problem with linear regression and the second with \acp{GA}.

\begin{algorithm}[htb]
\begin{algorithmic}
\Function{J}{$D,\left[m_1,\ldots,m_k\right]$}\Comment{Cost, or error, of the polynomial $m_1,\ldots,m_k$ in the dataset $D$}
	\State\Return $err \gets \Call{lm}{D,\left[m_1,\ldots,m_k\right]}$ \Comment{Error in Linear Regression}
\EndFunction
\Statex
\Function{GAPoly}{$D,pop_0,\epsilon$}
	\State $pop \gets pop_0$
	\While{$err > \epsilon$}
		\State $pop \gets \Call{IterateGA}{pop}$\Comment{Build next generation}
		\State $pop \gets \Call{Sort}{pop, key = \textsc{Error}}$\Comment{Sort population by error}
		\State $err \gets \textrm{error of }\Call{First}{pop}$
	\EndWhile
	\State\Return $\Call{First}{pop}$
\EndFunction
\end{algorithmic}
\caption{\ac{GAPoly} uses linear regression to find the monomials coefficients and genetic algorithms to explore the space of monomials.}\label{alg:gapoly}
\end{algorithm}

[EVALUATION STRATEGY]

[RESULTS]

\subsection{Polinomial Encoding}

The chromosome coding will as follow:
\begin{enumerate}
\item an initial segment detailing which monomials are active (the 1st monomial is always active), these are represented in unary description
\item the chromosome is split into $m$ sets of bits of equal size 
  (a monomial)
\item each monomial is split into $n$ sets of $d$ size each, i.e., a variable
\item for each variable, the remaining bits give the binary description of the variable degree, i.e., the maximum exponent is given by $2^d-1$
\end{enumerate}

Eg: consider polynomial $x_1^3 x_3 + x_3^7 + x_1 x_2$ with $m = 4, n = 3$ and $d=3$. One possible encoding would be:
$$110 - 011,000,001~~;~~000,000,111~~;~~001,001,000~~;~~110,010,101$$

(for reading purposes the semicolons separate monomials, the commas separate variables)

The first $3$ bits inform that the second and third monomials are active while the fourth is not (as said, the first monomial is always active). This last monomial does not enter neither in the polynomial regression nor in the fitness evaluation. However, it acts as a kind of junk DNA, becoming active when the third bit of the entire sequence flips from 0 to 1.

Let's interpret the first monomial description, $011,000,001$. It is divided by three since $n=3$. The first triple $011$ is the binary description of the exponent of variable $x_1$ which is 3, so the first monomial includes $x_1^3$. The second triple, $000$, means that $x_2$ is not part of the monomial. The third triple $001$ says that variable $x_3$ has exponent $1$, so the first monomial consists of $x_1^3 x_3$. All the remaining sets of $9$ bits are interpreted the same way and we get the previous polynomial.

Notice that all binary descriptions give rise to valid polynomials. However, if the coding consists of entirely zeros, by convention, it describes polynomial $x_1$. This has to do with the execution of the polynomial regression that would fail if we interpret it as the zero polynomial. Anyway, for progressive larger  binary descriptions, the chances of getting this zero description decrease exponentially, so it does not impact in any meaningful way in the algorithm's performance.

\subsection{Cost Function}

The polynomial fitness considered so far is based on the ability to predict the test set after the polynomial regression has found the appropriate coefficients $\theta_i$ for each one of the polynomial's monomials $m_i$.

This fitness function tends to prefer more complex polynomials, namely in the number of monomials which provides the regression algorithm for more fitting possibilities. One way to balance this is to provide a regularization term into the fitness function. Our proposal is to include a multiplicative factor into the fitness function proportional to the number of monomials. Thus, the fitness function becomes,

$$
J_{fit}\left( \Theta; D \right) = \lambda^{m} \sqrt{ \frac{1}{n} \sum_{i=1}^n \left( y^{(i)} - h_{\Theta}\left(x_1^{(i)}, \ldots, x_m^{(i)}\right) \right)^2 }
$$

where $m$ is the number of monomials the polynomial has. A $\lambda$ greater than one punishes polynomials with more monomials.

Somewhat unexpectedly after some experiences it was found that lower values for $\lambda$ sometimes provide better, even if marginal, results. Figure~\ref{Abalone_dataset_lambdas} shows results for the Abalone dataset with five runs for each $\lambda$ (cf. Experimental Results includes information about these datasets).

\begin{figure}[htb]\begin{center}
\begin{tikzpicture}
\begin{axis}[width=\textwidth,height=6cm,
	boxplot/draw direction=y,
	xtick={1,2,3,4,5,6,7,8,9,10,11,12,13},
	xticklabels={0.70, 0.75, 0.80, 0.85, 0.90, 0.95, 1.00, 1.05, 1.10, 1.15, 1.20, 1.25, 1.30},
	xlabel={regularisation exponent ($\lambda$)},
	ymin=2.0,
	ymax=3.0,
	ylabel={Error distribution (\ac{rmse})},
	title={Error distribution by regularisation exponent (\textbf{Abalone dataset})},
	mark options={scale=0.5},
	]
	\addplot[only marks, mark=o,color=black] table [x expr=1+0.2*rand,y index=0] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=2+0.2*rand,y index=1] {data/figure1.dat};		
	\addplot[only marks, mark=o,color=black] table [x expr=3+0.2*rand,y index=2] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=4+0.2*rand,y index=3] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=5+0.2*rand,y index=4] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=6+0.2*rand,y index=5] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=7+0.2*rand,y index=6] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=8+0.2*rand,y index=7] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=9+0.2*rand,y index=8] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=10+0.2*rand,y index=9] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=11+0.2*rand,y index=10] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=12+0.2*rand,y index=11] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=13+0.2*rand,y index=12] {data/figure1.dat};

	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
	\addplot[boxplot] table [y index=0] {data/figure1.dat};
	\addplot[boxplot] table [y index=1] {data/figure1.dat};
	\addplot[boxplot] table [y index=2] {data/figure1.dat};
	\addplot[boxplot] table [y index=3] {data/figure1.dat};
	\addplot[boxplot] table [y index=4] {data/figure1.dat};
	\addplot[boxplot] table [y index=5] {data/figure1.dat};
	\addplot[boxplot] table [y index=6] {data/figure1.dat};
	\addplot[boxplot] table [y index=7] {data/figure1.dat};		
	\addplot[boxplot] table [y index=8] {data/figure1.dat};
	\addplot[boxplot] table [y index=9] {data/figure1.dat};
	\addplot[boxplot] table [y index=10] {data/figure1.dat};
	\addplot[boxplot] table [y index=11] {data/figure1.dat};
	\addplot[boxplot] table [y index=12] {data/figure1.dat};

\end{axis}
\end{tikzpicture}
\includegraphics[height=6cm,width=\textwidth,angle=0]{figures/Abalone_dataset_lambdas.eps}

\caption{Fitness progress for the Abalone dataset.}\label{Abalone_dataset_lambdas}

\end{center}\end{figure}

%\begin{figure}[htb]
%	\begin{center}
%		\includegraphics[height=6cm,width=10cm,angle=0]
%			{figures/Abalone_dataset_lambdas.eps}
%		\caption{fitness progress for Abalone dataset}
%		\label{Abalone_dataset_lambdas}
%	\end{center}
%\end{figure}

Figure~\ref{Auto-Mpg_dataset_lambdas} presents regularization results for the Auto-MPG dataset:

\begin{figure}[htb]
	\begin{center}
		\includegraphics[height=6cm,width=10cm,angle=0]
			{figures/Auto-Mpg_dataset_lambdas.eps}
		\caption{fitness progress for Abalone dataset}
		\label{Auto-Mpg_dataset_lambdas}
	\end{center}
\end{figure}

The typical inflection point is around $\lambda = 0.8$. We'll use the regularisation parameter with this value.

\subsection{Genetic Operators}

\emph{Define the genetic operators}

\subsection{Other Search Parameters}

\emph{Explain the remaining search parameters}

\section{Experimental Results}

\begin{itemize}
\item Measured quantities

	\begin{itemize}

	\item error
	\item number of iterations to convergence
	\item memory usage
	\item F1, ROC, ?

	\end{itemize}

\item Selection of datasets and regression algorithms
\item Summary Figures and Numeric results
\end{itemize}
These results were found using R's programming language~\cite{R}.

To compare this paper's algorithm we applied the exact same train and test samples with other learning algorithms, namely: Linear Regression, Support Vector Machines~\cite{Meyer12}, Regression Trees~\cite{Therneau13}, Conditional Inference Trees~\cite{Hothorn06, Strobl07, Strobl08} and Random Forests~\cite{Liaw02}

In order to train and test the performance of \ac{GAPoly} it was used several datasets (see below). For each dataset, we selected 70\% for training purposes and the remaining observations to make the test set in order to compute the estimated \ac{rmse}. To achieve more robust results, for most datasets there were 25 runs, each one with different samples for the train and test sets. For the datasets with attribute values of different magnitudes, a preliminary scaling was executed. The results below are boxplots for the test set error predictions over these different runs.

\textbf{Artificial}: this is an artificial dataset with four numeric features, $x_1, \ldots x_4$, where $x_1,x_3$ are outcomes from Poisson random variables, and $x_2,x_4$ from Normal random variables. The dependent variable $y$ is given by expression $x_2x_4^2 + x_1^2x_3 + 5$. The dataset includes 50 observations.

\begin{figure}[htb]
	\begin{center}
		\includegraphics[height=6cm,width=10cm,angle=0]
			{figures/artificial_dataset1_lambda1_0.eps}
		\caption{results for Artificial dataset}
		\label{artificial_dataset1_lambda1.0}
	\end{center}
\end{figure}

This dataset was used in order to verify if \ac{GAPoly} was able to find the polynomial relation, which the algorithm did (cf. figure~\ref{artificial_dataset1_lambda1.0}). In this case, all runs used a population of 100 with 50 iterations for each run.

In the next datasets, the population has size 250 with 100 iterations for each run.

\textbf{Housing}: This data set concerns the task of predicting housing values in areas of Boston. There are 13 continuous attributes and the dependent variable is the median value of owner-occupied homes in \$1000's\footnote{This and the remaining regression datasets were selected from Luis Torgo's data repository, http://www.dcc.fc.up.pt/~ltorgo/Regression/DataSets.html. Most of these set originally come from UCI ML repository, http://archive.ics.uci.edu/ml/}. There are 506 observations. Figure~\ref{Housing_dataset_lambda0.8_25runs} presents the results for this dataset.

\begin{figure}[htb]
	\begin{center}
		\includegraphics[height=6cm,width=10cm,angle=0]
			{figures/Housing_dataset_lambda0_8_25runs.eps}
		\caption{results for Boston Housing dataset}
		\label{Housing_dataset_lambda0.8_25runs}
	\end{center}
\end{figure}

Just as an example of the model the \ac{GAPoly} algorithm outputs: in this dataset the best polynomial, with a \ac{rmse} of 0.38 for the test set, was the following:

$$y = -0.12 x_6 x_9^4 + 0.78 x_6 - 0.4 x_9^2 x_{13} + 0.17 x_{13}^2 - 0.044$$

where the attributes mean:
$x_6$, RM average number of rooms per dwelling;
$x_9$, RAD index of accessibility to radial highways;
$x_{13}$, Lower status of the population.

\textbf{Abalone}: This dataset can be used to predict the age of a abalone shell using the given 8 numeric attributes concerning several physical measurements. There are 4177 observations. Figure~\ref{Abalone_dataset_lambda0.8_25runs} presents the results for this dataset.

\begin{figure}[htb]
	\begin{center}
		\includegraphics[height=6cm,width=10cm,angle=0]
			{figures/Abalone_dataset_lambda0_8_25runs.eps}
		\caption{results for Abalone dataset}
		\label{Abalone_dataset_lambda0.8_25runs}
	\end{center}
\end{figure}

% best polynomial, rsme 0.65
% $$y = 0.52 s_1^2 x_5 - 0.13 x_2 x_8 -0.75 x_1^2 x_6 + 0.78 x_8 + 0.18 $$

\textbf{Auto-MPG}: This dataset is used to predict fuel consumption in miles per gallon, based on two discrete and five continuous attributes. There are 398 observations. Figure~\ref{Auto-Mpg_dataset_lambda0.8_25runs} presents the results for this dataset.

\begin{figure}[htb]
	\begin{center}
		\includegraphics[height=6cm,width=10cm,angle=0]
			{figures/Auto-Mpg_dataset_lambda0_8_25runs.eps}
		\caption{results for Auto-MPG dataset}
		\label{Auto-Mpg_dataset_lambda0.8_25runs}
	\end{center}
\end{figure}

% best polynomial, \ac{rmse} 0.32
%$$ y = 0.4 x_6 + 0.02 x_5^3 x_6 - 0.85 x_4 + 0.23 x_4^2 - 0.26 $$

\textbf{Kinematics}: This dataset is concerned with the realistic simulation of the forward kinematics of an 8 link robot arm. The task is to predict the distance of the end-effector from a target using 8 continuous attributes. There are 8192 observations. Figure~\ref{Kinematics300_lambda0.8_25runs} presents the results for this dataset.

\begin{figure}[htb]
	\begin{center}
		\includegraphics[height=6cm,width=10cm,angle=0]
			{figures/Kinematics300_lambda0_8_25runs.eps}
		\caption{results for Kinematics dataset}
		\label{Kinematics300_lambda0.8_25runs}
	\end{center}
\end{figure}

% best polynomial, \ac{rmse} 0.74
% $$ y = 0.24 x_5 - 0.23 x_5 x_6 x_7 - 0.19 x_4 x_7 - 0.5 x_3 $$

\subsection{Convergence speed}

The \ac{GA} quickly proceeds in the first 50 to 100 generations to reasonable error rates. Then, it proceeds slower achieving best solutions with marginal error reduction. Since the entire learning process takes some time, in the current R implementation\footnote{The processing time can probably be speeded by one to two orders in magnitude if the process is implemented in a low level programming language like C++. However, speed optimisation was not the focus of this article}, placing a limit between 50 to 100 generations already achieves good results, relative to higher iteration values. Figure~\ref{Abalone_fitnessProgress} shows a typical error evolution for the dataset Abalone given two different values for $\lambda$.

\begin{figure}[htb]
	\begin{center}
		\includegraphics[height=6cm,width=10cm,angle=0]
			{figures/Abalone_fitnessProgress.eps}
		\caption{fitness progress for Abalone dataset}
		\label{Abalone_fitnessProgress}
	\end{center}
\end{figure}

\section{Conclusion}

\subsection{Discussion}

The proposed method has competitive results comparing with some well-known regression methods. Only Random Forest outperforms \ac{GAPoly} systematically (it outperforms all the others) except in the Auto-MPG test and, of course, in the artificial dataset that uses a straightforward polynomial relation.

\section{Bibliography}

\bibliographystyle{plain}
\bibliography{fullbib}

\end{document}

### 2013-07-17 texts/data/process-this-shit.py

@@ -1,0 +1,48 @@
+import numpy as np
+
+
+def by_rows(d):
+	s = ""
+	ks = sorted(d.keys())
+	for k in ks:
+		s += "%05.3f "%eval(k)
+		for x in d[k]:
+			s+= "%05.3f "%x
+
+		s += "\n"
+	s += "\n"
+	return s
+
+def by_cols(d):
+	s = ""
+	ks = sorted(d.keys())
+	for k in ks:
+		 s += "%04.3f "%eval(k) 
+	s += "\n"
+	n = len( d[ ks[0] ] )
+
+	for i in range(n):
+		for k in ks:
+			s += "%04.3f "%d[k][i]
+		s += "\n"
+	s += "\n"
+	return s
+
+filename = "fc-abalone_lambda-errors.txt"
+
+f = file(filename,"r")
+
+# skip head
+f.readline()
+
+data = dict()
+
+for line in f.readlines():
+	line = line.replace("\n","").split(" ")
+	k = line[2]
+	if k in data.keys():
+		data[k].append( eval(line[1]) )
+	else:
+		data[k] = [ eval(line[1]) ]
+
+print by_cols(data)

### 2013-07-18 texts/paper-draft.tex
Explaining better the introduction of section 2.

@@
 \usepackage{amsfonts}
+
+\usepackage{amsthm}
+\newtheorem{theorem}{Theorem}
+
 \usepackage{graphicx}
@@
 \date{[{\sc draft} \today]} % delete this line to display the current date
+
+\newcommand{\sample}[1]{\ensuremath{^{\left(#1\right)}}}
 
@@
 
-With notable exceptions (\emph{e.g.} neural networks), most machine learning regression techniques are based on linear models.
-This assumption has many advantages including reduced computational complexity and strong theoretical framework.
+With notable exceptions (\emph{e.g.} neural networks) machine learning regression techniques are based on linear models.
+The linearity assumption has many advantages including reduced computational complexity and strong theoretical framework.
 However nonlinearity is unavoidable in many application scenarios, specially those with phase transitions or feedback loops, so common in ecology, cybernetics, robotics and other areas.
 
-Polynomials, one of the most studied subjects in mathematics, generalize linear functions and define, perhaps, the simplest and most used nonlinear models.
-Applications of polynomial include colorimetric calibration \cite{mendes2005adaptive}, explicit \emph{formulae} for turbulent pipe flows \cite{davidson1999method}, computational linguistics \cite{sanchez2009obtaining} and more recently, analytical techniques for cultural heritage materials \cite{csefalvayova2010use}, liquid epoxy molding process \cite{chan2011modeling}, B-spline surface reconstruction \cite{galvez2012iterative}, product design \cite{chan2012development} or forecasting cyanotoxins presence in water reservoirs \cite{garcia2013hybrid}.
-These examples not only illustrate the wide spectrum of applications but also the work in each one of these polynomial models uses, at some point, a genetic algorithm.
+Polynomials, one of the most studied subjects in mathematics, generalize li\-ne\-ar functions and define, perhaps, the simplest and most used nonlinear models.
+Applications include colorimetric calibration \cite{mendes2005adaptive}, explicit formulae for turbulent pipe flows \cite{davidson1999method}, computational linguistics \cite{sanchez2009obtaining} and more recently, analytical techniques for cultural heritage materials \cite{csefalvayova2010use}, liquid epoxy molding process \cite{chan2011modeling}, B-spline surface reconstruction \cite{galvez2012iterative}, product design \cite{chan2012development} or forecasting cyanotoxins presence in water reservoirs \cite{garcia2013hybrid}.
+This not only illustrates the wide spectrum of applications but, additionally, work in each one of these polynomial models uses, at some point, a genetic algorithm.
 
-\ac{GA} where, arguably, one the hottest topics of research in the recent decades and with good reason since they outline an optimisation scheme easy to conceptualise and with very broad application.
-If a nonlinear (or otherwise) model requires parameterisation, \acp{GA} provide a simple and often effective approach to search for locally optimal parameters.
-Research related to genetic algorithms abound and spans from the 1950s seminal work of Nils Aall Barricelli \cite{barricelli1962numerical} in the Institute for Advanced Study of Princeton to today's principal area of study of thousands of researchers, covered in hundreds of conferences, workshops and other meetings.
+\ac{GA} where, arguably, one the hottest topics of research in the recent decades and with good reason since they outline an optimization scheme easy to conceptualize and with very broad application.
+If a nonlinear (or otherwise) model requires parameterization, \acp{GA} provide a simple and often effective approach to search for locally optimal parameters.
+Research related to genetic algorithms abound and spans from the 1950s seminal work of Nils Aall Barricelli \cite{barricelli1962numerical} in the Institute for Advanced Study of Princeton to today's principal area of study for thousands of researchers, covered in hundreds of conferences, workshops and other meetings.
 Perhaps the key impulse to \acp{GA} come from John Holland's work and his book ``Adaptation in Natural and Artificial Systems'' \cite{holland1975adaptation}.
 
-One interesting ``flavour" of genetic algorithms, named \emph{genetic programming} by John Koza \cite{koza1992genetic}, proposed the use of \acp{GA} to search the syntatic stucture of complex functions.
+One interesting take on genetic algorithms, named \emph{genetic programming} by John Koza \cite{koza1992genetic}, proposed the use of \acp{GA} to search the syntactic structure of complex functions.
 This syntatic structure search is keen to the central ideas of deep learning \cite{bengio2013representation, bengio2009learning}, the subarea of machine learning actually producing the most promising results (\emph{e.g.} \cite{tarlowstochastic}).
-It is also related to the work presented in this paper in the sense that, unlike linear models that have a simple structure ($y=\sum_i \beta_i x_i$) nonlinear (in particular polynomial) models pose an additional ``structure" search problem.
+It is also related to the work presented in this paper in the sense that, unlike linear models that have a simple structure, $y=\sum_i \beta_i x_i$, nonlinear (in particular polynomial) models pose an additional ``structure" search problem.
 
@@
 In line with that research this work describes a general method to find a polynomial regression of a given dataset.
-The optimal regression minimises a cost function that accounts for both the \ac{rmse} and a regularisation factor to avoid over-fitting.
+The optimal regression minimises a cost function that accounts for both the \ac{rmse} and a regularization factor to avoid over-fitting.
 
@@
 
-The remainder of this paper is organised as usual: the next section describes the details of our method and is followed by a presentation of some performance results.
+The remainder of this paper is organized as usual: the next section describes the details of our method and is followed by a presentation of some performance results.
 The last section draws some conclusions and points future research tasks.
 
-\section{Polynomial Regression with Genetic Algorithms}
+\section{Genetic Algorithms for Polynomials}
 
-This section starts with a brief description of the polynomial representation used to support the encoding of polynomial instances.
-This encoding is required to define the populations the \acp{GA} will evolve.
-Then is defined the cost function, including regularisation for over-fitting.
+This section is dedicated to describe an algorithm to find a polynomial regression from a given dataset.
+It starts with a brief introduction and description of the algorithm and proceeds into core details as the encoding used to represent individual polynomial instances in the \ac{GA} populations and the regularization of the cost function.
 
@@
 $$
-p\left( x_1, \ldots, x_k\right) = \sum_i \beta_i m_i
+p\left( x_1, \ldots, x_k\right) = \sum_i \theta_i m_i
 $$
-where each $m_i$ is a monomial, $m_i = \prod_{j\in A_i} x_j^{\alpha_{ij}}$, the exponents are non-negative integers, $\alpha_{ij}\in\mathbb{N}_0$, and the coefficients are real valued, $\beta_i \in \mathbb{R}$.
-For example $p\left( x_1, x_2, x_3 \right) = 2 x_1 + x_2 x_3 + \frac{1}{2} x_1^2 x_3$ has monomials $m_1 = x_1, m_2 = x_2 x_3$ and $m_3 = x_1^2 x_3$ coefficients $\beta_1 = 2, \beta_2 = 1$ and $\beta_3 = 1/2$ and exponents $\alpha_{1,1} = 1, \alpha_{2,2}=1, \alpha_{2,3} = 1, \alpha_{3,1} = 2, \alpha_{3,3} = 1$ and all other $\alpha_{ij} = 0$.
-
-By using the exponents alone we have a matrix that defines the monomial structure of the polynomial, $A = \left[\alpha_{ij}\right]$. For the example above
+where each $m_i$ is a monomial, $m_i = \prod_{j} x_j^{\alpha_{ij}}$, the exponents are non-negative integers, $\alpha_{ij}\in\mathbb{N}_0$, and the coefficients are real valued, $\theta_i \in \mathbb{R}$.
+For example $p\left( x_1, x_2, x_3 \right) = 2 x_1 + x_2 x_3 + \frac{1}{2} x_1^2 x_3$ has monomials $m_1 = x_1, m_2 = x_2 x_3$ and $m_3 = x_1^2 x_3$ coefficients $\theta_1 = 2, \theta_2 = 1$ and $\theta_3 = 1/2$ and exponents $\alpha_{1,1} = 1, \alpha_{2,2}=1, \alpha_{2,3} = 1, \alpha_{3,1} = 2, \alpha_{3,3} = 1$ and all other $\alpha_{ij} = 0$.
+The exponents alone are a matrix that defines the monomial structure of the polynomial, $A = \left[\alpha_{ij}\right]$.
+For the example above
 $$
@@
 Changing the order of the rows doesn't change the polynomial whereas changing the order of the columns corresponds to changing the respective variables.
-%Notice that if we interpret matrix $A$ as a single column of monomials and set $B = \left[ \begin{array}{ccc} 2 & 1 & \frac{1}{2} \end{array} \right ]$ then
+%Notice that if we interpret matrix $A$ as a single column of monomials and set the coefficients $T = \left[ \begin{array}{ccc} 2 & 1 & \frac{1}{2} \end{array} \right ]$ then
 %$$
-%p\left( x_1, x_2, x_3 \right) = BA = \left[ \begin{array}{ccc} 2 & 1 & \frac{1}{2} \end{array} \right ]\left[ 
+%p\left( x_1, x_2, x_3 \right) = TA = \left[ \begin{array}{ccc} 2 & 1 & \frac{1}{2} \end{array} \right ]\left[ 
 %\begin{array}{ccc}
@@
 But more importantly, the polynomial regression problem can be naturally split into two subproblems:
+%
+\begin{enumerate}
+\item for a given set of monomials $m_1, \ldots, m_R$, find the regression coefficients $\theta_1,\ldots,\theta_R$ that minimize the \ac{rmse} on a given dataset;
 
-\begin{enumerate}
-\item for a given set of monomials $q_1, \ldots, q_r$, find regression coefficients $\theta_1,\ldots,\theta_r$ that minimize an adequate cost function;
-\item find the fittest set of monomials;
+\item find the fittest set of monomials, \emph{i.e.} the polynomial that minimizes the \ac{rmse} on the same dataset;
 \end{enumerate}
+%
+More precisely, concerning the first problem, let $\mathcal{D}$ be a dataset with $N$ observations of variables $Y, X_1,\ldots,X_n$ and $\mathcal{M} = \left\lbrace m_1,\ldots, m_R\right\rbrace$ a set of $R$ monomial expressions over $X_1,\ldots,X_n$.
+Define the hypothesis\footnote{The expression $m|_{X=x}$ reads ``\emph{replace all instances of $X$ by $x$ in $m$}''.}
+%
+\begin{eqnarray*}
+h_{\Theta,\mathcal{M}}\left(x_1,\ldots,x_n\right) &=& \sum_{j = 1}^R \theta_j m_j|_{X_i=x_i,\forall 1 \leq i \leq n}
+\end{eqnarray*}
+%
+and let the cost
+%
+\begin{eqnarray}
+J\left(\Theta;\mathcal{M},\mathcal{D}\right) &=& \sqrt{\frac{1}{N}\sum_{i=1}^N \left( y\sample{i} - h_{\Theta,\mathcal{M}}\left( x_1\sample{i},\ldots,x_n\sample{i} \right) \right)^2 }\label{eq:rmse}
+\end{eqnarray}
+%
+be the usual \acf{rmse} function. Now the first problem can be stated as:
+%
+\begin{quote}
+Given a dataset $\mathcal{D}$ and a set of monomials $\mathcal{M}$, find parameters $\Theta$ that minimize $J\left(\Theta;\mathcal{M},\mathcal{D}\right)$.
+\end{quote}
+%
+It turns out that this problem can be solved as a usual linear regression problem by expanding $\mathcal{D}$ with columns that replicate the monomials in $\mathcal{M}$.
 
-Algorithm \ref{alg:gapoly} solves the first problem with linear regression and the second with \acp{GA}.
-
+The second problem is treated in the \ac{GA} setting: Let $\mathcal{D}$ be a dataset as above and $\mathcal{P}$ a set of polynomials.
+For each polynomial $p\in \mathcal{P}$ let $\mathcal{M}_p$ be the set of monomials in $p$ (without the coefficients) and compute the fitness
+%
+\begin{eqnarray*}
+\phi_p &=& \min_\Theta J\left(\Theta;\mathcal{M}_p,\mathcal{D}\right)
+\end{eqnarray*}
+%
+by solving the first problem.
+With a fitness of every instance, a \ac{GA} will apply genetic operators (usually mutation and crossover) to evolve the population $\mathcal{P}$ until a reasonable approximation of a local minimum is found.
+%
 \begin{algorithm}[htb]
 \begin{algorithmic}
-\Function{J}{$D,\left[m_1,\ldots,m_k\right]$}\Comment{Cost, or error, of the polynomial $m_1,\ldots,m_k$ in the dataset $D$}
-	\State\Return $err \gets \Call{lm}{D,\left[m_1,\ldots,m_k\right]}$ \Comment{Error in Linear Regression}
-\EndFunction
-\Statex
 \Function{GAPoly}{$D,pop_0,\epsilon$}
-	\State $pop \gets pop_0$
+	\State $pop \gets pop_0;\: err \gets 1.0+\epsilon$
 	\While{$err > \epsilon$}
 		\State $pop \gets \Call{IterateGA}{pop}$\Comment{Build next generation}
-		\State $pop \gets \Call{Sort}{pop, key = \textsc{Error}}$\Comment{Sort population by error}
-		\State $err \gets \textrm{error of }\Call{First}{pop}$
+		\State $pop \gets \Call{Sort}{pop, key = \textsc{Lm.rmse}}$\Comment{Sort by \ac{rmse} of linear regression}
+		\State $err \gets \textsc{rmse}\left(\Call{First}{pop}\right)$
 	\EndWhile
@@
 \end{algorithmic}
-\caption{\ac{GAPoly} uses linear regression to find the monomials coefficients and genetic algorithms to explore the space of monomials.}\label{alg:gapoly}
+\caption{\ac{GAPoly} uses linear regression to find monomial coefficients that minimize the \ac{rmse} over a dataset and \acp{GA} to explore the space of polynomials.
+At exit the \ac{rmse} of the fittest instance is bounded by $\epsilon$.}\label{alg:gapoly}
 \end{algorithm} 
-
-[EVALUATION STRATEGY]
-
-[RESULTS]
-
-\subsection{Polinomial Encoding}
+%
+\begin{theorem}
+The composition of \acp{GA} with linear regression, as defined in Algorithm \ref{alg:gapoly}, converges to a polynomial that is a local minimum of the \ac{rmse} function, encapsulated in the fitness function $J$.
+\end{theorem}
+%
+\begin{proof}
+It follows from the properties of \acp{GA} and linear regression.
+\end{proof}
+%
+Subsection \ref{subs:polynomial.encoding} details of the encoding of individual polynomial instances as chromosomes and other parameters of the \ac{GA} implementation used.
+The regularization of the cost function is discussed in subsection \ref{subs:cost.function}.
+\subsection{Polinomial Encoding}\label{subs:polynomial.encoding}
 
@@
 
-\subsection{Cost Function}
+\subsection{Cost Function}\label{subs:cost.function}
 

### 2013-07-19 tests_Housing.Rmd
change Pictures, clean folder

@@
 ########################################################
+# check what attributes Random Forest select has the most importants
+
+train.p.size <- 0.7 # percentage of training set
+n.vars       <- ncol(my.data)-1
+
+inTrain   <- sample(1:nrow(my.data), train.p.size * nrow(my.data))
+train.set <- my.data[inTrain,]
+test.set  <- my.data[-inTrain,]
+
+rf.model <- randomForest(y ~ ., data=train.set, importance=TRUE, do.trace=100, ntree=100)
+rf.model$importance
 

### 2013-07-19 tests_test.Rmd
change Pictures, clean folder

@@
 source("gapolyfuncs.R")
-report <- make.report(my.data, population=100, iterations=50, mutation.rate=0.05)
+report <- make.report(my.data, population=100, iterations=50, mutation.rate=0.15)
 
@@
 
-# df <- data.frame(GA.Poly = report$ga.error,
-#                  Lin.Reg = report$lm.error,
-#                  SVM = report$svm.error,
-#                  RPART = report$rpart.error,
-#                  Rnd.For = report$rf.error,
-#                  CI.Tree = report$citree.error)
-# 
-# write.table(df, "Articial_dataset.txt")
+df <- data.frame(GA.Poly = report$ga.error,
+                 Lin.Reg = report$lm.error,
+                 SVM = report$svm.error,
+                 RPART = report$rpart.error,
+                 Rnd.For = report$rf.error,
+                 CI.Tree = report$citree.error)
+
+write.table(df, "Articial_dataset.txt")
 

### 2013-07-19 texts/paper-draft.tex
change Pictures, clean folder

@@
 
-\ac{GA} where, arguably, one the hottest topics of research in the recent decades and with good reason since they outline an optimization scheme easy to conceptualize and with very broad application.
+\ac{GA} where, arguably, one of the hottest topics of research in the recent decades and with good reason since they outline an optimization scheme easy to conceptualize and with very broad application.
 If a nonlinear (or otherwise) model requires parameterization, \acp{GA} provide a simple and often effective approach to search for locally optimal parameters.
@@
 
-The chromosome coding will as follow:
+The encoding for any polynomial will be as follows:
 \begin{enumerate}
-\item an initial segment detailing which monomials are active (the 1st monomial is always active), these are represented in unary description
-\item the chromosome is split into $m$ sets of bits of equal size 
-  (a monomial)
+\item an initial segment detailing which monomials are active (the 1st monomial is always active), this is represented in unary description, i.e., each monomial is identified by a single bit
+\item the remaining is split into $m$ sets of bits of equal size, each one representing a monomial
 \item each monomial is split into $n$ sets of $d$ size each, i.e., a variable
@@
 
-Eg: consider polynomial $x_1^3 x_3 + x_3^7 + x_1 x_2$ with $m = 4, n = 3$ and $d=3$. One possible encoding would be:
+Let's see an example: consider polynomial $x_1^3 x_3 + x_3^7 + x_1 x_2$ with $m = 4, n = 3$ and $d=3$. One possible encoding would be:
 $$110 - 011,000,001~~;~~000,000,111~~;~~001,001,000~~;~~110,010,101$$
@@
 
-The first $3$ bits inform that the second and third monomials are active while the fourth is not (as said, the first monomial is always active).
+The first three bits inform that the second and third monomials are active while the fourth is not (as said, the first monomial is always active).
 This last monomial does not enter neither in the polynomial regression nor in the fitness evaluation.
-However, it acts as a kind of junk DNA, becoming active when the third bit of the entire sequence flips from 0 to 1.
+However, it acts as a kind of junk DNA, becoming active when, in a future mutation or crossover, the third bit of the entire sequence flips from 0 to 1.
 
@@
 The third triple $001$ says that variable $x_3$ has exponent $1$, so the first monomial consists of $x_1^3 x_3$.
-All the remaining sets of $9$ bits are interpreted the same way and we get the previous polynomial.
+All the remaining sets of nine bits are interpreted the same way and we get the previous polynomial.
 
 Notice that all binary descriptions give rise to valid polynomials.
-However, if the coding consists of entirely zeros, by convention, it describes polynomial $x_1$.
+Notice however that this is not a bijective mapping.
+For each polynomial there are multiple representations.
+For example, $x_1+x_2$ and $x_2+x_1$ have different representations.
+The authors considered that more complex mappings in order to force a one to one mapping would impact negatively in the algorithm performance without giving anything in return.
+
+If the coding consists of entirely zeros, by convention, it describes polynomial $x_1$.
 This has to do with the execution of the polynomial regression that would fail if we interpret it as the zero polynomial.
@@
 Somewhat unexpectedly after some experiences it was found that lower values for $\lambda$ sometimes provide better, even if marginal, results.
-Figure~\ref{Abalone_dataset_lambdas} shows results for the Abalone dataset with five runs for each $\lambda$ (cf.
-Experimental Results includes information about these datasets).
+Figure~\ref{Abalone_dataset_lambdas} shows results for the Abalone dataset with ten runs for each $\lambda$, and figure~\ref{Auto-MPG_dataset_lambdas} presents regularization results for the Auto-MPG dataset.
+The following section Experimental Results includes information about these datasets.
 
@@
 \end{tikzpicture}
-\includegraphics[height=6cm,width=\textwidth,angle=0]{figures/Abalone_dataset_lambdas.eps}
 
-\caption{Fitness progress for the Abalone dataset.}\label{Abalone_dataset_lambdas}
+% \includegraphics[height=6cm,width=\textwidth,angle=0]{figures/Abalone_dataset_lambdas.eps}
+
+\caption{Error distribution by regularisation exponent for the Abalone dataset.}\label{Abalone_dataset_lambdas}
 
@@
 
-%\begin{figure}[htb]
-%	\begin{center}
-%		\includegraphics[height=6cm,width=10cm,angle=0]
-%			{figures/Abalone_dataset_lambdas.eps}
-%		\caption{fitness progress for Abalone dataset}
-%		\label{Abalone_dataset_lambdas}
-%	\end{center}
-%\end{figure}
+\begin{figure}[htb]\begin{center}
+\begin{tikzpicture}
+\begin{axis}[width=\textwidth,height=6cm,
+	boxplot/draw direction=y,
+	xtick={1,2,3,4,5,6,7,8,9,10,11,12,13},
+	xticklabels={0.70, 0.75, 0.80, 0.85, 0.90, 0.95, 1.00, 1.05, 1.10, 1.15, 1.20, 1.25, 1.30},
+	xlabel={regularisation exponent ($\lambda$)},
+	ymin=2.0,
+	ymax=5.0,
+	ylabel={Error distribution (\ac{rmse})},
+	title={Error distribution by regularisation exponent (\textbf{Auto-MPG dataset})},
+	mark options={scale=0.5},
+	]
+	\addplot[only marks, mark=o,color=black] table [x expr=1+0.2*rand,y index=0] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=2+0.2*rand,y index=1] {data/lambda.errors.Auto-Mpg.dat};		
+	\addplot[only marks, mark=o,color=black] table [x expr=3+0.2*rand,y index=2] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=4+0.2*rand,y index=3] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=5+0.2*rand,y index=4] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=6+0.2*rand,y index=5] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=7+0.2*rand,y index=6] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=8+0.2*rand,y index=7] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=9+0.2*rand,y index=8] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=10+0.2*rand,y index=9] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=11+0.2*rand,y index=10] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=12+0.2*rand,y index=11] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=13+0.2*rand,y index=12] {data/lambda.errors.Auto-Mpg.dat};
 
-Figure~\ref{Auto-Mpg_dataset_lambdas} presents regularization results for the Auto-MPG dataset:
+	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
+	\addplot[boxplot] table [y index=0] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[boxplot] table [y index=1] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[boxplot] table [y index=2] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[boxplot] table [y index=3] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[boxplot] table [y index=4] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[boxplot] table [y index=5] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[boxplot] table [y index=6] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[boxplot] table [y index=7] {data/lambda.errors.Auto-Mpg.dat};		
+	\addplot[boxplot] table [y index=8] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[boxplot] table [y index=9] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[boxplot] table [y index=10] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[boxplot] table [y index=11] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[boxplot] table [y index=12] {data/lambda.errors.Auto-Mpg.dat};
 
-\begin{figure}[htb]
-	\begin{center}
-		\includegraphics[height=6cm,width=10cm,angle=0]
-			{figures/Auto-Mpg_dataset_lambdas.eps}
-		\caption{fitness progress for Abalone dataset}
-		\label{Auto-Mpg_dataset_lambdas}
-	\end{center}
-\end{figure}
+\end{axis}
+\end{tikzpicture}
 
-The typical inflection point is around $\lambda = 0.8$. We'll use the regularisation parameter with this value.
+% \includegraphics[height=6cm,width=\textwidth,angle=0]{figures/Abalone_dataset_lambdas.eps}
+
+\caption{Error distribution by regularisation exponent for the Auto-MPG dataset.}\label{Auto-MPG_dataset_lambdas}
+
+\end{center}\end{figure}
+
+The typical inflection point lies around $\lambda = 0.8$. The following results use the regularisation parameter with this value.
 
@@
 
-\emph{Define the genetic operators}
-
-\subsection{Other Search Parameters}
-
-\emph{Explain the remaining search parameters}
+It was used the R package genalg~\cite{willighage2012genalg} for the execution of genetic algorithm.
+The operators were the standard ones: (a) crossover, i.e., a pair of solutions from the previous generations are combined by splitting and mixing their respective representations, and (b) mutation, changing the values of single bits; the mutation chance applied in the datasets was 5\%.
+There was also elitism between generations, i.e., 20\% of the best solutions are kept in the next generation.
 
@@
 
-To compare this paper's algorithm we applied the exact same train and test samples with other learning algorithms, namely: Linear Regression, Support Vector Machines~\cite{Meyer12}, Regression Trees~\cite{Therneau13}, Conditional Inference Trees~\cite{Hothorn06, Strobl07, Strobl08} and Random Forests~\cite{Liaw02}
+To compare this paper's proposed algorithm we applied the exact same train and test samples using other learning algorithms, namely: Linear Regression, Support Vector Machines~\cite{Meyer12}, Regression Trees~\cite{Therneau13}, Conditional Inference Trees~\cite{Hothorn06, Strobl07, Strobl08} and Random Forests~\cite{Liaw02}
 
 In order to train and test the performance of \ac{GAPoly} it was used several datasets (see below).
-For each dataset, we selected 70\% for training purposes and the remaining observations to make the test set in order to compute the estimated \ac{rmse}.
-To achieve more robust results, for most datasets there were 25 runs, each one with different samples for the train and test sets.
+For each dataset, we selected 70\% for training purposes and the remaining observations to make a test set in order to compute the estimated \ac{rmse}.
+To achieve more robust results, each dataset were processed 25 times, each one with different samples for the train and test sets.
 For the datasets with attribute values of different magnitudes, a preliminary scaling was executed.
@@
 The dependent variable $y$ is given by expression $x_2x_4^2 + x_1^2x_3 + 5$.
-The dataset includes 50 observations.
+The dataset includes 50 observations and, only in this case, there were executed just 10 runs.
 
-\begin{figure}[htb]
-	\begin{center}
-		\includegraphics[height=6cm,width=10cm,angle=0]
-			{figures/artificial_dataset1_lambda1_0.eps}
+\begin{figure}[htb]\begin{center}
+\begin{tikzpicture}
+\begin{axis}[width=\textwidth,height=6cm,
+	boxplot/draw direction=y,
+	xtick={1,2,3,4,5,6},
+	xticklabels={GA.Poly,Lin.Reg,SVM,RPART,Rnd.For,CI.Tree},
+	xlabel={Regression Method},
+	ymin=0,
+	ymax=300,
+	ylabel={Error distribution (\ac{rmse})},
+	title={Regression results for \textbf{Artificial dataset}},
+	mark options={scale=0.5},
+	]
+	\addplot[only marks, mark=o,color=black] table [x expr=1+0.2*rand,y index=0] {data/Articial_dataset.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=2+0.2*rand,y index=1] {data/Articial_dataset.dat};		
+	\addplot[only marks, mark=o,color=black] table [x expr=3+0.2*rand,y index=2] {data/Articial_dataset.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=4+0.2*rand,y index=3] {data/Articial_dataset.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=5+0.2*rand,y index=4] {data/Articial_dataset.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=6+0.2*rand,y index=5] {data/Articial_dataset.dat};
+
+	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
+	\addplot[boxplot] table [y index=0] {data/Articial_dataset.dat};
+	\addplot[boxplot] table [y index=1] {data/Articial_dataset.dat};
+	\addplot[boxplot] table [y index=2] {data/Articial_dataset.dat};
+	\addplot[boxplot] table [y index=3] {data/Articial_dataset.dat};
+	\addplot[boxplot] table [y index=4] {data/Articial_dataset.dat};
+	\addplot[boxplot] table [y index=5] {data/Articial_dataset.dat};
+
+\end{axis}
+\end{tikzpicture}
+
+% \includegraphics[height=6cm,width=\textwidth,angle=0]{figures/Abalone_dataset_lambdas.eps}
+
 		\caption{results for Artificial dataset}
 		\label{artificial_dataset1_lambda1.0}
-	\end{center}
-\end{figure}
+
+\end{center}\end{figure}
 
@@
 figure~\ref{artificial_dataset1_lambda1.0}).
-In this case, all runs used a population of 100 with 50 iterations for each run.
+In this case, we used a population of 100 solutions with 50 iterations for each run.
 
@@
 There are 13 continuous attributes and the dependent variable is the median value of owner-occupied homes in \$1000's\footnote{This and the remaining regression datasets were selected from Luis Torgo's data repository, http://www.dcc.fc.up.pt/~ltorgo/Regression/DataSets.html.
-Most of these set originally come from UCI ML repository, http://archive.ics.uci.edu/ml/}.
+Most of these datasets originally come from UCI ML repository, http://archive.ics.uci.edu/ml/}.
 There are 506 observations.
@@
 
-\begin{figure}[htb]
-	\begin{center}
-		\includegraphics[height=6cm,width=10cm,angle=0]
-			{figures/Housing_dataset_lambda0_8_25runs.eps}
+\begin{figure}[htb]\begin{center}
+\begin{tikzpicture}
+\begin{axis}[width=\textwidth,height=6cm,
+	boxplot/draw direction=y,
+	xtick={1,2,3,4,5,6},
+	xticklabels={GA.Poly,Lin.Reg,SVM,RPART,Rnd.For,CI.Tree},
+	xlabel={Regression Method},
+	ymin=0.25,
+	ymax=0.85,
+	ylabel={Error distribution (\ac{rmse})},
+	title={Regression results for \textbf{Housing dataset}},
+	mark options={scale=0.5},
+	]
+	\addplot[only marks, mark=o,color=black] table [x expr=1+0.2*rand,y index=0] {data/Housing300_lambda0.8_25runs.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=2+0.2*rand,y index=1] {data/Housing300_lambda0.8_25runs.dat};		
+	\addplot[only marks, mark=o,color=black] table [x expr=3+0.2*rand,y index=2] {data/Housing300_lambda0.8_25runs.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=4+0.2*rand,y index=3] {data/Housing300_lambda0.8_25runs.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=5+0.2*rand,y index=4] {data/Housing300_lambda0.8_25runs.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=6+0.2*rand,y index=5] {data/Housing300_lambda0.8_25runs.dat};
+
+	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
+	\addplot[boxplot] table [y index=0] {data/Housing300_lambda0.8_25runs.dat};
+	\addplot[boxplot] table [y index=1] {data/Housing300_lambda0.8_25runs.dat};
+	\addplot[boxplot] table [y index=2] {data/Housing300_lambda0.8_25runs.dat};
+	\addplot[boxplot] table [y index=3] {data/Housing300_lambda0.8_25runs.dat};
+	\addplot[boxplot] table [y index=4] {data/Housing300_lambda0.8_25runs.dat};
+	\addplot[boxplot] table [y index=5] {data/Housing300_lambda0.8_25runs.dat};
+
+\end{axis}
+\end{tikzpicture}
+
+% \includegraphics[height=6cm,width=\textwidth,angle=0]{figures/Abalone_dataset_lambdas.eps}
+
 		\caption{results for Boston Housing dataset}
 		\label{Housing_dataset_lambda0.8_25runs}
-	\end{center}
-\end{figure}
 
-Just as an example of the model the \ac{GAPoly} algorithm outputs: in this dataset the best polynomial, with a \ac{rmse} of 0.38 for the test set, was the following:
+\end{center}\end{figure}
+
+Just as an example of the model the \ac{GAPoly} algorithm outputs: in this dataset the best polynomial was the following:
 
@@
 
+For comparisation if we access the mean decrease in accuracy found by Random Forests, the most important attributes are -- in decreasing order -- $x_{13}, x_6, x_5$ (but $x_5$ is already considered $4$ times less important than $x_6$).
+Both algorithms agree in two of their three most important attributes.
+
 \textbf{Abalone}: This dataset can be used to predict the age of a abalone shell using the given 8 numeric attributes concerning several physical measurements.
@@
 
-\begin{figure}[htb]
-	\begin{center}
-		\includegraphics[height=6cm,width=10cm,angle=0]
-			{figures/Abalone_dataset_lambda0_8_25runs.eps}
-		\caption{results for Abalone dataset}
-		\label{Abalone_dataset_lambda0.8_25runs}
-	\end{center}
-\end{figure}
+\begin{figure}[htb]\begin{center}
+\begin{tikzpicture}
+\begin{axis}[width=\textwidth,height=6cm,
+	boxplot/draw direction=y,
+	xtick={1,2,3,4,5,6},
+	xticklabels={GA.Poly,Lin.Reg,SVM,RPART,Rnd.For,CI.Tree},
+	xlabel={Regression Method},
+	ymin=0.64,
+	ymax=0.79,
+	ylabel={Error distribution (\ac{rmse})},
+	title={Regression results for \textbf{Abalone dataset}},
+	mark options={scale=0.5},
+	]
+	\addplot[only marks, mark=o,color=black] table [x expr=1+0.2*rand,y index=0] {data/Abalone300_lambda0.8_25runs.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=2+0.2*rand,y index=1] {data/Abalone300_lambda0.8_25runs.dat};		
+	\addplot[only marks, mark=o,color=black] table [x expr=3+0.2*rand,y index=2] {data/Abalone300_lambda0.8_25runs.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=4+0.2*rand,y index=3] {data/Abalone300_lambda0.8_25runs.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=5+0.2*rand,y index=4] {data/Abalone300_lambda0.8_25runs.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=6+0.2*rand,y index=5] {data/Abalone300_lambda0.8_25runs.dat};
+
+	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
+	\addplot[boxplot] table [y index=0] {data/Abalone300_lambda0.8_25runs.dat};
+	\addplot[boxplot] table [y index=1] {data/Abalone300_lambda0.8_25runs.dat};
+	\addplot[boxplot] table [y index=2] {data/Abalone300_lambda0.8_25runs.dat};
+	\addplot[boxplot] table [y index=3] {data/Abalone300_lambda0.8_25runs.dat};
+	\addplot[boxplot] table [y index=4] {data/Abalone300_lambda0.8_25runs.dat};
+	\addplot[boxplot] table [y index=5] {data/Abalone300_lambda0.8_25runs.dat};
+
+\end{axis}
+\end{tikzpicture}
+
+% \includegraphics[height=6cm,width=\textwidth,angle=0]{figures/Abalone_dataset_lambdas.eps}
+
+\caption{results for Abalone dataset} \label{Abalone_dataset_lambda0.8_25runs}
+
+\end{center}\end{figure}
 
@@
 
-\begin{figure}[htb]
-	\begin{center}
-		\includegraphics[height=6cm,width=10cm,angle=0]
-			{figures/Auto-Mpg_dataset_lambda0_8_25runs.eps}
-		\caption{results for Auto-MPG dataset}
-		\label{Auto-Mpg_dataset_lambda0.8_25runs}
-	\end{center}
-\end{figure}
+\begin{figure}[htb]\begin{center}
+\begin{tikzpicture}
+\begin{axis}[width=\textwidth,height=6cm,
+	boxplot/draw direction=y,
+	xtick={1,2,3,4,5,6},
+	xticklabels={GA.Poly,Lin.Reg,SVM,RPART,Rnd.For,CI.Tree},
+	xlabel={Regression Method},
+	ymin=0.3,
+	ymax=0.55,
+	ylabel={Error distribution (\ac{rmse})},
+	title={Regression results for \textbf{Auto-MPG dataset}},
+	mark options={scale=0.5},
+	]
+	\addplot[only marks, mark=o,color=black] table [x expr=1+0.2*rand,y index=0] {data/Auto-Mpg300_lambda0.8_25runs.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=2+0.2*rand,y index=1] {data/Auto-Mpg300_lambda0.8_25runs.dat};		
+	\addplot[only marks, mark=o,color=black] table [x expr=3+0.2*rand,y index=2] {data/Auto-Mpg300_lambda0.8_25runs.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=4+0.2*rand,y index=3] {data/Auto-Mpg300_lambda0.8_25runs.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=5+0.2*rand,y index=4] {data/Auto-Mpg300_lambda0.8_25runs.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=6+0.2*rand,y index=5] {data/Auto-Mpg300_lambda0.8_25runs.dat};
+
+	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
+	\addplot[boxplot] table [y index=0] {data/Auto-Mpg300_lambda0.8_25runs.dat};
+	\addplot[boxplot] table [y index=1] {data/Auto-Mpg300_lambda0.8_25runs.dat};
+	\addplot[boxplot] table [y index=2] {data/Auto-Mpg300_lambda0.8_25runs.dat};
+	\addplot[boxplot] table [y index=3] {data/Auto-Mpg300_lambda0.8_25runs.dat};
+	\addplot[boxplot] table [y index=4] {data/Auto-Mpg300_lambda0.8_25runs.dat};
+	\addplot[boxplot] table [y index=5] {data/Auto-Mpg300_lambda0.8_25runs.dat};
+
+\end{axis}
+\end{tikzpicture}
+
+% \includegraphics[height=6cm,width=\textwidth,angle=0]{figures/Abalone_dataset_lambdas.eps}
+
+\caption{results for Auto-MPG dataset}	\label{Auto-Mpg_dataset_lambda0.8_25runs}
+
+\end{center}\end{figure}
 
@@
 
-\begin{figure}[htb]
-	\begin{center}
-		\includegraphics[height=6cm,width=10cm,angle=0]
-			{figures/Kinematics300_lambda0_8_25runs.eps}
+\begin{figure}[htb]\begin{center}
+\begin{tikzpicture}
+\begin{axis}[width=\textwidth,height=7cm,
+	boxplot/draw direction=y,
+	xtick={1,2,3,4,5,6},
+	xticklabels={GA.Poly,Lin.Reg,SVM,RPART,Rnd.For,CI.Tree},
+	xlabel={Regression Method},
+	ymin=0.53,
+	ymax=0.85,
+	ylabel={Error distribution (\ac{rmse})},
+	title={Regression results for \textbf{Kinematics dataset}},
+	mark options={scale=0.5},
+	]
+	\addplot[only marks, mark=o,color=black] table [x expr=1+0.2*rand,y index=0] {data/Kinematics300_lambda0.8_25runs.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=2+0.2*rand,y index=1] {data/Kinematics300_lambda0.8_25runs.dat};		
+	\addplot[only marks, mark=o,color=black] table [x expr=3+0.2*rand,y index=2] {data/Kinematics300_lambda0.8_25runs.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=4+0.2*rand,y index=3] {data/Kinematics300_lambda0.8_25runs.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=5+0.2*rand,y index=4] {data/Kinematics300_lambda0.8_25runs.dat};
+	\addplot[only marks, mark=o,color=black] table [x expr=6+0.2*rand,y index=5] {data/Kinematics300_lambda0.8_25runs.dat};
+
+	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
+	\addplot[boxplot] table [y index=0] {data/Kinematics300_lambda0.8_25runs.dat};
+	\addplot[boxplot] table [y index=1] {data/Kinematics300_lambda0.8_25runs.dat};
+	\addplot[boxplot] table [y index=2] {data/Kinematics300_lambda0.8_25runs.dat};
+	\addplot[boxplot] table [y index=3] {data/Kinematics300_lambda0.8_25runs.dat};
+	\addplot[boxplot] table [y index=4] {data/Kinematics300_lambda0.8_25runs.dat};
+	\addplot[boxplot] table [y index=5] {data/Kinematics300_lambda0.8_25runs.dat};
+
+\end{axis}
+\end{tikzpicture}
+
 \caption{results for Kinematics dataset}
 		\label{Kinematics300_lambda0.8_25runs}
-	\end{center}
-\end{figure}
+
+\end{center}\end{figure}
 
@@
 
-\subsection{Discussion}
-
 The proposed method has competitive results comparing with some well-known regression methods.
-Only Random Forest outperforms \ac{GAPoly} systematically (it outperforms all the others) except in the Auto-MPG test and, of course, in the artificial dataset that uses a straightforward polynomial relation.
+Only Random Forest outperforms \ac{GAPoly} systematically (it also outperforms all the other regression algorithms).
+One exception -- the artificial dataset that uses a straightforward polynomial relation -- is the Auto-MPG datatest where \ac{GAPoly} has comparable results.
+This is evidence that applying standard genetic operators for polynomial model searching is a viable tool for regression purposes.
 

### 2013-07-19 old/gapolyfuncs - times.R

@@ -1,170 +1,0 @@
-#########################################################################
-# initial values, can be overriden
-n.vars     <- 3  # x1, x2,... 
-max.compts <- 4  # poly = compt1 + compt2 + ...
-max.degree <- 3  # meaning exponents up to 2^max.degree - 1
-
-#########################################################################
-# Root mean square error
-rsme <- function(v1, v2=c(0)) {
-  sqrt(mean((v1-v2)^2))
-}
-
-#########################################################################
-# convert a binary vector into an integer
-bin2dec <- function(x) 
-  sum(2^(which(rev(unlist(strsplit(as.character(x), "")) == 1))-1))
-
-#########################################################################
-# The chromosome coding will as follow:
-# + the chromosome is split into `max.compts` sets of bits of equal size 
-#   (a component)
-# + each component is split into `n.vars` sets of `max.degree` size each 
-#   (a variable)
-# + for each variable, the number of 1s gives the variable's degree 
-#   (this means that a given polynomial may have more than one possible 
-#   representation)
-# 
-# Eg: consider polynomial $x_1^3 \times x_3 + x_3^7 + x_1 \times x_2$
-# 
-# One possible coding -- for the values in the previous code snippet -- would be:
-# 
-# $$011,000,001;000,000,111;001,001,000;000,000,000$$
-# 
-# (the semicolons separate components, the commas seperate variables)
-# 
-# The next functions transform the binary vector into a polynomial formula 
-# that `lm()` understands:
-
-make.component <- function(bits, n.vars) {
-  vars <- paste0("x",1:n.vars)
-  elements <- split(bits, ceiling(seq_along(bits)/(length(bits)/n.vars)))
-  str <- ""
-  for(i in 1:length(elements)) {
-    power <- bin2dec(elements[[i]])
-    if (power == 1)
-      if (str=="")
-        str = vars[i]
-    else  
-      str = paste0(str, "*", vars[i])
-    if (power > 1)
-      if (str=="")
-        str = paste0(vars[i], "^", power)
-    else  
-      str = paste0(str, "*", vars[i], "^", power)
-  }
-  str
-}
-
-make.formula <- function(bits, n.vars, max.compts) {
-  components <- split(bits, ceiling(seq_along(bits)/(length(bits)/max.compts)))
-  str <- ""
[…]

### 2013-07-19 old/gapolyfuncs - without regularization.R

@@ -1,171 +1,0 @@
-#########################################################################
-# initial values, can be overriden
-n.vars     <- 3  # x1, x2,... 
-max.compts <- 4  # poly = compt1 + compt2 + ...
-max.degree <- 3  # meaning exponents up to 2^max.degree - 1
-
-#########################################################################
-# Root mean square error
-rsme <- function(v1, v2=c(0)) {
-  sqrt(mean((v1-v2)^2))
-}
-
-#########################################################################
-# convert a binary vector into an integer
-bin2dec <- function(x) 
-  sum(2^(which(rev(unlist(strsplit(as.character(x), "")) == 1))-1))
-
-#########################################################################
-# The chromosome coding will as follow:
-# + the chromosome is split into `max.compts` sets of bits of equal size 
-#   (a component)
-# + each component is split into `n.vars` sets of `max.degree` size each 
-#   (a variable)
-# + for each variable, the number of 1s gives the variable's degree 
-#   (this means that a given polynomial may have more than one possible 
-#   representation)
-# 
-# Eg: consider polynomial $x_1^3 \times x_3 + x_3^7 + x_1 \times x_2$
-# 
-# One possible coding -- for the values in the previous code snippet -- would be:
-# 
-# $$011,000,001;000,000,111;001,001,000;000,000,000$$
-# 
-# (the semicolons separate components, the commas seperate variables)
-# 
-# The next functions transform the binary vector into a polynomial formula 
-# that `lm()` understands:
-
-make.component <- function(bits, n.vars) {
-  vars <- paste0("x",1:n.vars)
-  elements <- split(bits, ceiling(seq_along(bits)/(length(bits)/n.vars)))
-  str <- ""
-  for(i in 1:length(elements)) {
-    power <- bin2dec(elements[[i]])
-    if (power == 1)
-      if (str=="")
-        str = vars[i]
-    else  
-      str = paste0(str, "*", vars[i])
-    if (power > 1)
-      if (str=="")
-        str = paste0(vars[i], "^", power)
-    else  
-      str = paste0(str, "*", vars[i], "^", power)
-  }
-  str
-}
-
-make.formula <- function(bits, n.vars, max.compts) {
-  components <- split(bits, ceiling(seq_along(bits)/(length(bits)/max.compts)))
-  str <- ""
[…]

### 2013-07-19 texts/data/process-this-shit.py

@@ -28,7 +28,7 @@ def by_cols(d):
 	s += "\n"
 	return s
 
-filename = "fc-abalone_lambda-errors.txt"
+filename = "lambda.errors.Auto-Mpg.txt"
 
 f = file(filename,"r")
 

### 2013-07-22 texts/paper-draft.tex
it works….

@@
 \documentclass[12pt]{article}
+
+%
+%	FILE ENCODING
+%
 \usepackage[latin1]{inputenc}
-\usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
-\geometry{a4paper} % or letter or a5paper or ... etc
-% \geometry{landscape} % rotated page geometry
+%
+
+%
+%	PAGE GEOMETRY
+%
+\usepackage{geometry}
+\geometry{a4paper}
+%
+
+%
+%	AMS STUFF
+%
 \usepackage{amsfonts}
-
 \usepackage{amsthm}
 \newtheorem{theorem}{Theorem}
+%
 
+%
+%	ACRONYMS
+%
+\usepackage{acronym}
+\acrodef{GA}{Genetic algorithms}
+\acrodef{rmse}[error]{root-mean-square error}
+\acrodef{GAPoly}[{\sc GAPoly}]{Genetic Algorithms for Polynomials}
+%
+
+%
+%	FIGURE AND STUFF ENVIRONMENT
+%
 \usepackage{graphicx}
-\usepackage{acronym}
+%
+
+%
+%	TIKZ AND RELATED STUFF
+%
 \usepackage{pgfplots}
-\pgfplotsset{compat=1.8}
+\pgfplotsset{compat=newest}
 \usepgfplotslibrary{statistics}
+%
+%	avoid processing figures every time
+\usepgfplotslibrary{external}
+\tikzexternalize
+% ENABLE NEXT LINE TO TYPESET FIGURES
+\tikzset{external/force remake}
+%
 
+%
+%	ALGORITHMS PACKAGES
+%
 \usepackage{algpseudocode}
 \usepackage{algorithm}
+%
+
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 
@@
 \author{Francisco Coelho \and João Neto}
-\date{[{\sc draft} \today]} % delete this line to display the current date
+\date{[{\sc draft} \today]}
 
+%
+%	DOCUMENT COMMANDS
+%
 \newcommand{\sample}[1]{\ensuremath{^{\left(#1\right)}}}
+%
 
-\acrodef{GA}{Genetic algorithms}
-\acrodef{rmse}{root-mean-square error}
-\acrodef{GAPoly}[{\sc GAPoly}]{Genetic Algorithms for Polynomials}
-
+%
+%
 %%% BEGIN DOCUMENT
+%
+%
 \begin{document}
@@
 \ac{GA} where, arguably, one of the hottest topics of research in the recent decades and with good reason since they outline an optimization scheme easy to conceptualize and with very broad application.
-If a nonlinear (or otherwise) model requires parameterization, \acp{GA} provide a simple and often effective approach to search for locally optimal parameters.
+If a nonlinear (or otherwise) model requires parameterization \acp{GA} provide a simple and often effective approach to search for locally optimal parameters.
 Research related to genetic algorithms abound and spans from the 1950s seminal work of Nils Aall Barricelli \cite{barricelli1962numerical} in the Institute for Advanced Study of Princeton to today's principal area of study for thousands of researchers, covered in hundreds of conferences, workshops and other meetings.
@@
 One interesting take on genetic algorithms, named \emph{genetic programming} by John Koza \cite{koza1992genetic}, proposed the use of \acp{GA} to search the syntactic structure of complex functions.
-This syntatic structure search is keen to the central ideas of deep learning \cite{bengio2013representation, bengio2009learning}, the subarea of machine learning actually producing the most promising results (\emph{e.g.} \cite{tarlowstochastic}).
+This syntatic structure search is keen to the central ideas of deep learning \cite{bengio2013representation, bengio2009learning}, a subarea of machine learning actually producing quite promising results (\emph{e.g.} \cite{tarlowstochastic}).
 It is also related to the work presented in this paper in the sense that, unlike linear models that have a simple structure, $y=\sum_i \beta_i x_i$, nonlinear (in particular polynomial) models pose an additional ``structure" search problem.
@@
 
-This section is dedicated to describe an algorithm to find a polynomial regression from a given dataset.
-It starts with a brief introduction and description of the algorithm and proceeds into core details as the encoding used to represent individual polynomial instances in the \ac{GA} populations and the regularization of the cost function.
+This section is dedicated to the description of an algorithm to find a polynomial regression from a given dataset.
+It starts with a brief introduction and outline of the algorithm and proceeds into core details as the encoding used to represent individual polynomial instances in the \ac{GA} populations and the regularization of the cost function.
 
@@
 where each $m_i$ is a monomial, $m_i = \prod_{j} x_j^{\alpha_{ij}}$, the exponents are non-negative integers, $\alpha_{ij}\in\mathbb{N}_0$, and the coefficients are real valued, $\theta_i \in \mathbb{R}$.
-For example $p\left( x_1, x_2, x_3 \right) = 2 x_1 + x_2 x_3 + \frac{1}{2} x_1^2 x_3$ has monomials $m_1 = x_1, m_2 = x_2 x_3$ and $m_3 = x_1^2 x_3$ coefficients $\theta_1 = 2, \theta_2 = 1$ and $\theta_3 = 1/2$ and exponents $\alpha_{1,1} = 1, \alpha_{2,2}=1, \alpha_{2,3} = 1, \alpha_{3,1} = 2, \alpha_{3,3} = 1$ and all other $\alpha_{ij} = 0$.
+For example $p\left( x_1, x_2, x_3 \right) = 2 x_1 + x_2 x_3 + \frac{1}{2} x_1^2 x_3$ has monomials $m_1 = x_1, m_2 = x_2 x_3$ and $m_3 = x_1^2 x_3$,    coefficients $\theta_1 = 2, \theta_2 = 1$ and $\theta_3 = 1/2$ and exponents $\alpha_{1,1} = 1, \alpha_{2,2}=1, \alpha_{2,3} = 1, \alpha_{3,1} = 2, \alpha_{3,3} = 1$ and all other $\alpha_{ij} = 0$.
 The exponents alone are a matrix that defines the monomial structure of the polynomial, $A = \left[\alpha_{ij}\right]$.
@@
 %
-\begin{algorithm}[htb]
+\begin{algorithm}[tb]
 \begin{algorithmic}
@@
 %
-\begin{theorem}
-The composition of \acp{GA} with linear regression, as defined in Algorithm \ref{alg:gapoly}, converges to a polynomial that is a local minimum of the \ac{rmse} function, encapsulated in the fitness function $J$.
-\end{theorem}
+Notice that the properties of \acp{GA} and linear regression entail that
+the composition of \acp{GA} with linear regression, as defined in Algorithm \ref{alg:gapoly}, converges to a polynomial that is a local minimum of the \ac{rmse} function, encapsulated in the fitness function $J$.
 %
-\begin{proof}
-It follows from the properties of \acp{GA} and linear regression.
-\end{proof}
+
 %
@@
 The regularization of the cost function is discussed in subsection \ref{subs:cost.function}.
+
+%
 \subsection{Polinomial Encoding}\label{subs:polynomial.encoding}
-
+%
 The encoding for any polynomial will be as follows:
@@
 Our proposal is to include a multiplicative factor into the fitness function proportional to the number of monomials.
-Thus, the fitness function becomes,
-
-$$
-J_{fit}\left( \Theta; D \right) = \lambda^{m} \sqrt{ \frac{1}{n} \sum_{i=1}^n \left( y^{(i)} - h_{\Theta}\left(x_1^{(i)}, \ldots, x_m^{(i)}\right) \right)^2 }
-$$
-
+Thus, the fitness function from equation \ref{eq:rmse} becomes
+\begin{eqnarray}
+J_{fit}\left( \Theta; D \right) &=& \lambda^{m} \sqrt{ \frac{1}{n} \sum_{i=1}^n \left( y^{(i)} - h_{\Theta}\left(x_1^{(i)}, \ldots, x_m^{(i)}\right) \right)^2 }
+\label{eq:rmse-reg}
+\end{eqnarray}
+%
 where $m$ is the number of monomials the polynomial has. A $\lambda$ greater than one punishes polynomials with more monomials.
@@
 
-\begin{figure}[htb]\begin{center}
+\begin{figure}[tb]\begin{center}
+\begin{tabular}{cc}
 \begin{tikzpicture}
-\begin{axis}[width=\textwidth,height=6cm,
-	boxplot/draw direction=y,
-	xtick={1,2,3,4,5,6,7,8,9,10,11,12,13},
-	xticklabels={0.70, 0.75, 0.80, 0.85, 0.90, 0.95, 1.00, 1.05, 1.10, 1.15, 1.20, 1.25, 1.30},
-	xlabel={regularisation exponent ($\lambda$)},
-	ymin=2.0,
-	ymax=3.0,
-	ylabel={Error distribution (\ac{rmse})},
-	title={Error distribution by regularisation exponent (\textbf{Abalone dataset})},
-	mark options={scale=0.5},
+\begin{axis}[
+	width=0.5\textwidth,height=6cm,
+	boxplot/draw direction=x,
+	ytick={1,2,3,4,5,6,7},
+	yticklabels={0.7, 0.8, 0.9, 1.0, 1.1, 1.2, 1.3},
+	title = {Abalone Dataset},
+	xlabel = {Error (Abalone)},
+	ylabel={Regularization ($\lambda$)},
 	]	
-	\addplot[only marks, mark=o,color=black] table [x expr=1+0.2*rand,y index=0] {data/figure1.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=2+0.2*rand,y index=1] {data/figure1.dat};		
-	\addplot[only marks, mark=o,color=black] table [x expr=3+0.2*rand,y index=2] {data/figure1.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=4+0.2*rand,y index=3] {data/figure1.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=5+0.2*rand,y index=4] {data/figure1.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=6+0.2*rand,y index=5] {data/figure1.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=7+0.2*rand,y index=6] {data/figure1.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=8+0.2*rand,y index=7] {data/figure1.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=9+0.2*rand,y index=8] {data/figure1.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=10+0.2*rand,y index=9] {data/figure1.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=11+0.2*rand,y index=10] {data/figure1.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=12+0.2*rand,y index=11] {data/figure1.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=13+0.2*rand,y index=12] {data/figure1.dat};
-
 	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
 	\addplot[boxplot] table [y index=0] {data/figure1.dat};
-	\addplot[boxplot] table [y index=1] {data/figure1.dat};
 	\addplot[boxplot] table [y index=2] {data/figure1.dat};
-	\addplot[boxplot] table [y index=3] {data/figure1.dat};
 	\addplot[boxplot] table [y index=4] {data/figure1.dat};
-	\addplot[boxplot] table [y index=5] {data/figure1.dat};
 	\addplot[boxplot] table [y index=6] {data/figure1.dat};		
-	\addplot[boxplot] table [y index=7] {data/figure1.dat};		
 	\addplot[boxplot] table [y index=8] {data/figure1.dat};
-	\addplot[boxplot] table [y index=9] {data/figure1.dat};
 	\addplot[boxplot] table [y index=10] {data/figure1.dat};
-	\addplot[boxplot] table [y index=11] {data/figure1.dat};
 	\addplot[boxplot] table [y index=12] {data/figure1.dat};
+\end{axis}	
+\end{tikzpicture}
+%
+% RIGHT PANE
+&
+%
+%
+\begin{tikzpicture}
+\begin{axis}[
+	width=0.5\textwidth,height=6cm,
+	boxplot/draw direction=x,
+	ytick={1,2,3,4,5,6,7},
+	yticklabels={0.7, 0.8, 0.9, 1.0, 1.1, 1.2, 1.3},
+	title = {Auto MPG Dataset},
+	xlabel={Error (Auto MPG)},
+	]	
+	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
+	\addplot[boxplot] table [y index=0] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[boxplot] table [y index=2] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[boxplot] table [y index=4] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[boxplot] table [y index=6] {data/lambda.errors.Auto-Mpg.dat};		
+	\addplot[boxplot] table [y index=8] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[boxplot] table [y index=10] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[boxplot] table [y index=12] {data/lambda.errors.Auto-Mpg.dat};
 
@@
 
-% \includegraphics[height=6cm,width=\textwidth,angle=0]{figures/Abalone_dataset_lambdas.eps}
+\end{tabular}
 
@@
 \end{center}\end{figure}
+%
+%
+\end{document}
 
-\begin{figure}[htb]\begin{center}
+\begin{figure}[tb]\begin{center}
 \begin{tikzpicture}
@@
 	]	
-	\addplot[only marks, mark=o,color=black] table [x expr=1+0.2*rand,y index=0] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=2+0.2*rand,y index=1] {data/lambda.errors.Auto-Mpg.dat};		
-	\addplot[only marks, mark=o,color=black] table [x expr=3+0.2*rand,y index=2] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=4+0.2*rand,y index=3] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=5+0.2*rand,y index=4] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=6+0.2*rand,y index=5] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=7+0.2*rand,y index=6] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=8+0.2*rand,y index=7] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=9+0.2*rand,y index=8] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=10+0.2*rand,y index=9] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=11+0.2*rand,y index=10] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=12+0.2*rand,y index=11] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=13+0.2*rand,y index=12] {data/lambda.errors.Auto-Mpg.dat};
-
 	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
@@
 \begin{center}
-		\includegraphics[height=6cm,width=10cm,angle=0]
-			{figures/Abalone_fitnessProgress.eps}
-		\caption{fitness progress for Abalone dataset}
+\begin{tikzpicture}
+
+\begin{axis}[
+	width=0.95\textwidth,
+	height=6cm,
+	xmin=-5,
+	xmax=255,
+	mark options={scale=0.5},
+	line width=1pt,
+	xlabel={number of iterations},
+	ylabel={\ac{rmse}},
+	title={Error (\ac{rmse}) vs. number of iterations (\textbf{Abalone dataset})},
+	]
+
+	\addplot[solid,color=black] table [x expr=\coordindex, y index = 0] {data/lambdas.dat};\addlegendentry{$\lambda=0.975$};
+	\addplot[dashed,color=black] table [x expr=\coordindex, y index = 1] {data/lambdas.dat};\addlegendentry{$\lambda=1.000$};
+\end{axis}
+\end{tikzpicture}
+\caption{fitness progress for Abalone dataset (population $200$)}
 \label{Abalone_fitnessProgress}
+
 \end{center}

### 2013-07-22 texts/data/lambdas.py

@@ -1,0 +1,507 @@
+lambda095=[2.473536,
+2.414038,
+2.414038,
+2.414038,
+2.406747,
+2.406747,
+2.316709,
+2.350614,
+2.330452,
+2.313667,
+2.251034,
+2.251034,
+2.251034,
+2.275441,
+2.275441,
+2.275441,
+2.275307,
+2.275307,
+2.275307,
+2.268287,
+2.268287,
+2.268287,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.225473,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
+2.200263,
[…]

### 2013-07-22 texts/data/proc-lambdas.py

@@ -1,0 +1,7 @@
+f = file("lambdas.dat")
+
+for line in f.readlines():
+	if len(line) > 0:
+		eval(line)
+
+print dir()

### 2013-07-22 texts/paper-draft.tex
adicionados ficheiros com imagens

@@
 \begin{eqnarray}
-J\left(\Theta;\mathcal{M},\mathcal{D}\right) &=& \sqrt{\frac{1}{N}\sum_{i=1}^N \left( y\sample{i} - h_{\Theta,\mathcal{M}}\left( x_1\sample{i},\ldots,x_n\sample{i} \right) \right)^2 }\label{eq:rmse}
+J_{fit}\left(\Theta;\mathcal{M},\mathcal{D}\right) &=& \sqrt{\frac{1}{N}\sum_{i=1}^N \left( y\sample{i} - h_{\Theta,\mathcal{M}}\left( x_1\sample{i},\ldots,x_n\sample{i} \right) \right)^2 }\label{eq:rmse}
 \end{eqnarray}
@@
 Our proposal is to include a multiplicative factor into the fitness function proportional to the number of monomials.
-Thus, the fitness function from equation \ref{eq:rmse} becomes
+Thus, the fitness function from equation \ref{eq:rmse} defines
 \begin{eqnarray}
-J_{fit}\left( \Theta; D \right) &=& \lambda^{m} \sqrt{ \frac{1}{n} \sum_{i=1}^n \left( y^{(i)} - h_{\Theta}\left(x_1^{(i)}, \ldots, x_m^{(i)}\right) \right)^2 }
-\label{eq:rmse-reg}
+J_{reg}\left(\Theta;\mathcal{M},\mathcal{D}\right) &=& \lambda^{m} J_{fit}\left(\Theta;\mathcal{M},\mathcal{D}\right)\label{eq:rmse-reg}
 \end{eqnarray}
@@
 \caption{Error distribution by regularisation exponent for the Abalone dataset.}\label{Abalone_dataset_lambdas}
-
-\end{center}\end{figure}
-%
-%
-\end{document}
-
-\begin{figure}[tb]\begin{center}
-\begin{tikzpicture}
-\begin{axis}[width=\textwidth,height=6cm,
-	boxplot/draw direction=y,
-	xtick={1,2,3,4,5,6,7,8,9,10,11,12,13},
-	xticklabels={0.70, 0.75, 0.80, 0.85, 0.90, 0.95, 1.00, 1.05, 1.10, 1.15, 1.20, 1.25, 1.30},
-	xlabel={regularisation exponent ($\lambda$)},
-	ymin=2.0,
-	ymax=5.0,
-	ylabel={Error distribution (\ac{rmse})},
-	title={Error distribution by regularisation exponent (\textbf{Auto-MPG dataset})},
-	mark options={scale=0.5},
-	]	
-	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
-	\addplot[boxplot] table [y index=0] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot] table [y index=1] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot] table [y index=2] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot] table [y index=3] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot] table [y index=4] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot] table [y index=5] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot] table [y index=6] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot] table [y index=7] {data/lambda.errors.Auto-Mpg.dat};		
-	\addplot[boxplot] table [y index=8] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot] table [y index=9] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot] table [y index=10] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot] table [y index=11] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot] table [y index=12] {data/lambda.errors.Auto-Mpg.dat};
-
-\end{axis}
-\end{tikzpicture}
-
-% \includegraphics[height=6cm,width=\textwidth,angle=0]{figures/Abalone_dataset_lambdas.eps}
-
-\caption{Error distribution by regularisation exponent for the Auto-MPG dataset.}\label{Auto-MPG_dataset_lambdas}
 

### 2013-07-22 texts/paper-draft.tex
test sync

@@
-\documentclass[12pt]{article}
+\documentclass{elsarticle}
 
@@
 %
-\usepackage{geometry}
-\geometry{a4paper}
+%\usepackage{geometry}
+%\geometry{a4paper}
 %
@@
 % ENABLE NEXT LINE TO TYPESET FIGURES
-\tikzset{external/force remake}
+%\tikzset{external/force remake}
 %
@@
 %
-
-\title{Selection of Polynomial Features by Genetic Algorithms}
-\author{Francisco Coelho \and João Neto}
-\date{[{\sc draft} \today]}
 
@@
 
-\maketitle
+\begin{frontmatter}
+
+\title{Selection of Polynomial Features by Genetic Algorithms}
+\author[ue,labmag]{Francisco Coelho}
+\ead{fc@di.uevora.pt}
+
+\author[fcul,labmag]{João Pedro Neto}
+\ead{jpn@di.fc.ul.pt}
+
+\address[ue]{Dept. Informática, Universidade de Évora, Rua Romão Ramalho 58 Évora 7000-671 Évora}
+\address[fcul]{Dept. Informática, Faculdade de Ciências da Universidade de Lisboa, Campo Grande 1749-016 Lisboa}
+\address[labmag]{Laboratory of Agent Modelling (LabMAg)}
+
+\date{[{\sc draft} \today]}
+
 \begin{abstract}
@@
 \end{abstract}
+
+\end{frontmatter}
 
@@
 \begin{axis}[
-	width=0.5\textwidth,height=6cm,
+	width=0.49\textwidth,height=6cm,
 	boxplot/draw direction=x,
@@
 \begin{axis}[
-	width=0.5\textwidth,height=6cm,
+	width=0.49\textwidth,height=6cm,
 	boxplot/draw direction=x,
 	ytick={1,2,3,4,5,6,7},
-	yticklabels={0.7, 0.8, 0.9, 1.0, 1.1, 1.2, 1.3},
+	%yticklabels={0.7, 0.8, 0.9, 1.0, 1.1, 1.2, 1.3},
+	yticklabels={},
 	title = {Auto MPG Dataset},
@@
 
-\textbf{Artificial}: this is an artificial dataset with four numeric features, $x_1, \ldots x_4$, where $x_1,x_3$ are outcomes from Poisson random variables, and $x_2,x_4$ from Normal random variables.
-The dependent variable $y$ is given by expression $x_2x_4^2 + x_1^2x_3 + 5$.
-The dataset includes 50 observations and, only in this case, there were executed just 10 runs.
-
-\begin{figure}[htb]\begin{center}
+\begin{figure}[tb]\begin{center}
 \begin{tikzpicture}
-\begin{axis}[width=\textwidth,height=6cm,
-	boxplot/draw direction=y,
-	xtick={1,2,3,4,5,6},
-	xticklabels={GA.Poly,Lin.Reg,SVM,RPART,Rnd.For,CI.Tree},
-	xlabel={Regression Method},
-	ymin=0,
-	ymax=300,
-	ylabel={Error distribution (\ac{rmse})},
+\begin{axis}[width=0.90\textwidth,height=6cm,
+	boxplot/draw direction=x,
+	ytick={1,2,3,4,5,6},
+	yticklabels={GA.Poly,Lin.Reg,SVM,RPART,Rnd.For,CI.Tree},
+	ylabel={Regression Method},
+	xmin=-5,
+	xmax=300,
+	xlabel={Error distribution (\ac{rmse})},
 	title={Regression results for \textbf{Artificial dataset}},
-	mark options={scale=0.5},
 	]
-	\addplot[only marks, mark=o,color=black] table [x expr=1+0.2*rand,y index=0] {data/Articial_dataset.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=2+0.2*rand,y index=1] {data/Articial_dataset.dat};		
-	\addplot[only marks, mark=o,color=black] table [x expr=3+0.2*rand,y index=2] {data/Articial_dataset.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=4+0.2*rand,y index=3] {data/Articial_dataset.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=5+0.2*rand,y index=4] {data/Articial_dataset.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=6+0.2*rand,y index=5] {data/Articial_dataset.dat};
-
 	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
@@
 
-% \includegraphics[height=6cm,width=\textwidth,angle=0]{figures/Abalone_dataset_lambdas.eps}
-
 \caption{results for Artificial dataset}
@@
 \end{center}\end{figure}
+
+\textbf{Artificial}: this is an artificial dataset with four numeric features, $x_1, \ldots x_4$, where $x_1,x_3$ are outcomes from Poisson random variables, and $x_2,x_4$ from Normal random variables.
+The dependent variable $y$ is given by expression $x_2x_4^2 + x_1^2x_3 + 5$.
+The dataset includes 50 observations and, only in this case, there were executed just 10 runs.
 
@@
 \begin{tikzpicture}
-\begin{axis}[width=\textwidth,height=6cm,
+\begin{axis}[width=0.98\textwidth,height=6cm,
 	boxplot/draw direction=y,
@@
 	]
-	\addplot[only marks, mark=o,color=black] table [x expr=1+0.2*rand,y index=0] {data/Housing300_lambda0.8_25runs.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=2+0.2*rand,y index=1] {data/Housing300_lambda0.8_25runs.dat};		
-	\addplot[only marks, mark=o,color=black] table [x expr=3+0.2*rand,y index=2] {data/Housing300_lambda0.8_25runs.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=4+0.2*rand,y index=3] {data/Housing300_lambda0.8_25runs.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=5+0.2*rand,y index=4] {data/Housing300_lambda0.8_25runs.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=6+0.2*rand,y index=5] {data/Housing300_lambda0.8_25runs.dat};
-
 	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
@@
 	]
-	\addplot[only marks, mark=o,color=black] table [x expr=1+0.2*rand,y index=0] {data/Abalone300_lambda0.8_25runs.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=2+0.2*rand,y index=1] {data/Abalone300_lambda0.8_25runs.dat};		
-	\addplot[only marks, mark=o,color=black] table [x expr=3+0.2*rand,y index=2] {data/Abalone300_lambda0.8_25runs.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=4+0.2*rand,y index=3] {data/Abalone300_lambda0.8_25runs.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=5+0.2*rand,y index=4] {data/Abalone300_lambda0.8_25runs.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=6+0.2*rand,y index=5] {data/Abalone300_lambda0.8_25runs.dat};
 
@@
 	]
-	\addplot[only marks, mark=o,color=black] table [x expr=1+0.2*rand,y index=0] {data/Auto-Mpg300_lambda0.8_25runs.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=2+0.2*rand,y index=1] {data/Auto-Mpg300_lambda0.8_25runs.dat};		
-	\addplot[only marks, mark=o,color=black] table [x expr=3+0.2*rand,y index=2] {data/Auto-Mpg300_lambda0.8_25runs.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=4+0.2*rand,y index=3] {data/Auto-Mpg300_lambda0.8_25runs.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=5+0.2*rand,y index=4] {data/Auto-Mpg300_lambda0.8_25runs.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=6+0.2*rand,y index=5] {data/Auto-Mpg300_lambda0.8_25runs.dat};
 
@@
 \begin{tikzpicture}
-\begin{axis}[width=\textwidth,height=7cm,
+\begin{axis}[width=0.98\textwidth,height=7cm,
 	boxplot/draw direction=y,
@@
 	]
-	\addplot[only marks, mark=o,color=black] table [x expr=1+0.2*rand,y index=0] {data/Kinematics300_lambda0.8_25runs.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=2+0.2*rand,y index=1] {data/Kinematics300_lambda0.8_25runs.dat};		
-	\addplot[only marks, mark=o,color=black] table [x expr=3+0.2*rand,y index=2] {data/Kinematics300_lambda0.8_25runs.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=4+0.2*rand,y index=3] {data/Kinematics300_lambda0.8_25runs.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=5+0.2*rand,y index=4] {data/Kinematics300_lambda0.8_25runs.dat};
-	\addplot[only marks, mark=o,color=black] table [x expr=6+0.2*rand,y index=5] {data/Kinematics300_lambda0.8_25runs.dat};
 
@@
 \begin{axis}[
-	width=0.95\textwidth,
+	width=0.98\textwidth,
 	height=6cm,
@@
 
-\section{Bibliography}
+\section*{Acknowledgements}
+
+The authors thank the Fundação para a Ciência e Tecnologia (FCT) and the  laboratório de I\&D LabMAg the financial support given to this work under the strategic project PEst-­OE/EEI/UI0434/2011.
+
+%\section{Bibliography}
 

### 2013-07-22 texts/paper-draft.tex
- Arrumei as figuras; - Uniformizei as variáveis (**atenção: monómios -> q**);

@@
 %	avoid processing figures every time
-\usepgfplotslibrary{external}
-\tikzexternalize
+%\usepgfplotslibrary{external}
+%\tikzexternalize
 % ENABLE NEXT LINE TO TYPESET FIGURES
@@
 $$
-p\left( x_1, \ldots, x_k\right) = \sum_i \theta_i m_i
+p\left( x_1, \ldots, x_m\right) = \sum_i \theta_i q_i
 $$
-where each $m_i$ is a monomial, $m_i = \prod_{j} x_j^{\alpha_{ij}}$, the exponents are non-negative integers, $\alpha_{ij}\in\mathbb{N}_0$, and the coefficients are real valued, $\theta_i \in \mathbb{R}$.
-For example $p\left( x_1, x_2, x_3 \right) = 2 x_1 + x_2 x_3 + \frac{1}{2} x_1^2 x_3$ has monomials $m_1 = x_1, m_2 = x_2 x_3$ and $m_3 = x_1^2 x_3$,    coefficients $\theta_1 = 2, \theta_2 = 1$ and $\theta_3 = 1/2$ and exponents $\alpha_{1,1} = 1, \alpha_{2,2}=1, \alpha_{2,3} = 1, \alpha_{3,1} = 2, \alpha_{3,3} = 1$ and all other $\alpha_{ij} = 0$.
+where each $q_i$ is a monomial, $q_i = \prod_{j} x_j^{\alpha_{ij}}$, the exponents are non-negative integers, $\alpha_{ij}\in\mathbb{N}_0$, and the coefficients are real valued, $\theta_i \in \mathbb{R}$.
+For example $p\left( x_1, x_2, x_3 \right) = 2 x_1 + x_2 x_3 + \frac{1}{2} x_1^2 x_3$ has monomials $q_1 = x_1, q_2 = x_2 x_3$ and $q_3 = x_1^2 x_3$, coefficients $\theta_1 = 2, \theta_2 = 1$ and $\theta_3 = 1/2$ and exponents $\alpha_{1,1} = 1, \alpha_{2,2}=1, \alpha_{2,3} = 1, \alpha_{3,1} = 2, \alpha_{3,3} = 1$ and all other $\alpha_{ij} = 0$.
 The exponents alone are a matrix that defines the monomial structure of the polynomial, $A = \left[\alpha_{ij}\right]$.
@@
 \begin{enumerate}
-\item for a given set of monomials $m_1, \ldots, m_R$, find the regression coefficients $\theta_1,\ldots,\theta_R$ that minimize the \ac{rmse} on a given dataset;
+\item for a given set of monomials $\mathcal{P} = \left\lbrace q_1, \ldots, q_k\right\rbrace$, find the regression coefficients $\theta_1,\ldots,\theta_k$ that minimize the \ac{rmse} on a given dataset;
 
@@
 %
-More precisely, concerning the first problem, let $\mathcal{D}$ be a dataset with $N$ observations of variables $Y, X_1,\ldots,X_n$ and $\mathcal{M} = \left\lbrace m_1,\ldots, m_R\right\rbrace$ a set of $R$ monomial expressions over $X_1,\ldots,X_n$.
-Define the hypothesis\footnote{The expression $m|_{X=x}$ reads ``\emph{replace all instances of $X$ by $x$ in $m$}''.}
+More precisely, concerning the first problem, let $\mathcal{D}$ be a dataset with $n$ observations of variables $Y, X_1,\ldots,X_m$ and $\mathcal{P} = \left\lbrace q_1,\ldots, q_k\right\rbrace$ a set of $k$ monomial expressions over $X_1,\ldots,X_m$.
+Define the hypothesis\footnote{The expression $q|_{X=x}$ reads ``\emph{replace all instances of $X$ by $x$ in $q$}''.}
 %
 \begin{eqnarray*}
-h_{\Theta,\mathcal{M}}\left(x_1,\ldots,x_n\right) &=& \sum_{j = 1}^R \theta_j m_j|_{X_i=x_i,\forall 1 \leq i \leq n}
+h_{\Theta,\mathcal{P}}\left(x_1,\ldots,x_m\right) &=& \sum_{j = 1}^k \theta_j q_j|_{X_i=x_i,\forall 1 \leq i \leq m}
 \end{eqnarray*}
@@
 \begin{eqnarray}
-J_{fit}\left(\Theta;\mathcal{M},\mathcal{D}\right) &=& \sqrt{\frac{1}{N}\sum_{i=1}^N \left( y\sample{i} - h_{\Theta,\mathcal{M}}\left( x_1\sample{i},\ldots,x_n\sample{i} \right) \right)^2 }\label{eq:rmse}
+J_{\textrm{fit}}\left(\Theta;\mathcal{P},\mathcal{D}\right) &=& \sqrt{\frac{1}{n}\sum_{i=1}^n \left( y\sample{i} - h_{\Theta,\mathcal{P}}\left( x_1\sample{i},\ldots,x_m\sample{i} \right) \right)^2 }\label{eq:rmse}
 \end{eqnarray}
@@
 \begin{quote}
-Given a dataset $\mathcal{D}$ and a set of monomials $\mathcal{M}$, find parameters $\Theta$ that minimize $J\left(\Theta;\mathcal{M},\mathcal{D}\right)$.
+Given a dataset $\mathcal{D}$ and a set of monomials $\mathcal{P}$, find parameters $\Theta$ that minimize $J\left(\Theta;\mathcal{P},\mathcal{D}\right)$.
 \end{quote}
@@
 
-The second problem is treated in the \ac{GA} setting: Let $\mathcal{D}$ be a dataset as above and $\mathcal{P}$ a set of polynomials.
-For each polynomial $p\in \mathcal{P}$ let $\mathcal{M}_p$ be the set of monomials in $p$ (without the coefficients) and compute the fitness
+The second problem is treated in the \ac{GA} setting: Let $\mathcal{D}$ be a dataset as above and $\mathcal{Q}$ a set of polynomials.
+For each polynomial $p\in \mathcal{Q}$ let $\mathcal{M}_p$ be the set of monomials in $p$ (without the coefficients) and compute the fitness
 %
@@
 Notice that the properties of \acp{GA} and linear regression entail that
-the composition of \acp{GA} with linear regression, as defined in Algorithm \ref{alg:gapoly}, converges to a polynomial that is a local minimum of the \ac{rmse} function, encapsulated in the fitness function $J$.
+the composition of \acp{GA} with linear regression, as defined in Algorithm \ref{alg:gapoly}, converges to a polynomial that is a local minimum of the fitness function, encapsulated in the \ac{rmse} function $J_{\textrm{fit}}$.
 %
@@
 %
-Subsection \ref{subs:polynomial.encoding} details of the encoding of individual polynomial instances as chromosomes and other parameters of the \ac{GA} implementation used.
+Subsection \ref{subs:polynomial.encoding} describes the encoding of individual polynomial instances as chromosomes and other parameters of the \ac{GA} implementation used.
 The regularization of the cost function is discussed in subsection \ref{subs:cost.function}.
@@
 \begin{enumerate}
-\item an initial segment detailing which monomials are active (the 1st monomial is always active), this is represented in unary description, i.e., each monomial is identified by a single bit
-\item the remaining is split into $m$ sets of bits of equal size, each one representing a monomial
-\item each monomial is split into $n$ sets of $d$ size each, i.e., a variable
-\item for each variable, the remaining bits give the binary description of the variable degree, i.e., the maximum exponent is given by $2^d-1$
+\item an initial segment detailing which monomials are active (the first monomial is always active), this is represented in unary description, i.e., each monomial is identified by a single bit
+\item the remaining is split into $k$ sets of bits of equal size, each one representing a monomial
+\item each monomial is split into $m$ sets of $d$ size each, \emph{i.e.}, a variable
+\item for each variable, the remaining bits give the binary description of the variable degree, \emph{i.e.}, the maximum exponent is given by $2^d-1$
 \end{enumerate}
 
-Let's see an example: consider polynomial $x_1^3 x_3 + x_3^7 + x_1 x_2$ with $m = 4, n = 3$ and $d=3$. One possible encoding would be:
+Let's see an example: consider polynomial $x_1^3 x_3 + x_3^7 + x_1 x_2$ with $k = 4, m = 3$ and $d=3$. One possible encoding would be:
 $$110 - 011,000,001~~;~~000,000,111~~;~~001,001,000~~;~~110,010,101$$
@@
 The first three bits inform that the second and third monomials are active while the fourth is not (as said, the first monomial is always active).
-This last monomial does not enter neither in the polynomial regression nor in the fitness evaluation.
+This last monomial does not enter neither in the polynomial regression nor in the \ac{rmse} (fitness) evaluation.
 However, it acts as a kind of junk DNA, becoming active when, in a future mutation or crossover, the third bit of the entire sequence flips from 0 to 1.
@@
 Let's interpret the first monomial description, $011,000,001$.
-It is divided by three since $n=3$.
+It is divided by three since $m=3$.
 The first triple $011$ is the binary description of the exponent of variable $x_1$ which is 3, so the first monomial includes $x_1^3$.
@@
 
-The polynomial fitness considered so far is based on the ability to predict the test set after the polynomial regression has found the appropriate coefficients $\theta_i$ for each one of the polynomial's monomials $m_i$.
+The polynomial regression \ac{rmse} considered so far is based on the ability to predict the test set after the polynomial regression has found the appropriate coefficients $\theta_i$ for each one of the polynomial's monomials $q_i$.
 
-This fitness function tends to prefer more complex polynomials, namely in the number of monomials which provides the regression algorithm for more fitting possibilities.
-One way to balance this is to provide a regularization term into the fitness function.
-Our proposal is to include a multiplicative factor into the fitness function proportional to the number of monomials.
-Thus, the fitness function from equation \ref{eq:rmse} defines
+This \ac{rmse} function tends to prefer more complex polynomials, namely in the number of monomials which provides the regression algorithm for more fitting possibilities.
+One way to balance this is to provide a regularization term into the \ac{rmse} function.
+Our proposal is to include a multiplicative factor proportional to the number of monomials.
+Thus, the error from equation \ref{eq:rmse} defines
 \begin{eqnarray}
-J_{reg}\left(\Theta;\mathcal{M},\mathcal{D}\right) &=& \lambda^{m} J_{fit}\left(\Theta;\mathcal{M},\mathcal{D}\right)\label{eq:rmse-reg}
+J_{\textrm{reg}}\left(\Theta;\mathcal{P},\mathcal{D}\right) &=& \lambda^{k} J_{\textrm{fit}}\left(\Theta;\mathcal{P},\mathcal{D}\right)\label{eq:rmse-reg}
 \end{eqnarray}
 %
-where $m$ is the number of monomials the polynomial has. A $\lambda$ greater than one punishes polynomials with more monomials.
+where $k$ is the number of monomials in the polynomial. When $\lambda > 1$ polynomials with more monomials are penalized.
 
@@
 	yticklabels={0.7, 0.8, 0.9, 1.0, 1.1, 1.2, 1.3},
-	title = {Abalone Dataset},
-	xlabel = {Error (Abalone)},
-	ylabel={Regularization ($\lambda$)},
+	title = {Dataset: \textsc{Abalone}},
+	xlabel = {Error distribution},
+	ylabel={Regularization weight ($\lambda$)},
+	xmax=2.7,
 	]	
-	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
+	%\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
 	\addplot[boxplot] table [y index=0] {data/figure1.dat};
-	\addplot[boxplot] table [y index=2] {data/figure1.dat};
+	\addplot[boxplot][fill=gray!50!white] table [y index=2] {data/figure1.dat};
 	\addplot[boxplot] table [y index=4] {data/figure1.dat};
@@
 	yticklabels={},
-	title = {Auto MPG Dataset},
-	xlabel={Error (Auto MPG)},
+	title = {Dataset: \textsc{Auto MPG}},
+	xlabel = {Error distribution},
+	ylabel={},
 	]	
-	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
+	%\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
 	\addplot[boxplot] table [y index=0] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot] table [y index=2] {data/lambda.errors.Auto-Mpg.dat};
+	\addplot[boxplot][fill=gray!50!white] table [y index=2] {data/lambda.errors.Auto-Mpg.dat};
 	\addplot[boxplot] table [y index=4] {data/lambda.errors.Auto-Mpg.dat};
@@
 \end{tikzpicture}
-
 \end{tabular}
 
-\caption{Error distribution by regularisation exponent for the Abalone dataset.}\label{Abalone_dataset_lambdas}
+\caption{Error distribution by regularization exponent for the Abalone dataset.}\label{Abalone_dataset_lambdas}\label{Auto-MPG_dataset_lambdas}
 
@@
 
-The typical inflection point lies around $\lambda = 0.8$. The following results use the regularisation parameter with this value.
+The typical inflection point lies around $\lambda = 0.8$. The following results use the regularization parameter with this value.
 
@@
 The operators were the standard ones: (a) crossover, i.e., a pair of solutions from the previous generations are combined by splitting and mixing their respective representations, and (b) mutation, changing the values of single bits; the mutation chance applied in the datasets was 5\%.
-There was also elitism between generations, i.e., 20\% of the best solutions are kept in the next generation.
+There was also elitism between generations, i.e., 20\% of the best solutions are kept in (fc: survive to?) the next generation.
 
@@
 	ytick={1,2,3,4,5,6},
-	yticklabels={GA.Poly,Lin.Reg,SVM,RPART,Rnd.For,CI.Tree},
+	%yticklabels={GA.Poly,Rnd.For,Lin.Reg,SVM,RPART,CI.Tree},
 	ylabel={Regression Method},
@@
 	xmax=300,
-	xlabel={Error distribution (\ac{rmse})},
-	title={Regression results for \textbf{Artificial dataset}},
+	xlabel={Error distribution},
+	title={Dataset: \textsc{Artificial}},
 	]
-	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
-	\addplot[boxplot] table [y index=0] {data/Articial_dataset.dat};
+	%\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
+	\addplot[boxplot][fill=gray!50!white] table [y index=0] {data/Articial_dataset.dat};
+	\addplot[boxplot] table [y index=4] {data/Articial_dataset.dat};
 	\addplot[boxplot] table [y index=1] {data/Articial_dataset.dat};
@@
 	\addplot[boxplot] table [y index=3] {data/Articial_dataset.dat};
-	\addplot[boxplot] table [y index=4] {data/Articial_dataset.dat};
 	\addplot[boxplot] table [y index=5] {data/Articial_dataset.dat};
@@
 
-\caption{results for Artificial dataset}
+\caption{results for Artificial dataset \emph{must} highlight method \#1: GAPoly}
 \label{artificial_dataset1_lambda1.0}
@@
 
-\textbf{Artificial}: this is an artificial dataset with four numeric features, $x_1, \ldots x_4$, where $x_1,x_3$ are outcomes from Poisson random variables, and $x_2,x_4$ from Normal random variables.
+\newcommand{\axpl}[3]{
+\begin{axis}[width=0.49\textwidth,height=4cm,
+	ticks=major,
+	boxplot/draw direction=x,
+	ytick={1,2,3,4,5,6},
+	%yticklabels={GA.Poly, Lin.Reg, SVM, RPART,Rnd.For,CI.Tree},
+	title={Dataset: \textsc{#2}},
+	ylabel={Method},
+	xlabel={Error},
+	separate axis lines,
+	axis lines = box,
+%	axis y line = #3,
+	]
+	%\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
+	\addplot[boxplot][fill=gray!50!white] table [y index=0] {data/#1.dat};
+	\addplot[boxplot] table [y index=4] {data/#1.dat};
+	\addplot[boxplot] table [y index=1] {data/#1.dat};
+	\addplot[boxplot] table [y index=2] {data/#1.dat};
+	\addplot[boxplot] table [y index=3] {data/#1.dat};
+	\addplot[boxplot] table [y index=5] {data/#1.dat};
+
+\end{axis}
+}
+
+\begin{figure}[tb]\begin{center}
+\begin{tabular}{cc}
+\begin{tikzpicture}
+	\axpl{Housing300_lambda0.8_25runs}{Housing}{left}
+\end{tikzpicture}
+%
+&
+%
+\begin{tikzpicture}
+	\axpl{Abalone300_lambda0.8_25runs}{Abalone}{right}
+\end{tikzpicture}
+%
+\\
+%
+\begin{tikzpicture}
+	\axpl{Auto-Mpg300_lambda0.8_25runs}{Auto MPG}{left}
+\end{tikzpicture}
+%
+&
+%
+\begin{tikzpicture}
+	\axpl{Kinematics300_lambda0.8_25runs}{Kinematics}{right}
+\end{tikzpicture}
+
+\end{tabular}
+
+\caption{results for Housing, Abalone, Auto MPG and Kinematics datasets. Methods are: 1: \ac{GAPoly}, 2: Random Forests 3: Linear Regression, 4: SVM, 5: RPART and 6: CI.Tree}
+\label{Housing_dataset_lambda0.8_25runs}
+\label{Abalone_dataset_lambda0.8_25runs}
+\label{Auto-Mpg_dataset_lambda0.8_25runs}
+\label{Kinematics300_lambda0.8_25runs}
+\end{center}\end{figure}
+
+\begin{description}
+\item{\textsc{Artificial}} this is an artificial dataset with four numeric features, $x_1, \ldots x_4$, where $x_1,x_3$ are outcomes from Poisson random variables, and $x_2,x_4$ from Normal random variables.
 The dependent variable $y$ is given by expression $x_2x_4^2 + x_1^2x_3 + 5$.
-The dataset includes 50 observations and, only in this case, there were executed just 10 runs.
+The dataset includes $n=50$ observations and, only in this case, there were executed just 10 runs.
 
@@
 figure~\ref{artificial_dataset1_lambda1.0}).
-In this case, we used a population of 100 solutions with 50 iterations for each run.
+In this case, we used a population of $N=100$ solutions with 50 iterations for each run.
+\textbf{Não entendo}
 
-In the next datasets, the population has size 250 with 100 iterations for each run.
+In the next datasets, the population has size $n=250$ with 100 iterations for each run.
 
-\textbf{Housing}: This data set concerns the task of predicting housing values in areas of Boston.
-There are 13 continuous attributes and the dependent variable is the median value of owner-occupied homes in \$1000's\footnote{This and the remaining regression datasets were selected from Luis Torgo's data repository, http://www.dcc.fc.up.pt/~ltorgo/Regression/DataSets.html.
+\item{\textsc{Housing}}: This data set concerns the task of predicting housing values in areas of Boston.
+There are $m=13$ continuous attributes and the dependent variable is the median value of owner-occupied homes in \$1000's\footnote{This and the remaining regression datasets were selected from Luis Torgo's data repository, http://www.dcc.fc.up.pt/~ltorgo/Regression/DataSets.html.
 Most of these datasets originally come from UCI ML repository, http://archive.ics.uci.edu/ml/}.
-There are 506 observations.
+There are $N=506$ observations.
 Figure~\ref{Housing_dataset_lambda0.8_25runs} presents the results for this dataset.
 
-\begin{figure}[htb]\begin{center}
-\begin{tikzpicture}
-\begin{axis}[width=0.98\textwidth,height=6cm,
-	boxplot/draw direction=y,
-	xtick={1,2,3,4,5,6},
-	xticklabels={GA.Poly,Lin.Reg,SVM,RPART,Rnd.For,CI.Tree},
-	xlabel={Regression Method},
-	ymin=0.25,
-	ymax=0.85,
-	ylabel={Error distribution (\ac{rmse})},
-	title={Regression results for \textbf{Housing dataset}},
-	mark options={scale=0.5},
-	]
-	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
-	\addplot[boxplot] table [y index=0] {data/Housing300_lambda0.8_25runs.dat};
-	\addplot[boxplot] table [y index=1] {data/Housing300_lambda0.8_25runs.dat};
-	\addplot[boxplot] table [y index=2] {data/Housing300_lambda0.8_25runs.dat};
-	\addplot[boxplot] table [y index=3] {data/Housing300_lambda0.8_25runs.dat};
-	\addplot[boxplot] table [y index=4] {data/Housing300_lambda0.8_25runs.dat};
-	\addplot[boxplot] table [y index=5] {data/Housing300_lambda0.8_25runs.dat};
-
-\end{axis}
-\end{tikzpicture}
-
-% \includegraphics[height=6cm,width=\textwidth,angle=0]{figures/Abalone_dataset_lambdas.eps}
-
-		\caption{results for Boston Housing dataset}
-		\label{Housing_dataset_lambda0.8_25runs}
-
-\end{center}\end{figure}
-
-Just as an example of the model the \ac{GAPoly} algorithm outputs: in this dataset the best polynomial was the following:
+Just as an example of the model the \ac{GAPoly} algorithm outputs: in this dataset the best polynomial,
 
@@
 
-For comparisation if we access the mean decrease in accuracy found by Random Forests, the most important attributes are -- in decreasing order -- $x_{13}, x_6, x_5$ (but $x_5$ is already considered $4$ times less important than $x_6$).
+For comparation, if we access the mean decrease in accuracy found by Random Forests, the most important attributes are --- in decreasing order --- $x_{13}, x_6, x_5$ (but $x_5$ is already considered $4$ times less important than $x_6$).
 Both algorithms agree in two of their three most important attributes.
 
-\textbf{Abalone}: This dataset can be used to predict the age of a abalone shell using the given 8 numeric attributes concerning several physical measurements.
-There are 4177 observations.
+\item{\textsc{Abalone}} This dataset can be used to predict the age of a abalone shell using the given $m=8$ numeric attributes concerning several physical measurements.
+There are $n=4177$ observations.
 Figure~\ref{Abalone_dataset_lambda0.8_25runs} presents the results for this dataset.
-
-\begin{figure}[htb]\begin{center}
-\begin{tikzpicture}
-\begin{axis}[width=0.98\textwidth,height=6cm,
-	boxplot/draw direction=y,
-	xtick={1,2,3,4,5,6},
-	xticklabels={GA.Poly,Lin.Reg,SVM,RPART,Rnd.For,CI.Tree},
-	xlabel={Regression Method},
-	ymin=0.64,
-	ymax=0.79,
-	ylabel={Error distribution (\ac{rmse})},
-	title={Regression results for \textbf{Abalone dataset}},
-	mark options={scale=0.5},
-	]
-
-	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
-	\addplot[boxplot] table [y index=0] {data/Abalone300_lambda0.8_25runs.dat};
-	\addplot[boxplot] table [y index=1] {data/Abalone300_lambda0.8_25runs.dat};
-	\addplot[boxplot] table [y index=2] {data/Abalone300_lambda0.8_25runs.dat};
-	\addplot[boxplot] table [y index=3] {data/Abalone300_lambda0.8_25runs.dat};
-	\addplot[boxplot] table [y index=4] {data/Abalone300_lambda0.8_25runs.dat};
-	\addplot[boxplot] table [y index=5] {data/Abalone300_lambda0.8_25runs.dat};
-
-\end{axis}
-\end{tikzpicture}
-
-% \includegraphics[height=6cm,width=\textwidth,angle=0]{figures/Abalone_dataset_lambdas.eps}
-
-\caption{results for Abalone dataset} \label{Abalone_dataset_lambda0.8_25runs}
-
-\end{center}\end{figure}
 
@@
 
-\textbf{Auto-MPG}: This dataset is used to predict fuel consumption in miles per gallon, based on two discrete and five continuous attributes.
-There are 398 observations.
+\item{\textsc{Auto-MPG}} This dataset is used to predict fuel consumption in miles per gallon, based on two discrete and five continuous attributes ($m=7$).
+There are $n=398$ observations.
 Figure~\ref{Auto-Mpg_dataset_lambda0.8_25runs} presents the results for this dataset.
-
-\begin{figure}[htb]\begin{center}
-\begin{tikzpicture}
-\begin{axis}[width=0.98\textwidth,height=6cm,
-	boxplot/draw direction=y,
-	xtick={1,2,3,4,5,6},
-	xticklabels={GA.Poly,Lin.Reg,SVM,RPART,Rnd.For,CI.Tree},
-	xlabel={Regression Method},
-	ymin=0.3,
-	ymax=0.55,
-	ylabel={Error distribution (\ac{rmse})},
-	title={Regression results for \textbf{Auto-MPG dataset}},
-	mark options={scale=0.5},
-	]
-
-	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
-	\addplot[boxplot] table [y index=0] {data/Auto-Mpg300_lambda0.8_25runs.dat};
-	\addplot[boxplot] table [y index=1] {data/Auto-Mpg300_lambda0.8_25runs.dat};
-	\addplot[boxplot] table [y index=2] {data/Auto-Mpg300_lambda0.8_25runs.dat};
-	\addplot[boxplot] table [y index=3] {data/Auto-Mpg300_lambda0.8_25runs.dat};
-	\addplot[boxplot] table [y index=4] {data/Auto-Mpg300_lambda0.8_25runs.dat};
-	\addplot[boxplot] table [y index=5] {data/Auto-Mpg300_lambda0.8_25runs.dat};
-
-\end{axis}
-\end{tikzpicture}
-
-% \includegraphics[height=6cm,width=\textwidth,angle=0]{figures/Abalone_dataset_lambdas.eps}
-
-\caption{results for Auto-MPG dataset}	\label{Auto-Mpg_dataset_lambda0.8_25runs}
-
-\end{center}\end{figure}
 
@@
 
-\textbf{Kinematics}: This dataset is concerned with the realistic simulation of the forward kinematics of an 8 link robot arm.
-The task is to predict the distance of the end-effector from a target using 8 continuous attributes.
-There are 8192 observations.
+\item{\textsc{Kinematics}} This dataset is concerned with the realistic simulation of the forward kinematics of an 8 link robot arm.
+The task is to predict the distance of the end-effector from a target using $m=8$ continuous attributes.
+There are $n=8192$ observations.
 Figure~\ref{Kinematics300_lambda0.8_25runs} presents the results for this dataset.
-
-\begin{figure}[htb]\begin{center}
-\begin{tikzpicture}
-\begin{axis}[width=0.98\textwidth,height=7cm,
-	boxplot/draw direction=y,
-	xtick={1,2,3,4,5,6},
-	xticklabels={GA.Poly,Lin.Reg,SVM,RPART,Rnd.For,CI.Tree},
-	xlabel={Regression Method},
-	ymin=0.53,
-	ymax=0.85,
-	ylabel={Error distribution (\ac{rmse})},
-	title={Regression results for \textbf{Kinematics dataset}},
-	mark options={scale=0.5},
-	]
-
-	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
-	\addplot[boxplot] table [y index=0] {data/Kinematics300_lambda0.8_25runs.dat};
-	\addplot[boxplot] table [y index=1] {data/Kinematics300_lambda0.8_25runs.dat};
-	\addplot[boxplot] table [y index=2] {data/Kinematics300_lambda0.8_25runs.dat};
-	\addplot[boxplot] table [y index=3] {data/Kinematics300_lambda0.8_25runs.dat};
-	\addplot[boxplot] table [y index=4] {data/Kinematics300_lambda0.8_25runs.dat};
-	\addplot[boxplot] table [y index=5] {data/Kinematics300_lambda0.8_25runs.dat};
-
-\end{axis}
-\end{tikzpicture}
-
-\caption{results for Kinematics dataset}
-		\label{Kinematics300_lambda0.8_25runs}
-
-\end{center}\end{figure}
 
@@
 % $$ y = 0.24 x_5 - 0.23 x_5 x_6 x_7 - 0.19 x_4 x_7 - 0.5 x_3 $$
+
+\end{description}
 
@@
 Since the entire learning process takes some time, in the current R implementation\footnote{The processing time can probably be speeded by one to two orders in magnitude if the process is implemented in a low level programming language like C++.
-However, speed optimisation was not the focus of this article}, placing a limit between 50 to 100 generations already achieves good results, relative to higher iteration values.
+However, speed optimization was not the focus of this article}, placing a limit between 50 to 100 generations already achieves good results, relative to higher iteration values.
 Figure~\ref{Abalone_fitnessProgress} shows a typical error evolution for the dataset Abalone given two different values for $\lambda$.
 
-\begin{figure}[htb]
+\begin{figure}[tb]
 \begin{center}
@@
 	xlabel={number of iterations},
-	ylabel={\ac{rmse}},
-	title={Error (\ac{rmse}) vs. number of iterations (\textbf{Abalone dataset})},
+	ylabel={error},
+	title={Error \emph{vs.} number of iterations (Dataset: \textsc{Abalone})},
 	]
@@
 \end{tikzpicture}
-\caption{fitness progress for Abalone dataset (population $200$)}
+\caption{Error progress for Abalone dataset (population $200$)}
 \label{Abalone_fitnessProgress}
@@
 
-The authors thank the Fundação para a Ciência e Tecnologia (FCT) and the  laboratório de I\&D LabMAg the financial support given to this work under the strategic project PEst-­OE/EEI/UI0434/2011.
+The authors are grateful to the \textsc{Fundação para a Ciência e Tecnologia} (FCT) and the  R\&D laboratory \textsc{LabMAg} for the financial support given to this work, under the strategic project \textsc{PEst-­OE/EEI/UI0434/2011}.
 

### 2013-07-23 texts/paper-draft.tex
added legends, future work, removed typos

@@
 
-This representation of polynomials makes the problem of structure search very clear: except for the trivial cases, the number of possible monomials given $n$ variables and a maximum joint degree $d$ grows exponentialy with either $n$ or $d$.
+This representation of polynomials makes the problem of structure search very clear: except for the trivial cases, the number of possible monomials given $n$ variables and a maximum joint degree $d$ grows exponentially with either $n$ or $d$.
 But more importantly, the polynomial regression problem can be naturally split into two subproblems:
@@
 %
-\subsection{Polinomial Encoding}\label{subs:polynomial.encoding}
+\subsection{Polynomial Encoding}\label{subs:polynomial.encoding}
 %
@@
 \item an initial segment detailing which monomials are active (the first monomial is always active), this is represented in unary description, i.e., each monomial is identified by a single bit
-\item the remaining is split into $k$ sets of bits of equal size, each one representing a monomial
+\item the remaining bits are split into $k$ sets of equal size, each one representing a monomial
 \item each monomial is split into $m$ sets of $d$ size each, \emph{i.e.}, a variable
-\item for each variable, the remaining bits give the binary description of the variable degree, \emph{i.e.}, the maximum exponent is given by $2^d-1$
+\item for each variable, the remaining bits are the binary description of the variable degree, \emph{i.e.}, the maximum exponent is given by $2^d-1$
 \end{enumerate}
@@
 This last monomial does not enter neither in the polynomial regression nor in the \ac{rmse} (fitness) evaluation.
-However, it acts as a kind of junk DNA, becoming active when, in a future mutation or crossover, the third bit of the entire sequence flips from 0 to 1.
+However, it acts as a kind of junk DNA, becoming active when, in a future mutation or crossover, the third bit of the entire sequence flips from~0 to~1.
 
@@
 The second triple, $000$, means that $x_2$ is not part of the monomial.
-The third triple $001$ says that variable $x_3$ has exponent $1$, so the first monomial consists of $x_1^3 x_3$.
-All the remaining sets of nine bits are interpreted the same way and we get the previous polynomial.
+The third triple $001$ says that variable $x_3$ has exponent $1$.
+The first monomial is~$x_1^3 x_3$.
 
 Notice that all binary descriptions give rise to valid polynomials.
-Notice however that this is not a bijective mapping.
+However this is not a bijective mapping.
 For each polynomial there are multiple representations.
 For example, $x_1+x_2$ and $x_2+x_1$ have different representations.
-The authors considered that more complex mappings in order to force a one to one mapping would impact negatively in the algorithm performance without giving anything in return.
+The authors considered that more complex mappings in order to force a one to one mapping would impact negatively in the algorithm's performance.
 
-If the coding consists of entirely zeros, by convention, it describes polynomial $x_1$.
+% If the coding consists of entirely zeros, by convention, it describes polynomial $x_1$.
 This has to do with the execution of the polynomial regression that would fail if we interpret it as the zero polynomial.
@@
 
-The polynomial regression \ac{rmse} considered so far is based on the ability to predict the test set after the polynomial regression has found the appropriate coefficients $\theta_i$ for each one of the polynomial's monomials $q_i$.
+The polynomial regression \ac{rmse} considered so far is based on the ability to predict the test set after the polynomial regression has found the appropriate coefficients $\theta_i$ for each one of the monomials $q_i$.
 
@@
 
-Somewhat unexpectedly after some experiences it was found that lower values for $\lambda$ sometimes provide better, even if marginal, results.
-Figure~\ref{Abalone_dataset_lambdas} shows results for the Abalone dataset with ten runs for each $\lambda$, and figure~\ref{Auto-MPG_dataset_lambdas} presents regularization results for the Auto-MPG dataset.
-The following section Experimental Results includes information about these datasets.
+Somewhat unexpectedly after some experiences it was found that lower values for $\lambda$ provide better, even if marginal, results.
+Figure~\ref{Abalone_dataset_lambdas} shows regularization results for the Abalone and Auto-MPG datasets with ten runs for each $\lambda$.
+The following section  includes information about these datasets.
 
@@
 
-\caption{Error distribution by regularization exponent for the Abalone dataset.}\label{Abalone_dataset_lambdas}\label{Auto-MPG_dataset_lambdas}
+\caption{Error distribution by regularization exponent for the Abalone and Auto-MPG datasets.
+The box plots summarize the error values of 10 runs for each value of $\lambda$.
+The smallest overall error is shown in gray and, in both cases, was achieved with $\lambda=0.8$.}
+\label{Abalone_dataset_lambdas}\label{Auto-MPG_dataset_lambdas}
 
@@
 
-The typical inflection point lies around $\lambda = 0.8$. The following results use the regularization parameter with this value.
+The typical inflection point lies around $\lambda = 0.8$. The dataset results for applying the proposed regression method use the regularization parameter with this value.
 
@@
 
-It was used the R package genalg~\cite{willighage2012genalg} for the execution of genetic algorithm.
-The operators were the standard ones: (a) crossover, i.e., a pair of solutions from the previous generations are combined by splitting and mixing their respective representations, and (b) mutation, changing the values of single bits; the mutation chance applied in the datasets was 5\%.
-There was also elitism between generations, i.e., 20\% of the best solutions are kept in (fc: survive to?) the next generation.
+To perform the genetic algorithm it was used the R package genalg~\cite{willighage2012genalg}.
+The operators were the standard ones: (a)~crossover, \emph{i.e.}, a pair of solutions from the previous generations are combined by splitting and mixing their respective representations, and (b)~mutation, changing the values of single bits; the mutation chance applied in the datasets was 5\%.
+There was also elitism between generations, \emph{i.e.}, 20\% of the best solutions survive to the next generation.
 
@@
 
-\begin{itemize}
-\item Measured quantities
+The results were found using R's programming language~\cite{R}.
 
-	\begin{itemize}
+To compare this paper's proposed algorithm we applied the exact same train and test samples using several well-known learning algorithms for regression, namely: Linear Regression, Support Vector Machines~\cite{Meyer12}, Regression Trees~\cite{Therneau13}, Conditional Inference Trees~\cite{Hothorn06, Strobl07, Strobl08} and Random Forests~\cite{Liaw02}
 
-	\item error
-	\item number of iterations to convergence
-	\item memory usage
-	\item F1, ROC, ?
-
-	\end{itemize}
-
-\item Selection of datasets and regression algorithms
-\item Summary Figures and Numeric results
-\end{itemize}
-These results were found using R's programming language~\cite{R}.
-
-To compare this paper's proposed algorithm we applied the exact same train and test samples using other learning algorithms, namely: Linear Regression, Support Vector Machines~\cite{Meyer12}, Regression Trees~\cite{Therneau13}, Conditional Inference Trees~\cite{Hothorn06, Strobl07, Strobl08} and Random Forests~\cite{Liaw02}
-
-In order to train and test the performance of \ac{GAPoly} it was used several datasets (see below).
+In order to train and test the performance of \ac{GAPoly} several mainstream datasets were used.
 For each dataset, we selected 70\% for training purposes and the remaining observations to make a test set in order to compute the estimated \ac{rmse}.
-To achieve more robust results, each dataset were processed 25 times, each one with different samples for the train and test sets.
+To achieve more robust results, each dataset were processed $25$~times, each one with different samples for the train and test sets.
 For the datasets with attribute values of different magnitudes, a preliminary scaling was executed.
-The results below are boxplots for the test set error predictions over these different runs.
+The results below are box plots for the test set error predictions over these different runs.
 
@@
 
-\caption{results for Artificial dataset \emph{must} highlight method \#1: GAPoly}
+\caption{results for Artificial dataset.
+Methods are: 1: \ac{GAPoly}, 2: Random Forests 3: Linear Regression, 4: SVM, 5: RPART and 6: CI.Tree.
+As shown, \ac{GAPoly} was able to find the exact polynomial structure that generates the dataset, reducing the test set prediction error to zero.}
 \label{artificial_dataset1_lambda1.0}
@@
 
-\caption{results for Housing, Abalone, Auto MPG and Kinematics datasets. Methods are: 1: \ac{GAPoly}, 2: Random Forests 3: Linear Regression, 4: SVM, 5: RPART and 6: CI.Tree}
+\caption{results for Housing, Abalone, Auto MPG and Kinematics datasets.
+Methods are: 1: \ac{GAPoly} (the grey box plot), 2: Random Forests 3: Linear Regression, 4: SVM, 5: RPART and 6: CI.Tree.
+Random Forests, an ensemble method, has the better results.
+However \ac{GAPoly} achieves similar errors in two of the datasets.
+For the remaining two \ac{GAPoly} produces similar errors to the other classic regression methods.}
 \label{Housing_dataset_lambda0.8_25runs}
@@
 \item{\textsc{Artificial}} this is an artificial dataset with four numeric features, $x_1, \ldots x_4$, where $x_1,x_3$ are outcomes from Poisson random variables, and $x_2,x_4$ from Normal random variables.
-The dependent variable $y$ is given by expression $x_2x_4^2 + x_1^2x_3 + 5$.
-The dataset includes $n=50$ observations and, only in this case, there were executed just 10 runs.
+The dependent variable~$y$ is given by expression $x_2x_4^2 + x_1^2x_3 + 5$.
+The dataset includes $n=50$~observations.
 
 This dataset was used in order to verify if \ac{GAPoly} was able to find the polynomial relation, which the algorithm did (cf.
-figure~\ref{artificial_dataset1_lambda1.0}).
-In this case, we used a population of $N=100$ solutions with 50 iterations for each run.
-\textbf{Não entendo}
+figure~\ref{artificial_dataset1_lambda1.0}).The genetic algorithm run with a population of $N=100$~solutions with 50~iterations for each run.
 
-In the next datasets, the population has size $n=250$ with 100 iterations for each run.
+In the next datasets, the population of the genetic algorithm had size~$n=250$ with 100~iterations for each run.
 
 \item{\textsc{Housing}}: This data set concerns the task of predicting housing values in areas of Boston.
-There are $m=13$ continuous attributes and the dependent variable is the median value of owner-occupied homes in \$1000's\footnote{This and the remaining regression datasets were selected from Luis Torgo's data repository, http://www.dcc.fc.up.pt/~ltorgo/Regression/DataSets.html.
-Most of these datasets originally come from UCI ML repository, http://archive.ics.uci.edu/ml/}.
+There are $m=13$ continuous attributes and the dependent variable is the median value of owner-occupied homes in \$1000's.
 There are $N=506$ observations.
-Figure~\ref{Housing_dataset_lambda0.8_25runs} presents the results for this dataset.
 
@@
 
-For comparation, if we access the mean decrease in accuracy found by Random Forests, the most important attributes are --- in decreasing order --- $x_{13}, x_6, x_5$ (but $x_5$ is already considered $4$ times less important than $x_6$).
+For comparison, if we access the mean decrease in accuracy found by Random Forests, the most important attributes are --- in decreasing order --- $x_{13}, x_6, x_5$ (but $x_5$ is already considered $4$ times less important than $x_6$).
 Both algorithms agree in two of their three most important attributes.
@@
 There are $n=4177$ observations.
-Figure~\ref{Abalone_dataset_lambda0.8_25runs} presents the results for this dataset.
 
@@
 
-\item{\textsc{Auto-MPG}} This dataset is used to predict fuel consumption in miles per gallon, based on two discrete and five continuous attributes ($m=7$).
-There are $n=398$ observations.
-Figure~\ref{Auto-Mpg_dataset_lambda0.8_25runs} presents the results for this dataset.
+\item{\textsc{Auto-MPG}} This dataset is used to predict fuel consumption in miles per gallon, based on two discrete and five continuous attributes ($m=7$). There are $n=398$ observations.
 
@@
 There are $n=8192$ observations.
-Figure~\ref{Kinematics300_lambda0.8_25runs} presents the results for this dataset.
 
@@
 
-The \ac{GA} quickly proceeds in the first 50 to 100 generations to reasonable error rates.
+The \ac{GA} quickly proceeds in the first~50 to~100 generations to reasonable error rates.
 Then, it proceeds slower achieving best solutions with marginal error reduction.
-Since the entire learning process takes some time, in the current R implementation\footnote{The processing time can probably be speeded by one to two orders in magnitude if the process is implemented in a low level programming language like C++.
-However, speed optimization was not the focus of this article}, placing a limit between 50 to 100 generations already achieves good results, relative to higher iteration values.
+Since the entire learning process takes some time, in the current R implementation, placing a limit between 50 to 100 generations already achieves good results, relative to higher iteration values.
 Figure~\ref{Abalone_fitnessProgress} shows a typical error evolution for the dataset Abalone given two different values for $\lambda$.
@@
 	xmin=-5,
-	xmax=255,
+	xmax=245,
 	mark options={scale=0.5},
@@
 \end{tikzpicture}
-\caption{Error progress for Abalone dataset (population $200$)}
+\caption{Error progress for Abalone dataset during a single execution of the genetic algorithm.
+The figure shows the fitness evolution for two different regularization values.
+The population for both consisted of $200$ polynomials.
+The error values seem to stabilize around iteration~250.}
 \label{Abalone_fitnessProgress}
@@
 
-\section{Conclusion}
+\section{Conclusion and Future Work}
 
@@
 
+\ac{GAPoly} demands some processing time.
+On a current quad-core computer, processing the Kinematics dataset (with 8k~observations) takes approximately 5~minutes.
+The processing time can probably be speeded by one to two orders in magnitude if the process is implemented in a low level programming language like C++.
+However, speed optimization was not the focus of this article.
+
+A cross-validation procedure can be implemented to refine the appropriate parameter values to achieve better errors.
+Namely, the regularization parameter, $\lambda$, can be tested with several values, instead of being fixed at $0.8$.
+Other parameters like mutation chance or the amount of elitism could also be tested.
+However, these type of tests need a low-level, fast implementation of \ac{GAPoly}.
+
 \section*{Acknowledgements}
@@
 The authors are grateful to the \textsc{Fundação para a Ciência e Tecnologia} (FCT) and the  R\&D laboratory \textsc{LabMAg} for the financial support given to this work, under the strategic project \textsc{PEst-­OE/EEI/UI0434/2011}.
+
+The datasets used herein were selected from Luis Torgo's data  repository, www.dcc.fc.up.pt/~ltorgo/Regression/DataSets.html.
+Most of these datasets originally come from UCI ML repository, http://archive.ics.uci.edu/ml/.
+
+The authors wish to thank professor André Falcão for all the useful discussions around the article.
 

### 2013-07-23 texts/paper_spfga-final.tex
Pre-submission: * renamed and reencoded main .tex file; * archived old text

--- 2013-07-23 texts/paper-draft.tex
@@
 %
-\usepackage[latin1]{inputenc}
+\usepackage[utf8]{inputenc}
 %
@@
 %
+%	HYPERREFERENCES
+%
+\usepackage{hyperref}
+%
+
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \title{Selection of Polynomial Features by Genetic Algorithms}
+
 \author[ue,labmag]{Francisco Coelho}
@@
 
-\author[fcul,labmag]{João Pedro Neto}
+\author[fcul,labmag]{João Pedro Neto\corref{jpn}}
 \ead{jpn@di.fc.ul.pt}
+\cortext[jpn]{Corresponding author}
 
-\address[ue]{Dept. Informática, Universidade de Évora, Rua Romão Ramalho 58 Évora 7000-671 Évora}
+\address[ue]{Dept. Informática, Universidade de Évora, Rua Romão Ramalho 58, 7000-671 Évora}
 \address[fcul]{Dept. Informática, Faculdade de Ciências da Universidade de Lisboa, Campo Grande 1749-016 Lisboa}
@@
 
+\begin{keyword}
+polynomial regression \sep genetic algorithm 
+\end{keyword}
 \end{frontmatter}
@@
 One interesting take on genetic algorithms, named \emph{genetic programming} by John Koza \cite{koza1992genetic}, proposed the use of \acp{GA} to search the syntactic structure of complex functions.
-This syntatic structure search is keen to the central ideas of deep learning \cite{bengio2013representation, bengio2009learning}, a subarea of machine learning actually producing quite promising results (\emph{e.g.} \cite{tarlowstochastic}).
+This syntactic structure search is keen to the central ideas of deep learning \cite{bengio2013representation, bengio2009learning}, a subarea of machine learning actually producing quite promising results (\emph{e.g.} \cite{tarlowstochastic}).
 It is also related to the work presented in this paper in the sense that, unlike linear models that have a simple structure, $y=\sum_i \beta_i x_i$, nonlinear (in particular polynomial) models pose an additional ``structure" search problem.
@@
 In line with that research this work describes a general method to find a polynomial regression of a given dataset.
-The optimal regression minimises a cost function that accounts for both the \ac{rmse} and a regularization factor to avoid over-fitting.
+The optimal regression minimizes a cost function that accounts for both the \ac{rmse} and a regularization factor to avoid over-fitting.
 
@@
 \begin{quote}
-Given a dataset $\mathcal{D}$ and a set of monomials $\mathcal{P}$, find parameters $\Theta$ that minimize $J\left(\Theta;\mathcal{P},\mathcal{D}\right)$.
+Given a dataset $\mathcal{D}$ and a set of monomials $\mathcal{P}$, find parameters $\Theta$ that minimize $J_{\textrm{fit}}\left(\Theta;\mathcal{P},\mathcal{D}\right)$.
 \end{quote}
@@
 
-The second problem is treated in the \ac{GA} setting: Let $\mathcal{D}$ be a dataset as above and $\mathcal{Q}$ a set of polynomials.
-For each polynomial $p\in \mathcal{Q}$ let $\mathcal{M}_p$ be the set of monomials in $p$ (without the coefficients) and compute the fitness
+The second problem is treated in the \ac{GA} setting: Let $\mathcal{D}$ be a dataset as above and $Q$ a set of polynomials.
+For each polynomial $p\in Q$ let $\mathcal{P}_p$ be the set of monomials in $p$ (without the coefficients) and compute the fitness
 %
 \begin{eqnarray*}
-\phi_p &=& \min_\Theta J\left(\Theta;\mathcal{M}_p,\mathcal{D}\right)
+\phi_p &=& \min_\Theta J_{\textrm{fit}}\left(\Theta;\mathcal{P}_p,\mathcal{D}\right)
 \end{eqnarray*}
@@
 by solving the first problem.
-With a fitness of every instance, a \ac{GA} will apply genetic operators (usually mutation and crossover) to evolve the population $\mathcal{P}$ until a reasonable approximation of a local minimum is found.
+With a fitness of every instance, a \ac{GA} will apply genetic operators (usually mutation and crossover) to evolve the population $Q$ until a reasonable approximation of a local minimum is found.
 %
@@
 This dataset was used in order to verify if \ac{GAPoly} was able to find the polynomial relation, which the algorithm did (cf.
-figure~\ref{artificial_dataset1_lambda1.0}).The genetic algorithm run with a population of $N=100$~solutions with 50~iterations for each run.
+figure~\ref{artificial_dataset1_lambda1.0}).
+The genetic algorithm run with a population of $n=100$~solutions with $50$~iterations for each run.
 
@@
 There are $m=13$ continuous attributes and the dependent variable is the median value of owner-occupied homes in \$1000's.
-There are $N=506$ observations.
+There are $n=506$ observations.
 
@@
 Only Random Forest outperforms \ac{GAPoly} systematically (it also outperforms all the other regression algorithms).
-One exception -- the artificial dataset that uses a straightforward polynomial relation -- is the Auto-MPG datatest where \ac{GAPoly} has comparable results.
+One exception --- the artificial dataset that uses a straightforward polynomial relation --- is the Auto-MPG dataset where \ac{GAPoly} has comparable results.
 This is evidence that applying standard genetic operators for polynomial model searching is a viable tool for regression purposes.
 
-\ac{GAPoly} demands some processing time.
+For complexity considerations \ac{GAPoly} demands some processing time.
 On a current quad-core computer, processing the Kinematics dataset (with 8k~observations) takes approximately 5~minutes.
@@
 
-The authors are grateful to the \textsc{Fundação para a Ciência e Tecnologia} (FCT) and the  R\&D laboratory \textsc{LabMAg} for the financial support given to this work, under the strategic project \textsc{PEst-­OE/EEI/UI0434/2011}.
+The authors are grateful to the \textsc{Fundação para a Ciência e Tecnologia} (FCT) and the  R\&D laboratory \textsc{LabMAg} for the financial support given to this work, under the strategic project \textsc{PEst-OE/EEI/UI0434/2011}.
 
-The datasets used herein were selected from Luis Torgo's data  repository, www.dcc.fc.up.pt/~ltorgo/Regression/DataSets.html.
-Most of these datasets originally come from UCI ML repository, http://archive.ics.uci.edu/ml/.
+The datasets used herein were selected from Luis Torgo's data  repository, \url{http://www.dcc.fc.up.pt/~ltorgo/Regression/DataSets.html}.
+Most come originally from UCI ML repository, \url{http://archive.ics.uci.edu/ml/}.
 

### 2013-07-23 texts/paper_spfga-final.tex
[creio que] o estilo do artigo, com as sub-opções, e o estilo das referências, estão OK;
0.5 de mim pensa que está pronto para submeter :D
0.5 de mim sabe que há erros por todo o lado :O

@@
-\documentclass{elsarticle}
+\documentclass[draft,authoryear,3p,times,twocolumn]{elsarticle}
 
@@
 Polynomials, one of the most studied subjects in mathematics, generalize li\-ne\-ar functions and define, perhaps, the simplest and most used nonlinear models.
-Applications include colorimetric calibration \cite{mendes2005adaptive}, explicit formulae for turbulent pipe flows \cite{davidson1999method}, computational linguistics \cite{sanchez2009obtaining} and more recently, analytical techniques for cultural heritage materials \cite{csefalvayova2010use}, liquid epoxy molding process \cite{chan2011modeling}, B-spline surface reconstruction \cite{galvez2012iterative}, product design \cite{chan2012development} or forecasting cyanotoxins presence in water reservoirs \cite{garcia2013hybrid}.
+Applications include colorimetric calibration \citep{mendes2005adaptive}, explicit formulae for turbulent pipe flows \citep{davidson1999method}, computational linguistics \citep{sanchez2009obtaining} and more recently, analytical techniques for cultural heritage materials \citep{csefalvayova2010use}, liquid epoxy molding process \citep{chan2011modeling}, B-spline surface reconstruction \citep{galvez2012iterative}, product design \citep{chan2012development} or forecasting cyanotoxins presence in water reservoirs \citep{garcia2013hybrid}.
 This not only illustrates the wide spectrum of applications but, additionally, work in each one of these polynomial models uses, at some point, a genetic algorithm.
@@
 If a nonlinear (or otherwise) model requires parameterization \acp{GA} provide a simple and often effective approach to search for locally optimal parameters.
-Research related to genetic algorithms abound and spans from the 1950s seminal work of Nils Aall Barricelli \cite{barricelli1962numerical} in the Institute for Advanced Study of Princeton to today's principal area of study for thousands of researchers, covered in hundreds of conferences, workshops and other meetings.
-Perhaps the key impulse to \acp{GA} come from John Holland's work and his book ``Adaptation in Natural and Artificial Systems'' \cite{holland1975adaptation}.
+Research related to genetic algorithms abound and spans from the 1950s seminal work of Nils Aall Barricelli \citep{barricelli1962numerical} in the Institute for Advanced Study of Princeton to today's principal area of study for thousands of researchers, covered in hundreds of conferences, workshops and other meetings.
+Perhaps the key impulse to \acp{GA} come from John Holland's work and his book ``Adaptation in Natural and Artificial Systems'' \citep{holland1975adaptation}.
 
-One interesting take on genetic algorithms, named \emph{genetic programming} by John Koza \cite{koza1992genetic}, proposed the use of \acp{GA} to search the syntactic structure of complex functions.
-This syntactic structure search is keen to the central ideas of deep learning \cite{bengio2013representation, bengio2009learning}, a subarea of machine learning actually producing quite promising results (\emph{e.g.} \cite{tarlowstochastic}).
+One interesting take on genetic algorithms, named \emph{genetic programming} by John Koza \citep{koza1992genetic}, proposed the use of \acp{GA} to search the syntactic structure of complex functions.
+This syntactic structure search is keen to the central ideas of deep learning \citep{bengio2013representation, bengio2009learning}, a subarea of machine learning actually producing quite promising results (\emph{e.g.} in \cite{Tarlow:2013fk}).
 It is also related to the work presented in this paper in the sense that, unlike linear models that have a simple structure, $y=\sum_i \beta_i x_i$, nonlinear (in particular polynomial) models pose an additional ``structure" search problem.
 
-The idea of using \acp{GA} to find a polynomial regression is not new \cite{maertens2006genetic, yu2008optimal, wu2009novel} but still generates original research \cite{hofwing2011optimal, cetisli2011polynomial}.
+The idea of using \acp{GA} to find a polynomial regression is not new \citep{maertens2006genetic, yu2008optimal, wu2009novel} but still generates original research \citep{hofwing2011optimal, cetisli2011polynomial}.
 In line with that research this work describes a general method to find a polynomial regression of a given dataset.
@@
 \begin{enumerate}
-\item for a given set of monomials $\mathcal{P} = \left\lbrace q_1, \ldots, q_k\right\rbrace$, find the regression coefficients $\theta_1,\ldots,\theta_k$ that minimize the \ac{rmse} on a given dataset;
+\item For a given set of monomials $\mathcal{P} = \left\lbrace q_1, \ldots, q_k\right\rbrace$, find the regression coefficients $\theta_1,\ldots,\theta_k$ that minimize the \ac{rmse} on a given dataset;
 
-\item find the fittest set of monomials, \emph{i.e.} the polynomial that minimizes the \ac{rmse} on the same dataset;
+\item Find the fittest set of monomials, \emph{i.e.} the polynomial that minimizes the \ac{rmse} on the same dataset;
 \end{enumerate}
@@
 \begin{eqnarray}
-J_{\textrm{fit}}\left(\Theta;\mathcal{P},\mathcal{D}\right) &=& \sqrt{\frac{1}{n}\sum_{i=1}^n \left( y\sample{i} - h_{\Theta,\mathcal{P}}\left( x_1\sample{i},\ldots,x_m\sample{i} \right) \right)^2 }\label{eq:rmse}
+J_{\textrm{fit}}\left(\Theta;\mathcal{P},\mathcal{D}\right)  & =  \nonumber \\
+\sqrt{\frac{1}{n}\sum_{i=1}^n \left( y\sample{i} - h_{\Theta,\mathcal{P}}\left( x_1\sample{i},\ldots,x_m\sample{i} \right) \right)^2 }\label{eq:rmse}
 \end{eqnarray}
@@
 	\While{$err > \epsilon$}
-		\State $pop \gets \Call{IterateGA}{pop}$\Comment{Build next generation}
-		\State $pop \gets \Call{Sort}{pop, key = \textsc{Lm.rmse}}$\Comment{Sort by \ac{rmse} of linear regression}
+		\State $pop \gets \Call{IterateGA}{pop}$
+		\State $pop \gets \Call{Sort}{pop, key = \textsc{Lm.rmse}}$
 		\State $err \gets \textsc{rmse}\left(\Call{First}{pop}\right)$
@@
 %
-Subsection \ref{subs:polynomial.encoding} describes the encoding of individual polynomial instances as chromosomes and other parameters of the \ac{GA} implementation used.
+Subsection \ref{subs:polynomial.encoding} describes the encoding of individual polynomial instances as chromosomes and other parameters of the utilized \ac{GA} implementation.
 The regularization of the cost function is discussed in subsection \ref{subs:cost.function}.
@@
 Let's see an example: consider polynomial $x_1^3 x_3 + x_3^7 + x_1 x_2$ with $k = 4, m = 3$ and $d=3$. One possible encoding would be:
-$$110 - 011,000,001~~;~~000,000,111~~;~~001,001,000~~;~~110,010,101$$
+\begin{verbatim}
+110 - 011,000,001 ; 000,000,111 ;
+      001,001,000 ; 110,010,101
+\end{verbatim}
 
@@
 
-Let's interpret the first monomial description, $011,000,001$.
+Let's interpret the first monomial description, \verb!011,000,001!.
 It is divided by three since $m=3$.
-The first triple $011$ is the binary description of the exponent of variable $x_1$ which is 3, so the first monomial includes $x_1^3$.
-The second triple, $000$, means that $x_2$ is not part of the monomial.
-The third triple $001$ says that variable $x_3$ has exponent $1$.
+The first triple \verb!011!
+is the binary description of the exponent of variable $x_1$ which is 3, so the first monomial includes $x_1^3$.
+The second triple, \verb!000!, means that $x_2$ is not part of the monomial.
+The third triple \verb!001!
+says that variable $x_3$ has exponent $1$.
 The first monomial is~$x_1^3 x_3$.
@@
 Somewhat unexpectedly after some experiences it was found that lower values for $\lambda$ provide better, even if marginal, results.
-Figure~\ref{Abalone_dataset_lambdas} shows regularization results for the Abalone and Auto-MPG datasets with ten runs for each $\lambda$.
+Figure~\ref{Abalone_dataset_lambdas} shows regularization results for the Abalone and Auto MPG datasets with ten runs for each $\lambda$.
 The following section  includes information about these datasets.
 
-\begin{figure}[tb]\begin{center}
+\begin{figure*}[tb]\begin{center}
 \begin{tabular}{cc}
@@
 \begin{axis}[
-	width=0.49\textwidth,height=6cm,
+	width=0.49\textwidth,height=4cm,
 	boxplot/draw direction=x,
@@
 
-\caption{Error distribution by regularization exponent for the Abalone and Auto-MPG datasets.
+\caption{Error distribution by regularization exponent for the Abalone and Auto MPG datasets.
 The box plots summarize the error values of 10 runs for each value of $\lambda$.
@@
 
-\end{center}\end{figure}
+\end{center}\end{figure*}
 
@@
 
-To perform the genetic algorithm it was used the R package genalg~\cite{willighage2012genalg}.
+To perform the genetic algorithm it was used the R package genalg~\citep{willighage2012genalg}.
 The operators were the standard ones: (a)~crossover, \emph{i.e.}, a pair of solutions from the previous generations are combined by splitting and mixing their respective representations, and (b)~mutation, changing the values of single bits; the mutation chance applied in the datasets was 5\%.
@@
 
-The results were found using R's programming language~\cite{R}.
-
-To compare this paper's proposed algorithm we applied the exact same train and test samples using several well-known learning algorithms for regression, namely: Linear Regression, Support Vector Machines~\cite{Meyer12}, Regression Trees~\cite{Therneau13}, Conditional Inference Trees~\cite{Hothorn06, Strobl07, Strobl08} and Random Forests~\cite{Liaw02}
+The results were found using \texttt{R} programming language~\citep{R}.
+To compare this paper's proposed algorithm we applied the exact same train and test samples using several well-known learning algorithms for regression, namely: Linear Regression, Support Vector Machines~\citep{Meyer12}, Regression Trees~\citep{Therneau13}, Conditional Inference Trees~\citep{Hothorn06, Strobl07, Strobl08} and Random Forests~\citep{Liaw02}
 
@@
 \begin{tikzpicture}
-\begin{axis}[width=0.90\textwidth,height=6cm,
+\begin{axis}[width=0.48\textwidth,height=4cm,
 	boxplot/draw direction=x,
@@
 	%yticklabels={GA.Poly,Rnd.For,Lin.Reg,SVM,RPART,CI.Tree},
-	ylabel={Regression Method},
-	xmin=-5,
-	xmax=300,
+	ylabel={Method index},
+	xmin=-25,
+	xmax=325,
 	xlabel={Error distribution},
@@
 
-\caption{results for Artificial dataset.
-Methods are: 1: \ac{GAPoly}, 2: Random Forests 3: Linear Regression, 4: SVM, 5: RPART and 6: CI.Tree.
-As shown, \ac{GAPoly} was able to find the exact polynomial structure that generates the dataset, reducing the test set prediction error to zero.}
+\caption{Results for Artificial dataset.
+As shown, \ac{GAPoly} was able to find the exact polynomial structure that generates the dataset, reducing the test set prediction error to zero.
+The regression methods depicted are: 1.
+\ac{GAPoly}, 2.
+Random Forests, 3.
+Linear Regression, 4.
+SVM, 5.
+Regression Trees and 6.
+Conditional Inference Trees}
 \label{artificial_dataset1_lambda1.0}
@@
 	title={Dataset: \textsc{#2}},
-	ylabel={Method},
-	xlabel={Error},
+	ylabel={Method index},
+	xlabel={Error distribution},
 	separate axis lines,
@@
 
-\begin{figure}[tb]\begin{center}
+\begin{figure*}[tb]\begin{center}
 \begin{tabular}{cc}
@@
 
-\caption{results for Housing, Abalone, Auto MPG and Kinematics datasets.
-Methods are: 1: \ac{GAPoly} (the grey box plot), 2: Random Forests 3: Linear Regression, 4: SVM, 5: RPART and 6: CI.Tree.
-Random Forests, an ensemble method, has the better results.
+\caption{Summary results for different regression methods on the Housing, Abalone, Auto MPG and Kinematics datasets.
+Random Forests, an ensemble method (``2'' in the plots), has the better results.
 However \ac{GAPoly} achieves similar errors in two of the datasets.
-For the remaining two \ac{GAPoly} produces similar errors to the other classic regression methods.}
+For the remaining two \ac{GAPoly} produces similar errors to the other classic regression methods.
+The regression methods depicted in these figures are: 1.
+\ac{GAPoly}, 2.
+Random Forests, 3.
+Linear Regression, 4.
+SVM, 5.
+Regression Trees and 6.
+Conditional Inference Trees}
 \label{Housing_dataset_lambda0.8_25runs}
@@
 \label{Kinematics300_lambda0.8_25runs}
-\end{center}\end{figure}
+\end{center}\end{figure*}
 
@@
 
-\item{\textsc{Auto-MPG}} This dataset is used to predict fuel consumption in miles per gallon, based on two discrete and five continuous attributes ($m=7$). There are $n=398$ observations.
+\item{\textsc{Auto MPG}} This dataset is used to predict fuel consumption in miles per gallon, based on two discrete and five continuous attributes ($m=7$). There are $n=398$ observations.
 
@@
 
-\begin{figure}[tb]
+\begin{figure*}[tb]
 \begin{center}
@@
 	width=0.98\textwidth,
-	height=6cm,
+	height=4cm,
 	xmin=-5,
 	xmax=245,
-	mark options={scale=0.5},
-	line width=1pt,
-	xlabel={number of iterations},
-	ylabel={error},
-	title={Error \emph{vs.} number of iterations (Dataset: \textsc{Abalone})},
+	%line width=1pt,
+	xlabel={Number of iterations},
+	ylabel={Error},
+	title={Error evolution with number of iterations (Dataset: \textsc{Abalone})},
 	]
@@
 \end{center}
-\end{figure}
+\end{figure*}
 
@@
 Only Random Forest outperforms \ac{GAPoly} systematically (it also outperforms all the other regression algorithms).
-One exception --- the artificial dataset that uses a straightforward polynomial relation --- is the Auto-MPG dataset where \ac{GAPoly} has comparable results.
+One exception --- the artificial dataset that uses a straightforward polynomial relation --- is the Auto MPG dataset where \ac{GAPoly} has comparable results.
 This is evidence that applying standard genetic operators for polynomial model searching is a viable tool for regression purposes.
@@
 On a current quad-core computer, processing the Kinematics dataset (with 8k~observations) takes approximately 5~minutes.
-The processing time can probably be speeded by one to two orders in magnitude if the process is implemented in a low level programming language like C++.
+The processing time can probably be speeded by one to two orders in magnitude if the process is implemented in a low level programming language like \texttt{C++}.
 However, speed optimization was not the focus of this article.
@@
 
-The authors are grateful to the \textsc{Fundação para a Ciência e Tecnologia} (FCT) and the  R\&D laboratory \textsc{LabMAg} for the financial support given to this work, under the strategic project \textsc{PEst-OE/EEI/UI0434/2011}.
+The authors are grateful to the Fundação para a Ciência e Tecnologia (FCT) and the  R\&D laboratory LabMAg for the financial support given to this work, under the strategic project \textsc{PEst-OE/EEI/UI0434/2011}.
 
-The datasets used herein were selected from Luis Torgo's data  repository, \url{http://www.dcc.fc.up.pt/~ltorgo/Regression/DataSets.html}.
+The datasets used herein were selected from Luís Torgo's data  repository, \url{http://www.dcc.fc.up.pt/~ltorgo/Regression/DataSets.html}.
 Most come originally from UCI ML repository, \url{http://archive.ics.uci.edu/ml/}.
 
-The authors wish to thank professor André Falcão for all the useful discussions around the article.
+The authors wish to thank professor André Falcão for motivation and useful discussions around the article.
 
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{model2-names}
 \bibliography{fullbib}

### 2013-07-25 texts/paper_spfga-draft.tex
A quantidade de porcaria que os editores são capazes de inventar é assombrosa.

--- 2013-07-23 texts/paper_spfga-final.tex
@@
 
-\author[ue,labmag]{Francisco Coelho}
+\author[ue,labmag]{Francisco Coelho\corref{fc}}
 \ead{fc@di.uevora.pt}
+\cortext[fc]{Corresponding author}
 
-\author[fcul,labmag]{João Pedro Neto\corref{jpn}}
+\author[fcul,labmag]{João Pedro Neto}
 \ead{jpn@di.fc.ul.pt}
-\cortext[jpn]{Corresponding author}
 
@@
 Polynomials, one of the most studied subjects in mathematics, generalize li\-ne\-ar functions and define, perhaps, the simplest and most used nonlinear models.
-Applications include colorimetric calibration \citep{mendes2005adaptive}, explicit formulae for turbulent pipe flows \citep{davidson1999method}, computational linguistics \citep{sanchez2009obtaining} and more recently, analytical techniques for cultural heritage materials \citep{csefalvayova2010use}, liquid epoxy molding process \citep{chan2011modeling}, B-spline surface reconstruction \citep{galvez2012iterative}, product design \citep{chan2012development} or forecasting cyanotoxins presence in water reservoirs \citep{garcia2013hybrid}.
+Applications include colorimetric calibration \citep{mendes2005adaptive}, explicit formul\ae\ for turbulent pipe flows \citep{davidson1999method}, computational linguistics \citep{sanchez2009obtaining} and more recently, analytical techniques for cultural heritage materials \citep{csefalvayova2010use}, liquid epoxy molding process \citep{chan2011modeling}, B-spline surface reconstruction \citep{galvez2012iterative}, product design \citep{chan2012development} or forecasting cyanotoxins presence in water reservoirs \citep{garcia2013hybrid}.
 This not only illustrates the wide spectrum of applications but, additionally, work in each one of these polynomial models uses, at some point, a genetic algorithm.
@@
 Only Random Forest outperforms \ac{GAPoly} systematically (it also outperforms all the other regression algorithms).
-One exception --- the artificial dataset that uses a straightforward polynomial relation --- is the Auto MPG dataset where \ac{GAPoly} has comparable results.
+One exception --- besides the artificial dataset that uses a straightforward polynomial relation --- is the Auto MPG dataset where \ac{GAPoly} has comparable results.
 This is evidence that applying standard genetic operators for polynomial model searching is a viable tool for regression purposes.

### 2013-07-25 texts/submission_01-prl/paper_spfga-final.tex
A quantidade de porcaria que os editores são capazes de inventar é assombrosa.

--- 2013-07-25 texts/paper_spfga-draft.tex
@@
-\documentclass[draft,authoryear,3p,times,twocolumn]{elsarticle}
+\documentclass[final,authoryear,3p,times,twocolumn]{elsarticle}
 
@@
 \acrodef{rmse}[error]{root-mean-square error}
-\acrodef{GAPoly}[{\sc GAPoly}]{Genetic Algorithms for Polynomials}
+\acrodef{GAPoly}[\textsc{GAPoly}]{Genetic Algorithms for Polynomials}
 %
@@
 \usepackage{graphicx}
-%
-
-%
-%	TIKZ AND RELATED STUFF
-%
-\usepackage{pgfplots}
-\pgfplotsset{compat=newest}
-\usepgfplotslibrary{statistics}
-%
-%	avoid processing figures every time
-%\usepgfplotslibrary{external}
-%\tikzexternalize
-% ENABLE NEXT LINE TO TYPESET FIGURES
-%\tikzset{external/force remake}
 %
@@
 %
-\begin{quote}
-Given a dataset $\mathcal{D}$ and a set of monomials $\mathcal{P}$, find parameters $\Theta$ that minimize $J_{\textrm{fit}}\left(\Theta;\mathcal{P},\mathcal{D}\right)$.
-\end{quote}
+\emph{Given a dataset $\mathcal{D}$ and a set of monomials $\mathcal{P}$, find parameters $\Theta$ that minimize $J_{\textrm{fit}}\left(\Theta;\mathcal{P},\mathcal{D}\right)$.}
+
 %
@@
 
+%
 \begin{figure*}[tb]\begin{center}
 \begin{tabular}{cc}
-\begin{tikzpicture}
-\begin{axis}[
-	width=0.49\textwidth,height=4cm,
-	boxplot/draw direction=x,
-	ytick={1,2,3,4,5,6,7},
-	yticklabels={0.7, 0.8, 0.9, 1.0, 1.1, 1.2, 1.3},
-	title = {Dataset: \textsc{Abalone}},
-	xlabel = {Error distribution},
-	ylabel={Regularization weight ($\lambda$)},
-	xmax=2.7,
-	]	
-	%\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
-	\addplot[boxplot] table [y index=0] {data/figure1.dat};
-	\addplot[boxplot][fill=gray!50!white] table [y index=2] {data/figure1.dat};
-	\addplot[boxplot] table [y index=4] {data/figure1.dat};
-	\addplot[boxplot] table [y index=6] {data/figure1.dat};		
-	\addplot[boxplot] table [y index=8] {data/figure1.dat};
-	\addplot[boxplot] table [y index=10] {data/figure1.dat};
-	\addplot[boxplot] table [y index=12] {data/figure1.dat};
-\end{axis}	
-\end{tikzpicture}
+\includegraphics[width=0.49\textwidth]{figure_1a.pdf}
 %
@@
 %
-\begin{tikzpicture}
-\begin{axis}[
-	width=0.49\textwidth,height=4cm,
-	boxplot/draw direction=x,
-	ytick={1,2,3,4,5,6,7},
-	%yticklabels={0.7, 0.8, 0.9, 1.0, 1.1, 1.2, 1.3},
-	yticklabels={},
-	title = {Dataset: \textsc{Auto MPG}},
-	xlabel = {Error distribution},
-	ylabel={},
-	]	
-	%\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
-	\addplot[boxplot] table [y index=0] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot][fill=gray!50!white] table [y index=2] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot] table [y index=4] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot] table [y index=6] {data/lambda.errors.Auto-Mpg.dat};		
-	\addplot[boxplot] table [y index=8] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot] table [y index=10] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot] table [y index=12] {data/lambda.errors.Auto-Mpg.dat};
-
-\end{axis}
-\end{tikzpicture}
+\includegraphics[width=0.49\textwidth]{figure_1b.pdf}
 \end{tabular}
@@
 
-The results were found using \texttt{R} programming language~\citep{R}.
+The results were found using \texttt{R} programming language~\citep{R}\footnote{The datasets and the \texttt{R} code used to produce the results and plots in this paper are available online at \url{https://github.com/jpneto/GenAlgPoly}.}.
 To compare this paper's proposed algorithm we applied the exact same train and test samples using several well-known learning algorithms for regression, namely: Linear Regression, Support Vector Machines~\citep{Meyer12}, Regression Trees~\citep{Therneau13}, Conditional Inference Trees~\citep{Hothorn06, Strobl07, Strobl08} and Random Forests~\citep{Liaw02}
@@
 \begin{figure}[tb]\begin{center}
-\begin{tikzpicture}
-\begin{axis}[width=0.48\textwidth,height=4cm,
-	boxplot/draw direction=x,
-	ytick={1,2,3,4,5,6},
-	%yticklabels={GA.Poly,Rnd.For,Lin.Reg,SVM,RPART,CI.Tree},
-	ylabel={Method index},
-	xmin=-25,
-	xmax=325,
-	xlabel={Error distribution},
-	title={Dataset: \textsc{Artificial}},
-	]
-	%\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
-	\addplot[boxplot][fill=gray!50!white] table [y index=0] {data/Articial_dataset.dat};
-	\addplot[boxplot] table [y index=4] {data/Articial_dataset.dat};
-	\addplot[boxplot] table [y index=1] {data/Articial_dataset.dat};
-	\addplot[boxplot] table [y index=2] {data/Articial_dataset.dat};
-	\addplot[boxplot] table [y index=3] {data/Articial_dataset.dat};
-	\addplot[boxplot] table [y index=5] {data/Articial_dataset.dat};
-
-\end{axis}
-\end{tikzpicture}
+\includegraphics[width=0.48\textwidth]{figure_2.pdf}
 
@@
 
-\newcommand{\axpl}[3]{
-\begin{axis}[width=0.49\textwidth,height=4cm,
-	ticks=major,
-	boxplot/draw direction=x,
-	ytick={1,2,3,4,5,6},
-	%yticklabels={GA.Poly, Lin.Reg, SVM, RPART,Rnd.For,CI.Tree},
-	title={Dataset: \textsc{#2}},
-	ylabel={Method index},
-	xlabel={Error distribution},
-	separate axis lines,
-	axis lines = box,
-%	axis y line = #3,
-	]
-	%\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
-	\addplot[boxplot][fill=gray!50!white] table [y index=0] {data/#1.dat};
-	\addplot[boxplot] table [y index=4] {data/#1.dat};
-	\addplot[boxplot] table [y index=1] {data/#1.dat};
-	\addplot[boxplot] table [y index=2] {data/#1.dat};
-	\addplot[boxplot] table [y index=3] {data/#1.dat};
-	\addplot[boxplot] table [y index=5] {data/#1.dat};
-
-\end{axis}
-}
-
 \begin{figure*}[tb]\begin{center}
 \begin{tabular}{cc}
-\begin{tikzpicture}
-	\axpl{Housing300_lambda0.8_25runs}{Housing}{left}
-\end{tikzpicture}
+\includegraphics[width=0.48\textwidth]{figure_3a.pdf}
 %
@@
 %
-\begin{tikzpicture}
-	\axpl{Abalone300_lambda0.8_25runs}{Abalone}{right}
-\end{tikzpicture}
+\includegraphics[width=0.48\textwidth]{figure_3b.pdf}
 %
@@
 %
-\begin{tikzpicture}
-	\axpl{Auto-Mpg300_lambda0.8_25runs}{Auto MPG}{left}
-\end{tikzpicture}
+\includegraphics[width=0.48\textwidth]{figure_3c.pdf}
 %
@@
 %
-\begin{tikzpicture}
-	\axpl{Kinematics300_lambda0.8_25runs}{Kinematics}{right}
-\end{tikzpicture}
-
+\includegraphics[width=0.48\textwidth]{figure_3d.pdf}
+%
 \end{tabular}
@@
 \begin{center}
-\begin{tikzpicture}
-
-\begin{axis}[
-	width=0.98\textwidth,
-	height=4cm,
-	xmin=-5,
-	xmax=245,
-	%line width=1pt,
-	xlabel={Number of iterations},
-	ylabel={Error},
-	title={Error evolution with number of iterations (Dataset: \textsc{Abalone})},
-	]
-
-	\addplot[solid,color=black] table [x expr=\coordindex, y index = 0] {data/lambdas.dat};\addlegendentry{$\lambda=0.975$};
-	\addplot[dashed,color=black] table [x expr=\coordindex, y index = 1] {data/lambdas.dat};\addlegendentry{$\lambda=1.000$};
-\end{axis}
-\end{tikzpicture}
+\includegraphics[width=0.98\textwidth]{figure_4.pdf}
 \caption{Error progress for Abalone dataset during a single execution of the genetic algorithm.

### 2013-07-29 texts/submission_01-prl/paper_spfga-review.tex
Acertos para a submissão.
mais correcções para a submissão.

--- 2013-07-25 texts/submission_01-prl/paper_spfga-final.tex
@@
-\documentclass[final,authoryear,3p,times,twocolumn]{elsarticle}
+\documentclass[review,preprint,authoryear,12pt]{elsarticle}
+
+%
+%	LINE NUMBERING FOR REVISION
+%
+\usepackage{lineno}
+%
 
@@
 \usepackage{amsfonts}
-\usepackage{amsthm}
+%\usepackage{amsthm}
 \newtheorem{theorem}{Theorem}
@@
 \begin{document}
-
 \begin{frontmatter}
@@
 
-\author[ue,labmag]{Francisco Coelho\corref{fc}}
+\author[ue,labmag]{Francisco ~Coelho\corref{fc}}
 \ead{fc@di.uevora.pt}
@@
 
-\author[fcul,labmag]{João Pedro Neto}
+\author[fcul,labmag]{João ~Pedro ~Neto}
 \ead{jpn@di.fc.ul.pt}
@@
 \address[labmag]{Laboratory of Agent Modelling (LabMAg)}
-
-\date{[{\sc draft} \today]}
-
+%
+%\date{[{\sc draft} \today]}
+%
 \begin{abstract}
@@
 \end{abstract}
-
+%
 \begin{keyword}
-polynomial regression \sep genetic algorithm 
+polynomial regression \sep genetic algorithm \sep feature extraction \sep dimensionality reduction 
 \end{keyword}
 \end{frontmatter}
-
+%
+%	Line numbers for revision
+%
+%
+\linenumbers
+%
 \section{Introduction}
@@
 Applications include colorimetric calibration \citep{mendes2005adaptive}, explicit formul\ae\ for turbulent pipe flows \citep{davidson1999method}, computational linguistics \citep{sanchez2009obtaining} and more recently, analytical techniques for cultural heritage materials \citep{csefalvayova2010use}, liquid epoxy molding process \citep{chan2011modeling}, B-spline surface reconstruction \citep{galvez2012iterative}, product design \citep{chan2012development} or forecasting cyanotoxins presence in water reservoirs \citep{garcia2013hybrid}.
-This not only illustrates the wide spectrum of applications but, additionally, work in each one of these polynomial models uses, at some point, a genetic algorithm.
+These examples not only illustrate the wide spectrum of applications but, additionally, work in each one uses, at some point, a genetic algorithm.
 
@@
 
-A usual representation of polynomials is
+An usual representation of polynomials is
 $$
@@
 The exponents alone are a matrix that defines the monomial structure of the polynomial, $A = \left[\alpha_{ij}\right]$.
-For the example above
+
+For the example above the matrix of monomials is
 $$
@@
 		\State $pop \gets \Call{IterateGA}{pop}$
-		\State $pop \gets \Call{Sort}{pop, key = \textsc{Lm.rmse}}$
-		\State $err \gets \textsc{rmse}\left(\Call{First}{pop}\right)$
+		\State $pop \gets \Call{Sort}{pop, key = J}$\Comment{Sort population by regression error}
+		\State $err \gets J\left(\Call{First}{pop}\right)$
 	\EndWhile
@@
 \caption{\ac{GAPoly} uses linear regression to find monomial coefficients that minimize the \ac{rmse} over a dataset and \acp{GA} to explore the space of polynomials.
+The role of the linear regression step is to produce a fitting error that sorts the population.
 At exit the \ac{rmse} of the fittest instance is bounded by $\epsilon$.}\label{alg:gapoly}
@@
 Let's see an example: consider polynomial $x_1^3 x_3 + x_3^7 + x_1 x_2$ with $k = 4, m = 3$ and $d=3$. One possible encoding would be:
-\begin{verbatim}
-110 - 011,000,001 ; 000,000,111 ;
-      001,001,000 ; 110,010,101
-\end{verbatim}
+\begin{center}
+\texttt{110 - 011,000,001 ; 000,000,111 ; 001,001,000 ; 110,010,101}
+\end{center}
 
@@
 
-\begin{figure}[tb]\begin{center}
-\includegraphics[width=0.48\textwidth]{figure_2.pdf}
+\begin{figure*}[tb]\begin{center}
+\includegraphics[width=0.98\textwidth]{figure_2.pdf}
 

### 2013-09-05 texts/paper_spfga-final.tex
migrações

@@
 %
+%	EMBEDDED DATASETS
+%
+\begin{filecontents}{inlambdas.dat}
+2.474 2.460
+2.414 2.392
+2.414 2.392
+2.414 2.392
+2.407 2.392
+2.407 2.392
+2.317 2.331
+2.351 2.331
+2.330 2.331
+2.314 2.318
+2.251 2.318
+2.251 2.318
+2.251 2.318
+2.275 2.318
+2.275 2.318
+2.275 2.318
+2.275 2.318
+2.275 2.296
+2.275 2.296
+2.268 2.296
+2.268 2.296
+2.268 2.296
+2.225 2.296
+2.225 2.296
+2.225 2.296
+2.225 2.296
+2.225 2.296
+2.225 2.296
+2.225 2.268
+2.225 2.268
+2.225 2.268
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.261
+2.225 2.247
+2.225 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.247
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.200 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.238
+2.185 2.221
+2.185 2.221
+\end{filecontents}
+
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 
@@
 
+%
+\newcommand{\boxp}[6]{\addplot
+	[boxplot prepared={
+		lower whisker=#1,
+		lower quartile=#2,
+		median=#3,
+		upper quartile=#4,
+		upper whisker=#5,
+	},][fill=#6!50!white] coordinates {}}
+%
 \begin{figure*}[tb]\begin{center}
@@
 	ylabel={Regularization weight ($\lambda$)},
-	xmax=2.7,
 	]	
-	%\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
-	\addplot[boxplot] table [y index=0] {data/figure1.dat};
-	\addplot[boxplot][fill=gray!50!white] table [y index=2] {data/figure1.dat};
-	\addplot[boxplot] table [y index=4] {data/figure1.dat};
-	\addplot[boxplot] table [y index=6] {data/figure1.dat};		
-	\addplot[boxplot] table [y index=8] {data/figure1.dat};
-	\addplot[boxplot] table [y index=10] {data/figure1.dat};
-	\addplot[boxplot] table [y index=12] {data/figure1.dat};
+	%
+	%	Original dataset: data/figure1.dat
+	%	y indexes: 0,2,4,6,8,10,12
+	%
+	\boxp{2.18}{2.21}{2.27}{2.33}{2.36}{white};	
+	\boxp{2.12}{2.15}{2.16}{2.23}{2.34}{gray};						
+	\boxp{2.18}{2.20}{2.23}{2.29}{2.33}{white};
+	\boxp{2.15}{2.16}{2.20}{2.29}{2.31}{white};
+	\boxp{2.46}{2.465}{2.47}{2.473}{2.476}{white};
+	\boxp{2.44}{2.48}{2.62}{2.65}{2.655}{white};
+	\boxp{2.36}{2.47}{2.55}{2.56}{2.61}{white};	
 \end{axis}	
@@
 	ytick={1,2,3,4,5,6,7},
-	%yticklabels={0.7, 0.8, 0.9, 1.0, 1.1, 1.2, 1.3},
 	yticklabels={},
@@
 	]
-	%\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
-	\addplot[boxplot] table [y index=0] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot][fill=gray!50!white] table [y index=2] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot] table [y index=4] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot] table [y index=6] {data/lambda.errors.Auto-Mpg.dat};		
-	\addplot[boxplot] table [y index=8] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot] table [y index=10] {data/lambda.errors.Auto-Mpg.dat};
-	\addplot[boxplot] table [y index=12] {data/lambda.errors.Auto-Mpg.dat};
-
+	%
+	%	Original dataset: data/lambda.errors.Auto-Mpg.dat
+	%	y indexes: 0,2,4,6,8,10,12
+	%
+	\boxp{2.27}{2.63}{2.79}{2.87}{3.09}{white};
+	\boxp{2.41}{2.55}{2.65}{2.757}{2.88}{gray};	
+	\boxp{2.45}{2.58}{2.76}{2.88}{3.02}{white};	
+	\boxp{2.73}{2.731}{2.82}{2.86}{2.87}{white};	
+	\boxp{2.69}{2.86}{3.04}{3.27}{3.53}{white};
+	\boxp{2.68}{3.12}{3.37}{3.67}{3.86}{white};
+	\boxp{3.84}{4.22}{4.33}{4.5}{4.8}{white};
 \end{axis}
@@
 	ytick={1,2,3,4,5,6},
-	%yticklabels={GA.Poly,Rnd.For,Lin.Reg,SVM,RPART,CI.Tree},
 	ylabel={Method index},
 	xmin=-25,
-	xmax=325,
+	xmax=350,
 	xlabel={Error distribution},
@@
 	]
-	%\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
-	\addplot[boxplot][fill=gray!50!white] table [y index=0] {data/Articial_dataset.dat};
-	\addplot[boxplot] table [y index=4] {data/Articial_dataset.dat};
-	\addplot[boxplot] table [y index=1] {data/Articial_dataset.dat};
-	\addplot[boxplot] table [y index=2] {data/Articial_dataset.dat};
-	\addplot[boxplot] table [y index=3] {data/Articial_dataset.dat};
-	\addplot[boxplot] table [y index=5] {data/Articial_dataset.dat};
+	%
+	%	Original dataset: data/Articial_dataset.dat
+	%	y indexes: 0,4,1,2,3,5
+	%	cols #3 and #5 have the same values
+	%
+	\boxp{-0.005}{-0.001}{0.0}{0.001}{0.005}{gray};	
+	\boxp{152.18}{195.59}{213.88}{246.72}{314.78}{white};	
+	\boxp{63.69}{82.96}{87.07}{103.68}{134.66}{white};	
+	\boxp{56.10}{81.39}{87.70}{114.82}{157.83}{white};		
+	\boxp{137.46}{166.65}{196.00}{225.99}{284.15}{white};
+	\boxp{137.46}{166.65}{196.00}{225.99}{284.15}{white};
 
@@
 
-\newcommand{\axpl}[3]{
+\newcommand{\axpla}[3]{
 \begin{axis}[width=0.49\textwidth,height=4cm,
@@
 }
+\newcommand{\axpl}[2]{
+\begin{axis}[width=0.49\textwidth,height=4cm,
+	ticks=major,
+	boxplot/draw direction=x,
+	ytick={1,2,3,4,5,6},
+	title={Dataset: \textsc{#1}},
+	ylabel={Method index},
+	xlabel={Error distribution},
+	separate axis lines,
+	axis lines = box,
+	]
+#2
+\end{axis}
+}
 
@@
 \begin{tikzpicture}
-	\axpl{Housing300_lambda0.8_25runs}{Housing}{left}
+	\axpl{Housing}{
+	\boxp{0.39}{0.50}{0.54}{0.64}{0.85}{gray};
+	\boxp{0.29}{0.34}{0.37}{0.40}{0.48}{white};
+	\boxp{0.46}{0.51}{0.55}{0.56}{0.61}{white};
+	\boxp{0.46}{0.53}{0.57}{0.60}{0.67}{white};
+	\boxp{0.40}{0.48}{0.54}{0.57}{0.66}{white};
+	\boxp{0.41}{0.46}{0.51}{0.55}{0.62}{white};
+	}
 \end{tikzpicture}
@@
 \begin{tikzpicture}
-	\axpl{Abalone300_lambda0.8_25runs}{Abalone}{right}
+	\axpl{Abalone}{
+	\boxp{0.6497}{0.6743}{0.685}{0.6948}{0.7213}{gray};
+	\boxp{0.6494}{0.6684}{0.6745}{0.6812}{0.698}{white};
+	\boxp{0.6797}{0.6874}{0.6937}{0.6966}{0.7092}{white};
+	\boxp{0.6736}{0.6959}{0.7043}{0.7111}{0.7249}{white};
+	\boxp{0.7201}{0.7459}{0.7551}{0.7638}{0.7876}{white};
+	\boxp{0.6829}{0.7038}{0.7179}{0.7274}{0.7587}{white};
+	}
 \end{tikzpicture}
@@
 \begin{tikzpicture}
-	\axpl{Auto-Mpg300_lambda0.8_25runs}{Auto MPG}{left}
-\end{tikzpicture}
-%
+	\axpl{Auto MPG}{
+	\boxp{0.3234}{0.3423}{0.3744}{0.3919}{0.4264}{gray};
+	\boxp{0.3056}{0.3311}{0.3607}{0.3731}{0.4169}{white};
+	\boxp{0.371}{0.4042}{0.4153}{0.4341}{0.4763}{white};
+	\boxp{0.3696}{0.4122}{0.4286}{0.4476}{0.491}{white};
+	\boxp{0.3887}{0.4234}{0.445}{0.479}{0.5106}{white};
+	\boxp{0.3636}{0.3927}{0.4245}{0.4635}{0.4969}{white};
+	}
+\end{tikzpicture}%
 &
@@
 \begin{tikzpicture}
-	\axpl{Kinematics300_lambda0.8_25runs}{Kinematics}{right}
+	\axpl{Kinematics}{
+	\boxp{0.7434}{0.759}{0.7641}{0.7804}{0.7947}{gray};
+	\boxp{0.564}{0.5686}{0.5717}{0.5771}{0.587}{white};
+	\boxp{0.7473}{0.7625}{0.7676}{0.776}{0.7821}{white};
+	\boxp{0.7535}{0.7697}{0.7781}{0.7827}{0.792}{white};
+	\boxp{0.7974}{0.8097}{0.8211}{0.8292}{0.8377}{white};
+	\boxp{0.7328}{0.744}{0.7555}{0.7633}{0.778}{white};
+	}
 \end{tikzpicture}
-
+%
 \end{tabular}
 
+%
+%	Generate "boxplot prepared" parameters
+%
+%\begin{tikzpicture}
+%\newcommand{\dataset}{Kinematics300_lambda0.8_25runs}
+%\newcommand{\colindex}{5} % 0 4 1 2 3 5
+%\begin{axis}[
+%	boxplot/draw direction=x,
+%	ymax=2,
+%	%xmin=0.79,xmax=0.85,
+%	]
+%	\addplot[boxplot][fill=green!50!white] table [y index=\colindex] {data/\dataset.dat} [above]
+%	node at (boxplot box cs: \boxplotvalue{median},1)
+%  		{\{\pgfmathprintnumber[precision=4]{\boxplotvalue{lower whisker}}\}\{\pgfmathprintnumber[precision=4]{\boxplotvalue{lower quartile}}\}\{\pgfmathprintnumber[precision=4]{\boxplotvalue{median}}\}\{\pgfmathprintnumber[precision=4]{\boxplotvalue{upper quartile}}\}\{\pgfmathprintnumber[precision=4]{\boxplotvalue{upper whisker}}\}};
+%
+%\end{axis}
+%
+%\end{tikzpicture}
 \caption{Summary results for different regression methods on the Housing, Abalone, Auto MPG and Kinematics datasets.
@@
 
-	\addplot[solid,color=black] table [x expr=\coordindex, y index = 0] {data/lambdas.dat};\addlegendentry{$\lambda=0.975$};
-	\addplot[dashed,color=black] table [x expr=\coordindex, y index = 1] {data/lambdas.dat};\addlegendentry{$\lambda=1.000$};
+	\addplot[solid,color=black] table [x expr=\coordindex, y index = 0] {inlambdas.dat};\addlegendentry{$\lambda=0.975$};
+	\addplot[dashed,color=black] table [x expr=\coordindex, y index = 1] {inlambdas.dat};\addlegendentry{$\lambda=1.000$};
 \end{axis}

### 2013-10-30 texts/submission_02-asc/paper_spfga-review.tex
- Second Journal - Added paywalled journals

--- 2013-07-29 texts/submission_01-prl/paper_spfga-review.tex
@@
 \usepackage{acronym}
+%
 \acrodef{GA}{Genetic algorithms}
@@
 
-The proposed method has competitive results comparing with some well-known regression methods.
+Of all the non-ensemble regression methods considered, the proposed method is the one that shows, with one exception, the lowest error.
 Only Random Forest outperforms \ac{GAPoly} systematically (it also outperforms all the other regression algorithms).
-One exception --- besides the artificial dataset that uses a straightforward polynomial relation --- is the Auto MPG dataset where \ac{GAPoly} has comparable results.
+The single non-ensemble exception --- besides the artificial dataset that uses a straightforward polynomial relation --- is the Auto MPG dataset where \ac{GAPoly} has comparable results.
 This is evidence that applying standard genetic operators for polynomial model searching is a viable tool for regression purposes.

### 2013-11-15 tests_Housing.Rmd
add comparisation between reg & non reg GA.Poly

@@
 ########################################################
+# Compare versions of GA.Poly (with vs without regularization)
+
+source("gapolyfuncs.R")
+report <- test.regularization(my.data, population=30, iterations=25, lambda=0.8, runs=5)
+
+save(report, file="Housing300_lambda0.8.vs.no.reg.Rda") # recover with load("Housing300.Rda")
+
+########################################################
 # Lambda Test

### 2013-11-15 gapolyfuncs.R

@@ -292,3 +292,87 @@ monitorEvalFactory <- function(train.set, n.vars, max.monoids) {
     print(eval(bestSolution))
   }
 }
+
+#########################################################################
+# Compare GA.Poly with vs. without normalization
+
+# pre: my.data must have columns named x1...xn and the last one is y (the output)
+test.regularization <- function(my.data, 
+                                population=100,     # GA population
+                                iterations=25,      # GA iterations
+                                runs=10,            # number of runs
+                                mutation.rate=0.05, # GA mutation rate
+                                lambda=1.0,         # regularization variable
+                                verbose=TRUE) {
+  
+  library(genalg)
+  
+  train.p.size <- 0.7 # percentage of training set
+  n.vars       <- ncol(my.data)-1
+  
+  # vectors keeping the rsme errors for each regression method
+  ga.error     <- rep(0,runs)
+  ga.reg.error <- rep(0,runs)
+  
+  best.ga.model <- NULL  # the best GA model 
+  best.ga.error <- 10e10 # & its error
+
+  best.ga.reg.model <- NULL  # the best GA model 
+  best.ga.reg.error <- 10e10 # & its error
+  
+  for (i in 1:runs) {
+    
+    # make train & test set (ie, each run has a different train+test sets)
+    inTrain   <- sample(1:nrow(my.data), train.p.size * nrow(my.data))
+    train.set <- my.data[inTrain,]
+    test.set  <- my.data[-inTrain,]
+
+    # test GA Poly without regularization
+    
+    GAmodel <- rbga.bin(size = max.monoids + max.degree*n.vars*max.monoids, 
+                        popSize = population, 
+                        iters = iterations, 
+                        mutationChance = mutation.rate, 
+                        elitism = TRUE, 
+                        evalFunc = evalFuncFactory(train.set, n.vars, max.monoids, 0))
+    
+    best.solution <- GAmodel$population[1,]
+    best.formula <- paste0("y ~ ", make.formula(best.solution, n.vars, max.monoids))
+    ga.model <- lm(best.formula, data=train.set)
+    ga.pred  <- predict(ga.model, test.set[,-ncol(test.set)]) 
+    ga.error[i] <- rsme(ga.pred,test.set[,ncol(test.set)])
+    
+    if(ga.error[i] < best.ga.error) { # save best ga.model
+      best.ga.model <- ga.model
[…]

### 2013-11-15 gapolyfuncs.R
organized article references

@@ -368,7 +368,7 @@ test.regularization <- function(my.data,
     }
     
     if (verbose)
-      cat(paste(i,"."))
+      cat(paste0(i,"."))
   }
   
   list(ga.error=ga.error,         # make the errors report into a list

### 2013-11-15 tests_Housing.Rmd
organized article references
minot stuff

@@
 source("gapolyfuncs.R")
-report <- test.regularization(my.data, population=30, iterations=25, lambda=0.8, runs=5)
+report <- test.regularization(my.data, population=300, iterations=100, lambda=0.8, runs=25)
 
@@
 ########################################################
-# check what attributes Random Forest select has the most importants
+# check what attributes Random Forest select as the most importants
 

### 2013-11-15 texts/submission_02-asc/paper_spfga-review.tex
Regularizations

@@
 \acrodef{GAPoly}[\textsc{GAPoly}]{Genetic Algorithms for Polynomials}
+\acrodef{EPR}{Evolutionary Polynomial Regression}
 %
@@
 \newcommand{\sample}[1]{\ensuremath{^{\left(#1\right)}}}
+\newcommand{\note}[1]{\textbf{[~#1~]}}
 %
@@
 
-\title{Selection of Polynomial Features by Genetic Algorithms}
+\title{A Method for Regularisation of Evolutionary Polynomial Regressions using Support Vector Machines}
 
@@
 \begin{abstract}
-Many applications require models that have no acceptable linear approximation and many nonlinear models are defined by polynomials.
-The use of genetic algorithms to find polynomial models is decades old but still poses challenges due to the complexity of the search and different definitions of optimal solution.
+Many applications require models that have no acceptable linear approximation and the simpler nonlinear models are defined by polynomials.
+The use of genetic algorithms to find polynomial models, \ac{EPR}, provides an automatic method to build such model from data but still poses challenges due to the complexity of the search and different definitions of optimal solution.
 
@@
 \begin{keyword}
-polynomial regression \sep genetic algorithm \sep feature extraction \sep dimensionality reduction 
+evolutionary polynomial regression \sep regularization \sep feature extraction \sep dimensionality reduction 
 \end{keyword}
@@
 The linearity assumption has many advantages including reduced computational complexity and strong theoretical framework.
-However nonlinearity is unavoidable in many application scenarios, specially those with phase transitions or feedback loops, so common in ecology, cybernetics, robotics and other areas.
+However nonlinearity is unavoidable in many application scenarios, specially those with phase transitions or feedback loops, so common in engineering, ecology, cybernetics and other areas.
+The kernel trick in Support Vector Machines alleviates this problem by allowing a non-linear transformation of  the feature-space.
+\note{Find references.}
 
@@
 The idea of using \acp{GA} to find a polynomial regression is not new \citep{maertens2006genetic, yu2008optimal, wu2009novel} but still generates original research \citep{hofwing2011optimal, cetisli2011polynomial}.
-In line with that research this work describes a general method to find a polynomial regression of a given dataset.
+\note{Insert references and resume of EPR; Redo the purpose of this work.} In line with that research this work describes a general method to find a polynomial regression of a given dataset.
 The optimal regression minimizes a cost function that accounts for both the \ac{rmse} and a regularization factor to avoid over-fitting.

### 2013-11-16 gapolyfuncs.R
fixed problem with previous reg vs non.reg result

@@ -327,14 +327,13 @@ test.regularization <- function(my.data,
     train.set <- my.data[inTrain,]
     test.set  <- my.data[-inTrain,]
 
-    # test GA Poly without regularization
-    
+    # test GA Poly without regularization    
     GAmodel <- rbga.bin(size = max.monoids + max.degree*n.vars*max.monoids, 
                         popSize = population, 
                         iters = iterations, 
                         mutationChance = mutation.rate, 
                         elitism = TRUE, 
-                        evalFunc = evalFuncFactory(train.set, n.vars, max.monoids, 0))
+                        evalFunc = evalFuncFactory(train.set, n.vars, max.monoids, 1))
     
     best.solution <- GAmodel$population[1,]
     best.formula <- paste0("y ~ ", make.formula(best.solution, n.vars, max.monoids))

### 2013-11-20 texts/submission_02-asc/paper_spfga-review.tex
New Abstract

@@
 \acrodef{EPR}{Evolutionary Polynomial Regression}
+\acrodef{EPRR}[EPR$^2$]{Evolutionary Polynomial Regression with Regularisation}
 %
@@
 \begin{abstract}
-Many applications require models that have no acceptable linear approximation and the simpler nonlinear models are defined by polynomials.
-The use of genetic algorithms to find polynomial models, \ac{EPR}, provides an automatic method to build such model from data but still poses challenges due to the complexity of the search and different definitions of optimal solution.
+While many applications require models that have no acceptable linear approximation the simpler nonlinear models are defined by polynomials.
+The use of genetic algorithms to find polynomial models, \ac{EPR}, provides an automatic method to build such models from data but still poses challenges due to the complexity of the search and different definitions of optimal solution.
 
-This paper describes a two-step method that uses genetic algorithms and linear regression to find empirical polynomial regressions.
-Experiments on common datasets show that, discounting the training computational effort, this method is quite competitive.
+This paper describes \ac{EPRR}, an algorithm that extends the general \ac{EPR} method with regularisation and a set of experiences to compare both flavours of \ac{EPR} and other usual regression methods including Random Forests, Linear Regression and Support Vector Machines on common datasets.
+The empiric conclusion of those experiments is that \ac{EPRR} achieves better fitting errors and that, with respect to the plain \ac{EPR}, computation time is shorter.
 \end{abstract}
@@
 \begin{keyword}
-evolutionary polynomial regression \sep regularization \sep feature extraction \sep dimensionality reduction 
+evolutionary polynomial regression \sep regularisation \sep feature extraction \sep dimensionality reduction 
 \end{keyword}
@@
 
-Polynomials, one of the most studied subjects in mathematics, generalize li\-ne\-ar functions and define, perhaps, the simplest and most used nonlinear models.
-Applications include colorimetric calibration \citep{mendes2005adaptive}, explicit formul\ae\ for turbulent pipe flows \citep{davidson1999method}, computational linguistics \citep{sanchez2009obtaining} and more recently, analytical techniques for cultural heritage materials \citep{csefalvayova2010use}, liquid epoxy molding process \citep{chan2011modeling}, B-spline surface reconstruction \citep{galvez2012iterative}, product design \citep{chan2012development} or forecasting cyanotoxins presence in water reservoirs \citep{garcia2013hybrid}.
+Polynomials, one of the most studied subjects in mathematics, generalise li\-ne\-ar functions and define, perhaps, the simplest and most used nonlinear models.
+Applications include colorimetric calibration \citep{mendes2005adaptive}, explicit formul\ae\ for turbulent pipe flows \citep{davidson1999method}, computational linguistics \citep{sanchez2009obtaining} and more recently, analytical techniques for cultural heritage materials \citep{csefalvayova2010use}, liquid epoxy moulding process \citep{chan2011modeling}, B-spline surface reconstruction \citep{galvez2012iterative}, product design \citep{chan2012development} or forecasting cyanotoxins presence in water reservoirs \citep{garcia2013hybrid}.
 These examples not only illustrate the wide spectrum of applications but, additionally, work in each one uses, at some point, a genetic algorithm.
 
-\ac{GA} where, arguably, one of the hottest topics of research in the recent decades and with good reason since they outline an optimization scheme easy to conceptualize and with very broad application.
+\ac{GA} where, arguably, one of the hottest topics of research in the recent decades and with good reason since they outline an optimisation scheme easy to conceptualise and with very broad application.
 If a nonlinear (or otherwise) model requires parameterization \acp{GA} provide a simple and often effective approach to search for locally optimal parameters.

### 2013-11-20 texts/submission_02-asc/paper_spfga-review.tex
Completed abstract and almost completed introduction.

@@
-\documentclass[review,preprint,authoryear,12pt]{elsarticle}
+%\documentclass[review,preprint,authoryear,12pt]{elsarticle}
+\documentclass[preprint,authoryear,12pt]{elsarticle}
 
@@
 %
-\usepackage{lineno}
+%\usepackage{lineno}
 %
@@
 \acrodef{EPRR}[EPR$^2$]{Evolutionary Polynomial Regression with Regularisation}
+\acrodef{SVM}{Support Vector Machine}
 %
@@
 \newcommand{\sample}[1]{\ensuremath{^{\left(#1\right)}}}
+\newcommand{\at}[1]{\ensuremath{\!\left(#1\right)}}
 \newcommand{\note}[1]{\textbf{[~#1~]}}
@@
 While many applications require models that have no acceptable linear approximation the simpler nonlinear models are defined by polynomials.
-The use of genetic algorithms to find polynomial models, \ac{EPR}, provides an automatic method to build such models from data but still poses challenges due to the complexity of the search and different definitions of optimal solution.
+The use of genetic algorithms to find polynomial models from data is known as \ac{EPR}.
+%
+This paper introduces \ac{EPRR}, an algorithm that extends the \ac{EPR} method and describes a set of experiences on common datasets that compare both flavours of \ac{EPR} and other methods including Random Forests, Linear and Support Vector Regression.
 
-This paper describes \ac{EPRR}, an algorithm that extends the general \ac{EPR} method with regularisation and a set of experiences to compare both flavours of \ac{EPR} and other usual regression methods including Random Forests, Linear Regression and Support Vector Machines on common datasets.
-The empiric conclusion of those experiments is that \ac{EPRR} achieves better fitting errors and that, with respect to the plain \ac{EPR}, computation time is shorter.
+The empiric conclusion of those experiments is that \ac{EPRR} achieves better errors than the other non-ensemble methods and shorter computation time than plain \ac{EPR}.
 \end{abstract}
@@
 %
-\linenumbers
+%\linenumbers
 %
@@
 
-With notable exceptions (\emph{e.g.} neural networks) machine learning regression techniques are based on linear models.
-The linearity assumption has many advantages including reduced computational complexity and strong theoretical framework.
+With notable exceptions (\emph{e.g.} neural networks) machine learning regression techniques produce linear models.
+The linearity assumption has many advantages including reduced computational complexity and strong the\-o\-re\-ti\-cal framework.
 However nonlinearity is unavoidable in many application scenarios, specially those with phase transitions or feedback loops, so common in engineering, ecology, cybernetics and other areas.
-The kernel trick in Support Vector Machines alleviates this problem by allowing a non-linear transformation of  the feature-space.
-\note{Find references.}
+The kernel trick in \acp{SVM} (\cite{scholkopf1997kernel,liang2012eigen,Bao:2013aa}) alleviates this problem by allowing special non-linear transformations of the feature-space.
+The condition such transformations must meet is known as the \emph{kernel trick}, $k\at{x,x'} = \left< \varphi\at{x}, \varphi\at{x'} \right>$, where $\varphi$ is the feature-space transformation and $\left<\cdot,\cdot\right>$ denotes the inner product.
+The ``trick'' consists on computing the kernel $k\at{x,x'}$ while avoiding the computation of the inner product and the transformations $\varphi\at{x}, \varphi\at{x'}$.
+A special case of polynomial transformation, the \emph{polynomial kernel}, $k\at{x,x'} = \left<x,x'\right>^d$ is commonly used in regression and classification tasks with \acp{SVM}.
+However general polynomial transformations do not verify the kernel trick.
 
 Polynomials, one of the most studied subjects in mathematics, generalise li\-ne\-ar functions and define, perhaps, the simplest and most used nonlinear models.
-Applications include colorimetric calibration \citep{mendes2005adaptive}, explicit formul\ae\ for turbulent pipe flows \citep{davidson1999method}, computational linguistics \citep{sanchez2009obtaining} and more recently, analytical techniques for cultural heritage materials \citep{csefalvayova2010use}, liquid epoxy moulding process \citep{chan2011modeling}, B-spline surface reconstruction \citep{galvez2012iterative}, product design \citep{chan2012development} or forecasting cyanotoxins presence in water reservoirs \citep{garcia2013hybrid}.
+Applications include colorimetric calibration \citep{Mendes:2005aa}, explicit formul\ae\ for turbulent pipe flows
+ \citep{Davidson:1999aa}, computational linguistics \citep{Sanchez:2009aa} and more recently analytical techniques for cultural heritage materials \citep{Csefalvayova:2010aa}, liquid epoxy moulding process \citep{Chan:2011aa}, B-spline surface reconstruction \citep{Galvez:2012aa}, product design \citep{Chan:2012aa} or forecasting cyanotoxins presence in water reservoirs \citep{Garcia-Nieto:2013aa}.
 These examples not only illustrate the wide spectrum of applications but, additionally, work in each one uses, at some point, a genetic algorithm.
@@
 If a nonlinear (or otherwise) model requires parameterization \acp{GA} provide a simple and often effective approach to search for locally optimal parameters.
-Research related to genetic algorithms abound and spans from the 1950s seminal work of Nils Aall Barricelli \citep{barricelli1962numerical} in the Institute for Advanced Study of Princeton to today's principal area of study for thousands of researchers, covered in hundreds of conferences, workshops and other meetings.
-Perhaps the key impulse to \acp{GA} come from John Holland's work and his book ``Adaptation in Natural and Artificial Systems'' \citep{holland1975adaptation}.
+Research related to genetic algorithms abound and spans from the 1950s seminal work of Nils Aall Barricelli \citep{Barricelli:1962aa} in the Institute for Advanced Study of Princeton to today's principal area of study for thousands of researchers, covered in hundreds of conferences, workshops and other meetings.
+Perhaps the key impulse to \acp{GA} come from John Holland's work and his book ``Adaptation in Natural and Artificial Systems'' \citep{Holland:1975aa}.
 
-One interesting take on genetic algorithms, named \emph{genetic programming} by John Koza \citep{koza1992genetic}, proposed the use of \acp{GA} to search the syntactic structure of complex functions.
-This syntactic structure search is keen to the central ideas of deep learning \citep{bengio2013representation, bengio2009learning}, a subarea of machine learning actually producing quite promising results (\emph{e.g.} in \cite{Tarlow:2013fk}).
+One interesting take on genetic algorithms, named \emph{genetic programming} by John Koza \citep{Koza:1992aa}, proposed the use of \acp{GA} to search the syntactic structure of complex functions.
+This syntactic structure search is keen to the central ideas of deep learning \citep{Bengio:2009aa,Bengio:2013aa}, a subarea of machine learning actually producing quite promising results (\emph{e.g.} in \cite{Tarlow:2013fk}).
 It is also related to the work presented in this paper in the sense that, unlike linear models that have a simple structure, $y=\sum_i \beta_i x_i$, nonlinear (in particular polynomial) models pose an additional ``structure" search problem.
 
-The idea of using \acp{GA} to find a polynomial regression is not new \citep{maertens2006genetic, yu2008optimal, wu2009novel} but still generates original research \citep{hofwing2011optimal, cetisli2011polynomial}.
-\note{Insert references and resume of EPR; Redo the purpose of this work.} In line with that research this work describes a general method to find a polynomial regression of a given dataset.
+The idea of using \acp{GA} to find a polynomial regression is not new \citep{Maertens:2006aa, Yu:2008aa, Wu:2009aa} but still generates original research \citep{Hofwing:2011aa,Cetisli:2011aa}.
+The modern formulation of the use of \ac{GA} to find polynomial models is known as \acf{EPR} and systematization can be traced back to the work of Davidson, Savic and Walters (\cite{Davidson:2003aa}).
+Further developements include multi-objective optimizations (\cite{Giustolisi:2009aa}).
+
+\note{Redo the purpose of this work.} In line with that research this work describes a general method to find a polynomial regression of a given dataset.
 The optimal regression minimizes a cost function that accounts for both the \ac{rmse} and a regularization factor to avoid over-fitting.
@@
 
-It turns out that, discarding the computational cost of training, the polynomial regression method presented here, \ac{GAPoly}, provides a quite competitive regression method.
-Indeed, it is only systematically out-performed by random forests, an \emph{ensemble} method.
+It turns out that, discarding the computational cost of training, the polynomial regression method presented here, \acf{EPRR}, provides a quite effective regression method.
+Indeed, according to the results of common regression methods on several real-world datasets it is only systematically out-performed by random forests, an \emph{ensemble} method.
 
@@
 
+\note{Discutir a localidade do regressor}
+
 Of all the non-ensemble regression methods considered, the proposed method is the one that shows, with one exception, the lowest error.

### 2013-11-20 texts/submission_02-asc/paper_spfga-review.tex
Completed introduction.

@@
 The kernel trick in \acp{SVM} (\cite{scholkopf1997kernel,liang2012eigen,Bao:2013aa}) alleviates this problem by allowing special non-linear transformations of the feature-space.
-The condition such transformations must meet is known as the \emph{kernel trick}, $k\at{x,x'} = \left< \varphi\at{x}, \varphi\at{x'} \right>$, where $\varphi$ is the feature-space transformation and $\left<\cdot,\cdot\right>$ denotes the inner product.
+The condition such transformations must meet is known as the \emph{kernel trick}, $k\at{x,x'} = \left< \varphi\at{x}, \varphi\at{x'} \right>$, where $\varphi$ is the feature-space transformation and $\left<\cdot,\cdot\right>$ denotes inner product.
 The ``trick'' consists on computing the kernel $k\at{x,x'}$ while avoiding the computation of the inner product and the transformations $\varphi\at{x}, \varphi\at{x'}$.
@@
  \citep{Davidson:1999aa}, computational linguistics \citep{Sanchez:2009aa} and more recently analytical techniques for cultural heritage materials \citep{Csefalvayova:2010aa}, liquid epoxy moulding process \citep{Chan:2011aa}, B-spline surface reconstruction \citep{Galvez:2012aa}, product design \citep{Chan:2012aa} or forecasting cyanotoxins presence in water reservoirs \citep{Garcia-Nieto:2013aa}.
-These examples not only illustrate the wide spectrum of applications but, additionally, work in each one uses, at some point, a genetic algorithm.
+These examples not only illustrate the wide spectrum of applications but, additionally, in each one uses, at some point, a \ac{GA}.
 
-\ac{GA} where, arguably, one of the hottest topics of research in the recent decades and with good reason since they outline an optimisation scheme easy to conceptualise and with very broad application.
+\acp{GA} where, arguably, one of the hottest topics of research in the recent decades and with good reason since they outline an optimisation scheme easy to conceptualise and with very broad application.
 If a nonlinear (or otherwise) model requires parameterization \acp{GA} provide a simple and often effective approach to search for locally optimal parameters.
@@
 
-One interesting take on genetic algorithms, named \emph{genetic programming} by John Koza \citep{Koza:1992aa}, proposed the use of \acp{GA} to search the syntactic structure of complex functions.
+One interesting variation of genetic algorithms, named \emph{genetic programming} by John Koza \citep{Koza:1992aa}, proposed the use of \acp{GA} to search the syntactic structure of complex functions.
 This syntactic structure search is keen to the central ideas of deep learning \citep{Bengio:2009aa,Bengio:2013aa}, a subarea of machine learning actually producing quite promising results (\emph{e.g.} in \cite{Tarlow:2013fk}).
-It is also related to the work presented in this paper in the sense that, unlike linear models that have a simple structure, $y=\sum_i \beta_i x_i$, nonlinear (in particular polynomial) models pose an additional ``structure" search problem.
+It is also related to the work presented in this paper in the sense that, unlike linear models that have a simple structure, $y=\sum_i \beta_i x_i$, nonlinear (in particular polynomial) models pose an additional structure search problem.
 
@@
 
-\note{Redo the purpose of this work.} In line with that research this work describes a general method to find a polynomial regression of a given dataset.
-The optimal regression minimizes a cost function that accounts for both the \ac{rmse} and a regularization factor to avoid over-fitting.
+\note{Redo the purpose of this work.} This paper describes an extension of the general \ac{EPR} method to find a \emph{regularized} polynomial regression of a given dataset.
+The optimal regression results from a cost function that accounts for both the \ac{rmse} and a regularization factor to avoid over-fitting.
 
@@
 
-It turns out that, discarding the computational cost of training, the polynomial regression method presented here, \acf{EPRR}, provides a quite effective regression method.
-Indeed, according to the results of common regression methods on several real-world datasets it is only systematically out-performed by random forests, an \emph{ensemble} method.
+It turns out that, discarding the computational cost of training, the regression method presented here, \acf{EPRR}, prooves to be quite effective.
+Indeed, according to the results of common  methods on several real-world datasets it is only systematically out-performed by random forests, an \emph{ensemble} method.
 

### 2013-11-21 texts/submission_02-asc/paper_spfga-review.tex
start working on motivation

@@
 
-The empiric conclusion of those experiments is that \ac{EPRR} achieves better errors than the other non-ensemble methods and shorter computation time than plain \ac{EPR}.
+The empiric conclusion of those experiments is that \ac{EPRR} achieves better fitting than the other non-ensemble methods and shorter computation time than plain \ac{EPR}.
 \end{abstract}
@@
 However nonlinearity is unavoidable in many application scenarios, specially those with phase transitions or feedback loops, so common in engineering, ecology, cybernetics and other areas.
-The kernel trick in \acp{SVM} (\cite{scholkopf1997kernel,liang2012eigen,Bao:2013aa}) alleviates this problem by allowing special non-linear transformations of the feature-space.
+The kernel trick in \acp{SVM} (\cite{scholkopf1997kernel, liang2012eigen, Bao:2013aa}) alleviates this problem by allowing special non-linear transformations of the feature-space.
 The condition such transformations must meet is known as the \emph{kernel trick}, $k\at{x,x'} = \left< \varphi\at{x}, \varphi\at{x'} \right>$, where $\varphi$ is the feature-space transformation and $\left<\cdot,\cdot\right>$ denotes inner product.

### 2013-11-21 tests_Housing.Rmd
more data points for reg vs nonreg

@@
 ```{r}
-set.seed(101)
+set.seed(204)
 
@@
 
-save(report, file="Housing300_lambda0.8.vs.no.reg.Rda") # recover with load("Housing300.Rda")
+save(report, file="Housing300_lambda0.8.vs.no.reg_3.Rda") # recover with load("Housing300.Rda")
+
+df <- data.frame(GA.Poly = report$ga.error,
+                 GA.Poly.reg = report$ga.reg.error)
+
+write.table(df, "Housing300_lambda0.8.vs.no.reg_3.txt")
 

### 2013-11-22 tests_Housing.Rmd
add datapoints: optimizing with less steps

@@
 ```{r}
-set.seed(204)
+set.seed(304)
 
@@
 source("gapolyfuncs.R")
-report <- test.regularization(my.data, population=300, iterations=100, lambda=0.8, runs=25)
+report <- test.regularization(my.data, population=150, iterations=75, lambda=0.8, runs=50)
 
-save(report, file="Housing300_lambda0.8.vs.no.reg_3.Rda") # recover with load("Housing300.Rda")
+save(report, file="Housing300_lambda0.8.vs.no.reg_pop200.Rda") # recover with load("Housing300.Rda")
 
@@
 
-write.table(df, "Housing300_lambda0.8.vs.no.reg_3.txt")
+write.table(df, "Housing300_lambda0.8.vs.no.reg_pop200.txt")
 

### 2013-11-22 tests_Housing.Rmd
added 2nd version of regularization
FC suggestion: make a more standard regularization: add to the rsme the
sum of the squared weigths (of the linear regression)

@@
 ########################################################
-# Lambda Test
+# Lambda Test (1st variant of regularization)
 lambda.values <- seq(0.5,0.999,length.out=20)
@@
   ggtitle("Housing dataset")
+
+########################################################
+# Lambda Test (2nd variant of regularization)
+size <- 31
+lambda.values <- rep(NA, size)
+
+val <- 0.001
+for(i in 1:size) {
+  lambda.values[i] = val
+  val = val * 2^(1/3)  # duplicate every three steps
+}
+
+n.runs <- 5
+
+source("gapolyfuncs.R")
+errors <- test.lambda(my.data, lambda.values, runs=n.runs, population=5, iterations=15)
+
+df <- data.frame(error = as.vector(matrix(errors,nrow=1)),
+                 lambda = as.factor(rep(round(lambda.values,3), each=n.runs)))
+
+write.table(df, "lambda.errors.Housing.txt")
+
+df <- read.table("lambda.errors.Housing.txt", header=TRUE)
+
+df$lambda <- as.factor(df$lambda)
+
+library(ggplot2)
+qplot(lambda,error,data=df,geom=c("jitter","boxplot")) + 
+  ylab("root-mean-square error") + 
+  xlab("lambda value") +
+ # coord_cartesian(ylim = c(0, 50))  +
+  ggtitle("Housing dataset")
+
 ##############################################

### 2013-11-22 gapolyfuncs.R

@@ -88,10 +88,16 @@ evalFuncFactory <- function(df, n.vars, max.monoids, lambda=1) {
   function(chromosome) {
     formula <- paste0("y ~ ", make.formula(chromosome, n.vars, max.monoids))
     model <- lm(formula, data=df)
-    # for regularization: find how many monoids are there:
-    n.monoids <- length(strsplit(formula,"[+]")[[1]]) + 1 # check how many '+' are there
-    # rbga.bin() minimizes, so the rmse will be smaller as the interations advance
-    return( sqrt(mean(residuals(model)^2)) * lambda^n.monoids ) 
+
+#     # These steps refer to the first version of regularization, using #monoids
+#     # for regularization: find how many monoids are there:
+#     n.monoids <- length(strsplit(formula,"[+]")[[1]]) + 1 # check how many '+' are there
+#     # rbga.bin() minimizes, so the rmse will be smaller as the interations advance
+#     return( sqrt(mean(residuals(model)^2)) * lambda^n.monoids ) 
+    
+    # These steps refer to the 2nd version of regularization, using the sum of
+    # squared weigths outputted by lm()
+    return( sqrt(mean(residuals(model)^2)) + lambda * sum(model$coefficients^2) )
   }
 }
 

### 2013-11-22 tests_Housing.Rmd
some extra data

@@
 ```{r}
-set.seed(304)
+set.seed(121)
 
@@
 # Compare versions of GA.Poly (with vs without regularization)
+set.seed(765)
+source("gapolyfuncs.R")
+report <- test.regularization(my.data, population=300, iterations=75, lambda=0.01, runs=10)
 
-source("gapolyfuncs.R")
-report <- test.regularization(my.data, population=150, iterations=75, lambda=0.8, runs=50)
-
-save(report, file="Housing300_lambda0.8.vs.no.reg_pop200.Rda") # recover with load("Housing300.Rda")
+save(report, file="Housing300_lambda0.01.vs.no.reg2_2.Rda")  
+# report <- load("Housing300_lambda0.01.vs.no.reg2_1.Rda")  # recover with
 
@@
 
-write.table(df, "Housing300_lambda0.8.vs.no.reg_pop200.txt")
+write.table(df, "Housing300_lambda0.01.vs.no.reg2_2.txt")
 
@@
 # Lambda Test (2nd variant of regularization)
-size <- 31
+size <- 20
 lambda.values <- rep(NA, size)
 
-val <- 0.001
+val <- 0.003
 for(i in 1:size) {
@@
 
-n.runs <- 5
+n.runs <- 10
 
 source("gapolyfuncs.R")
-errors <- test.lambda(my.data, lambda.values, runs=n.runs, population=5, iterations=15)
+errors <- test.lambda(my.data, lambda.values, runs=n.runs, population=50, iterations=40)
 

### 2013-11-22 gapolyfuncs.R

@@ -308,7 +308,7 @@ test.regularization <- function(my.data,
                                 iterations=25,      # GA iterations
                                 runs=10,            # number of runs
                                 mutation.rate=0.05, # GA mutation rate
-                                lambda=1.0,         # regularization variable
+                                lambda=0.0,         # regularization variable
                                 verbose=TRUE) {
   
   library(genalg)
@@ -339,7 +339,7 @@ test.regularization <- function(my.data,
                         iters = iterations, 
                         mutationChance = mutation.rate, 
                         elitism = TRUE, 
-                        evalFunc = evalFuncFactory(train.set, n.vars, max.monoids, 1))
+                        evalFunc = evalFuncFactory(train.set, n.vars, max.monoids, 0)) # reg v.2
     
     best.solution <- GAmodel$population[1,]
     best.formula <- paste0("y ~ ", make.formula(best.solution, n.vars, max.monoids))
@@ -373,7 +373,7 @@ test.regularization <- function(my.data,
     }
     
     if (verbose)
-      cat(paste0(i,"."))
+      cat(paste0(i,". non reg: ",ga.error[i], "reg: ", ga.reg.error[i],"\n"))
   }
   
   list(ga.error=ga.error,         # make the errors report into a list

### 2013-11-23 tests_Housing.Rmd
added non regularized data (Abalone & Housing)

@@
 ```{r}
-set.seed(121)
+set.seed(321)
 
@@
 source("gapolyfuncs.R")
-report <- make.report(my.data, population=300, iterations=100, lambda=0.8, runs=25)
+report <- make.report(my.data, population=300, iterations=100, lambda=1, runs=25)
 
@@
 
-save(report, file="Housing300_lambda0.8.Rda") # recover with load("Housing300.Rda")
+save(report, file="Housing300_lambda1.Rda") # recover with load("Housing300.Rda")
 
@@
 
-write.table(df, "Housing300_lambda0.8.txt")
+write.table(df, "Housing300_lambda1.txt")
 

### 2013-11-23 gapolyfuncs.R

@@ -89,15 +89,15 @@ evalFuncFactory <- function(df, n.vars, max.monoids, lambda=1) {
     formula <- paste0("y ~ ", make.formula(chromosome, n.vars, max.monoids))
     model <- lm(formula, data=df)
 
-#     # These steps refer to the first version of regularization, using #monoids
-#     # for regularization: find how many monoids are there:
-#     n.monoids <- length(strsplit(formula,"[+]")[[1]]) + 1 # check how many '+' are there
-#     # rbga.bin() minimizes, so the rmse will be smaller as the interations advance
-#     return( sqrt(mean(residuals(model)^2)) * lambda^n.monoids ) 
+    # These steps refer to the first version of regularization, using #monoids
+    # for regularization: find how many monoids are there:
+    n.monoids <- length(strsplit(formula,"[+]")[[1]]) + 1 # check how many '+' are there
+    # rbga.bin() minimizes, so the rmse will be smaller as the interations advance
+    return( sqrt(mean(residuals(model)^2)) * lambda^n.monoids ) 
     
-    # These steps refer to the 2nd version of regularization, using the sum of
-    # squared weigths outputted by lm()
-    return( sqrt(mean(residuals(model)^2)) + lambda * sum(model$coefficients^2) )
+#     # These steps refer to the 2nd version of regularization, using the sum of
+#     # squared weigths outputted by lm()
+#     return( sqrt(mean(residuals(model)^2)) + lambda * sum(model$coefficients^2) )
   }
 }
 
@@ -180,7 +180,7 @@ make.report <- function(my.data,
     citree.error[i] <- rsme(citree.pred,test.set[,ncol(test.set)])    
     
     if (verbose)
-      cat(i)
+      cat(paste0(i,": error ga: ",ga.error[i]))
   }
   
   list(ga.error=ga.error,         # make the errors report into a list
@@ -339,8 +339,7 @@ test.regularization <- function(my.data,
                         iters = iterations, 
                         mutationChance = mutation.rate, 
                         elitism = TRUE, 
-                        evalFunc = evalFuncFactory(train.set, n.vars, max.monoids, 0)) # reg v.2
-    
+                        evalFunc = evalFuncFactory(train.set, n.vars, max.monoids, 1))
     best.solution <- GAmodel$population[1,]
     best.formula <- paste0("y ~ ", make.formula(best.solution, n.vars, max.monoids))
     ga.model <- lm(best.formula, data=train.set)

### 2013-11-24 tests_Auto-Mpg.Rmd
added non red data (Auto-MPG & Kinetics)
Auto-MPG does not have practically any diif
Kinetics non-reg is a little bit *better* :-(

@@
 source("gapolyfuncs.R")
-report <- make.report(my.data, population=300, iterations=50, lambda=0.8)
+report <- make.report(my.data, population=300, iterations=100, lambda=1, runs=25)
 
@@
 
-save(report, file="Auto-Mpg300.Rda") # recover with load("Auto-Mpg300.Rda")
+save(report, file="Auto-Mpg300_lambda1.Rda") # recover with load("Auto-Mpg300.Rda")
 
@@
 
-write.table(df, "Auto-Mpg300.txt")
+write.table(df, "Auto-Mpg300_lambda1.txt")
 

### 2013-11-24 tests_Kinematics.Rmd
added all data results on latex file
added non red data (Auto-MPG & Kinetics)
Auto-MPG does not have practically any diif
Kinetics non-reg is a little bit *better* :-(

@@
 source("gapolyfuncs.R")
-report <- make.report(my.data, population=300, iterations=100, runs=25, lambda=0.8)
+report <- make.report(my.data, population=300, iterations=100, runs=25, lambda=1)
 
@@
 
-save(report, file="Kinematics300_lambda0.8_25runs.Rda") # recover with load("Kinematics.Rda")
+save(report, file="Kinematics300_lambda1.Rda") # recover with load("Kinematics.Rda")
 
@@
 
-write.table(df, "Kinematics200.txt")
+write.table(df, "Kinematics300_lambda1.txt")
 

### 2013-12-03 gapolyfuncs.R
added datapoints with 'classic' EPR
ie, no regularization and all monoids are active. This is the closest
version we are able to in accordance to the algorithms from the
references.

@@ -60,6 +60,11 @@ make.formula <- function(bits, n.vars, max.monoids) {
   bits          <- bits[-active.bits]      # the remainder is the info about the monoids
   monoids       <- split(bits, ceiling(seq_along(bits)/(length(bits)/max.monoids)))
   formula <- ""
+  
+  # this next line simulates the EPR using all monoids
+  # it should be uncommented when we want to get results without any type of regulatization
+  active.monoid <- rep(1, max.monoids)
+  
   for(i in 1:length(monoids)) {
     if (active.monoid[i]==1) {
       monoid <- make.monoid(monoids[[i]], n.vars, vars)
@@ -98,6 +103,7 @@ evalFuncFactory <- function(df, n.vars, max.monoids, lambda=1) {
 #     # These steps refer to the 2nd version of regularization, using the sum of
 #     # squared weigths outputted by lm()
 #     return( sqrt(mean(residuals(model)^2)) + lambda * sum(model$coefficients^2) )
+#     # However, the results were worse than the 1st version...
   }
 }
 

### 2013-12-10 gapolyfuncs.R
added new variant
where the formula includes all linear terms

@@ -27,7 +27,7 @@ bin2dec <- function(x)
 # 
 # One possible coding -- for the values in the previous code snippet -- would be:
 # 
-# $$ 1111 011,000,001;000,000,111;001,001,000;000,000,000$$
+# $$ 1111 011,000,001;000,000,111;001,001,000;000,000,000 $$
 # 
 # (the semicolons separate monoids, the commas separate variables)
 # 
@@ -63,7 +63,7 @@ make.formula <- function(bits, n.vars, max.monoids) {
   
   # this next line simulates the EPR using all monoids
   # it should be uncommented when we want to get results without any type of regulatization
-  active.monoid <- rep(1, max.monoids)
+  # active.monoid <- rep(1, max.monoids)
   
   for(i in 1:length(monoids)) {
     if (active.monoid[i]==1) {
@@ -75,8 +75,13 @@ make.formula <- function(bits, n.vars, max.monoids) {
           formula = paste0(formula," + I(", monoid,")")
     }
   }
+  
+  # This next line adds all the linear terms into the expression
+  formula <- paste(paste0("I(x",1:n.vars,") + "), collapse="") + formula
+  
   if (formula=="")     # the extreme case where the polynomial is empty
     formula = "I(x1)"  # default polynomial formula (by convention)
+  
   formula
 }
 

### 2013-12-10 gapolyfuncs.R
add new data points using linear add-ons

@@ -76,11 +76,11 @@ make.formula <- function(bits, n.vars, max.monoids) {
     }
   }
   
-  # This next line adds all the linear terms into the expression
-  formula <- paste(paste0("I(x",1:n.vars,") + "), collapse="") + formula
-  
   if (formula=="")     # the extreme case where the polynomial is empty
     formula = "I(x1)"  # default polynomial formula (by convention)
+
+  # This next line adds all the linear terms into the expression
+  formula <- paste0( paste(paste0("I(x",1:n.vars,") + "), collapse=""), formula)
   
   formula
 }

### 2013-12-13 code/exploratory/gap.py
Started python version.

@@ -1,0 +1,190 @@
+#!/usr/bin/env python
+
+#
+#   Fitness strategy for sklearn
+#
+#
+#       1. Define a searcher:
+#           searcher = GridSearchCV(
+#               estimator,
+#               param_grid,
+#               
+#               )
+import numpy as np
+from sklearn import cross_validation as cv
+from sklearn import svm
+from sklearn import preprocessing
+from sklearn import datasets
+from deap import algorithms, base, creator, tools
+#
+#   Supress deprecated warnings
+#
+import warnings
+warnings.filterwarnings("ignore")
+#       
+def save_data(filename, data):
+    f = open(filename, 'w')
+    f.write('Id,Solution\n')
+    i=1
+    for x in data:
+        f.write('%d,%d\n' % (i,x))
+        i += 1
+    f.close()
+#    
+def load_data(root_folder = '../../datasets/'):
+    '''
+    Loads data from project data folder and returns a triple (train, labels, test) where:
+        - train and labels are used to setup classifiers
+        - test is used to produce predictions
+    '''
+    load_file = lambda f,s: np.loadtxt( open(f, 'r'), delimiter = ',', skiprows = s )
+    data_folder = root_folder + 'Auto-Mpg/'    
+    data = load_file( data_folder + 'auto_no_missing.data', 0)
+    
+    return data
+    
+class PolyTransform:
+    '''Represent a set of monomial transformations.
+    
+    It is assumed that a set of n variables, X1 .. Xn, is fixed.
+    An instance of this class defines a set of transformations, Z1 .. Zm, where
+        - each Zj = (X1 ** E1j) * .. *(Xn ** Enj) and
+        - the exponents Eij are integers or fractions.
+    
+    Example:
+    
+        Given the dataset X1,X2
+        
+        X1  X2
+        0   1
+        1   1
+        2   2
+        
+        and the PolyTransform [Z1 = [1,0], Z2 = [2,2], Z3 = [2,0]] we obtain
+        Z1  Z2  Z3
+        0   0   0
+        1   1   1
+        2   16  4
+    '''
+    def __init__(self, degrees, dtype = np.int8):
+        '''Defines a monomial transformation.
+        
+        Args:
+            degrees (list of lists of integers/fractions): the set of transformations.
+                - each element of degrees represents a single monomial transformation;
+                - all the elements of degrees are expected to have the same size;
+
[…]

### 2013-12-13 tests_Auto-Mpg.Rmd
added datapoints with 'classic' EPR
ie, no regularization and all monoids are active. This is the closest
version we are able to in accordance to the algorithms from the
references.

--- 2013-07-15 tests_Auto-Mpg.Rmd
@@
 source("gapolyfuncs.R")
-report <- make.report(my.data, population=300, iterations=50, lambda=0.8)
+report <- make.report(my.data, population=300, iterations=100, lambda=0.8, runs=25)
 
@@
 
-save(report, file="Auto-Mpg300.Rda") # recover with load("Auto-Mpg300.Rda")
+save(report, file="Auto-Mpg300_lambda0.8_add_linears.Rda") # recover with load("Auto-Mpg300.Rda")
 
@@
 
-write.table(df, "Auto-Mpg300.txt")
+write.table(df, "Auto-Mpg300_lambda0.8_add_linears.txt")
 

### 2013-12-13 texts/submission_02-asc/paper_spfga-review.tex
collect data & paper debug

@@
 \acrodef{GAPoly}[\textsc{GAPoly}]{Genetic Algorithms for Polynomials}
+\acrodef{GAPolyreg}[\textsc{GAPoly.reg}]{Genetic Algorithms for Polynomials with Regularization}
 \acrodef{EPR}{Evolutionary Polynomial Regression}
 \acrodef{EPRR}[EPR$^2$]{Evolutionary Polynomial Regression with Regularisation}
-\acrodef{SVM}{Support Vector Machine}
 %
@@
 
-\title{A Method for Regularisation of Evolutionary Polynomial Regressions using Support Vector Machines}
+\title{A Method for Regularisation of Evolutionary Polynomial Regressions}
 
@@
 %
-This paper introduces \ac{EPRR}, an algorithm that extends the \ac{EPR} method and describes a set of experiences on common datasets that compare both flavours of \ac{EPR} and other methods including Random Forests, Linear and Support Vector Regression.
+This paper introduces \ac{EPRR}, an algorithm that extends the \ac{EPR} method and describes a set of experiences on common datasets that compare both flavours of \ac{EPR} and other methods including Linear Regression, Regression Trees and Support Vector Regression.
 
@@
 
-\acp{GA} where, arguably, one of the hottest topics of research in the recent decades and with good reason since they outline an optimisation scheme easy to conceptualise and with very broad application.
-If a nonlinear (or otherwise) model requires parameterization \acp{GA} provide a simple and often effective approach to search for locally optimal parameters.
+\acp{GA} were, arguably, one of the hottest topics of research in the recent decades and with good reason since they outline an optimisation scheme easy to conceptualise and with very broad application.
+If a nonlinear (or otherwise) model requires parameterization, \acp{GA} provide a simple and often effective approach to search for locally optimal parameters.
 Research related to genetic algorithms abound and spans from the 1950s seminal work of Nils Aall Barricelli \citep{Barricelli:1962aa} in the Institute for Advanced Study of Princeton to today's principal area of study for thousands of researchers, covered in hundreds of conferences, workshops and other meetings.
@@
 The modern formulation of the use of \ac{GA} to find polynomial models is known as \acf{EPR} and systematization can be traced back to the work of Davidson, Savic and Walters (\cite{Davidson:2003aa}).
-Further developements include multi-objective optimizations (\cite{Giustolisi:2009aa}).
+Further developments include multi-objective optimizations (\cite{Giustolisi:2009aa}).
 
-\note{Redo the purpose of this work.} This paper describes an extension of the general \ac{EPR} method to find a \emph{regularized} polynomial regression of a given dataset.
+This paper describes an extension of the general \ac{EPR} method to find a \emph{regularized} polynomial regression of a given dataset.
 The optimal regression results from a cost function that accounts for both the \ac{rmse} and a regularization factor to avoid over-fitting.
+Another important change is in the coding mechanism which corresponds to two new features: (a) inclusion of the linear terms in the polynomial expression, and (b) the activation/deactivation of monoids during evolution to allow the increase/decrease of polynomial expressions.
 
@@
 
-It turns out that, discarding the computational cost of training, the regression method presented here, \acf{EPRR}, prooves to be quite effective.
+% It turns out that, discarding the computational cost of training, the regression method presented here, \acf{EPRR}, proves to be quite effective.
 Indeed, according to the results of common  methods on several real-world datasets it is only systematically out-performed by random forests, an \emph{ensemble} method.
 
-The remainder of this paper is organized as usual: the next section describes the details of our method and is followed by a presentation of some performance results.
-The last section draws some conclusions and points future research tasks.
+The next section describes the details of our method and is followed by a presentation of some performance results. The last section draws some conclusions and points future research tasks.
 
@@
 	\State $pop \gets pop_0;\: err \gets 1.0+\epsilon$
-	\While{$err > \epsilon$}
+	\While{$err > \epsilon \land \neg\Call{limit}{iterations}$}
 		\State $pop \gets \Call{IterateGA}{pop}$
@@
 The role of the linear regression step is to produce a fitting error that sorts the population.
-At exit the \ac{rmse} of the fittest instance is bounded by $\epsilon$.}\label{alg:gapoly}
+At exit the \ac{rmse} of the fittest instance is bounded by $\epsilon$ or the maximum number of allowed iterations.}\label{alg:gapoly}
 \end{algorithm} 
@@
 
-Let's interpret the first monomial description, \verb!011,000,001!.
+% Let's interpret the first monomial description, \verb!011,000,001!.
 It is divided by three since $m=3$.
@@
 The authors considered that more complex mappings in order to force a one to one mapping would impact negatively in the algorithm's performance.
+
+In the experiments that led to this article, our tests concluded that the inclusion of the linear terms achieved a significant reduction of the error rates.
+So, considering the previous example, the final polynomial would be $x_1^3 x_3 + x_3^7 + x_1 x_2 + x_1 + x_2 + x_3$.
 
@@
 
+We'll denote by \ac{GAPolyreg} the proposed extension of the EPR algorithm which includes the regularized cost function together with the new features in polynomial coding.
+
 \subsection{Genetic Operators}
@@
 The results were found using \texttt{R} programming language~\citep{R}\footnote{The datasets and the \texttt{R} code used to produce the results and plots in this paper are available online at \url{https://github.com/jpneto/GenAlgPoly}.}.
-To compare this paper's proposed algorithm we applied the exact same train and test samples using several well-known learning algorithms for regression, namely: Linear Regression, Support Vector Machines~\citep{Meyer12}, Regression Trees~\citep{Therneau13}, Conditional Inference Trees~\citep{Hothorn06, Strobl07, Strobl08} and Random Forests~\citep{Liaw02}
+To compare this paper's proposed algorithm we applied the exact same train and test samples using several well-known learning algorithms for regression, namely: the classic \ac{GAPoly}, Linear Regression, Support Vector Machines~\citep{Meyer12}, Regression Trees~\citep{Therneau13} and Conditional Inference Trees~\citep{Hothorn06, Strobl07, Strobl08}.
 
-In order to train and test the performance of \ac{GAPoly} several mainstream datasets were used.
+In order to train and test the performance of \ac{GAPolyreg} several mainstream datasets were used.
 For each dataset, we selected 70\% for training purposes and the remaining observations to make a test set in order to compute the estimated \ac{rmse}.
@@
 \caption{Results for Artificial dataset.
-As shown, \ac{GAPoly} was able to find the exact polynomial structure that generates the dataset, reducing the test set prediction error to zero.
+As shown, \ac{GAPolyreg} was able to find the exact polynomial structure that generates the dataset, reducing the test set prediction error to zero.
 The regression methods depicted are: 1.
-\ac{GAPoly}, 2.
+\ac{GAPolyreg}, 2.
 Random Forests, 3.

### 2013-12-13 texts/submission_02-asc/paper_spfga-review.tex
Updated figures.

@@
 \caption{Results for Artificial dataset.
-As shown, \ac{GAPolyreg} was able to find the exact polynomial structure that generates the dataset, reducing the test set prediction error to zero.
+As shown, polynomial regression finds the exact polynomial structure that generates the dataset, reducing the test set prediction error to zero.
 The regression methods depicted are: 1.
 \ac{GAPolyreg}, 2.
-Random Forests, 3.
-Linear Regression, 4.
-SVM, 5.
-Regression Trees and 6.
+Linear Regression, 3.
+SVM, 4.
+Regression Trees and 5.
 Conditional Inference Trees}
@@
 The regression methods depicted in these figures are: 1.
-\ac{GAPoly}, 2.
-Random Forests, 3.
+\ac{GAPolyreg}, 2.
+\ac{GAPoly}, 3.
 Linear Regression, 4.
@@
 There are $n=506$ observations.
-
-Just as an example of the model the \ac{GAPoly} algorithm outputs: in this dataset the best polynomial,
-
-$$y = -0.12 x_6 x_9^4 + 0.78 x_6 - 0.4 x_9^2 x_{13} + 0.17 x_{13}^2 - 0.044$$
-
-where the attributes mean:
-$x_6$, RM average number of rooms per dwelling;
-$x_9$, RAD index of accessibility to radial highways;
-$x_{13}$, Lower status of the population.
-
-For comparison, if we access the mean decrease in accuracy found by Random Forests, the most important attributes are --- in decreasing order --- $x_{13}, x_6, x_5$ (but $x_5$ is already considered $4$ times less important than $x_6$).
-Both algorithms agree in two of their three most important attributes.
 

### 2013-12-13 code/exploratory/gap.py

@@ -15,12 +15,13 @@ from sklearn import cross_validation as cv
 from sklearn import svm
 from sklearn import preprocessing
 from sklearn import datasets
+from sklearn import linear_model
 from deap import algorithms, base, creator, tools
 #
 #   Supress deprecated warnings
 #
 import warnings
-warnings.filterwarnings("ignore")
+warnings.filterwarnings('ignore')
 #       
 def save_data(filename, data):
     f = open(filename, 'w')
@@ -43,7 +44,7 @@ def load_data(root_folder = '../../datasets/'):
     
     return data
     
-class PolyTransform:
+class PolyTerms:
     '''Represent a set of monomial transformations.
     
     It is assumed that a set of n variables, X1 .. Xn, is fixed.
@@ -60,7 +61,7 @@ class PolyTransform:
         1   1
         2   2
         
-        and the PolyTransform [Z1 = [1,0], Z2 = [2,2], Z3 = [2,0]] we obtain
+        and the PolyTerms [Z1 = [1,0], Z2 = [2,2], Z3 = [2,0]] we obtain
         Z1  Z2  Z3
         0   0   0
         1   1   1
@@ -78,7 +79,7 @@ class PolyTransform:
             dtype (numpy type): the numpy.array type to hold the degrees
             
         Returns:
-            an instance of PolyTransform
+            an instance of PolyTerms
             
         Defines:
             .num_terms (int): the number of monomial terms given in degrees
@@ -116,74 +117,96 @@ class PolyTransform:
         '''
         return self.terms.__repr__()
 
-class AbstractScoreProcess:
-    def __init__(self, dataset):
-        self.dataset = dataset
-        self.score = None
-        self._changed_ = False
+class AbstractPolyScorer:
+    def __init__(self):
+        self.poly = None
+        self.scores = None
         
-    def _compute_score_(self):
-        self.score = cv.cross_val_score(estimator, X_train, y_train, cv = 5)
+    def compute_score(self):
+        return 0.0
         
-    def score(self):
-        if self._changed_:
-            self._compute_score_()
-            self._changed_ = False
-            
-        return score
+    def score(self, poly):
+        self.poly = poly
+        self.scores = None
[…]

### 2013-12-13 tests_Housing.Rmd
added datapoints with 'classic' EPR
ie, no regularization and all monoids are active. This is the closest
version we are able to in accordance to the algorithms from the
references.
added 'load' commands to tests_*.Rmd

--- 2013-11-22 tests_Housing.Rmd
@@
 
-save(report, file="Housing300_lambda0.8.Rda") # recover with load("Housing300.Rda")
+save(report, file="Housing300_lambda0.8_add_linears.Rda") # recover with load("Housing300.Rda")
+
+load(file="Housing300_lambda0.8_add_linears.Rda")
+
+load(file="Housing300_lambda1_allmonoids.Rda")
+
+summary(report$ga.error)
 
@@
 
-write.table(df, "Housing300_lambda0.8.txt")
+write.table(df, "Housing300_lambda0.8_add_linears.txt")
 

### 2013-12-13 tests_Kinematics.Rmd
added datapoints with 'classic' EPR
ie, no regularization and all monoids are active. This is the closest
version we are able to in accordance to the algorithms from the
references.
add new data points using linear add-ons
added 'load' commands to tests_*.Rmd

--- 2013-07-15 tests_Kinematics.Rmd
@@
 
+load(file="Kinematics300_lambda0.8_add_linears.Rda")
+load(file="Kinematics300_lambda1_allmonoids.Rda")
+summary(report$ga.error)
+
 # See Results
 
-save(report, file="Kinematics300_lambda0.8_25runs.Rda") # recover with load("Kinematics.Rda")
+save(report, file="Kinematics300_lambda0.8_add_linears.Rda") # recover with load("Kinematics.Rda")
 
@@
 
-write.table(df, "Kinematics200.txt")
+write.table(df, "Kinematics300_lambda0.8_add_linears.txt")
 

### 2013-12-17 tests_Auto-Mpg.Rmd
added 'load' commands to tests_*.Rmd
testing tuning Auto-MPG

@@
 report <- make.report(my.data, population=300, iterations=100, lambda=0.8, runs=25)
+
+load(file="Auto-Mpg300_lambda0.8_add_linears.Rda")
+load(file="Auto-Mpg300_lambda1_allmonoids.Rda")
+summary(report$ga.error)
 
@@
 ########################################################
+# Tuning SVM & RPART
+
+rsme <- function(v1, v2=c(0)) {
+  sqrt(mean((v1-v2)^2))
+}
+
+library(e1071)
+
+head(my.data,8)
+
+n.runs <- 25
+train.p.size <- 0.7 # percentage of training set
+
+svm.error <- rep(NA,n.runs)
+rpart.error <- rep(NA,runs)
+
+for (i in 1:n.runs) {
+  inTrain   <- sample(1:nrow(my.data), train.p.size * nrow(my.data))
+  train.set <- my.data[inTrain,]
+  test.set  <- my.data[-inTrain,]
+
+  svm.tune <- tune.svm(y ~ ., data=train.set)
+  svm.pred  <- predict(svm.tune$best.model, test.set[,-ncol(test.set)]) 
+  svm.error[i] <- rsme(svm.pred,test.set[,ncol(test.set)])
+
+  rpart.tune <- tune.rpart(y ~ ., data=train.set)
+  rpart.pred  <- predict(rpart.tune$best.model, test.set[,-ncol(test.set)]) 
+  rpart.error[i] <- rsme(rpart.pred,test.set[,ncol(test.set)])
+}
+
+########################################################
 

### 2013-12-17 tests_Housing.Rmd
added tuned algs

@@
 
+########################################################
+# Tuning SVM & RPART
+
+rsme <- function(v1, v2=c(0)) {
+  sqrt(mean((v1-v2)^2))
+}
+
+library(e1071)
+
+head(my.data,8)
+
+n.runs <- 25
+train.p.size <- 0.7 # percentage of training set
+
+svm.error <- rep(NA,n.runs)
+rpart.error <- rep(NA,n.runs)
+
+for (i in 1:n.runs) {
+  inTrain   <- sample(1:nrow(my.data), train.p.size * nrow(my.data))
+  train.set <- my.data[inTrain,]
+  test.set  <- my.data[-inTrain,]
+
+  svm.tune <- tune.svm(y ~ ., data=train.set)
+  svm.pred  <- predict(svm.tune$best.model, test.set[,-ncol(test.set)]) 
+  svm.error[i] <- rsme(svm.pred,test.set[,ncol(test.set)])
+
+  rpart.tune <- tune.rpart(y ~ ., data=train.set)
+  rpart.pred  <- predict(rpart.tune$best.model, test.set[,-ncol(test.set)]) 
+  rpart.error[i] <- rsme(rpart.pred,test.set[,ncol(test.set)])
+
+  cat(i)
+}
+
+summary(svm.error)
+summary(rpart.error)
+
 ##############################################

### 2013-12-17 tests_Kinematics.Rmd
added tuned algs

@@
 ########################################################
+# Tuning SVM & RPART
+
+rsme <- function(v1, v2=c(0)) {
+  sqrt(mean((v1-v2)^2))
+}
+
+library(e1071)
+
+head(my.data,8)
+
+n.runs <- 25
+train.p.size <- 0.7 # percentage of training set
+
+svm.error <- rep(NA,n.runs)
+rpart.error <- rep(NA,n.runs)
+
+for (i in 1:n.runs) {
+  inTrain   <- sample(1:nrow(my.data), train.p.size * nrow(my.data))
+  train.set <- my.data[inTrain,]
+  test.set  <- my.data[-inTrain,]
+
+  svm.tune <- tune.svm(y ~ ., data=train.set)
+  svm.pred  <- predict(svm.tune$best.model, test.set[,-ncol(test.set)]) 
+  svm.error[i] <- rsme(svm.pred,test.set[,ncol(test.set)])
+
+  rpart.tune <- tune.rpart(y ~ ., data=train.set)
+  rpart.pred  <- predict(rpart.tune$best.model, test.set[,-ncol(test.set)]) 
+  rpart.error[i] <- rsme(rpart.pred,test.set[,ncol(test.set)])
+}
+
+summary(svm.error)
+summary(rpart.error)
+
+########################################################
 

### 2013-12-17 texts/submission_02-asc/paper_spfga-review.tex
added tuned algs

@@
 \acrodef{rmse}[error]{root-mean-square error}
-\acrodef{GAPoly}[\textsc{GAPoly}]{Genetic Algorithms for Polynomials}
-\acrodef{GAPolyreg}[\textsc{GAPoly.reg}]{Genetic Algorithms for Polynomials with Regularization}
+%\acrodef{GAPoly}[\textsc{GAPoly}]{Genetic Algorithms for Polynomials}
+%\acrodef{GAPolyreg}[\textsc{GAPoly.reg}]{Genetic Algorithms for Polynomials with Regularization}
+
+\acrodef{GAPoly}[EPR]{Evolutionary Polynomial Regression}
+\acrodef{GAPolyreg}[EPR$^2$]{Evolutionary Polynomial Regression with Regularisation}
+
 \acrodef{EPR}{Evolutionary Polynomial Regression}
 \acrodef{EPRR}[EPR$^2$]{Evolutionary Polynomial Regression with Regularisation}
+
+\acrodef{SVM}[SVM]{Support Vector Machines}
 %
@@
 \begin{abstract}
-While many applications require models that have no acceptable linear approximation the simpler nonlinear models are defined by polynomials.
+While many applications require models that have no acceptable linear approximation, the simpler nonlinear models are defined by polynomials.
 The use of genetic algorithms to find polynomial models from data is known as \ac{EPR}.
@@
 
-The empiric conclusion of those experiments is that \ac{EPRR} achieves better fitting than the other non-ensemble methods and shorter computation time than plain \ac{EPR}.
+The empiric conclusion of those experiments is that \ac{EPRR} is able to achieve better fitting than other non-ensemble methods and it has shorter computation time than plain \ac{EPR}.
 \end{abstract}
@@
 However nonlinearity is unavoidable in many application scenarios, specially those with phase transitions or feedback loops, so common in engineering, ecology, cybernetics and other areas.
-The kernel trick in \acp{SVM} (\cite{scholkopf1997kernel, liang2012eigen, Bao:2013aa}) alleviates this problem by allowing special non-linear transformations of the feature-space.
+The kernel trick in \ac{SVM} (\cite{scholkopf1997kernel, liang2012eigen, Bao:2013aa}) alleviates this problem by allowing special non-linear transformations of the feature-space.
 The condition such transformations must meet is known as the \emph{kernel trick}, $k\at{x,x'} = \left< \varphi\at{x}, \varphi\at{x'} \right>$, where $\varphi$ is the feature-space transformation and $\left<\cdot,\cdot\right>$ denotes inner product.
@@
 The optimal regression results from a cost function that accounts for both the \ac{rmse} and a regularization factor to avoid over-fitting.
-Another important change is in the coding mechanism which corresponds to two new features: (a) inclusion of the linear terms in the polynomial expression, and (b) the activation/deactivation of monoids during evolution to allow the increase/decrease of polynomial expressions.
+Another important change is in the coding mechanism which introduces two new features: (a) inclusion of the linear terms in the polynomial expression, and (b) the activation/deactivation of monoids during evolution to allow the increase/decrease of polynomial expressions.
 
@@
 
-The next section describes the details of our method and is followed by a presentation of some performance results. The last section draws some conclusions and points future research tasks.
+The next section describes the method's details and is followed by a presentation of some performance results. The last section draws some conclusions and points future research tasks.
 
@@
 \begin{algorithmic}
-\Function{GAPoly}{$D,pop_0,\epsilon$}
+\Function{EPR}{$D,pop_0,\epsilon$}
 	\State $pop \gets pop_0;\: err \gets 1.0+\epsilon$
@@
 
-Let's see an example: consider polynomial $x_1^3 x_3 + x_3^7 + x_1 x_2$ with $k = 4, m = 3$ and $d=3$. One possible encoding would be:
+A short example: consider polynomial $x_1^3 x_3 + x_3^7 + x_1 x_2$ with $k = 4, m = 3$ and $d=3$.
+A possible encoding would be (for reading purposes the semicolons separate monomials, the commas separate variables):
 \begin{center}
@@
 
-(for reading purposes the semicolons separate monomials, the commas separate variables)
-
-The first three bits inform that the second and third monomials are active while the fourth is not (as said, the first monomial is always active).
-This last monomial does not enter neither in the polynomial regression nor in the \ac{rmse} (fitness) evaluation.
+The first three bits inform that the second and third monomials are active while the fourth is not (the first monomial is always active).
+This fourth monomial does not enter neither in the polynomial regression nor in the \ac{rmse} (fitness) evaluation.
 However, it acts as a kind of junk DNA, becoming active when, in a future mutation or crossover, the third bit of the entire sequence flips from~0 to~1.
@@
 
-In the experiments that led to this article, our tests concluded that the inclusion of the linear terms achieved a significant reduction of the error rates.
+In the experiments that led to this article, the tests showed that the inclusion of the linear terms achieved a significant reduction of error rates.
 So, considering the previous example, the final polynomial would be $x_1^3 x_3 + x_3^7 + x_1 x_2 + x_1 + x_2 + x_3$.
+The next section discusses this feature.
 
@@
 One way to balance this is to provide a regularization term into the \ac{rmse} function.
-Our proposal is to include a multiplicative factor proportional to the number of monomials.
+The article's proposal is to include a multiplicative factor proportional to the number of monomials.
 Thus, the error from equation \ref{eq:rmse} defines
@@
 
-Somewhat unexpectedly after some experiences it was found that lower values for $\lambda$ provide better, even if marginal, results.
-Figure~\ref{Abalone_dataset_lambdas} shows regularization results for the Abalone and Auto MPG datasets with ten runs for each $\lambda$.
-The following section  includes information about these datasets.
-
-%
 \begin{figure*}[tb]\begin{center}
@@
 
-The typical inflection point lies around $\lambda = 0.8$. The dataset results for applying the proposed regression method use the regularization parameter with this value.
+Figure~\ref{Abalone_dataset_lambdas} shows regularization results for the Abalone and Auto MPG datasets with ten runs for each $\lambda$ (the following section  includes information about these datasets).
+The typical inflection point lies around $\lambda = 0.8$.
+This somewhat paradoxical result -- since regularization tries to minimize the model's complexity in order to prevent overfitting -- is somewhat justified by the inclusion of the linear terms mentioned in the previous section.
+The linear terms `pushes' the model to linearity while a regularization term $\lambda < 1$ tries to add non-linearity terms.
+Within this tension, the algorithm is able to reduce the overall error, compared to the typical implementation \ac{EPR} found in the literature.
 
@@
 To compare this paper's proposed algorithm we applied the exact same train and test samples using several well-known learning algorithms for regression, namely: the classic \ac{GAPoly}, Linear Regression, Support Vector Machines~\citep{Meyer12}, Regression Trees~\citep{Therneau13} and Conditional Inference Trees~\citep{Hothorn06, Strobl07, Strobl08}.
+To achieve better error results, the \ac{SVM} and Regression Tree algorithms had their parameters tuned.
 
@@
 \end{center}\end{figure*}
+
+\begin{description}
+\item{\textsc{Artificial}} this is an artificial dataset with four numeric features, $x_1, \ldots x_4$, where $x_1,x_3$ are outcomes from Poisson random variables, and $x_2,x_4$ from Normal random variables.
+The dependent variable~$y$ is given by expression $x_2x_4^2 + x_1^2x_3 + 5$.
+The dataset includes $n=50$~observations.
+
+This dataset was used in order to verify if \ac{GAPoly} was able to find the polynomial relation, which the algorithm did (cf.
+figure~\ref{artificial_dataset1_lambda1.0}).
+The genetic algorithm run with a population of $n=100$~solutions with $50$~iterations for each run.
+
+In the next datasets, the population of the genetic algorithm had size~$n=300$ with 100~iterations for each run. \ac{GAPolyreg} was run with regularization parameter $lambda = 0.8$.
 
@@
 
-\begin{description}
-\item{\textsc{Artificial}} this is an artificial dataset with four numeric features, $x_1, \ldots x_4$, where $x_1,x_3$ are outcomes from Poisson random variables, and $x_2,x_4$ from Normal random variables.
-The dependent variable~$y$ is given by expression $x_2x_4^2 + x_1^2x_3 + 5$.
-The dataset includes $n=50$~observations.
-
-This dataset was used in order to verify if \ac{GAPoly} was able to find the polynomial relation, which the algorithm did (cf.
-figure~\ref{artificial_dataset1_lambda1.0}).
-The genetic algorithm run with a population of $n=100$~solutions with $50$~iterations for each run.
-
-In the next datasets, the population of the genetic algorithm had size~$n=250$ with 100~iterations for each run.
-
 \item{\textsc{Housing}}: This data set concerns the task of predicting housing values in areas of Boston.
@@
 Since the entire learning process takes some time, in the current R implementation, placing a limit between 50 to 100 generations already achieves good results, relative to higher iteration values.
-Figure~\ref{Abalone_fitnessProgress} shows a typical error evolution for the dataset Abalone given two different values for $\lambda$.
+Figure~\ref{Abalone_fitnessProgress} shows a typical error evolution for the dataset Abalone given a regularized version ($\lambda = 0.975$) and a non-regularized ($\lambda = 1.0$).
 
@@
 
-Of all the non-ensemble regression methods considered, the proposed method is the one that shows, with one exception, the lowest error.
+% Of all the regression methods considered, the proposed method is the one that shows, with one exception, the lowest error.
 Only Random Forest outperforms \ac{GAPoly} systematically (it also outperforms all the other regression algorithms).
@@
 
-For complexity considerations \ac{GAPoly} demands some processing time.
-On a current quad-core computer, processing the Kinematics dataset (with 8k~observations) takes approximately 5~minutes.
-The processing time can probably be speeded by one to two orders in magnitude if the process is implemented in a low level programming language like \texttt{C++}.
+Of the regression methods considered, \ac{SVM} achieved the best results in 3 out of 4 datasets.
+However, \ac{SVM} was tuned for each particular dataset while \ac{GAPolyreg} was executed with its parameters preset.
+Even so, \ac{GAPolyreg} had the best error in the Abalone dataset, and achieved competitive results in the Housing and Auto MPG datasets.
+
+Comparing \ac{GAPoly} and \ac{GAPolyreg} -- which is the main article's topic -- the regularized version achieved much better results at Abalone and especially Kinematics.
+The Housing improved errors when compared with \ac{GAPoly} in a difference in means, resulted in a 95\% HDI (Highest Density Interval) equal to [0.001, 0.119] which, while borderline, achieves statistical significance.
+Only in the Auto MOG dataset \ac{GAPoly} achieved better results, even if not that different from \ac{GAPolyreg}.
+
+For complexity considerations \ac{GAPoly} and \ac{GAPolyreg} demands some processing time.
+On a quad-core computer, processing the Kinematics dataset (with 8k~observations) takes approximately 5~minutes.
+The processing time can probably be speeded between one to two orders in magnitude if the process is implemented in a low level programming language like \texttt{C++}.
 However, speed optimization was not the focus of this article.
@@
 Other parameters like mutation chance or the amount of elitism could also be tested.
-However, these type of tests need a low-level, fast implementation of \ac{GAPoly}.
+However, these type of tests need a low-level, fast implementation of \ac{GAPoly} and were not considered.
 

### 2013-12-17 code/exploratory/gap.py
progresso em EPRE@python.

@@ -15,6 +15,7 @@ from sklearn import cross_validation as cv
 from sklearn import svm
 from sklearn import preprocessing
 from sklearn import datasets
+from sklearn import metrics
 from sklearn import linear_model
 from deap import algorithms, base, creator, tools
 #
@@ -67,7 +68,7 @@ class PolyTerms:
         1   1   1
         2   16  4
     '''
-    def __init__(self, degrees, dtype = np.int8):
+    def __init__(self, degrees, dtype = None):
         '''Defines a monomial transformation.
         
         Args:
@@ -87,8 +88,13 @@ class PolyTerms:
             
         No check is performed.
         '''
-        self.terms = np.array( degrees, dtype = dtype)
+        if dtype is None:
+            self.terms = np.array( degrees )
+        else:
+            self.terms = np.array( degrees, dtype = dtype)
         self.num_terms, self.num_vars = self.terms.shape
+        
+        self.coef_ = None
                 
     def __call__(self, X, extend = False):
         '''Evaluates the transformation on a given dataset, X.
@@ -116,7 +122,34 @@ class PolyTerms:
                 a string with the exponents matrix.
         '''
         return self.terms.__repr__()
+        
+def mutate_polyterm(p, num_mutations = 1, valid_degrees = [0,1,2]):
+    z = np.copy(p.terms)
+    rows, cols = z.shape
+    for _ in range(num_mutations):
+        r = np.random.randint(rows)
+        c = np.random.randint(cols)
+        z[r,c] = np.random.choice(a = valid_degrees)
+    return PolyTerms(z)
+        
 
+    
+def random_matrix(num_terms = 2, num_vars = 2, valid_degrees = [0,1,2]):
+    z = np.zeros( (num_terms, num_vars) )
+    for r in range(num_terms):
+        for c in range(num_vars):
+            z[r,c] = np.random.choice(a = valid_degrees)
+    
+    return z
+    
+def random_polyterm(num_terms = 2, num_vars = 2, valid_degrees = [0,1,2]):        
+    return PolyTerms(random_matrix(num_terms, num_vars, valid_degrees))
+        
+class RandomPolyTerms(PolyTerms):
+    def __init__(self,num_terms, num_vars, valid_degrees):
+        PolyTerms.__init__(self, np.zeros( (num_terms, num_vars)))
+        self.terms = random_polyterm(num_terms, num_vars, valid_degrees).terms
+        
 class AbstractPolyScorer:
[…]

### 2013-12-19 texts/submission_02-asc/paper_spfga-review.tex
- corrected minor issues with bibliography; - further gap.py evolution towards the R cousin;

@@
 \acp{GA} were, arguably, one of the hottest topics of research in the recent decades and with good reason since they outline an optimisation scheme easy to conceptualise and with very broad application.
-If a nonlinear (or otherwise) model requires parameterization, \acp{GA} provide a simple and often effective approach to search for locally optimal parameters.
+If a nonlinear (or otherwise) model requires parameterisation, \acp{GA} provide a simple and often effective approach to search for locally optimal parameters.
 Research related to genetic algorithms abound and spans from the 1950s seminal work of Nils Aall Barricelli \citep{Barricelli:1962aa} in the Institute for Advanced Study of Princeton to today's principal area of study for thousands of researchers, covered in hundreds of conferences, workshops and other meetings.
@@
 The idea of using \acp{GA} to find a polynomial regression is not new \citep{Maertens:2006aa, Yu:2008aa, Wu:2009aa} but still generates original research \citep{Hofwing:2011aa,Cetisli:2011aa}.
-The modern formulation of the use of \ac{GA} to find polynomial models is known as \acf{EPR} and systematization can be traced back to the work of Davidson, Savic and Walters (\cite{Davidson:2003aa}).
-Further developments include multi-objective optimizations (\cite{Giustolisi:2009aa}).
+The modern formulation of the use of \ac{GA} to find polynomial models is known as \acf{EPR} and systematisation can be traced back to the work of Davidson, Savic and Walters (\cite{Davidson:2003aa}).
+Further developments include multi-objective optimisations (\cite{Giustolisi:2009aa}).
 
 This paper describes an extension of the general \ac{EPR} method to find a \emph{regularized} polynomial regression of a given dataset.
-The optimal regression results from a cost function that accounts for both the \ac{rmse} and a regularization factor to avoid over-fitting.
+The optimal regression results from a cost function that accounts for both the \ac{rmse} and a regularisation factor to avoid over-fitting.
 Another important change is in the coding mechanism which introduces two new features: (a) inclusion of the linear terms in the polynomial expression, and (b) the activation/deactivation of monoids during evolution to allow the increase/decrease of polynomial expressions.
@@
 This section is dedicated to the description of an algorithm to find a polynomial regression from a given dataset.
-It starts with a brief introduction and outline of the algorithm and proceeds into core details as the encoding used to represent individual polynomial instances in the \ac{GA} populations and the regularization of the cost function.
+It starts with a brief introduction and outline of the algorithm and proceeds into core details as the encoding used to represent individual polynomial instances in the \ac{GA} populations and the regularisation of the cost function.
 
@@
 %
-Subsection \ref{subs:polynomial.encoding} describes the encoding of individual polynomial instances as chromosomes and other parameters of the utilized \ac{GA} implementation.
-The regularization of the cost function is discussed in subsection \ref{subs:cost.function}.
+Subsection \ref{subs:polynomial.encoding} describes the encoding of individual polynomial instances as chromosomes and other parameters of the utilised \ac{GA} implementation.
+The regularisation of the cost function is discussed in subsection \ref{subs:cost.function}.
 
@@
 This \ac{rmse} function tends to prefer more complex polynomials, namely in the number of monomials which provides the regression algorithm for more fitting possibilities.
-One way to balance this is to provide a regularization term into the \ac{rmse} function.
+One way to balance this is to provide a regularisation term into the \ac{rmse} function.
 The article's proposal is to include a multiplicative factor proportional to the number of monomials.
@@
 
-\caption{Error distribution by regularization exponent for the Abalone and Auto MPG datasets.
-The box plots summarize the error values of 10 runs for each value of $\lambda$.
-The smallest overall error is shown in gray and, in both cases, was achieved with $\lambda=0.8$.}
+\caption{Error distribution by regularisation exponent for the Abalone and Auto MPG datasets.
+The box plots summarise the error values of 10 runs for each value of $\lambda$.
+The smallest overall error is shown in grey and, in both cases, was achieved with $\lambda=0.8$.}
 \label{Abalone_dataset_lambdas}\label{Auto-MPG_dataset_lambdas}
@@
 
-Figure~\ref{Abalone_dataset_lambdas} shows regularization results for the Abalone and Auto MPG datasets with ten runs for each $\lambda$ (the following section  includes information about these datasets).
+Figure~\ref{Abalone_dataset_lambdas} shows regularisation results for the Abalone and Auto MPG datasets with ten runs for each $\lambda$ (the following section  includes information about these datasets).
 The typical inflection point lies around $\lambda = 0.8$.
-This somewhat paradoxical result -- since regularization tries to minimize the model's complexity in order to prevent overfitting -- is somewhat justified by the inclusion of the linear terms mentioned in the previous section.
-The linear terms `pushes' the model to linearity while a regularization term $\lambda < 1$ tries to add non-linearity terms.
+This somewhat paradoxical result -- since regularisation tries to minimize the model's complexity in order to prevent overfitting -- is somewhat justified by the inclusion of the linear terms mentioned in the previous section.
+The linear terms `pushes' the model to linearity while a regularisation term $\lambda < 1$ tries to add non-linearity terms.
 Within this tension, the algorithm is able to reduce the overall error, compared to the typical implementation \ac{EPR} found in the literature.
 
-We'll denote by \ac{GAPolyreg} the proposed extension of the EPR algorithm which includes the regularized cost function together with the new features in polynomial coding.
+We'll denote by \ac{GAPolyreg} the proposed extension of the EPR algorithm which includes the regularised cost function together with the new features in polynomial coding.
 
@@
 
-To perform the genetic algorithm it was used the R package genalg~\citep{willighage2012genalg}.
+To perform the genetic algorithm it was used the R package genalg~\citep{Willighagen:2012aa}.
 The operators were the standard ones: (a)~crossover, \emph{i.e.}, a pair of solutions from the previous generations are combined by splitting and mixing their respective representations, and (b)~mutation, changing the values of single bits; the mutation chance applied in the datasets was 5\%.
@@
 
-The results were found using \texttt{R} programming language~\citep{R}\footnote{The datasets and the \texttt{R} code used to produce the results and plots in this paper are available online at \url{https://github.com/jpneto/GenAlgPoly}.}.
-To compare this paper's proposed algorithm we applied the exact same train and test samples using several well-known learning algorithms for regression, namely: the classic \ac{GAPoly}, Linear Regression, Support Vector Machines~\citep{Meyer12}, Regression Trees~\citep{Therneau13} and Conditional Inference Trees~\citep{Hothorn06, Strobl07, Strobl08}.
+The results were found using \texttt{R} programming language~\citep{R-Core-Team:2013aa}\footnote{The datasets and the \texttt{R} code used to produce the results and plots in this paper are available online at \url{https://github.com/jpneto/GenAlgPoly}.}.
+To compare this paper's proposed algorithm we applied the exact same train and test samples using several well-known learning algorithms for regression, namely: the classic \ac{GAPoly}, Linear Regression, Support Vector Machines~\citep{Meyer:2012aa}, Regression Trees~\citep{Therneau:2013aa} and Conditional Inference Trees~\citep{Hothorn:2006aa,Strobl:2007aa,Strobl:2008aa}.
 To achieve better error results, the \ac{SVM} and Regression Tree algorithms had their parameters tuned.
@@
 
-In the next datasets, the population of the genetic algorithm had size~$n=300$ with 100~iterations for each run. \ac{GAPolyreg} was run with regularization parameter $lambda = 0.8$.
+In the next datasets, the population of the genetic algorithm had size~$n=300$ with 100~iterations for each run. \ac{GAPolyreg} was run with regularisation parameter $lambda = 0.8$.
 
@@
 Since the entire learning process takes some time, in the current R implementation, placing a limit between 50 to 100 generations already achieves good results, relative to higher iteration values.
-Figure~\ref{Abalone_fitnessProgress} shows a typical error evolution for the dataset Abalone given a regularized version ($\lambda = 0.975$) and a non-regularized ($\lambda = 1.0$).
+Figure~\ref{Abalone_fitnessProgress} shows a typical error evolution for the dataset Abalone given a regularised version ($\lambda = 0.975$) and a non-regularized ($\lambda = 1.0$).
 
@@
 \caption{Error progress for Abalone dataset during a single execution of the genetic algorithm.
-The figure shows the fitness evolution for two different regularization values.
+The figure shows the fitness evolution for two different regularisation values.
 The population for both consisted of $200$ polynomials.
-The error values seem to stabilize around iteration~250.}
+The error values seem to stabilise around iteration~250.}
 \label{Abalone_fitnessProgress}
@@
 The processing time can probably be speeded between one to two orders in magnitude if the process is implemented in a low level programming language like \texttt{C++}.
-However, speed optimization was not the focus of this article.
+However, speed optimisation was not the focus of this article.
 
 A cross-validation procedure can be implemented to refine the appropriate parameter values to achieve better errors.
-Namely, the regularization parameter, $\lambda$, can be tested with several values, instead of being fixed at $0.8$.
+Namely, the regularisation parameter, $\lambda$, can be tested with several values, instead of being fixed at $0.8$.
 Other parameters like mutation chance or the amount of elitism could also be tested.
@@
 \bibliographystyle{model2-names}
-\bibliography{fullbib}
+\bibliography{fullbib2}
 

### 2013-12-19 code/exploratory/gap.py

@@ -95,6 +95,7 @@ class PolyTerms:
         self.num_terms, self.num_vars = self.terms.shape
         
         self.coef_ = None
+        self.intercept_ = None
                 
     def __call__(self, X, extend = False):
         '''Evaluates the transformation on a given dataset, X.
@@ -114,15 +115,21 @@ class PolyTerms:
             return np.c_[X,Z]
         else:
             return Z
-            
+    def __termrepr__(self, t):
+        return "".join(["X_{%d}^{%d}"%(i+1,d) for (i,d) in enumerate(t) if not(d == 0)])
+        
     def __repr__(self):
         '''Computes a string representation of the instance.
             
             Returns:
                 a string with the exponents matrix.
         '''
-        return self.terms.__repr__()
-        
+        if self.coef_ is None:
+            return self.terms.__repr__()
+        else:
+            terms_str = [self.__termrepr__(t) for t in self.terms]
+            return "".join( ['%+5.2f%s'%p for p in zip(self.coef_, terms_str)]) + '%+5.2f'%self.intercept_
+                    
 def mutate_polyterm(p, num_mutations = 1, valid_degrees = [0,1,2]):
     z = np.copy(p.terms)
     rows, cols = z.shape
@@ -164,7 +171,7 @@ class AbstractPolyScorer:
         return self.compute_score()
         
 def test_loss(x,y):
-    return -np.dot(x - y, x - y)
+    return np.dot(x - y, x - y)
     
 rms_loss = metrics.make_scorer( test_loss,
     greater_is_better = False,
@@ -186,9 +193,10 @@ class PolyRegression_Scorer(AbstractPolyScorer):
         
     def compute_score(self):
         Z = self.poly(self.X)
-        Z_train, Z_test, y_train, y_test = cv.train_test_split(Z,self.y)
+        Z_train, Z_test, y_train, y_test = cv.train_test_split(Z,self.y, train_size = 0.05)
         self.estimator.fit(Z_train, y_train)
         self.poly.coef_ = self.estimator.coef_
+        self.poly.intercept_ = self.estimator.intercept_
         self.scores = cv.cross_val_score(
             self.estimator,
             Z_test,
@@ -321,24 +329,32 @@ def test(tests = ['basic', 'regression', 'search']):
         from deap import tools
         #
         max_num_terms = 4
-        valid_degrees = [0, 1, 2, 3, 4]
+        valid_degrees = [0, 1, 2]
[…]

### 2013-12-19 texts/submission_02-asc/paper_spfga-review.tex
Text revision upto subsection 2.1

--- 2013-12-17 texts/submission_02-asc/paper_spfga-review.tex
@@
 \acrodef{GA}{Genetic algorithms}
-\acrodef{rmse}[error]{root-mean-square error}
-%\acrodef{GAPoly}[\textsc{GAPoly}]{Genetic Algorithms for Polynomials}
-%\acrodef{GAPolyreg}[\textsc{GAPoly.reg}]{Genetic Algorithms for Polynomials with Regularization}
+\acrodef{rmse}[error]{root-mean-square}
+%\acrodef{EPR}[\textsc{EPR}]{Genetic Algorithms for Polynomials}
+%\acrodef{EPRR}[\textsc{EPR.reg}]{Genetic Algorithms for Polynomials with Regularization}
 
-\acrodef{GAPoly}[EPR]{Evolutionary Polynomial Regression}
-\acrodef{GAPolyreg}[EPR$^2$]{Evolutionary Polynomial Regression with Regularisation}
+\acrodef{EPR}[EPR]{Evolutionary Polynomial Regression}
+\acrodef{EPRR}[EPR$^2$]{Evolutionary Polynomial Regression with Regularization}
 
 \acrodef{EPR}{Evolutionary Polynomial Regression}
-\acrodef{EPRR}[EPR$^2$]{Evolutionary Polynomial Regression with Regularisation}
+\acrodef{EPRR}[EPR$^2$]{Evolutionary Polynomial Regression with Regularization}
 
@@
 \newcommand{\at}[1]{\ensuremath{\!\left(#1\right)}}
-\newcommand{\note}[1]{\textbf{[~#1~]}}
+\newcommand{\note}[1]{\textbf{\small [Note\footnote{\textbf{Note:} #1}]}}
+\newcommand{\nth}{\ensuremath{^{th}}}
 %
@@
 
-\title{A Method for Regularisation of Evolutionary Polynomial Regressions}
+\title{A Method for Regularization of Evolutionary Polynomial Regressions}
 
@@
 %
-This paper introduces \ac{EPRR}, an algorithm that extends the \ac{EPR} method and describes a set of experiences on common datasets that compare both flavours of \ac{EPR} and other methods including Linear Regression, Regression Trees and Support Vector Regression.
+This paper introduces \ac{EPRR}, an algorithm that extends the \ac{EPR} method and describes a set of experiences on common datasets that compare both flavors of \ac{EPR} and other methods including Linear Regression, Regression Trees and Support Vector Regression.
 
@@
 \begin{keyword}
-evolutionary polynomial regression \sep regularisation \sep feature extraction \sep dimensionality reduction 
+evolutionary polynomial regression \sep regularization \sep feature extraction \sep dimensionality reduction 
 \end{keyword}
@@
 
-Polynomials, one of the most studied subjects in mathematics, generalise li\-ne\-ar functions and define, perhaps, the simplest and most used nonlinear models.
+Polynomials, one of the most studied subjects in mathematics, generalize li\-ne\-ar functions and define, perhaps, the simplest and most used nonlinear models.
 Applications include colorimetric calibration \citep{Mendes:2005aa}, explicit formul\ae\ for turbulent pipe flows
@@
 
-\acp{GA} were, arguably, one of the hottest topics of research in the recent decades and with good reason since they outline an optimisation scheme easy to conceptualise and with very broad application.
-If a nonlinear (or otherwise) model requires parameterization, \acp{GA} provide a simple and often effective approach to search for locally optimal parameters.
+\acp{GA} were, arguably, one of the hottest topics of research in the recent decades and with good reason since they outline an optimization scheme easy to conceptualize and with very broad application.
+If a nonlinear (or otherwize) model requires parameterization, \acp{GA} provide a simple and often effective approach to search for locally optimal parameters.
 Research related to genetic algorithms abound and spans from the 1950s seminal work of Nils Aall Barricelli \citep{Barricelli:1962aa} in the Institute for Advanced Study of Princeton to today's principal area of study for thousands of researchers, covered in hundreds of conferences, workshops and other meetings.
@@
 
-One interesting variation of genetic algorithms, named \emph{genetic programming} by John Koza \citep{Koza:1992aa}, proposed the use of \acp{GA} to search the syntactic structure of complex functions.
-This syntactic structure search is keen to the central ideas of deep learning \citep{Bengio:2009aa,Bengio:2013aa}, a subarea of machine learning actually producing quite promising results (\emph{e.g.} in \cite{Tarlow:2013fk}).
+One interesting variation of genetic algorithms, named \emph{genetic programming} by John Koza \citep{Koza:1992aa}, proposes the use of \acp{GA} to search the syntactic structure of complex functions.
+Syntactic structure search is also keen to the central ideas of deep learning \citep{Bengio:2009aa,Bengio:2013aa}, a subarea of machine learning actually producing quite promising results (\emph{e.g.} in \cite{Tarlow:2013fk}).
 It is also related to the work presented in this paper in the sense that, unlike linear models that have a simple structure, $y=\sum_i \beta_i x_i$, nonlinear (in particular polynomial) models pose an additional structure search problem.
@@
 
-This paper describes an extension of the general \ac{EPR} method to find a \emph{regularized} polynomial regression of a given dataset.
-The optimal regression results from a cost function that accounts for both the \ac{rmse} and a regularization factor to avoid over-fitting.
-Another important change is in the coding mechanism which introduces two new features: (a) inclusion of the linear terms in the polynomial expression, and (b) the activation/deactivation of monoids during evolution to allow the increase/decrease of polynomial expressions.
+This paper describes an extension of the general \ac{EPR} method to find a regularized polynomial regression of a given dataset.
+The optimal regression results from a cost function that accounts for both the \ac{rmse} and a regularization factor to avoid overfitting.
+\note{Explain a bit more the overfitting avoiding method.}
 
@@
 
-This section is dedicated to the description of an algorithm to find a polynomial regression from a given dataset.
-It starts with a brief introduction and outline of the algorithm and proceeds into core details as the encoding used to represent individual polynomial instances in the \ac{GA} populations and the regularization of the cost function.
+This section starts with a brief introduction and outline of the evolutionary polynomial regression algorithm, \ac{EPR}, and proceeds into core details as the encoding used to represent individual polynomial instances in the \ac{GA} populations and the regularization of the cost function.
 
-An usual representation of polynomials is
+An usual representation of polynomials is through expressions of the form
 $$
@@
 $$
-where each $q_i$ is a monomial, $q_i = \prod_{j} x_j^{\alpha_{ij}}$, the exponents are non-negative integers, $\alpha_{ij}\in\mathbb{N}_0$, and the coefficients are real valued, $\theta_i \in \mathbb{R}$.
-For example $p\left( x_1, x_2, x_3 \right) = 2 x_1 + x_2 x_3 + \frac{1}{2} x_1^2 x_3$ has monomials $q_1 = x_1, q_2 = x_2 x_3$ and $q_3 = x_1^2 x_3$, coefficients $\theta_1 = 2, \theta_2 = 1$ and $\theta_3 = 1/2$ and exponents $\alpha_{1,1} = 1, \alpha_{2,2}=1, \alpha_{2,3} = 1, \alpha_{3,1} = 2, \alpha_{3,3} = 1$ and all other $\alpha_{ij} = 0$.
-The exponents alone are a matrix that defines the monomial structure of the polynomial, $A = \left[\alpha_{ij}\right]$.
+where each $q_i = \prod_{j} x_j^{\alpha_{ij}}$ is a monomial, the exponents $\alpha_{ij}\in\mathbb{N}_0$ are non-negative integers and the coefficients $\theta_i \in \mathbb{R}$ are real valued.
+For example $p\left( x_1, x_2, x_3 \right) = 2 x_1 + x_2 x_3 + \frac{1}{2} x_1^2 x_3$ has monomials $q_1 = x_1, q_2 = x_2 x_3$ and $q_3 = x_1^2 x_3$, exponents $\alpha_{1,1} = 1, \alpha_{2,2}=1, \alpha_{2,3} = 1, \alpha_{3,1} = 2, \alpha_{3,3} = 1$ and all other $\alpha_{ij} = 0$ and coefficients $\theta_1 = 2, \theta_2 = 1$ and $\theta_3 = 1/2$.
 
-For the example above the matrix of monomials is
+The exponents alone can be organized into a matrix $\left[\alpha_{ij}\right]$ that defines the monomial structure of the polynomial. For the example above the matrix representation of the monomials is
 $$
-A = \left[ 
+\left[ 
 \begin{array}{ccc}
@@
 This representation of polynomials makes the problem of structure search very clear: except for the trivial cases, the number of possible monomials given $n$ variables and a maximum joint degree $d$ grows exponentially with either $n$ or $d$.
-But more importantly, the polynomial regression problem can be naturally split into two subproblems:
+But more importantly, by separating the set of monomials from the coefficients, the polynomial regression problem can be naturally split into two subproblems:
 %
 \begin{enumerate}
-\item For a given set of monomials $\mathcal{P} = \left\lbrace q_1, \ldots, q_k\right\rbrace$, find the regression coefficients $\theta_1,\ldots,\theta_k$ that minimize the \ac{rmse} on a given dataset;
+\item For a given set of monomials $\mathcal{Q} = \left\lbrace q_1, \ldots, q_k\right\rbrace$ find the regression coefficients $\Theta = \left\lbrace \theta_1,\ldots,\theta_k \right\rbrace$ that minimize the \ac{rmse} on a given dataset;
 
@@
 %
-More precisely, concerning the first problem, let $\mathcal{D}$ be a dataset with $n$ observations of variables $Y, X_1,\ldots,X_m$ and $\mathcal{P} = \left\lbrace q_1,\ldots, q_k\right\rbrace$ a set of $k$ monomial expressions over $X_1,\ldots,X_m$.
-Define the hypothesis\footnote{The expression $q|_{X=x}$ reads ``\emph{replace all instances of $X$ by $x$ in $q$}''.}
+More precisely, concerning the first problem, let $\mathcal{D}$ be a dataset with $n$ observations of variables $Y, X_1,\ldots,X_m$ and $\mathcal{Q} = \left\lbrace q_1,\ldots, q_k\right\rbrace$ a set of $k$ monomial expressions over $X_1,\ldots,X_m$.
+Define the hypothesis\footnote{The expression ``$q|_{X=x}$'' reads ``\emph{$q$ with all instances of $X$ replaced by $x$}.''}
 %
 \begin{eqnarray*}
-h_{\Theta,\mathcal{P}}\left(x_1,\ldots,x_m\right) &=& \sum_{j = 1}^k \theta_j q_j|_{X_i=x_i,\forall 1 \leq i \leq m}
+h_{\Theta,\mathcal{Q}}\left(x_1,\ldots,x_m\right) &=& \sum_{j = 1}^k \theta_j q_j|_{X_i=x_i,\forall 1 \leq i \leq m}
 \end{eqnarray*}
 %
-and let the cost
+and let the error (as ``cost'') be
 %
 \begin{eqnarray}
-J_{\textrm{fit}}\left(\Theta;\mathcal{P},\mathcal{D}\right)  & =  \nonumber \\
-\sqrt{\frac{1}{n}\sum_{i=1}^n \left( y\sample{i} - h_{\Theta,\mathcal{P}}\left( x_1\sample{i},\ldots,x_m\sample{i} \right) \right)^2 }\label{eq:rmse}
+J_{\textrm{fit}}\left(\Theta;\mathcal{Q},\mathcal{D}\right)  & =  \nonumber \\
+\sqrt{\frac{1}{n}\sum_{i=1}^n \left( y\sample{i} - h_{\Theta,\mathcal{Q}}\left( x_1\sample{i},\ldots,x_m\sample{i} \right) \right)^2 }\label{eq:rmse}
 \end{eqnarray}
 %
-be the usual \acf{rmse} function. Now the first problem can be stated as:
+the usual \acf{rmse} function. Now the first problem can be stated as:
 %
-\emph{Given a dataset $\mathcal{D}$ and a set of monomials $\mathcal{P}$, find parameters $\Theta$ that minimize $J_{\textrm{fit}}\left(\Theta;\mathcal{P},\mathcal{D}\right)$.}
+\emph{Given a dataset $\mathcal{D}$ and a set of monomials $\mathcal{Q}$ find parameters $\Theta$ that minimize the cost $J_{\textrm{fit}}\left(\Theta;\mathcal{Q},\mathcal{D}\right)$.}
 
 %
-It turns out that this problem can be solved as a usual linear regression problem by expanding $\mathcal{D}$ with columns that replicate the monomials in $\mathcal{M}$.
-
-The second problem is treated in the \ac{GA} setting: Let $\mathcal{D}$ be a dataset as above and $Q$ a set of polynomials.
-For each polynomial $p\in Q$ let $\mathcal{P}_p$ be the set of monomials in $p$ (without the coefficients) and compute the fitness
+This is a simple linear regression problem obtained by expanding $\mathcal{D}$ with columns that replicate the monomials in $\mathcal{Q}$.
+The second problem is treated in the \ac{GA} setting: Let $\mathcal{D}$ be a dataset as above and $\mathcal{P}$ a set of polynomials.
+For each polynomial $p\in \mathcal{P}$ let $\mathcal{Q}_p$ be the set of monomials in $p$ (without the coefficients) and define the (anti) fitness
 %
 \begin{eqnarray*}
-\phi_p &=& \min_\Theta J_{\textrm{fit}}\left(\Theta;\mathcal{P}_p,\mathcal{D}\right)
+\phi_p &=& \min_\Theta J_{\textrm{fit}}\left(\Theta;\mathcal{Q}_p,\mathcal{D}\right)
 \end{eqnarray*}
@@
 by solving the first problem.
-With a fitness of every instance, a \ac{GA} will apply genetic operators (usually mutation and crossover) to evolve the population $Q$ until a reasonable approximation of a local minimum is found.
+With a fitness of every instance, the \ac{GA} genetic operators (usually mutation and crossover) evolve the population $\mathcal{P}$ until a reasonable approximation of a local minimum is found.
 %
-\begin{algorithm}[tb]
+\begin{algorithm}[t]
 \begin{algorithmic}
-\Function{EPR}{$D,pop_0,\epsilon$}
+\Function{EPR}{$D,pop_0,\epsilon, maxiter$}
 	\State $pop \gets pop_0;\: err \gets 1.0+\epsilon$
-	\While{$err > \epsilon \land \neg\Call{limit}{iterations}$}
+	\While{$err > \epsilon \land iterations < maxiter$}
 		\State $pop \gets \Call{IterateGA}{pop}$
@@
 \end{algorithmic}
-\caption{\ac{GAPoly} uses linear regression to find monomial coefficients that minimize the \ac{rmse} over a dataset and \acp{GA} to explore the space of polynomials.
-The role of the linear regression step is to produce a fitting error that sorts the population.
+\caption{This \ac{EPR} algorithm uses linear regression in the calculation of the \ac{rmse} $J$ and the space of polynomials is searched by the \acp{GA} iteration step.
 At exit the \ac{rmse} of the fittest instance is bounded by $\epsilon$ or the maximum number of allowed iterations.}\label{alg:gapoly}
@@
 %
-Notice that the properties of \acp{GA} and linear regression entail that
-the composition of \acp{GA} with linear regression, as defined in Algorithm \ref{alg:gapoly}, converges to a polynomial that is a local minimum of the fitness function, encapsulated in the \ac{rmse} function $J_{\textrm{fit}}$.
+Notice that the properties of \acp{GA} and linear regression entail that Algorithm \ref{alg:gapoly} converges to a polynomial that is a local minimum of the fitness function, encapsulated in the \ac{rmse} function $J_{\textrm{fit}}$.
 %
@@
 %
-Subsection \ref{subs:polynomial.encoding} describes the encoding of individual polynomial instances as chromosomes and other parameters of the utilized \ac{GA} implementation.
+Subsection \ref{subs:polynomial.encoding} describes the encoding of individual polynomial instances as chromosomes and other parameters used in the \ac{GA} implementation.
 The regularization of the cost function is discussed in subsection \ref{subs:cost.function}.
@@
 %
-The encoding for any polynomial will be as follows:
+%\begin{quotation}
+%\note{Next paragraph moved here from the introduction.}
+%Another important change is in the coding mechanism which introduces two new features: (a) inclusion of the linear terms in the polynomial expression, and (b) the activation/deactivation of monomials during evolution to allow the increase/decrease of polynomial expressions.
+%\end{quotation}
+%
+The specific encoding (representation) of a set of monomials is an important aspect in the implementation of \ac{EPR}.
+The choice described below permits active and inactive monomials for regression purposes.
+The active (or inactive) state of a monomial might change through mutation or crossover.
+This simple mechanism enhances variation in the complexity of polynomial expressions by evolutionary operations.
+
+Let  $\left\lbrace q_1, \ldots, q_k\right\rbrace$ be a set of monomials over the variables $X_1, \ldots, X_m$. The encoding of that set using $d$ bits per exponent is a binary list such that
 \begin{enumerate}
-\item an initial segment detailing which monomials are active (the first monomial is always active), this is represented in unary description, i.e., each monomial is identified by a single bit
-\item the remaining bits are split into $k$ sets of equal size, each one representing a monomial
-\item each monomial is split into $m$ sets of $d$ size each, \emph{i.e.}, a variable
-\item for each variable, the remaining bits are the binary description of the variable degree, \emph{i.e.}, the maximum exponent is given by $2^d-1$
+\item the initial segment of $k$ bits defines the active state of each monomial;
+\item the remaining bits are split into $k$ segments of size $m\times d$, each representing a monomial;
+\item the bits in each monomial segment are split into $m$ sub-segments of size $d$.
+The $j\nth$ sub-segment is the binary representation of the degree of the variable $X_j$ in the enclosing monomial segment;
 \end{enumerate}
+This encoding can also be viewed as the flattening of the binary exponents in the matrix representation prefixed by the activation segment.
+The set $\left\lbrace x_1^3 x_3, x_3^7, x_1 x_2\right\rbrace$ (where $m = 3$) has matrix representation
+$$\left[ 
+\begin{array}{ccc}
+3 & 0 & 1 \\
+0 & 0 & 7 \\
+1 & 1 & 0
+\end{array}
+\right] = \left[ 
+\begin{array}{ccc}
+011 & 000 & 001 \\
+000 & 000 & 111 \\
+001 & 001 & 000
+\end{array}
+\right]_{\left( 2 \right)}
+$$
+where the right matrix is in binary form with  $d = 3$ bits. An encoding where the extra monomial $x_1^6x_2^2x_3^5$ is inactive, setting $k = 4$, would be
+\begin{center}
+1110;~011, 000, 001;~000,000,111;~001,001,000;~110,010,101
+\end{center}
+where, for reading purposes, semicolons separate segments and commas separate variables. The first $k=4$ bits inform that the first, second and third monomials are active while the fourth is not.
 
-A short example: consider polynomial $x_1^3 x_3 + x_3^7 + x_1 x_2$ with $k = 4, m = 3$ and $d=3$.
-A possible encoding would be (for reading purposes the semicolons separate monomials, the commas separate variables):
-\begin{center}
-\texttt{110 - 011,000,001 ; 000,000,111 ; 001,001,000 ; 110,010,101}
-\end{center}
+While each valid encoding represents a set of monomials the map is not bijective: each set of monomials has multiple encodings, for example by changing $d$ or the order of monomial segments.
+However, considering the \ac{EPR} task, this is a minor issue and a bijective map would bring computational complexity and impact negatively the algorithm's performance.
 
-The first three bits inform that the second and third monomials are active while the fourth is not (the first monomial is always active).
-This fourth monomial does not enter neither in the polynomial regression nor in the \ac{rmse} (fitness) evaluation.
-However, it acts as a kind of junk DNA, becoming active when, in a future mutation or crossover, the third bit of the entire sequence flips from~0 to~1.
-
-% Let's interpret the first monomial description, \verb!011,000,001!.
-It is divided by three since $m=3$.
-The first triple \verb!011!
-is the binary description of the exponent of variable $x_1$ which is 3, so the first monomial includes $x_1^3$.
-The second triple, \verb!000!, means that $x_2$ is not part of the monomial.
-The third triple \verb!001!
-says that variable $x_3$ has exponent $1$.
-The first monomial is~$x_1^3 x_3$.
-
-Notice that all binary descriptions give rise to valid polynomials.
-However this is not a bijective mapping.
-For each polynomial there are multiple representations.
-For example, $x_1+x_2$ and $x_2+x_1$ have different representations.
-The authors considered that more complex mappings in order to force a one to one mapping would impact negatively in the algorithm's performance.
-
-In the experiments that led to this article, the tests showed that the inclusion of the linear terms achieved a significant reduction of error rates.
-So, considering the previous example, the final polynomial would be $x_1^3 x_3 + x_3^7 + x_1 x_2 + x_1 + x_2 + x_3$.
-The next section discusses this feature.
-
+There is one final remark concerning the encoding.
+As it is, the activation segment can become all zeros, representing the empty set of monomials.
+This situation can be avoided with a simple hack: Given an encoding, the first monomial is always considered active, thus restricting the syntactic form of encodings to binary strings starting with $1$.
+In practice, this means that the implementation of the encoding can omit the first bit.
 % If the coding consists of entirely zeros, by convention, it describes polynomial $x_1$.
@@
 
-This \ac{rmse} function tends to prefer more complex polynomials, namely in the number of monomials which provides the regression algorithm for more fitting possibilities.
-One way to balance this is to provide a regularization term into the \ac{rmse} function.
-The article's proposal is to include a multiplicative factor proportional to the number of monomials.
+A known problem of using a cost function based only in the dataset \ac{rmse} and with polynomial regressions in general is a tendency to overfit the training data.
+The excessive variance is can be reduced by a regularization of the \ac{rmse} function.
+Tnd the inclusion of a penalty factor proportional to the number of monomials is a possible regularization method.
 Thus, the error from equation \ref{eq:rmse} defines
 \begin{eqnarray}
-J_{\textrm{reg}}\left(\Theta;\mathcal{P},\mathcal{D}\right) &=& \lambda^{k} J_{\textrm{fit}}\left(\Theta;\mathcal{P},\mathcal{D}\right)\label{eq:rmse-reg}
+J_{\textrm{reg}}\left(\Theta;\mathcal{Q},\mathcal{D}\right) &=& \lambda^{k} J_{\textrm{fit}}\left(\Theta;\mathcal{Q},\mathcal{D}\right)\label{eq:rmse-reg}
 \end{eqnarray}
@@
 The box plots summarize the error values of 10 runs for each value of $\lambda$.
-The smallest overall error is shown in gray and, in both cases, was achieved with $\lambda=0.8$.}
+The smallest overall error is shown in grey and, in both cases, was achieved with $\lambda=0.8$.}
 \label{Abalone_dataset_lambdas}\label{Auto-MPG_dataset_lambdas}
@@
 
-We'll denote by \ac{GAPolyreg} the proposed extension of the EPR algorithm which includes the regularized cost function together with the new features in polynomial coding.
+We'll denote by \ac{EPRR} the proposed extension of the EPR algorithm which includes the regularized cost function together with the new features in polynomial coding.
 
@@
 
-The results were found using \texttt{R} programming language~\citep{R}\footnote{The datasets and the \texttt{R} code used to produce the results and plots in this paper are available online at \url{https://github.com/jpneto/GenAlgPoly}.}.
-To compare this paper's proposed algorithm we applied the exact same train and test samples using several well-known learning algorithms for regression, namely: the classic \ac{GAPoly}, Linear Regression, Support Vector Machines~\citep{Meyer12}, Regression Trees~\citep{Therneau13} and Conditional Inference Trees~\citep{Hothorn06, Strobl07, Strobl08}.
+\begin{quotation}
+\note{Moved from previous section.} In the experiments that led to this article, the tests showed that the inclusion of the linear terms achieved a significant reduction of error rates.
+So, considering the previous example, the final set of monomials would be $\left\lbrace x_1^3 x_3, x_3^7, x_1 x_2, x_1, x_2, x_3 \right\rbrace$.
+The next section discusses this feature.
+\end{quotation}
+
+The results were found using \texttt{R} programming language~\citep{R-Core-Team:2013aa}\footnote{The datasets and the \texttt{R} code used to produce the results and plots in this paper are available online at \url{https://github.com/jpneto/GenAlgPoly}.}.
+To compare this paper's proposed algorithm we applied the exact same train and test samples using several well-known learning algorithms for regression, namely: the classic \ac{EPR}, Linear Regression, Support Vector Machines~\citep{Meyer:2012aa}, Regression Trees~\citep{Therneau:2013aa} and Conditional Inference Trees~\citep{Hothorn:2006aa,Strobl:2007aa,Strobl:2008aa}.
 To achieve better error results, the \ac{SVM} and Regression Tree algorithms had their parameters tuned.
 
-In order to train and test the performance of \ac{GAPolyreg} several mainstream datasets were used.
+In order to train and test the performance of \ac{EPRR} several mainstream datasets were used.
 For each dataset, we selected 70\% for training purposes and the remaining observations to make a test set in order to compute the estimated \ac{rmse}.
@@
 \includegraphics[width=0.98\textwidth]{figure_2.pdf}
-
+\note{Plots of ``Regression Trees'' and ``Conditional Inference Trees'' look the same!}
 \caption{Results for Artificial dataset.
@@
 The regression methods depicted are: 1.
-\ac{GAPolyreg}, 2.
+\ac{EPRR}, 2.
 Linear Regression, 3.
@@
 \label{artificial_dataset1_lambda1.0}
-
 \end{center}\end{figure*}
@@
 
-This dataset was used in order to verify if \ac{GAPoly} was able to find the polynomial relation, which the algorithm did (cf.
+This dataset was used in order to verify if \ac{EPR} was able to find the polynomial relation, which the algorithm did (cf.
 figure~\ref{artificial_dataset1_lambda1.0}).
@@
 
-In the next datasets, the population of the genetic algorithm had size~$n=300$ with 100~iterations for each run. \ac{GAPolyreg} was run with regularization parameter $lambda = 0.8$.
+In the next datasets, the population of the genetic algorithm had size~$n=300$ with 100~iterations for each run. \ac{EPRR} was run with regularization parameter $lambda = 0.8$.
 
@@
 
-\caption{Summary results for different regression methods on the Housing, Abalone, Auto MPG and Kinematics datasets.
-Random Forests, an ensemble method (``2'' in the plots), has the better results.
-However \ac{GAPoly} achieves similar errors in two of the datasets.
-For the remaining two \ac{GAPoly} produces similar errors to the other classic regression methods.
+\caption{Summary results for different regression methods on the Housing, Abalone, Auto MPG and Kinematics.
+Although \ac{EPRR} not always achieves the smallest error, performance is on-par with more sophisticated methods.
 The regression methods depicted in these figures are: 1.
-\ac{GAPolyreg}, 2.
-\ac{GAPoly}, 3.
+\ac{EPRR}, 2.
+\ac{EPR}, 3.
 Linear Regression, 4.
@@
 % Of all the regression methods considered, the proposed method is the one that shows, with one exception, the lowest error.
-Only Random Forest outperforms \ac{GAPoly} systematically (it also outperforms all the other regression algorithms).
-The single non-ensemble exception --- besides the artificial dataset that uses a straightforward polynomial relation --- is the Auto MPG dataset where \ac{GAPoly} has comparable results.
+Only Random Forest outperforms \ac{EPR} systematically (it also outperforms all the other regression algorithms).
+The single non-ensemble exception --- besides the artificial dataset that uses a straightforward polynomial relation --- is the Auto MPG dataset where \ac{EPR} has comparable results.
 This is evidence that applying standard genetic operators for polynomial model searching is a viable tool for regression purposes.
@@
 Of the regression methods considered, \ac{SVM} achieved the best results in 3 out of 4 datasets.
-However, \ac{SVM} was tuned for each particular dataset while \ac{GAPolyreg} was executed with its parameters preset.
-Even so, \ac{GAPolyreg} had the best error in the Abalone dataset, and achieved competitive results in the Housing and Auto MPG datasets.
+However, \ac{SVM} was tuned for each particular dataset while \ac{EPRR} was executed with its parameters preset.
+Even so, \ac{EPRR} had the best error in the Abalone dataset, and achieved competitive results in the Housing and Auto MPG datasets.
 
-Comparing \ac{GAPoly} and \ac{GAPolyreg} -- which is the main article's topic -- the regularized version achieved much better results at Abalone and especially Kinematics.
-The Housing improved errors when compared with \ac{GAPoly} in a difference in means, resulted in a 95\% HDI (Highest Density Interval) equal to [0.001, 0.119] which, while borderline, achieves statistical significance.
-Only in the Auto MOG dataset \ac{GAPoly} achieved better results, even if not that different from \ac{GAPolyreg}.
+Comparing \ac{EPR} and \ac{EPRR} -- which is the main article's topic -- the regularized version achieved much better results at Abalone and especially Kinematics.
+The Housing improved errors when compared with \ac{EPR} in a difference in means, resulted in a 95\% HDI (Highest Density Interval) equal to [0.001, 0.119] which, while borderline, achieves statistical significance.
+Only in the Auto MOG dataset \ac{EPR} achieved better results, even if not that different from \ac{EPRR}.
 
-For complexity considerations \ac{GAPoly} and \ac{GAPolyreg} demands some processing time.
+For complexity considerations \ac{EPR} and \ac{EPRR} demands some processing time.
 On a quad-core computer, processing the Kinematics dataset (with 8k~observations) takes approximately 5~minutes.
@@
 Other parameters like mutation chance or the amount of elitism could also be tested.
-However, these type of tests need a low-level, fast implementation of \ac{GAPoly} and were not considered.
+However, these type of tests need a low-level, fast implementation of \ac{EPR} and were not considered.
 

### 2013-12-19 code/exploratory/gap.py
Testing another branch.

@@ -434,7 +434,7 @@ def test(tests = ['basic', 'regression', 'search']):
     #    
     #
     #
-    print('\nDone.')
+    print('\nAll Done.')
     #
     
 if __name__ == '__main__':

### 2013-12-19 texts/submission_02-asc/paper_spfga-review.tex
Text revision into subsection 2.2: regularization cost.

@@
 \acp{GA} were, arguably, one of the hottest topics of research in the recent decades and with good reason since they outline an optimization scheme easy to conceptualize and with very broad application.
-If a nonlinear (or otherwize) model requires parameterization, \acp{GA} provide a simple and often effective approach to search for locally optimal parameters.
+If a nonlinear (or otherwise) model requires parameterization, \acp{GA} provide a simple and often effective approach to search for locally optimal parameters.
 Research related to genetic algorithms abound and spans from the 1950s seminal work of Nils Aall Barricelli \citep{Barricelli:1962aa} in the Institute for Advanced Study of Princeton to today's principal area of study for thousands of researchers, covered in hundreds of conferences, workshops and other meetings.
@@
 This encoding can also be viewed as the flattening of the binary exponents in the matrix representation prefixed by the activation segment.
-The set $\left\lbrace x_1^3 x_3, x_3^7, x_1 x_2\right\rbrace$ (where $m = 3$) has matrix representation
+The set $\left\lbrace x_1^3 x_3, x_3^7, x_1 x_2\right\rbrace$ (then $m = 3$) has matrix representation
 $$\left[ 
@@
 $$
-where the right matrix is in binary form with  $d = 3$ bits. An encoding where the extra monomial $x_1^6x_2^2x_3^5$ is inactive, setting $k = 4$, would be
+where the right matrix is in binary form with  $d = 3$ bits. An encoding of this set of monomials with the extra monomial $x_1^6x_2^2x_3^5$ inactive, setting $k = 4$, would be
 \begin{center}
@@
 While each valid encoding represents a set of monomials the map is not bijective: each set of monomials has multiple encodings, for example by changing $d$ or the order of monomial segments.
-However, considering the \ac{EPR} task, this is a minor issue and a bijective map would bring computational complexity and impact negatively the algorithm's performance.
+However, considering the \ac{EPR} task, this is a minor issue and a bijective map would add computational complexity and negative impact to the algorithm's performance.
 
-There is one final remark concerning the encoding.
+There is one final remark concerning this encoding method.
 As it is, the activation segment can become all zeros, representing the empty set of monomials.
@@
 
-The polynomial regression \ac{rmse} considered so far is based on the ability to predict the test set after the polynomial regression has found the appropriate coefficients $\theta_i$ for each one of the monomials $q_i$.
-
-A known problem of using a cost function based only in the dataset \ac{rmse} and with polynomial regressions in general is a tendency to overfit the training data.
-The excessive variance is can be reduced by a regularization of the \ac{rmse} function.
-Tnd the inclusion of a penalty factor proportional to the number of monomials is a possible regularization method.
-Thus, the error from equation \ref{eq:rmse} defines
+The polynomial regression \ac{rmse} considered so far accounts for the ability to predict the transformed testset.
+A known problem of using a cost function based only in the dataset \ac{rmse} (and of polynomial regressions in general) is a tendency to overfit the training data.
+This excessive variance can be reduced by regularizing the \ac{rmse} function and one possible method is the inclusion of a penalty factor proportional to the number of monomials.
+Thus, the error from equation \ref{eq:rmse} defines the regularized version
 \begin{eqnarray}
@@
 
-\caption{Error distribution by regularization exponent for the Abalone and Auto MPG datasets.
-The box plots summarize the error values of 10 runs for each value of $\lambda$.
-The smallest overall error is shown in grey and, in both cases, was achieved with $\lambda=0.8$.}
+\caption{Error distribution by regularization exponent for two common datasets.
+The box plots summarize error values of ten simulations for each value of $\lambda$.
+The smallest overall error, in grey, is achieved in both datasets when $\lambda=0.8$.}
 \label{Abalone_dataset_lambdas}\label{Auto-MPG_dataset_lambdas}
@@
 
-The datasets used herein were selected from Luís Torgo's data  repository, \url{http://www.dcc.fc.up.pt/~ltorgo/Regression/DataSets.html}.
-Most come originally from UCI ML repository, \url{http://archive.ics.uci.edu/ml/}.
+Datasets used herein were selected from Luís Torgo's data repository, \url{http://www.dcc.fc.up.pt/~ltorgo/Regression/DataSets.html}.
+Most can also be found in the UCI ML repository at \url{http://archive.ics.uci.edu/ml/}.
 
@@
 
-%\section{Bibliography}
+\section*{Bibliography}
 

### 2013-12-20 texts/submission_02-asc/paper_spfga-review.tex
Completed Text revision upto subsection 2.2.

@@
 
-This representation of polynomials makes the problem of structure search very clear: except for the trivial cases, the number of possible monomials given $n$ variables and a maximum joint degree $d$ grows exponentially with either $n$ or $d$.
+This partial representation of polynomials makes the problem of structure search very clear: except for the trivial cases, the number of possible monomials given $n$ variables and a maximum joint degree $d$ grows exponentially with either $n$ or $d$.
 But more importantly, by separating the set of monomials from the coefficients, the polynomial regression problem can be naturally split into two subproblems:
@@
 This is a simple linear regression problem obtained by expanding $\mathcal{D}$ with columns that replicate the monomials in $\mathcal{Q}$.
+The resulting dataset, $\mathcal{D} \cup \mathcal{Q}\left( \mathcal{D} \right)$, adds the monomial transformations in $\mathcal{Q}$ to the original dataset $\mathcal{D}$.
+An alternative formulation would just replace $\mathcal{D}$ by $\mathcal{Q}\left( \mathcal{D} \right)$.
+It turns out that the first formulation is a special case of the second (by including the variables in the monomial set) and has better \ac{rmse} performance --- what is not surprising because it uses more features.
+
 The second problem is treated in the \ac{GA} setting: Let $\mathcal{D}$ be a dataset as above and $\mathcal{P}$ a set of polynomials.
@@
 \end{algorithmic}
-\caption{This \ac{EPR} algorithm uses linear regression in the calculation of the \ac{rmse} $J$ and the space of polynomials is searched by the \acp{GA} iteration step.
+\caption{This \ac{EPR} algorithm uses linear regression for the calculation of the \ac{rmse} $J$ and the space of polynomials is searched in the \acp{GA} iteration step.
 At exit the \ac{rmse} of the fittest instance is bounded by $\epsilon$ or the maximum number of allowed iterations.}\label{alg:gapoly}
@@
 This encoding can also be viewed as the flattening of the binary exponents in the matrix representation prefixed by the activation segment.
-The set $\left\lbrace x_1^3 x_3, x_3^7, x_1 x_2\right\rbrace$ (then $m = 3$) has matrix representation
+The set $\left\lbrace x_1^3 x_3, x_3^7, x_1 x_2\right\rbrace$ (with $m = 3$) has matrix representation
 $$\left[ 
@@
 $$
-where the right matrix is in binary form with  $d = 3$ bits. An encoding of this set of monomials with the extra monomial $x_1^6x_2^2x_3^5$ inactive, setting $k = 4$, would be
+where the right matrix is in binary form using $d = 3$ bits. An encoding of this set of monomials with the extra monomial $x_1^6x_2^2x_3^5$ inactive, setting $k = 4$, would be
 \begin{center}
@@
 
-The polynomial regression \ac{rmse} considered so far accounts for the ability to predict the transformed testset.
-A known problem of using a cost function based only in the dataset \ac{rmse} (and of polynomial regressions in general) is a tendency to overfit the training data.
+The polynomial regression \ac{rmse} considered so far accounts for the ability to predict the transformed test-set.
+A known problem of using a cost function based only in the dataset \ac{rmse} (and of polynomial regressions in general) is the tendency to overfit  training data.
 This excessive variance can be reduced by regularizing the \ac{rmse} function and one possible method is the inclusion of a penalty factor proportional to the number of monomials.
-Thus, the error from equation \ref{eq:rmse} defines the regularized version
+In this line the error from equation \ref{eq:rmse} defines a regularized version
 \begin{eqnarray}
-J_{\textrm{reg}}\left(\Theta;\mathcal{Q},\mathcal{D}\right) &=& \lambda^{k} J_{\textrm{fit}}\left(\Theta;\mathcal{Q},\mathcal{D}\right)\label{eq:rmse-reg}
+J_{\textrm{reg}}\left(\Theta, \lambda;\mathcal{Q},\mathcal{D}\right) &=& \lambda^{k} J_{\textrm{fit}}\left(\Theta;\mathcal{Q},\mathcal{D}\right)\label{eq:rmse-reg}
 \end{eqnarray}
 %
-where $k$ is the number of monomials in the polynomial. When $\lambda > 1$ polynomials with more monomials are penalized.
+where $k$ is the number of monomials in the polynomial.
+When $\lambda > 1$ polynomials with more monomials are penalized.
+The regularized extension of \ac{EPR}, denoted by \ac{EPRR}, includes the regularized cost function.
 
@@
 The box plots summarize error values of ten simulations for each value of $\lambda$.
-The smallest overall error, in grey, is achieved in both datasets when $\lambda=0.8$.}
-\label{Abalone_dataset_lambdas}\label{Auto-MPG_dataset_lambdas}
+The smallest overall error, in grey, is achieved in both datasets when $\lambda=0.8$.
+Performance of the non-regularized \ac{EPR} is plotted in the line $\lambda = 1$.}
+\label{fig:lambda.error-distribution}\label{Abalone_dataset_lambdas}\label{Auto-MPG_dataset_lambdas}
 
@@
 
-Figure~\ref{Abalone_dataset_lambdas} shows regularization results for the Abalone and Auto MPG datasets with ten runs for each $\lambda$ (the following section  includes information about these datasets).
-The typical inflection point lies around $\lambda = 0.8$.
-This somewhat paradoxical result -- since regularization tries to minimize the model's complexity in order to prevent overfitting -- is somewhat justified by the inclusion of the linear terms mentioned in the previous section.
-The linear terms `pushes' the model to linearity while a regularization term $\lambda < 1$ tries to add non-linearity terms.
-Within this tension, the algorithm is able to reduce the overall error, compared to the typical implementation \ac{EPR} found in the literature.
-
-We'll denote by \ac{EPRR} the proposed extension of the EPR algorithm which includes the regularized cost function together with the new features in polynomial coding.
+A simple exploration on the effect of the regularization parameter is depicted in Figure~\ref{fig:lambda.error-distribution} where it is possible to observe that the typical inflection point lies around $\lambda = 0.8$.
+This value, favoring ``larger'' polynomials is justified by the balance of the data's non-linearity and polynomial complexity: below $\lambda = 0.8$, even penalized, larger monomial sets achieve better \ac{rmse} performance than smaller ones while above that value the size of the monomial set is excessive.
+Within this tension the overall error results reduced when compared to the non-regularized \ac{EPR} version.
 
@@
 \section{Experimental Results}
-
-\begin{quotation}
-\note{Moved from previous section.} In the experiments that led to this article, the tests showed that the inclusion of the linear terms achieved a significant reduction of error rates.
-So, considering the previous example, the final set of monomials would be $\left\lbrace x_1^3 x_3, x_3^7, x_1 x_2, x_1, x_2, x_3 \right\rbrace$.
-The next section discusses this feature.
-\end{quotation}
 

### 2013-12-20 texts/submission_02-asc/paper_spfga-review.tex
Completed text revision in subsection 2.3.

@@
 A known problem of using a cost function based only in the dataset \ac{rmse} (and of polynomial regressions in general) is the tendency to overfit  training data.
-This excessive variance can be reduced by regularizing the \ac{rmse} function and one possible method is the inclusion of a penalty factor proportional to the number of monomials.
-In this line the error from equation \ref{eq:rmse} defines a regularized version
+
+Excessive variance of the estimation method can be reduced by regularizing the \ac{rmse} function with a penalty factor.
+Thus, to reduce polynomial complexity and variance by regularizing the size of the monomial set the \ac{rmse} function from equation \ref{eq:rmse} is multiplied by a factor $\lambda^{k}$
 \begin{eqnarray}
@@
 %
-where $k$ is the number of monomials in the polynomial.
-When $\lambda > 1$ polynomials with more monomials are penalized.
-The regularized extension of \ac{EPR}, denoted by \ac{EPRR}, includes the regularized cost function.
+where $k$ is the number of monomials in the polynomial. When $\lambda > 1$ polynomials with more monomials are penalized. The regularized extension of \ac{EPR} is denoted by \acf{EPRR}.
 
@@
 
-\subsection{Genetic Operators}
+\subsection{Genetic Algorithm Parameterization}
 
-To perform the genetic algorithm it was used the R package genalg~\citep{Willighagen:2012aa}.
-The operators were the standard ones: (a)~crossover, \emph{i.e.}, a pair of solutions from the previous generations are combined by splitting and mixing their respective representations, and (b)~mutation, changing the values of single bits; the mutation chance applied in the datasets was 5\%.
-There was also elitism between generations, \emph{i.e.}, 20\% of the best solutions survive to the next generation.
+In general \acp{GA} offer many possibilities with respect to the choice of genetic operators and respective application rates, population evolution, \emph{etc}.
+The results found here where obtained using the package \texttt{genalg}~\citep{Willighagen:2012aa} with default parameters, standard operators (crossover and mutation) and population evolution with 20\% elitism between generations.
 
@@
 \caption{Results for Artificial dataset.
-As shown, polynomial regression finds the exact polynomial structure that generates the dataset, reducing the test set prediction error to zero.
+As shown, polynomial regression finds the exact polynomial structure that generates the dataset: in line 1, the error box is centered in $0$ and has width $0$.
 The regression methods depicted are: 1.
@@
 The Housing improved errors when compared with \ac{EPR} in a difference in means, resulted in a 95\% HDI (Highest Density Interval) equal to [0.001, 0.119] which, while borderline, achieves statistical significance.
-Only in the Auto MOG dataset \ac{EPR} achieved better results, even if not that different from \ac{EPRR}.
+Only in the Auto MPG dataset \ac{EPR} achieved better results, even if not that different from \ac{EPRR}.
 

### 2013-12-21 texts/submission_02-asc/paper_spfga-review.tex
Completed text revision in section 3.

@@
 \newcommand{\at}[1]{\ensuremath{\!\left(#1\right)}}
-\newcommand{\note}[1]{\textbf{\small [Note\footnote{\textbf{Note:} #1}]}}
+\newcommand{\note}[1]{\textbf{\small{[Note: {#1}]}}}
 \newcommand{\nth}{\ensuremath{^{th}}}
@@
  \citep{Davidson:1999aa}, computational linguistics \citep{Sanchez:2009aa} and more recently analytical techniques for cultural heritage materials \citep{Csefalvayova:2010aa}, liquid epoxy moulding process \citep{Chan:2011aa}, B-spline surface reconstruction \citep{Galvez:2012aa}, product design \citep{Chan:2012aa} or forecasting cyanotoxins presence in water reservoirs \citep{Garcia-Nieto:2013aa}.
-These examples not only illustrate the wide spectrum of applications but, additionally, in each one uses, at some point, a \ac{GA}.
+These examples not only illustrate the wide spectrum of applications but, additionally, each one uses, at some point, \ac{GA}.
 
-\acp{GA} were, arguably, one of the hottest topics of research in the recent decades and with good reason since they outline an optimization scheme easy to conceptualize and with very broad application.
+Evolutionary algorithms, including \ac{GA}, were, arguably, one of the hottest topics of research in the recent decades and with good reason since they outline an optimization scheme easy to conceptualize and with very broad application.
 If a nonlinear (or otherwise) model requires parameterization, \acp{GA} provide a simple and often effective approach to search for locally optimal parameters.
-Research related to genetic algorithms abound and spans from the 1950s seminal work of Nils Aall Barricelli \citep{Barricelli:1962aa} in the Institute for Advanced Study of Princeton to today's principal area of study for thousands of researchers, covered in hundreds of conferences, workshops and other meetings.
+Related research abound and spans from the 1950s seminal work of Nils Aall Barricelli \citep{Barricelli:1962aa} in the Institute for Advanced Study of Princeton to today's principal area of study for thousands of researchers, covered in hundreds of conferences, workshops and other meetings.
 Perhaps the key impulse to \acp{GA} come from John Holland's work and his book ``Adaptation in Natural and Artificial Systems'' \citep{Holland:1975aa}.
@@
 This paper describes an extension of the general \ac{EPR} method to find a regularized polynomial regression of a given dataset.
-The optimal regression results from a cost function that accounts for both the \ac{rmse} and a regularization factor to avoid overfitting.
-\note{Explain a bit more the overfitting avoiding method.}
+The optimal regression results from a cost function that accounts for both the \ac{rmse} and a regularization factor to avoid overfit by penalysing polynomial complexity.
 
@@
 %
-Notice that the properties of \acp{GA} and linear regression entail that Algorithm \ref{alg:gapoly} converges to a polynomial that is a local minimum of the fitness function, encapsulated in the \ac{rmse} function $J_{\textrm{fit}}$.
+The properties of \acp{GA} and linear regression entail that Algorithm \ref{alg:gapoly} converges to a polynomial that is a local minimum of the fitness function, encapsulated in the \ac{rmse} function $J_{\textrm{fit}}$.
 %
@@
 
-The polynomial regression \ac{rmse} considered so far accounts for the ability to predict the transformed test-set.
+The polynomial regression \ac{rmse} considered so far accounts for the ability to predict the transformed testset.
 A known problem of using a cost function based only in the dataset \ac{rmse} (and of polynomial regressions in general) is the tendency to overfit  training data.
-
 Excessive variance of the estimation method can be reduced by regularizing the \ac{rmse} function with a penalty factor.
@@
 
-The results were found using \texttt{R} programming language~\citep{R-Core-Team:2013aa}\footnote{The datasets and the \texttt{R} code used to produce the results and plots in this paper are available online at \url{https://github.com/jpneto/GenAlgPoly}.}.
-To compare this paper's proposed algorithm we applied the exact same train and test samples using several well-known learning algorithms for regression, namely: the classic \ac{EPR}, Linear Regression, Support Vector Machines~\citep{Meyer:2012aa}, Regression Trees~\citep{Therneau:2013aa} and Conditional Inference Trees~\citep{Hothorn:2006aa,Strobl:2007aa,Strobl:2008aa}.
-To achieve better error results, the \ac{SVM} and Regression Tree algorithms had their parameters tuned.
+Here is described the experiment setup used to gather and summarize the empirical evidence that supports this comparative study of \ac{EPR} and \ac{EPRR}.
+Evaluation is focused in \ac{rmse} distribution and, besides \ac{EPR} and \ac{EPRR}, also uses several common regression methods and datasets easily accessible in \texttt{R}, the free software environment for statistical computing and graphics~\citep{R-Core-Team:2013aa}\footnote{The datasets and \texttt{R} code used to produce the results and plots in this paper are available online at \url{https://github.com/jpneto/GenAlgPoly}.}.
+A small consideration on the convergence speed concludes this section.
 
-In order to train and test the performance of \ac{EPRR} several mainstream datasets were used.
-For each dataset, we selected 70\% for training purposes and the remaining observations to make a test set in order to compute the estimated \ac{rmse}.
-To achieve more robust results, each dataset were processed $25$~times, each one with different samples for the train and test sets.
-For the datasets with attribute values of different magnitudes, a preliminary scaling was executed.
-The results below are box plots for the test set error predictions over these different runs.
+\subsection{Regression Methods and Datasets}
+
+The \ac{EPRR} method is ranked against several well-known learning algorithms for regression, namely: non-regularized \ac{EPR}, Linear Regression, Support Vector Machines~\citep{Meyer:2012aa}, Regression Trees~\citep{Therneau:2013aa} and Conditional Inference Trees~\citep{Hothorn:2006aa,Strobl:2007aa,Strobl:2008aa}.
+To achieve better error results the \ac{SVM} and Regression Tree parameters are tuned in each dataset.
+
+The performance of each method is evaluated on several common datasets.
+From each dataset 70\% of the observations are reserved for training purposes and the remaining observations used to estimate the \ac{rmse}.
+To enhance the robustness of results this process is repeated $25$~times, each time with different shuffling of the samples in the train and test sets.
+Some datasets with attribute values of different magnitudes have a pre-processing scaling transformation.
+The box plots in figures \ref{artificial_dataset1_lambda1.0} and \ref{fig:four.datasets.summary} resume the test set error distributions over these different runs.
+
+One of the used datasets, \textsc{Artificial}, has a special role: it is used to test if \ac{EPRR} is able to discover a polynomial model.
+The idea of this test is to generate a polynomial dependent variable and measure the \ac{EPRR} \ac{rmse} after fitting the dataset.
+The genetic algorithm parameterization for this dataset uses a population with size $n=100$ and evolves for $50$~generations.
+For the remaining datasets the population has size~$n=300$ and evolves for $100$~generations.
 
@@
 \includegraphics[width=0.98\textwidth]{figure_2.pdf}
-\note{Plots of ``Regression Trees'' and ``Conditional Inference Trees'' look the same!}
-\caption{Results for Artificial dataset.
-As shown, polynomial regression finds the exact polynomial structure that generates the dataset: in line 1, the error box is centered in $0$ and has width $0$.
+\caption{Testing polynomial discovery.
+The dataset is generated from a polynomial expression and, as shown, \ac{EPRR} finds the exact generator structure: in line 1, the error box is centered in $0$ and has width $0$.
 The regression methods depicted are: 1.
@@
 Conditional Inference Trees}
+\note{Plots of ``Regression Trees'' and ``Conditional Inference Trees'' look the same!} 
 \label{artificial_dataset1_lambda1.0}
-\end{center}\end{figure*}
+\end{center}
+\end{figure*}
 
 \begin{description}
-\item{\textsc{Artificial}} this is an artificial dataset with four numeric features, $x_1, \ldots x_4$, where $x_1,x_3$ are outcomes from Poisson random variables, and $x_2,x_4$ from Normal random variables.
-The dependent variable~$y$ is given by expression $x_2x_4^2 + x_1^2x_3 + 5$.
-The dataset includes $n=50$~observations.
+\item{\textsc{Artificial}} is a polynomial dataset with four numeric features, $x_1, \ldots x_4$, where $x_1,x_3$ are outcomes from Poisson random variables, and $x_2,x_4$ from Normal random variables.
+The dependent variable is given by the polynomial expression $y = x_2x_4^2 + x_1^2x_3 + 5$.
+The dataset includes $n=50$~observations;
 
-This dataset was used in order to verify if \ac{EPR} was able to find the polynomial relation, which the algorithm did (cf.
-figure~\ref{artificial_dataset1_lambda1.0}).
-The genetic algorithm run with a population of $n=100$~solutions with $50$~iterations for each run.
+\item{\textsc{Housing}} concerns the task of predicting housing values in areas of Boston.
+There are $n=506$ observations of $m=13$ continuous attributes and one dependent variable, the median value of owner-occupied homes in thousands of USD;
 
-In the next datasets, the population of the genetic algorithm had size~$n=300$ with 100~iterations for each run. \ac{EPRR} was run with regularization parameter $lambda = 0.8$.
+\item{\textsc{Abalone}} is used to predict the age of a abalone shell using $m=8$ numeric attributes concerning several physical measurements. There are $n=4177$ observations;
 
+% best polynomial, rsme 0.65
+% $$y = 0.52 s_1^2 x_5 - 0.13 x_2 x_8 -0.75 x_1^2 x_6 + 0.78 x_8 + 0.18 $$
+
+\item{\textsc{Auto MPG}} gathers fuel consumption in miles per gallon, based on two discrete and five continuous attributes ($m=7$). There are $n=398$ observations;
+
+% best polynomial, \ac{rmse} 0.32
+%$$ y = 0.4 x_6 + 0.02 x_5^3 x_6 - 0.85 x_4 + 0.23 x_4^2 - 0.26 $$
+
+\item{\textsc{Kinematics}} results from a realistic simulation of the forward kinematics of an 8 link robot arm.
+The task is to predict the distance of the end-effector from a target using $m=8$ continuous attributes.
+There are $n=8192$ observations;
+
+% best polynomial, \ac{rmse} 0.74
+% $$ y = 0.24 x_5 - 0.23 x_5 x_6 x_7 - 0.19 x_4 x_7 - 0.5 x_3 $$
+
+\end{description}
+%
 \begin{figure*}[tb]\begin{center}
@@
 \includegraphics[width=0.48\textwidth]{figure_3a.pdf}
+&
+\includegraphics[width=0.48\textwidth]{figure_3b.pdf}
+\\
+\includegraphics[width=0.48\textwidth]{figure_3c.pdf}
+&
+\includegraphics[width=0.48\textwidth]{figure_3d.pdf}
+\end{tabular}
 %
-&
-%
-\includegraphics[width=0.48\textwidth]{figure_3b.pdf}
-%
-\\
-%
-\includegraphics[width=0.48\textwidth]{figure_3c.pdf}
-%
-&
-%
-\includegraphics[width=0.48\textwidth]{figure_3d.pdf}
-%
-\end{tabular}
-
-\caption{Summary results for different regression methods on the Housing, Abalone, Auto MPG and Kinematics.
-Although \ac{EPRR} not always achieves the smallest error, performance is on-par with more sophisticated methods.
+\caption{Summary results for different regression methods on diverse datasets.
+Although \ac{EPRR} not always achieves the smallest expected error, performance is on-par with more sophisticated methods.
 The regression methods depicted in these figures are: 1.
-\ac{EPRR}, 2.
-\ac{EPR}, 3.
-Linear Regression, 4.
-SVM, 5.
-Regression Trees and 6.
-Conditional Inference Trees}
+\ac{EPRR}; 2.
+\ac{EPR}; 3.
+Linear Regression; 4.
+\ac{SVM}; 5.
+Regression Trees; 6.
+Conditional Inference Trees;}
+\label{fig:four.datasets.summary}
 \label{Housing_dataset_lambda0.8_25runs}
@@
 
-\item{\textsc{Housing}}: This data set concerns the task of predicting housing values in areas of Boston.
-There are $m=13$ continuous attributes and the dependent variable is the median value of owner-occupied homes in \$1000's.
-There are $n=506$ observations.
-
-\item{\textsc{Abalone}} This dataset can be used to predict the age of a abalone shell using the given $m=8$ numeric attributes concerning several physical measurements.
-There are $n=4177$ observations.
-
-% best polynomial, rsme 0.65
-% $$y = 0.52 s_1^2 x_5 - 0.13 x_2 x_8 -0.75 x_1^2 x_6 + 0.78 x_8 + 0.18 $$
-
-\item{\textsc{Auto MPG}} This dataset is used to predict fuel consumption in miles per gallon, based on two discrete and five continuous attributes ($m=7$). There are $n=398$ observations.
-
-% best polynomial, \ac{rmse} 0.32
-%$$ y = 0.4 x_6 + 0.02 x_5^3 x_6 - 0.85 x_4 + 0.23 x_4^2 - 0.26 $$
-
-\item{\textsc{Kinematics}} This dataset is concerned with the realistic simulation of the forward kinematics of an 8 link robot arm.
-The task is to predict the distance of the end-effector from a target using $m=8$ continuous attributes.
-There are $n=8192$ observations.
-
-% best polynomial, \ac{rmse} 0.74
-% $$ y = 0.24 x_5 - 0.23 x_5 x_6 x_7 - 0.19 x_4 x_7 - 0.5 x_3 $$
-
-\end{description}
-
 \subsection{Convergence speed}
-
-The \ac{GA} quickly proceeds in the first~50 to~100 generations to reasonable error rates.
-Then, it proceeds slower achieving best solutions with marginal error reduction.
-Since the entire learning process takes some time, in the current R implementation, placing a limit between 50 to 100 generations already achieves good results, relative to higher iteration values.
-Figure~\ref{Abalone_fitnessProgress} shows a typical error evolution for the dataset Abalone given a regularized version ($\lambda = 0.975$) and a non-regularized ($\lambda = 1.0$).
-
+%
 \begin{figure*}[tb]
@@
 \includegraphics[width=0.98\textwidth]{figure_4.pdf}
-\caption{Error progress for Abalone dataset during a single execution of the genetic algorithm.
+\caption{Error progress for the \textsc{Abalone} dataset during a single execution of the genetic algorithm.
 The figure shows the fitness evolution for two different regularization values.
-The population for both consisted of $200$ polynomials.
+The population for both consists of $200$ polynomials.
 The error values seem to stabilize around iteration~250.}
 \label{Abalone_fitnessProgress}
-
 \end{center}
 \end{figure*}
+%
+Since this work is oriented to the \ac{rmse} of the \ac{EPRR} model it is necessary to assess how this depends on the number of generations of the \ac{GA}.
+As illustrated in Figure~\ref{Abalone_fitnessProgress}, the \ac{rmse} quickly drops during the initial $50$ to $100$ generations.
+Then, it proceeds slower achieving better solutions only with marginal error reduction.
+
+%Since the entire learning process takes some time, in the current \texttt{R} implementation, placing a limit between 50 to 100 generations already achieves good results, relative to higher iteration values.
+Figure~\ref{Abalone_fitnessProgress} shows a typical error evolution for the dataset \textsc{Abalone} given a regularized version ($\lambda = 0.975$) and a non-regularized ($\lambda = 1.0$).\note{Why not $\lambda=0.8$?}
 
@@
 However, \ac{SVM} was tuned for each particular dataset while \ac{EPRR} was executed with its parameters preset.
-Even so, \ac{EPRR} had the best error in the Abalone dataset, and achieved competitive results in the Housing and Auto MPG datasets.
+Even so, \ac{EPRR} had the best error in the \textsc{Abalone} dataset, and achieved competitive results in the \textsc{Housing} and \textsc{Auto MPG} datasets.
 
-Comparing \ac{EPR} and \ac{EPRR} -- which is the main article's topic -- the regularized version achieved much better results at Abalone and especially Kinematics.
-The Housing improved errors when compared with \ac{EPR} in a difference in means, resulted in a 95\% HDI (Highest Density Interval) equal to [0.001, 0.119] which, while borderline, achieves statistical significance.
-Only in the Auto MPG dataset \ac{EPR} achieved better results, even if not that different from \ac{EPRR}.
+Comparing \ac{EPR} and \ac{EPRR} -- which is the main article's topic -- the regularized version achieved much better results at \textsc{Abalone} and especially \textsc{Kinematics}.
+The \textsc{Housing} dataset improved errors when compared with \ac{EPR} in a difference in means, resulted in a 95\% HDI (Highest Density Interval) equal to [0.001, 0.119] which, while borderline, achieves statistical significance.
+Only in the \textsc{Auto MPG} dataset \ac{EPR} achieved better results, even if not that different from \ac{EPRR}.
 
 For complexity considerations \ac{EPR} and \ac{EPRR} demands some processing time.
-On a quad-core computer, processing the Kinematics dataset (with 8k~observations) takes approximately 5~minutes.
+On a quad-core computer, processing the \textsc{Kinematics} dataset (with 8k~observations) takes approximately 5~minutes.
 The processing time can probably be speeded between one to two orders in magnitude if the process is implemented in a low level programming language like \texttt{C++}.

### 2013-12-22 texts/submission_02-asc/paper_spfga-review.tex
Pre-submit version.

@@
 From each dataset 70\% of the observations are reserved for training purposes and the remaining observations used to estimate the \ac{rmse}.
-To enhance the robustness of results this process is repeated $25$~times, each time with different shuffling of the samples in the train and test sets.
+To enhance the robustness of results this process is repeated $25$~times, each time with a different shuffling of the samples in the train and test sets.
 Some datasets with attribute values of different magnitudes have a pre-processing scaling transformation.
@@
 
-% Of all the regression methods considered, the proposed method is the one that shows, with one exception, the lowest error.
-Only Random Forest outperforms \ac{EPR} systematically (it also outperforms all the other regression algorithms).
-The single non-ensemble exception --- besides the artificial dataset that uses a straightforward polynomial relation --- is the Auto MPG dataset where \ac{EPR} has comparable results.
-This is evidence that applying standard genetic operators for polynomial model searching is a viable tool for regression purposes.
+Of the regression methods considered \ac{SVM} achieves the best results in 3 out of 4 datasets.
+However \ac{SVM} and Conditional Inference Trees are pre-trained, with parameters tuned for each particular dataset unlike \ac{EPRR}.
+Even so it is the best estimator for the \textsc{Abalone} dataset and in the remaining datasets it outperforms most of the other estimators.
 
-Of the regression methods considered, \ac{SVM} achieved the best results in 3 out of 4 datasets.
-However, \ac{SVM} was tuned for each particular dataset while \ac{EPRR} was executed with its parameters preset.
-Even so, \ac{EPRR} had the best error in the \textsc{Abalone} dataset, and achieved competitive results in the \textsc{Housing} and \textsc{Auto MPG} datasets.
-
-Comparing \ac{EPR} and \ac{EPRR} -- which is the main article's topic -- the regularized version achieved much better results at \textsc{Abalone} and especially \textsc{Kinematics}.
+Comparing \ac{EPR} and \ac{EPRR} --- the main article's topic --- the regularized version achieved much better results at \textsc{Abalone} and especially \textsc{Kinematics}.
 The \textsc{Housing} dataset improved errors when compared with \ac{EPR} in a difference in means, resulted in a 95\% HDI (Highest Density Interval) equal to [0.001, 0.119] which, while borderline, achieves statistical significance.
@@
 
-For complexity considerations \ac{EPR} and \ac{EPRR} demands some processing time.
+For complexity considerations \ac{EPR} and \ac{EPRR} demand some processing time.
 On a quad-core computer, processing the \textsc{Kinematics} dataset (with 8k~observations) takes approximately 5~minutes.
-The processing time can probably be speeded between one to two orders in magnitude if the process is implemented in a low level programming language like \texttt{C++}.
-However, speed optimization was not the focus of this article.
+The processing time can probably be speeded between one to two orders in magnitude if the process is implemented with computational speed in mind.
+However, errospeed optimization is not the focus of this article.
 

### 2013-12-23 texts/submission_02-asc/paper_spfga-review.tex
Pre-submit version with small corrections.

@@
 \acrodef{EPR}{Evolutionary Polynomial Regression}
-\acrodef{EPRR}[EPR$^2$]{Evolutionary Polynomial Regression with Regularization}
+\acrodef{EPRR}[EPRR]{Evolutionary Polynomial Regression with Regularization}
 
@@
 \includegraphics[width=0.98\textwidth]{figure_4.pdf}
-\caption{Error progress for the \textsc{Abalone} dataset during a single execution of the genetic algorithm.
+\caption{Learning curve: Error progress for the \textsc{Abalone} dataset during a single execution of the genetic algorithm.
 The figure shows the fitness evolution for two different regularization values.
@@
 
-\note{Discutir a localidade do regressor}
-
-Of the regression methods considered \ac{SVM} achieves the best results in 3 out of 4 datasets.
-However \ac{SVM} and Conditional Inference Trees are pre-trained, with parameters tuned for each particular dataset unlike \ac{EPRR}.
+Of the regression methods considered \ac{SVM} achieves the best results in three out of four datasets.
+However \ac{SVM} and Conditional Inference Trees are pre-trained, having parameters tuned for each particular dataset unlike \ac{EPRR}, that runs with the same parameterization on all datasets.
 Even so it is the best estimator for the \textsc{Abalone} dataset and in the remaining datasets it outperforms most of the other estimators.
 
-Comparing \ac{EPR} and \ac{EPRR} --- the main article's topic --- the regularized version achieved much better results at \textsc{Abalone} and especially \textsc{Kinematics}.
-The \textsc{Housing} dataset improved errors when compared with \ac{EPR} in a difference in means, resulted in a 95\% HDI (Highest Density Interval) equal to [0.001, 0.119] which, while borderline, achieves statistical significance.
-Only in the \textsc{Auto MPG} dataset \ac{EPR} achieved better results, even if not that different from \ac{EPRR}.
+Comparing \ac{EPR} and \ac{EPRR} --- the main article's topic --- the regularized version achieves much better results at \textsc{Abalone} and especially \textsc{Kinematics}.
+On the \textsc{Housing} dataset errors are improved  wrt \ac{EPR} in a difference in means, resulted in a 95\% HDI (Highest Density Interval) equal to $\left\lbrack 0.001, 0.119 \right\rbrack$ which, while borderline, achieves statistical significance.
+Only in the \textsc{Auto MPG} dataset \ac{EPR} achieves better results, even if not that different from \ac{EPRR}.
 
 For complexity considerations \ac{EPR} and \ac{EPRR} demand some processing time.
-On a quad-core computer, processing the \textsc{Kinematics} dataset (with 8k~observations) takes approximately 5~minutes.
-The processing time can probably be speeded between one to two orders in magnitude if the process is implemented with computational speed in mind.
-However, errospeed optimization is not the focus of this article.
+On a quad-core computer, processing the \textsc{Kinematics} dataset (with near $8K$~observations) takes approximately 5~minutes.
+Probably processing time can be reduced by one to two orders in magnitude if the algorithm is implemented with computational speed in mind.
+However, speed optimization is not the focus of this article.
 
@@
 Namely, the regularization parameter, $\lambda$, can be tested with several values, instead of being fixed at $0.8$.
-Other parameters like mutation chance or the amount of elitism could also be tested.
-However, these type of tests need a low-level, fast implementation of \ac{EPR} and were not considered.
+Other parameters like mutation chance or the amount of elitism can also be tested.
+However, these type of tests need a low-level, fast implementation of \ac{EPR} and are postponed to future investigation.
 
@@
 
-Datasets used herein were selected from Luís Torgo's data repository, \url{http://www.dcc.fc.up.pt/~ltorgo/Regression/DataSets.html}.
+Datasets used herein are selected from Luís Torgo's data repository, \url{http://www.dcc.fc.up.pt/~ltorgo/Regression/DataSets.html}.
 Most can also be found in the UCI ML repository at \url{http://archive.ics.uci.edu/ml/}.

### 2013-12-24 texts/EPRR_FILES-ASC/eprr.tex
Setup to submit to “Applied Soft Computing”.

--- 2013-12-23 texts/submission_02-asc/paper_spfga-review.tex
@@
-%\documentclass[review,preprint,authoryear,12pt]{elsarticle}
-\documentclass[preprint,authoryear,12pt]{elsarticle}
+%% Use the option review to obtain double line spacing
+\documentclass[times,review,preprint]{elsarticle}
 
-%
-%	LINE NUMBERING FOR REVISION
-%
-%\usepackage{lineno}
-%
+%% Use the options `twocolumn,final' to obtain the final layout
+%\documentclass[times,twocolumn,final,authoryear]{elsarticle}
 
-%
-%	FILE ENCODING
-%
 \usepackage[utf8]{inputenc}
-%
-
-%
-%	PAGE GEOMETRY
-%
-%\usepackage{geometry}
-%\geometry{a4paper}
-%
-
-%
-%	AMS STUFF
-%
-\usepackage{amsfonts}
-%\usepackage{amsthm}
-\newtheorem{theorem}{Theorem}
-%
-
-%
-%	ACRONYMS
-%
-\usepackage{acronym}
-%
-\acrodef{GA}{Genetic algorithms}
-\acrodef{rmse}[error]{root-mean-square}
-%\acrodef{EPR}[\textsc{EPR}]{Genetic Algorithms for Polynomials}
-%\acrodef{EPRR}[\textsc{EPR.reg}]{Genetic Algorithms for Polynomials with Regularization}
-
-\acrodef{EPR}[EPR]{Evolutionary Polynomial Regression}
-\acrodef{EPRR}[EPR$^2$]{Evolutionary Polynomial Regression with Regularization}
-
-\acrodef{EPR}{Evolutionary Polynomial Regression}
-\acrodef{EPRR}[EPRR]{Evolutionary Polynomial Regression with Regularization}
-
-\acrodef{SVM}[SVM]{Support Vector Machines}
-%
-
-%
-%	FIGURE AND STUFF ENVIRONMENT
-%
 \usepackage{graphicx}
-%
-
 %
@@
 %
-
-%
 %	HYPERREFERENCES
 %
-\usepackage{hyperref}
+\usepackage{url}
 %
-
 %
@@
 %
-
 %
@@
 %
+\usepackage{lineno}
+%
+%	ACRONYMS
+%
+\usepackage{acronym}
+%
+%
+%
+\def\TM{$^{\rm TM}$}
+
+\journal{Applied Soft Computing}
+
+\begin{document}
 
 %
-%
-%%% BEGIN DOCUMENT
-%
-%
-\begin{document}
+\acrodef{GA}{Genetic algorithms}
+\acrodef{rmse}[error]{root-mean-square}
+\acrodef{EPR}[EPR]{Evolutionary Polynomial Regression}
+\acrodef{EPRR}[EPR$^2$]{Evolutionary Polynomial Regression with Regularization}
+\acrodef{EPR}{Evolutionary Polynomial Regression}
+\acrodef{EPRR}[EPRR]{Evolutionary Polynomial Regression with Regularization}
+\acrodef{SVM}[SVM]{Support Vector Machines}
+
+\setcounter{page}{0}
+
 \begin{frontmatter}
 
-\title{A Method for Regularization of Evolutionary Polynomial Regressions}
+%% Title, authors and addresses
 
-\author[ue,labmag]{Francisco ~Coelho\corref{fc}}
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for the associated footnote;
+%% use the fnref command within \author or \address for footnotes;
+%% use the fntext command for the associated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for the associated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
+%%
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \address{Address\fnref{label3}}
+%% \fntext[label3]{}
+
+\title{A method for regularization of evolutionary polynomial regressions}
+
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{<author name>}
+%% \address[label1]{<address>}
+%% \address[label2]{<address>}
+
+\author[1]{Francisco Coelho\corref{cor1}} 
+\cortext[cor1]{Corresponding author: 
+  Tel.: +351-919-006-379}
 \ead{fc@di.uevora.pt}
-\cortext[fc]{Corresponding author}
+\author[2]{Jo\~{a}o Pedro Neto}
 
-\author[fcul,labmag]{João ~Pedro ~Neto}
-\ead{jpn@di.fc.ul.pt}
+\address[1]{Departamento de Informática, Universidade de Évora, Rua Romão Ramalho 58, 7000-671 Évora, Portugal}
+\address[2]{Departamento de Informática, Faculdade de Ciências da Universidade de Lisboa, Campo Grande, 1749-016 Lisboa, Portugal}
 
-\address[ue]{Dept. Informática, Universidade de Évora, Rua Romão Ramalho 58, 7000-671 Évora}
-\address[fcul]{Dept. Informática, Faculdade de Ciências da Universidade de Lisboa, Campo Grande 1749-016 Lisboa}
-\address[labmag]{Laboratory of Agent Modelling (LabMAg)}
-%
-%\date{[{\sc draft} \today]}
-%
 \begin{abstract}
 While many applications require models that have no acceptable linear approximation, the simpler nonlinear models are defined by polynomials.
-The use of genetic algorithms to find polynomial models from data is known as \ac{EPR}.
+The use of genetic algorithms to find polynomial models from data is known as Evolutionary Polynomial Regression.
 %
-This paper introduces \ac{EPRR}, an algorithm that extends the \ac{EPR} method and describes a set of experiences on common datasets that compare both flavors of \ac{EPR} and other methods including Linear Regression, Regression Trees and Support Vector Regression.
+This paper introduces Evolutionary Polynomial Regression with Regularization, an algorithm that extends the EPR method and describes a set of experiences on common datasets that compare both flavors of EPR and other methods including Linear Regression, Regression Trees and Support Vector Regression.
 
-The empiric conclusion of those experiments is that \ac{EPRR} is able to achieve better fitting than other non-ensemble methods and it has shorter computation time than plain \ac{EPR}.
+The empiric conclusion of those experiments is that EPR with regularization is able to achieve better fitting than other non-ensemble methods and it has shorter computation time than plain EPR.
 \end{abstract}
-%
+
 \begin{keyword}
-evolutionary polynomial regression \sep regularization \sep feature extraction \sep dimensionality reduction 
-\end{keyword}
+evolutionary polynomial regression \sep regularization \sep feature extraction \end{keyword}
+
 \end{frontmatter}
-%
-%	Line numbers for revision
-%
-%
-%\linenumbers
-%
+
+\linenumbers
+
 \section{Introduction}
-
 With notable exceptions (\emph{e.g.} neural networks) machine learning regression techniques produce linear models.
@@
 The idea of using \acp{GA} to find a polynomial regression is not new \citep{Maertens:2006aa, Yu:2008aa, Wu:2009aa} but still generates original research \citep{Hofwing:2011aa,Cetisli:2011aa}.
-The modern formulation of the use of \ac{GA} to find polynomial models is known as \acf{EPR} and systematization can be traced back to the work of Davidson, Savic and Walters (\cite{Davidson:2003aa}).
-Further developments include multi-objective optimizations (\cite{Giustolisi:2009aa}).
+The modern formulation of the use of \ac{GA} to find polynomial models is known as \acf{EPR} and systematization can be traced back to the work of Davidson, Savic and Walters \cite{Davidson:2003aa}.
+Further developments include multi-objective optimizations \cite{Giustolisi:2009aa}.
 
@@
 
-%A method that produces \emph{adequate} models from observed complex data has many uses.
-For example by a scientist to better understand the source of the data or by an autonomous agents adapting to the environment.
-
-% It turns out that, discarding the computational cost of training, the regression method presented here, \acf{EPRR}, proves to be quite effective.
-Indeed, according to the results of common  methods on several real-world datasets it is only systematically out-performed by random forests, an \emph{ensemble} method.
-
 The next section describes the method's details and is followed by a presentation of some performance results. The last section draws some conclusions and points future research tasks.
@@
 \section{Genetic Algorithms for Polynomials}
-
 This section starts with a brief introduction and outline of the evolutionary polynomial regression algorithm, \ac{EPR}, and proceeds into core details as the encoding used to represent individual polynomial instances in the \ac{GA} populations and the regularization of the cost function.
@@
 Changing the order of the rows doesn't change the polynomial whereas changing the order of the columns corresponds to changing the respective variables.
-%Notice that if we interpret matrix $A$ as a single column of monomials and set the coefficients $T = \left[ \begin{array}{ccc} 2 & 1 & \frac{1}{2} \end{array} \right ]$ then
-%$$
-%p\left( x_1, x_2, x_3 \right) = TA = \left[ \begin{array}{ccc} 2 & 1 & \frac{1}{2} \end{array} \right ]\left[ 
-%\begin{array}{ccc}
-%x_1 \\
-%x_2 x_3 \\
-%x_1^2 x_3
-%\end{array}
-%\right].
-%$$
 
@@
 \subsection{Polynomial Encoding}\label{subs:polynomial.encoding}
-%
-%\begin{quotation}
-%\note{Next paragraph moved here from the introduction.}
-%Another important change is in the coding mechanism which introduces two new features: (a) inclusion of the linear terms in the polynomial expression, and (b) the activation/deactivation of monomials during evolution to allow the increase/decrease of polynomial expressions.
-%\end{quotation}
-%
+
 The specific encoding (representation) of a set of monomials is an important aspect in the implementation of \ac{EPR}.
@@
 In practice, this means that the implementation of the encoding can omit the first bit.
-% If the coding consists of entirely zeros, by convention, it describes polynomial $x_1$.
-This has to do with the execution of the polynomial regression that would fail if we interpret it as the zero polynomial.
-Anyway, for progressive larger  binary descriptions, the chances of getting this zero description decrease exponentially, so it does not impact in any meaningful way in the algorithm's performance.
 
@@
 Conditional Inference Trees}
-\note{Plots of ``Regression Trees'' and ``Conditional Inference Trees'' look the same!} 
 \label{artificial_dataset1_lambda1.0}
@@
 
-% best polynomial, rsme 0.65
-% $$y = 0.52 s_1^2 x_5 - 0.13 x_2 x_8 -0.75 x_1^2 x_6 + 0.78 x_8 + 0.18 $$
-
 \item{\textsc{Auto MPG}} gathers fuel consumption in miles per gallon, based on two discrete and five continuous attributes ($m=7$). There are $n=398$ observations;
-
-% best polynomial, \ac{rmse} 0.32
-%$$ y = 0.4 x_6 + 0.02 x_5^3 x_6 - 0.85 x_4 + 0.23 x_4^2 - 0.26 $$
 
@@
 There are $n=8192$ observations;
-
-% best polynomial, \ac{rmse} 0.74
-% $$ y = 0.24 x_5 - 0.23 x_5 x_6 x_7 - 0.19 x_4 x_7 - 0.5 x_3 $$
 
@@
 
-%Since the entire learning process takes some time, in the current \texttt{R} implementation, placing a limit between 50 to 100 generations already achieves good results, relative to higher iteration values.
-Figure~\ref{Abalone_fitnessProgress} shows a typical error evolution for the dataset \textsc{Abalone} given a regularized version ($\lambda = 0.975$) and a non-regularized ($\lambda = 1.0$).\note{Why not $\lambda=0.8$?}
-
 \section{Conclusion and Future Work}
@@
 
-\section*{Bibliography}
-
-\bibliographystyle{model2-names}
+\bibliographystyle{elsarticle-num}
 \bibliography{fullbib2}

### 2013-12-27 code/exploratory/gap.py
Completed working class for EPRR

@@ -23,6 +23,8 @@ from deap import algorithms, base, creator, tools
 #
 import warnings
 warnings.filterwarnings('ignore')
+#
+#   Save a prediction in the Kaggle submission format
 #       
 def save_data(filename, data):
     f = open(filename, 'w')
@@ -32,6 +34,8 @@ def save_data(filename, data):
         f.write('%d,%d\n' % (i,x))
         i += 1
     f.close()
+#
+#   Load a dataset
 #    
 def load_data(root_folder = '../../datasets/'):
     '''
@@ -44,7 +48,9 @@ def load_data(root_folder = '../../datasets/'):
     data = load_file( data_folder + 'auto_no_missing.data', 0)
     
     return data
-    
+#
+#   Represent a set of monomial transformations
+#    
 class PolyTerms:
     '''Represent a set of monomial transformations.
     
@@ -89,7 +95,7 @@ class PolyTerms:
         No check is performed.
         '''
         if dtype is None:
-            self.terms = np.array( degrees )
+            self.terms = np.array( degrees, dtype = np.int8 )
         else:
             self.terms = np.array( degrees, dtype = dtype)
         self.num_terms, self.num_vars = self.terms.shape
@@ -97,6 +103,44 @@ class PolyTerms:
         self.coef_ = None
         self.intercept_ = None
                 
+    def simplify(self, epsilon = 1E-5):
+
+        if self.coef_ is None:
+            return self, None
+        #
+        #   remove terms with |coef| < epsilon
+        #
+        c = np.copy( self.coef_ )            
+        t = np.copy( self.terms )
+        big = np.abs(c) > epsilon
+        t = t[big,:]
+        c = c[big]
+        #
+        #   join common terms, summing the coefficients
+        #
+        n,_ = t.shape
+        cs = list()
+        open_terms = list(range(n))
+        closed_terms = list()
+        t_ = list()
+        c_ = list()
+        while len(open_terms) > 0:
+            i = open_terms.pop()
+            si = c[i]
+            closed_terms.append(i)
+            ti = t[i]
+            for j in open_terms:
+                if (t[j] == ti).all():
+                    open_terms.remove(j)
+                    closed_terms.append(j)
+                    si += c[j]
+            t_.append(list(ti))
+            c_.append(si)
+        p = PolyTerms(np.array(t_))
+        p.coef_ = c_
+        p.intercept_ = self.intercept_
+        return p
+        
     def __call__(self, X, extend = False):
[…]

### 2013-12-27 code/exploratory/test_gap.py

@@ -1,0 +1,128 @@
+#!/usr/bin/env python
+    
+from gap import *
+                   
+                   
+def test_basic():
+    print('\nTesting basic evaluation.')
+    poly = PolyTerms([
+        [0,1,0],
+        [2,0,0],
+        [1,1,1],
+        [0.5,3,1] ])
+    print('\tPolyTerms:\n%s'%poly)
+    print('\tVars:\t%d\n\tTerms\t:\t%d'%(poly.num_vars,poly.num_terms))
+    X = np.array([
+        [1,2,1],
+        [2,2,2 ]])
+    print('\tData:\n%s'%X)
+    print('\tTransform:\n%s'%poly(X))
+    #
+    print('\nTesting random polyterm generator.')
+    print('\tRandom PolyTerm:\n%s'%random_polyterm(4,2,[0,1,2,0.5]))
+    #
+    print('\nTesting polyterm simplification.')
+    p0 = PolyTerms([[0,1], [1,1], [2,1], [0,1]])
+    print('\tPolyTerms:\n%s'%p0)
+    p0.coef_ = [2, 3, 1E-6, 2]
+    p1 = p0.simplify()
+    print('\tReduced PolyTerms:\n%s'%p1)
+    
+
+def test_regression():
+    print('\nTesting regression.')
+    print('\n\tBoston House Prices\n')
+    #
+    scorer = PolyRegression_Scorer(
+        datasets.load_boston(),
+        linear_model.LinearRegression()
+        )
+    #
+    poly = random_polyterm(4, scorer.num_features, [0,1,2,0.5])
+    print('\tPolyTerms:\n%s'%poly)
+    #
+    scorer.score(poly)
+    print('\tEstimator coefs:\t%s'%(scorer.estimator.coef_))
+    print('\tEstimator intercept:\t%s'%(scorer.estimator.intercept_))
+    print('\tExpected score (error):\t%5f ±%5f'%(scorer.scores.mean(), scorer.scores.std()))
+
+def test_deap():
+    import array
+    import random
+
+    from deap import algorithms
+    from deap import base
+    from deap import creator
+    from deap import tools
+
+    creator.create("FitnessMax", base.Fitness, weights=(1.0,))
+    creator.create("Individual", array.array, typecode='b', fitness=creator.FitnessMax)
+
+    toolbox = base.Toolbox()
+
+    # Attribute generator
+    toolbox.register("attr_bool", random.randint, 0, 1)
+
+    # Structure initializers
[…]

### 2014-01-13 gapolyfuncs.R
added random grid search

@@ -64,7 +64,6 @@ make.formula <- function(bits, n.vars, max.monoids) {
   # this next line simulates the EPR using all monoids
   # it should be uncommented when we want to get results without any type of regulatization
   # active.monoid <- rep(1, max.monoids)
-  
   for(i in 1:length(monoids)) {
     if (active.monoid[i]==1) {
       monoid <- make.monoid(monoids[[i]], n.vars, vars)
@@ -391,3 +390,7 @@ test.regularization <- function(my.data,
        ga.model = best.ga.model,
        ga.reg.model = best.ga.reg.model)  # also include the best found ga.models
 }
+
+
+# cf: 
+# http://stats.stackexchange.com/questions/3328/given-a-10d-mcmc-chain-how-can-i-determine-its-posterior-modes-in-r

### 2014-01-13 gridfuncs.R

@@ -1,0 +1,111 @@
+######################################
+# Tuning GAPoly parameters
+
+source("gapolyfuncs.R")
+
+# Parameters:
+# + mutation.rate \in [0,0.5)
+# + elitism       percentage
+# + lambda        >= 0
+# + max.monoids   [1,9]
+# + max.degree    [2,9]
+
+#mutation.rate <- elitism.perc <- lambda <- max.monoids <- max.degree <- 0
+
+########################################
+# functions that compute next values (for each parameter)
+
+values.mutation  <- seq(0.01,0.5,len=50)
+next.mutation.rate <- function() { sample(values.mutation, 1) }
+
+values.elitism  <- seq(0,0.1,len=11)
+next.elitism    <- function() { sample(values.elitism, 1) }
+
+values.lambda   <- seq(0,2,len=81)
+next.lambda     <- function(){ sample(values.lambda,1) }
+
+next.max.monoid <- function(){ sample(2:9,1) }
+
+next.max.degree <- function(){ sample(2:4,1) }
+
+########################################
+
+tune.gapoly.params <- function(my.data,population=100,iterations=50,runs=10) {
+  
+  #initialize parameters
+  mutation.rate <- 0.05
+  elitism.perc  <- 0.01
+  lambda        <- 0.8
+  max.monoids   <- 4
+  max.degree    <- 3
+  
+  # prepare dataset
+  train.p.size <- 0.7 # percentage of training set
+  n.vars       <- ncol(my.data)-1
+  
+  # make train & test set (ie, each run has a different train+test sets)
+  inTrain   <- sample(1:nrow(my.data), train.p.size * nrow(my.data))
+  train.set <- my.data[inTrain,]
+  test.set  <- my.data[-inTrain,]
+  
+  report <- data.frame(mutation=rep(NA,runs),
+                       elitism=rep(NA,runs),
+                       lambda=rep(NA,runs),
+                       monoids=rep(NA,runs),
+                       degree=rep(NA,runs),
+                       error=rep(NA,runs),
+                       model=rep(NA,runs))
+  
+  for(run in 1:runs) {
+    # run GA.poly one time
+    GAmodel <- rbga.bin(size = max.monoids + max.degree*n.vars*max.monoids, 
+                        iters = iterations, 
+                        popSize = population, 
+                        mutationChance = mutation.rate, 
+                        elitism = TRUE,#elitism.perc * population, 
[…]

### 2014-01-14 gapolyfuncs.R
corrected elitism (missed the round())

@@ -392,5 +392,3 @@ test.regularization <- function(my.data,
 }
 
 
-# cf: 
-# http://stats.stackexchange.com/questions/3328/given-a-10d-mcmc-chain-how-can-i-determine-its-posterior-modes-in-r

### 2014-01-14 gridfuncs.R

@@ -62,7 +62,7 @@ tune.gapoly.params <- function(my.data,population=100,iterations=50,runs=10) {
                         iters = iterations, 
                         popSize = population, 
                         mutationChance = mutation.rate, 
-                        elitism = TRUE,#elitism.perc * population, 
+                        elitism = round(elitism.perc * population), 
                         evalFunc = evalFuncFactory(train.set, n.vars, max.monoids, lambda))
     
     best.solution <- GAmodel$population[1,]
@@ -109,3 +109,6 @@ head(my.data,8)
 
 report <- tune.gapoly.params(my.data, population=80, iterations=60, runs=25)
 report
+
+# cf: 
+# http://stats.stackexchange.com/questions/3328/given-a-10d-mcmc-chain-how-can-i-determine-its-posterior-modes-in-r

### 2014-01-14 gridfuncs.R
1st housing results

@@ -110,5 +110,7 @@ head(my.data,8)
 report <- tune.gapoly.params(my.data, population=80, iterations=60, runs=25)
 report
 
+write.table(report, "grid.search.Housing.txt")
+
 # cf: 
 # http://stats.stackexchange.com/questions/3328/given-a-10d-mcmc-chain-how-can-i-determine-its-posterior-modes-in-r

### 2014-01-14 gridfuncs.R
micro-change

@@ -107,10 +107,10 @@ names(my.data) <- c(paste0("x",1:(ncol(my.data)-1)),"y")
 my.data <- as.data.frame(scale(my.data))
 head(my.data,8)
 
-report <- tune.gapoly.params(my.data, population=80, iterations=60, runs=25)
+report <- tune.gapoly.params(my.data, population=100, iterations=75, runs=50)
 report
 
-write.table(report, "grid.search.Housing.txt")
+write.table(report, "grid.search.Housing2.txt")
 
 # cf: 
 # http://stats.stackexchange.com/questions/3328/given-a-10d-mcmc-chain-how-can-i-determine-its-posterior-modes-in-r

### 2014-01-14 gridfuncs.R
iterative grid search

@@ -3,15 +3,6 @@
 
 source("gapolyfuncs.R")
 
-# Parameters:
-# + mutation.rate \in [0,0.5)
-# + elitism       percentage
-# + lambda        >= 0
-# + max.monoids   [1,9]
-# + max.degree    [2,9]
-
-#mutation.rate <- elitism.perc <- lambda <- max.monoids <- max.degree <- 0
-
 ########################################
 # functions that compute next values (for each parameter)
 
@@ -29,7 +20,7 @@ next.max.monoid <- function(){ sample(2:9,1) }
 next.max.degree <- function(){ sample(2:4,1) }
 
 ########################################
-
+# Random Grid Search
 tune.gapoly.params <- function(my.data,population=100,iterations=50,runs=10) {
   
   #initialize parameters
@@ -84,13 +75,7 @@ tune.gapoly.params <- function(my.data,population=100,iterations=50,runs=10) {
     lambda        <- next.lambda()
     max.monoids   <- next.max.monoid()
     max.degree    <- next.max.degree()
-    
-#     print(mutation.rate)
-#     print(elitism.perc)
-#     print(lambda)
-#     print(max.monoids)
-#     print(max.degree)
-    
+
     # show progress
     cat(paste0(" ",run,"."))
   }
@@ -114,3 +99,243 @@ write.table(report, "grid.search.Housing2.txt")
 
 # cf: 
 # http://stats.stackexchange.com/questions/3328/given-a-10d-mcmc-chain-how-can-i-determine-its-posterior-modes-in-r
+
+#######################################
+# Iterative Search
+
+# In this search, we first test for one parameter, find the best result, then
+# pass to the next parameter, and so on until some convergence is seen
+
+tune.gapoly.iterative <- function(my.data,population=100,iterations=50,runs=10) {
+  
+  #initialize parameters
+  mutation.rate <- 0.05
+  elitism.perc  <- 0.01
+  lambda        <- 0.8
+  max.monoids   <- 4
+  max.degree    <- 3
+  
+  # prepare dataset
+  train.p.size <- 0.7 # percentage of training set
+  n.vars       <- ncol(my.data)-1
+  
+  # make train & test set (ie, each run has a different train+test sets)
+  inTrain   <- sample(1:nrow(my.data), train.p.size * nrow(my.data))
+  train.set <- my.data[inTrain,]
+  test.set  <- my.data[-inTrain,]
+  
+  report.obs <- 2*runs
+  report <- data.frame(mutation=rep(NA,report.obs),
[…]

### 2014-01-14 code/exploratory/eprr.py
Reprodução dos resultados mostrados no artigo.

@@ -1,6 +1,3 @@
-#
-#   TODO add regularisation to error
-#
 import numpy as np
 from sklearn import cross_validation as cv
 from sklearn import svm
@@ -8,6 +5,7 @@ from sklearn import preprocessing
 from sklearn import datasets
 from sklearn import metrics
 from sklearn import linear_model
+#import math
 
 from deap import algorithms, base, creator, tools
 
@@ -142,13 +140,16 @@ class PolyTerms:
         else:
             return Z
 
+    def poly_call(self, X):
+        return np.dot( self(X), self.coef_) + self.intercept_
+
     def score(self, X, y):
-        Z = self(X)
-        y_hat = np.dot(Z, self.coef_) + self.intercept_
-        y_mean = np.mean(y)
-        u = np.dot(y - y_hat, y - y_hat)
-        v = np.dot(y - y_mean, y - y_mean )
-        return (1. - u/v)
+        y_pred = self.poly_call(X)
+        #y_mean = np.mean(y)
+        #u = np.dot(y - y_hat, y - y_hat)
+        #v = np.dot(y - y_mean, y - y_mean )
+        #return (1. - u/v)
+        return metrics.r2_score(y, y_pred)
 
     def __repr__(self):
         '''Computes a string representation of the instance.
@@ -281,7 +282,8 @@ class EPRR:
                     y_test,
                     cv = self.cross_validations,
                     scoring = metrics.make_scorer(
-                        lambda x,y: penalty * np.dot(x-y,x-y),
+                        #lambda x,y: penalty * np.dot(x-y,x-y),
+                        lambda x,y: penalty * metrics.mean_squared_error(x,y),
                         greater_is_better = False,
                         needs_threshold = False ) )
             return scores
@@ -380,7 +382,7 @@ class EPRR:
 
     def predict(self, X):
         if self.__trained__:
-            return self.poly_(X)
+            return self.poly_.poly_call(X)
         else:
             return None
 

### 2014-01-14 code/exploratory/evaluate_estimators.py

@@ -1,0 +1,254 @@
+#!/opt/local/bin/python
+
+import sklearn as sk
+import sklearn.cross_validation as cv
+import numpy as np
+import eprr
+import os
+import glob
+
+#########################################################################################
+#
+#   GET DATASETS
+#
+#########################################################################################
+
+def get_dataset_abalone():
+    datafolder = '../../datasets'
+    dataset_name = 'Abalone'
+    datafile = glob.glob(os.path.join(
+        datafolder, dataset_name, '*.data'))[0]
+    D = np.loadtxt( open(datafile, 'r'), delimiter = ',', skiprows = 0, usecols = (1,2,3,4,5,6,7,8) )
+
+    D = sk.preprocessing.scale(D)
+    X = D[:,:-1]
+    y = D[:,-1]
+
+    return X,y
+
+def get_dataset_housing():
+    datafolder = '../../datasets'
+    dataset_name = 'Housing'
+    datafile = glob.glob(os.path.join(
+        datafolder, dataset_name, '*.data'))[0]
+    D = np.loadtxt( open(datafile, 'r'), delimiter = ',', skiprows = 0 )
+
+    D = sk.preprocessing.scale(D)
+    X = D[:,:-1]
+    y = D[:,-1]
+
+    return X,y
+
+
+def get_dataset_fc():
+    num_samples = 150
+    x1 = np.random.uniform(0,1, num_samples)
+    x2 = np.random.uniform(0,1, num_samples)
+    e = np.random.uniform(0,0.001, num_samples)    
+    y = (x1 ** 2) + (-2 * x1 * x2) + (3 * x2 ** 2) - 4 + e
+    X = np.c_[x1,x2]
+    return X,y    
+
+def get_dataset_jpn():    
+    num_samples = 100
+    x1 = np.random.poisson(1, num_samples)
+    x2 = np.random.poisson(1, num_samples)
+    x3 = np.random.uniform(0,1, num_samples)
+    x4 = np.random.uniform(0,1, num_samples)
+    y = x2 * (x4 ** 2) + (x1 ** 2) * x3 + 5
+    X = np.c_[x1,x2,x3,x4]
+    
+    return X,y       
+
+#########################################################################################
+#
+#   GET ESTIMATOR
+#
+#########################################################################################
+
+def get_estimator_tunedsvr_dataset(X,y):
+    X_train, X_test, y_train, y_test = cv.train_test_split(X,y, train_size = 0.66)
+    evaluator = cv.StratifiedKFold(y_train, n_folds = 3)
[…]

### 2014-01-14 code/exploratory/test_eprr.py

@@ -9,7 +9,8 @@ import glob
 def tester(f):
     def deco():
         print('\n********** START TESTING [%s] **********\n'%f.__name__)
-        f()
+        (p,s,e) = f()
+        print('\n\tbest poly:\t%s\n\twith score:\t%s\n\tand error:\t%s'%(p, s, e))
         print('\n********** END TESTING [%s] **********\n'%f.__name__)
     return deco
 
@@ -34,6 +35,7 @@ def test_polyterms():
     p0.coef_ = [2, 3, 1E-6, 2]
     p1 = p0.simplify()
     print('\tReduced PolyTerms:\n%s'%p1)
+    return (p1,0,0)
  
 #        
 def get_dataset_fc():
@@ -56,10 +58,10 @@ def get_dataset_jpn():
     X = np.c_[x1,x2,x3,x4]
     
     return X,y       
+
 #
 @tester
-def test_polysearch():
-    print("Search JPN poly")
+def test_polysearch_jpn():
     X,y = get_dataset_jpn()
 
     pop_size = 100
@@ -80,9 +82,10 @@ def test_polysearch():
     e = e.fit(X,y)
     error = e.poly_.fitness
     p = e.poly_.simplify()
-    print('\n\tbest poly:\t%s\n\twith score:\t%s\n\tand error:\t%s'%(p,p.score(X,y), error))
+    return (p,p.score(X,y), error)
 
-    print("Search FC poly")
+@tester
+def test_polysearch_fc():
     X,y = get_dataset_fc()
 
     pop_size = 100
@@ -103,7 +106,7 @@ def test_polysearch():
     e = e.fit(X,y)
     error = e.poly_.fitness
     p = e.poly_.simplify()
-    print('\n\tbest poly:\t%s\n\twith score:\t%s\n\tand error:\t%s'%(p,p.score(X,y), error))
+    return (p,p.score(X,y), error)
 
 
 def get_dataset_abalone():
@@ -123,10 +126,13 @@ def get_dataset_abalone():
 def test_abalone():
     X,y = get_dataset_abalone()
     num_samples, _ = X.shape
-    rand_rows = np.random.choice(num_samples, 100, replace = False)
+    rand_rows = np.random.choice(num_samples, 200, replace = False)
 
-    X = X[rand_rows,:]
-    y = y[rand_rows]
+    X_train = X[rand_rows,:]
+    y_train = y[rand_rows]
+
[…]

### 2014-01-14 gridfuncs.R
Housing dataset iterative search results

@@ -335,7 +335,7 @@ tune.gapoly.iterative <- function(my.data,population=100,iterations=50,runs=10) 
   report
 }
 
-report <- tune.gapoly.iterative(my.data, population=10, iterations=8, runs=3)
+report <- tune.gapoly.iterative(my.data, population=75, iterations=60, runs=5)
 report
 
 write.table(report, "grid.iterative.search.Housing.txt")

### 2015-07-24 texts/submission_03_asc/Figure_3-caption.tex
Início Revisão ASC.

--- 2013-07-25 texts/submission_01-prl/paper_spfga-final.tex
@@
-\documentclass[final,authoryear,3p,times,twocolumn]{elsarticle}
-
-%
-%	FILE ENCODING
-%
-\usepackage[utf8]{inputenc}
-%
-
-%
-%	PAGE GEOMETRY
-%
-%\usepackage{geometry}
-%\geometry{a4paper}
-%
-
-%
-%	AMS STUFF
-%
-\usepackage{amsfonts}
-\usepackage{amsthm}
-\newtheorem{theorem}{Theorem}
-%
-
-%
-%	ACRONYMS
-%
-\usepackage{acronym}
-\acrodef{GA}{Genetic algorithms}
-\acrodef{rmse}[error]{root-mean-square error}
-\acrodef{GAPoly}[\textsc{GAPoly}]{Genetic Algorithms for Polynomials}
-%
-
-%
-%	FIGURE AND STUFF ENVIRONMENT
-%
-\usepackage{graphicx}
-%
-
-%
-%	ALGORITHMS PACKAGES
-%
-\usepackage{algpseudocode}
-\usepackage{algorithm}
-%
-
-%
-%	HYPERREFERENCES
-%
-\usepackage{hyperref}
-%
-
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-
-%
-%	DOCUMENT COMMANDS
-%
-\newcommand{\sample}[1]{\ensuremath{^{\left(#1\right)}}}
-%
-
-%
-%
-%%% BEGIN DOCUMENT
-%
-%
-\begin{document}
-
-\begin{frontmatter}
-
-\title{Selection of Polynomial Features by Genetic Algorithms}
-
-\author[ue,labmag]{Francisco Coelho\corref{fc}}
-\ead{fc@di.uevora.pt}
-\cortext[fc]{Corresponding author}
-
-\author[fcul,labmag]{João Pedro Neto}
-\ead{jpn@di.fc.ul.pt}
-
-\address[ue]{Dept. Informática, Universidade de Évora, Rua Romão Ramalho 58, 7000-671 Évora}
-\address[fcul]{Dept. Informática, Faculdade de Ciências da Universidade de Lisboa, Campo Grande 1749-016 Lisboa}
-\address[labmag]{Laboratory of Agent Modelling (LabMAg)}
-
-\date{[{\sc draft} \today]}
-
-\begin{abstract}
-Many applications require models that have no acceptable linear approximation and many nonlinear models are defined by polynomials.
-The use of genetic algorithms to find polynomial models is decades old but still poses challenges due to the complexity of the search and different definitions of optimal solution.
-
-This paper describes a two-step method that uses genetic algorithms and linear regression to find empirical polynomial regressions.
-Experiments on common datasets show that, discounting the training computational effort, this method is quite competitive.
-\end{abstract}
-
-\begin{keyword}
-polynomial regression \sep genetic algorithm 
-\end{keyword}
-\end{frontmatter}
-
-\section{Introduction}
-
-With notable exceptions (\emph{e.g.} neural networks) machine learning regression techniques are based on linear models.
-The linearity assumption has many advantages including reduced computational complexity and strong theoretical framework.
-However nonlinearity is unavoidable in many application scenarios, specially those with phase transitions or feedback loops, so common in ecology, cybernetics, robotics and other areas.
-
-Polynomials, one of the most studied subjects in mathematics, generalize li\-ne\-ar functions and define, perhaps, the simplest and most used nonlinear models.
-Applications include colorimetric calibration \citep{mendes2005adaptive}, explicit formul\ae\ for turbulent pipe flows \citep{davidson1999method}, computational linguistics \citep{sanchez2009obtaining} and more recently, analytical techniques for cultural heritage materials \citep{csefalvayova2010use}, liquid epoxy molding process \citep{chan2011modeling}, B-spline surface reconstruction \citep{galvez2012iterative}, product design \citep{chan2012development} or forecasting cyanotoxins presence in water reservoirs \citep{garcia2013hybrid}.
-This not only illustrates the wide spectrum of applications but, additionally, work in each one of these polynomial models uses, at some point, a genetic algorithm.
-
-\ac{GA} where, arguably, one of the hottest topics of research in the recent decades and with good reason since they outline an optimization scheme easy to conceptualize and with very broad application.
-If a nonlinear (or otherwise) model requires parameterization \acp{GA} provide a simple and often effective approach to search for locally optimal parameters.
-Research related to genetic algorithms abound and spans from the 1950s seminal work of Nils Aall Barricelli \citep{barricelli1962numerical} in the Institute for Advanced Study of Princeton to today's principal area of study for thousands of researchers, covered in hundreds of conferences, workshops and other meetings.
-Perhaps the key impulse to \acp{GA} come from John Holland's work and his book ``Adaptation in Natural and Artificial Systems'' \citep{holland1975adaptation}.
-
-One interesting take on genetic algorithms, named \emph{genetic programming} by John Koza \citep{koza1992genetic}, proposed the use of \acp{GA} to search the syntactic structure of complex functions.
-This syntactic structure search is keen to the central ideas of deep learning \citep{bengio2013representation, bengio2009learning}, a subarea of machine learning actually producing quite promising results (\emph{e.g.} in \cite{Tarlow:2013fk}).
-It is also related to the work presented in this paper in the sense that, unlike linear models that have a simple structure, $y=\sum_i \beta_i x_i$, nonlinear (in particular polynomial) models pose an additional ``structure" search problem.
-
-The idea of using \acp{GA} to find a polynomial regression is not new \citep{maertens2006genetic, yu2008optimal, wu2009novel} but still generates original research \citep{hofwing2011optimal, cetisli2011polynomial}.
-In line with that research this work describes a general method to find a polynomial regression of a given dataset.
-The optimal regression minimizes a cost function that accounts for both the \ac{rmse} and a regularization factor to avoid over-fitting.
-
-%A method that produces \emph{adequate} models from observed complex data has many uses.
-For example by a scientist to better understand the source of the data or by an autonomous agents adapting to the environment.
-
-It turns out that, discarding the computational cost of training, the polynomial regression method presented here, \ac{GAPoly}, provides a quite competitive regression method.
-Indeed, it is only systematically out-performed by random forests, an \emph{ensemble} method.
-
-The remainder of this paper is organized as usual: the next section describes the details of our method and is followed by a presentation of some performance results.
-The last section draws some conclusions and points future research tasks.
-
-\section{Genetic Algorithms for Polynomials}
-
-This section is dedicated to the description of an algorithm to find a polynomial regression from a given dataset.
-It starts with a brief introduction and outline of the algorithm and proceeds into core details as the encoding used to represent individual polynomial instances in the \ac{GA} populations and the regularization of the cost function.
-
-A usual representation of polynomials is
-$$
-p\left( x_1, \ldots, x_m\right) = \sum_i \theta_i q_i
-$$
-where each $q_i$ is a monomial, $q_i = \prod_{j} x_j^{\alpha_{ij}}$, the exponents are non-negative integers, $\alpha_{ij}\in\mathbb{N}_0$, and the coefficients are real valued, $\theta_i \in \mathbb{R}$.
-For example $p\left( x_1, x_2, x_3 \right) = 2 x_1 + x_2 x_3 + \frac{1}{2} x_1^2 x_3$ has monomials $q_1 = x_1, q_2 = x_2 x_3$ and $q_3 = x_1^2 x_3$, coefficients $\theta_1 = 2, \theta_2 = 1$ and $\theta_3 = 1/2$ and exponents $\alpha_{1,1} = 1, \alpha_{2,2}=1, \alpha_{2,3} = 1, \alpha_{3,1} = 2, \alpha_{3,3} = 1$ and all other $\alpha_{ij} = 0$.
-The exponents alone are a matrix that defines the monomial structure of the polynomial, $A = \left[\alpha_{ij}\right]$.
-For the example above
-$$
-A = \left[ 
-\begin{array}{ccc}
-1 & 0 & 0 \\
-0 & 1 & 1 \\
-2 & 0 & 1
-\end{array}
-\right] \sim 
-\left[ 
-\begin{array}{ccc}
-x_1 \\
-x_2 x_3 \\
-x_1^2 x_3
-\end{array}
-\right]
-$$
-where each row defines a monomial and each column represents a variable.
-Changing the order of the rows doesn't change the polynomial whereas changing the order of the columns corresponds to changing the respective variables.
-%Notice that if we interpret matrix $A$ as a single column of monomials and set the coefficients $T = \left[ \begin{array}{ccc} 2 & 1 & \frac{1}{2} \end{array} \right ]$ then
-%$$
-%p\left( x_1, x_2, x_3 \right) = TA = \left[ \begin{array}{ccc} 2 & 1 & \frac{1}{2} \end{array} \right ]\left[ 
-%\begin{array}{ccc}
-%x_1 \\
-%x_2 x_3 \\
-%x_1^2 x_3
-%\end{array}
-%\right].
-%$$
-
-This representation of polynomials makes the problem of structure search very clear: except for the trivial cases, the number of possible monomials given $n$ variables and a maximum joint degree $d$ grows exponentially with either $n$ or $d$.
-But more importantly, the polynomial regression problem can be naturally split into two subproblems:
-%
-\begin{enumerate}
-\item For a given set of monomials $\mathcal{P} = \left\lbrace q_1, \ldots, q_k\right\rbrace$, find the regression coefficients $\theta_1,\ldots,\theta_k$ that minimize the \ac{rmse} on a given dataset;
-
-\item Find the fittest set of monomials, \emph{i.e.} the polynomial that minimizes the \ac{rmse} on the same dataset;
-\end{enumerate}
-%
-More precisely, concerning the first problem, let $\mathcal{D}$ be a dataset with $n$ observations of variables $Y, X_1,\ldots,X_m$ and $\mathcal{P} = \left\lbrace q_1,\ldots, q_k\right\rbrace$ a set of $k$ monomial expressions over $X_1,\ldots,X_m$.
-Define the hypothesis\footnote{The expression $q|_{X=x}$ reads ``\emph{replace all instances of $X$ by $x$ in $q$}''.}
-%
-\begin{eqnarray*}
-h_{\Theta,\mathcal{P}}\left(x_1,\ldots,x_m\right) &=& \sum_{j = 1}^k \theta_j q_j|_{X_i=x_i,\forall 1 \leq i \leq m}
-\end{eqnarray*}
-%
-and let the cost
-%
-\begin{eqnarray}
-J_{\textrm{fit}}\left(\Theta;\mathcal{P},\mathcal{D}\right)  & =  \nonumber \\
-\sqrt{\frac{1}{n}\sum_{i=1}^n \left( y\sample{i} - h_{\Theta,\mathcal{P}}\left( x_1\sample{i},\ldots,x_m\sample{i} \right) \right)^2 }\label{eq:rmse}
-\end{eqnarray}
-%
-be the usual \acf{rmse} function. Now the first problem can be stated as:
-%
-\emph{Given a dataset $\mathcal{D}$ and a set of monomials $\mathcal{P}$, find parameters $\Theta$ that minimize $J_{\textrm{fit}}\left(\Theta;\mathcal{P},\mathcal{D}\right)$.}
-
-%
-It turns out that this problem can be solved as a usual linear regression problem by expanding $\mathcal{D}$ with columns that replicate the monomials in $\mathcal{M}$.
-
-The second problem is treated in the \ac{GA} setting: Let $\mathcal{D}$ be a dataset as above and $Q$ a set of polynomials.
-For each polynomial $p\in Q$ let $\mathcal{P}_p$ be the set of monomials in $p$ (without the coefficients) and compute the fitness
-%
-\begin{eqnarray*}
-\phi_p &=& \min_\Theta J_{\textrm{fit}}\left(\Theta;\mathcal{P}_p,\mathcal{D}\right)
-\end{eqnarray*}
-%
-by solving the first problem.
-With a fitness of every instance, a \ac{GA} will apply genetic operators (usually mutation and crossover) to evolve the population $Q$ until a reasonable approximation of a local minimum is found.
-%
-\begin{algorithm}[tb]
-\begin{algorithmic}
-\Function{GAPoly}{$D,pop_0,\epsilon$}
-	\State $pop \gets pop_0;\: err \gets 1.0+\epsilon$
-	\While{$err > \epsilon$}
-		\State $pop \gets \Call{IterateGA}{pop}$
-		\State $pop \gets \Call{Sort}{pop, key = \textsc{Lm.rmse}}$
-		\State $err \gets \textsc{rmse}\left(\Call{First}{pop}\right)$
-	\EndWhile
-	\State\Return $\Call{First}{pop}$
-\EndFunction
-\end{algorithmic}
-\caption{\ac{GAPoly} uses linear regression to find monomial coefficients that minimize the \ac{rmse} over a dataset and \acp{GA} to explore the space of polynomials.
-At exit the \ac{rmse} of the fittest instance is bounded by $\epsilon$.}\label{alg:gapoly}
-\end{algorithm} 
-%
-Notice that the properties of \acp{GA} and linear regression entail that
-the composition of \acp{GA} with linear regression, as defined in Algorithm \ref{alg:gapoly}, converges to a polynomial that is a local minimum of the fitness function, encapsulated in the \ac{rmse} function $J_{\textrm{fit}}$.
-%
-
-%
-Subsection \ref{subs:polynomial.encoding} describes the encoding of individual polynomial instances as chromosomes and other parameters of the utilized \ac{GA} implementation.
-The regularization of the cost function is discussed in subsection \ref{subs:cost.function}.
-
-%
-\subsection{Polynomial Encoding}\label{subs:polynomial.encoding}
-%
-The encoding for any polynomial will be as follows:
-\begin{enumerate}
-\item an initial segment detailing which monomials are active (the first monomial is always active), this is represented in unary description, i.e., each monomial is identified by a single bit
-\item the remaining bits are split into $k$ sets of equal size, each one representing a monomial
-\item each monomial is split into $m$ sets of $d$ size each, \emph{i.e.}, a variable
-\item for each variable, the remaining bits are the binary description of the variable degree, \emph{i.e.}, the maximum exponent is given by $2^d-1$
-\end{enumerate}
-
-Let's see an example: consider polynomial $x_1^3 x_3 + x_3^7 + x_1 x_2$ with $k = 4, m = 3$ and $d=3$. One possible encoding would be:
-\begin{verbatim}
-110 - 011,000,001 ; 000,000,111 ;
-      001,001,000 ; 110,010,101
-\end{verbatim}
-
-(for reading purposes the semicolons separate monomials, the commas separate variables)
-
-The first three bits inform that the second and third monomials are active while the fourth is not (as said, the first monomial is always active).
-This last monomial does not enter neither in the polynomial regression nor in the \ac{rmse} (fitness) evaluation.
-However, it acts as a kind of junk DNA, becoming active when, in a future mutation or crossover, the third bit of the entire sequence flips from~0 to~1.
-
-Let's interpret the first monomial description, \verb!011,000,001!.
-It is divided by three since $m=3$.
-The first triple \verb!011!
-is the binary description of the exponent of variable $x_1$ which is 3, so the first monomial includes $x_1^3$.
-The second triple, \verb!000!, means that $x_2$ is not part of the monomial.
-The third triple \verb!001!
-says that variable $x_3$ has exponent $1$.
-The first monomial is~$x_1^3 x_3$.
-
-Notice that all binary descriptions give rise to valid polynomials.
-However this is not a bijective mapping.
-For each polynomial there are multiple representations.
-For example, $x_1+x_2$ and $x_2+x_1$ have different representations.
-The authors considered that more complex mappings in order to force a one to one mapping would impact negatively in the algorithm's performance.
-
-% If the coding consists of entirely zeros, by convention, it describes polynomial $x_1$.
-This has to do with the execution of the polynomial regression that would fail if we interpret it as the zero polynomial.
-Anyway, for progressive larger  binary descriptions, the chances of getting this zero description decrease exponentially, so it does not impact in any meaningful way in the algorithm's performance.
-
-\subsection{Cost Function}\label{subs:cost.function}
-
-The polynomial regression \ac{rmse} considered so far is based on the ability to predict the test set after the polynomial regression has found the appropriate coefficients $\theta_i$ for each one of the monomials $q_i$.
-
-This \ac{rmse} function tends to prefer more complex polynomials, namely in the number of monomials which provides the regression algorithm for more fitting possibilities.
-One way to balance this is to provide a regularization term into the \ac{rmse} function.
-Our proposal is to include a multiplicative factor proportional to the number of monomials.
-Thus, the error from equation \ref{eq:rmse} defines
-\begin{eqnarray}
-J_{\textrm{reg}}\left(\Theta;\mathcal{P},\mathcal{D}\right) &=& \lambda^{k} J_{\textrm{fit}}\left(\Theta;\mathcal{P},\mathcal{D}\right)\label{eq:rmse-reg}
-\end{eqnarray}
-%
-where $k$ is the number of monomials in the polynomial. When $\lambda > 1$ polynomials with more monomials are penalized.
-
-Somewhat unexpectedly after some experiences it was found that lower values for $\lambda$ provide better, even if marginal, results.
-Figure~\ref{Abalone_dataset_lambdas} shows regularization results for the Abalone and Auto MPG datasets with ten runs for each $\lambda$.
-The following section  includes information about these datasets.
-
-%
-\begin{figure*}[tb]\begin{center}
-\begin{tabular}{cc}
-\includegraphics[width=0.49\textwidth]{figure_1a.pdf}
-%
-% RIGHT PANE
-&
-%
-%
-\includegraphics[width=0.49\textwidth]{figure_1b.pdf}
-\end{tabular}
-
-\caption{Error distribution by regularization exponent for the Abalone and Auto MPG datasets.
-The box plots summarize the error values of 10 runs for each value of $\lambda$.
-The smallest overall error is shown in gray and, in both cases, was achieved with $\lambda=0.8$.}
-\label{Abalone_dataset_lambdas}\label{Auto-MPG_dataset_lambdas}
-
-\end{center}\end{figure*}
-
-The typical inflection point lies around $\lambda = 0.8$. The dataset results for applying the proposed regression method use the regularization parameter with this value.
-
-\subsection{Genetic Operators}
-
-To perform the genetic algorithm it was used the R package genalg~\citep{willighage2012genalg}.
-The operators were the standard ones: (a)~crossover, \emph{i.e.}, a pair of solutions from the previous generations are combined by splitting and mixing their respective representations, and (b)~mutation, changing the values of single bits; the mutation chance applied in the datasets was 5\%.
-There was also elitism between generations, \emph{i.e.}, 20\% of the best solutions survive to the next generation.
-
-\section{Experimental Results}
-
-The results were found using \texttt{R} programming language~\citep{R}\footnote{The datasets and the \texttt{R} code used to produce the results and plots in this paper are available online at \url{https://github.com/jpneto/GenAlgPoly}.}.
-To compare this paper's proposed algorithm we applied the exact same train and test samples using several well-known learning algorithms for regression, namely: Linear Regression, Support Vector Machines~\citep{Meyer12}, Regression Trees~\citep{Therneau13}, Conditional Inference Trees~\citep{Hothorn06, Strobl07, Strobl08} and Random Forests~\citep{Liaw02}
-
-In order to train and test the performance of \ac{GAPoly} several mainstream datasets were used.
-For each dataset, we selected 70\% for training purposes and the remaining observations to make a test set in order to compute the estimated \ac{rmse}.
-To achieve more robust results, each dataset were processed $25$~times, each one with different samples for the train and test sets.
-For the datasets with attribute values of different magnitudes, a preliminary scaling was executed.
-The results below are box plots for the test set error predictions over these different runs.
-
-\begin{figure}[tb]\begin{center}
-\includegraphics[width=0.48\textwidth]{figure_2.pdf}
-
-\caption{Results for Artificial dataset.
-As shown, \ac{GAPoly} was able to find the exact polynomial structure that generates the dataset, reducing the test set prediction error to zero.
-The regression methods depicted are: 1.
-\ac{GAPoly}, 2.
-Random Forests, 3.
-Linear Regression, 4.
-SVM, 5.
-Regression Trees and 6.
-Conditional Inference Trees}
-\label{artificial_dataset1_lambda1.0}
-
-\end{center}\end{figure}
-
-\begin{figure*}[tb]\begin{center}
-\begin{tabular}{cc}
-\includegraphics[width=0.48\textwidth]{figure_3a.pdf}
-%
-&
-%
-\includegraphics[width=0.48\textwidth]{figure_3b.pdf}
-%
-\\
-%
-\includegraphics[width=0.48\textwidth]{figure_3c.pdf}
-%
-&
-%
-\includegraphics[width=0.48\textwidth]{figure_3d.pdf}
-%
-\end{tabular}
-
-\caption{Summary results for different regression methods on the Housing, Abalone, Auto MPG and Kinematics datasets.
+Summary results for different regression methods on the Housing, Abalone, Auto MPG and Kinematics datasets.
 Random Forests, an ensemble method (``2'' in the plots), has the better results.
@@
 Regression Trees and 6.
-Conditional Inference Trees}
-\label{Housing_dataset_lambda0.8_25runs}
-\label{Abalone_dataset_lambda0.8_25runs}
-\label{Auto-Mpg_dataset_lambda0.8_25runs}
-\label{Kinematics300_lambda0.8_25runs}
-\end{center}\end{figure*}
-
-\begin{description}
-\item{\textsc{Artificial}} this is an artificial dataset with four numeric features, $x_1, \ldots x_4$, where $x_1,x_3$ are outcomes from Poisson random variables, and $x_2,x_4$ from Normal random variables.
-The dependent variable~$y$ is given by expression $x_2x_4^2 + x_1^2x_3 + 5$.
-The dataset includes $n=50$~observations.
-
-This dataset was used in order to verify if \ac{GAPoly} was able to find the polynomial relation, which the algorithm did (cf.
-figure~\ref{artificial_dataset1_lambda1.0}).
-The genetic algorithm run with a population of $n=100$~solutions with $50$~iterations for each run.
-
-In the next datasets, the population of the genetic algorithm had size~$n=250$ with 100~iterations for each run.
-
-\item{\textsc{Housing}}: This data set concerns the task of predicting housing values in areas of Boston.
-There are $m=13$ continuous attributes and the dependent variable is the median value of owner-occupied homes in \$1000's.
-There are $n=506$ observations.
-
-Just as an example of the model the \ac{GAPoly} algorithm outputs: in this dataset the best polynomial,
-
-$$y = -0.12 x_6 x_9^4 + 0.78 x_6 - 0.4 x_9^2 x_{13} + 0.17 x_{13}^2 - 0.044$$
-
-where the attributes mean:
-$x_6$, RM average number of rooms per dwelling;
-$x_9$, RAD index of accessibility to radial highways;
-$x_{13}$, Lower status of the population.
-
-For comparison, if we access the mean decrease in accuracy found by Random Forests, the most important attributes are --- in decreasing order --- $x_{13}, x_6, x_5$ (but $x_5$ is already considered $4$ times less important than $x_6$).
-Both algorithms agree in two of their three most important attributes.
-
-\item{\textsc{Abalone}} This dataset can be used to predict the age of a abalone shell using the given $m=8$ numeric attributes concerning several physical measurements.
-There are $n=4177$ observations.
-
-% best polynomial, rsme 0.65
-% $$y = 0.52 s_1^2 x_5 - 0.13 x_2 x_8 -0.75 x_1^2 x_6 + 0.78 x_8 + 0.18 $$
-
-\item{\textsc{Auto MPG}} This dataset is used to predict fuel consumption in miles per gallon, based on two discrete and five continuous attributes ($m=7$). There are $n=398$ observations.
-
-% best polynomial, \ac{rmse} 0.32
-%$$ y = 0.4 x_6 + 0.02 x_5^3 x_6 - 0.85 x_4 + 0.23 x_4^2 - 0.26 $$
-
-\item{\textsc{Kinematics}} This dataset is concerned with the realistic simulation of the forward kinematics of an 8 link robot arm.
-The task is to predict the distance of the end-effector from a target using $m=8$ continuous attributes.
-There are $n=8192$ observations.
-
-% best polynomial, \ac{rmse} 0.74
-% $$ y = 0.24 x_5 - 0.23 x_5 x_6 x_7 - 0.19 x_4 x_7 - 0.5 x_3 $$
-
-\end{description}
-
-\subsection{Convergence speed}
-
-The \ac{GA} quickly proceeds in the first~50 to~100 generations to reasonable error rates.
-Then, it proceeds slower achieving best solutions with marginal error reduction.
-Since the entire learning process takes some time, in the current R implementation, placing a limit between 50 to 100 generations already achieves good results, relative to higher iteration values.
-Figure~\ref{Abalone_fitnessProgress} shows a typical error evolution for the dataset Abalone given two different values for $\lambda$.
-
-\begin{figure*}[tb]
-\begin{center}
-\includegraphics[width=0.98\textwidth]{figure_4.pdf}
-\caption{Error progress for Abalone dataset during a single execution of the genetic algorithm.
-The figure shows the fitness evolution for two different regularization values.
-The population for both consisted of $200$ polynomials.
-The error values seem to stabilize around iteration~250.}
-\label{Abalone_fitnessProgress}
-
-\end{center}
-\end{figure*}
-
-\section{Conclusion and Future Work}
-
-The proposed method has competitive results comparing with some well-known regression methods.
-Only Random Forest outperforms \ac{GAPoly} systematically (it also outperforms all the other regression algorithms).
-One exception --- besides the artificial dataset that uses a straightforward polynomial relation --- is the Auto MPG dataset where \ac{GAPoly} has comparable results.
-This is evidence that applying standard genetic operators for polynomial model searching is a viable tool for regression purposes.
-
-For complexity considerations \ac{GAPoly} demands some processing time.
-On a current quad-core computer, processing the Kinematics dataset (with 8k~observations) takes approximately 5~minutes.
-The processing time can probably be speeded by one to two orders in magnitude if the process is implemented in a low level programming language like \texttt{C++}.
-However, speed optimization was not the focus of this article.
-
-A cross-validation procedure can be implemented to refine the appropriate parameter values to achieve better errors.
-Namely, the regularization parameter, $\lambda$, can be tested with several values, instead of being fixed at $0.8$.
-Other parameters like mutation chance or the amount of elitism could also be tested.
-However, these type of tests need a low-level, fast implementation of \ac{GAPoly}.
-
-\section*{Acknowledgements}
-
-The authors are grateful to the Fundação para a Ciência e Tecnologia (FCT) and the  R\&D laboratory LabMAg for the financial support given to this work, under the strategic project \textsc{PEst-OE/EEI/UI0434/2011}.
-
-The datasets used herein were selected from Luís Torgo's data  repository, \url{http://www.dcc.fc.up.pt/~ltorgo/Regression/DataSets.html}.
-Most come originally from UCI ML repository, \url{http://archive.ics.uci.edu/ml/}.
-
-The authors wish to thank professor André Falcão for motivation and useful discussions around the article.
-
-%\section{Bibliography}
-
-\bibliographystyle{model2-names}
-\bibliography{fullbib}
-
-\end{document}
+Conditional Inference Trees

### 2015-07-24 texts/submission_03_asc/eprr.tex
Início Revisão ASC.

--- 2013-12-24 texts/EPRR_FILES-ASC/eprr.tex
@@
 %% Use the option review to obtain double line spacing
-\documentclass[times,review,preprint]{elsarticle}
+\documentclass[review,preprint]{elsarticle}
 
@@
 \usepackage{graphicx}
+\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
+\usepackage{amssymb}
+\usepackage{latexsym}
 %
@@
 \newcommand{\nth}{\ensuremath{^{th}}}
+\newcommand{\review}[1]{\textcolor{blue}{~---~#1}}
 %
@@
 \journal{Applied Soft Computing}
+\bibliographystyle{elsarticle-num}
 
 \begin{document}
-
+%
 %
@@
 \acrodef{SVM}[SVM]{Support Vector Machines}
-
+%
 \setcounter{page}{0}
-
+%
 \begin{frontmatter}
-
-%% Title, authors and addresses
-
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for the associated footnote;
-%% use the fnref command within \author or \address for footnotes;
-%% use the fntext command for the associated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for the associated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-%%
-%% \title{Title\tnoteref{label1}}
-%% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \address{Address\fnref{label3}}
-%% \fntext[label3]{}
-
+%
 \title{A method for regularization of evolutionary polynomial regressions}
-
+%
 %% use optional labels to link authors explicitly to addresses:
@@
 %% \address[label2]{<address>}
-
-\author[1]{Francisco Coelho\corref{cor1}} 
-\cortext[cor1]{Corresponding author: 
-  Tel.: +351-919-006-379}
+%
+\author{Francisco Coelho\corref{corfc}} 
+\cortext[corfc]{Corresponding author: Tel.: +351-919-006-379}
 \ead{fc@di.uevora.pt}
-\author[2]{Jo\~{a}o Pedro Neto}
-
-\address[1]{Departamento de Informática, Universidade de Évora, Rua Romão Ramalho 58, 7000-671 Évora, Portugal}
-\address[2]{Departamento de Informática, Faculdade de Ciências da Universidade de Lisboa, Campo Grande, 1749-016 Lisboa, Portugal}
-
+\address{Departamento de Inform\'{a}tica, Universidade de \'{E}vora, Rua Rom\~{a}o Ramalho 58, 7000-671 \'{E}vora, Portugal}
+%
+\author{Jo\~{a}o Pedro Neto}
+\ead{jpn@di.fc.ul.pt}
+\address{Departamento de Inform\'{a}tica, Faculdade de Ci\^{e}ncias da Universidade de Lisboa, Campo Grande, 1749-016 Lisboa, Portugal}
+%
 \begin{abstract}
@@
 \end{abstract}
+%
+\begin{keyword}
+evolutionary polynomial regression \sep regularization \sep feature extraction
+\end{keyword}
+%\MSC 41A05\sep 41A10\sep 65D05\sep 65D17
+%\KWD
+%
+\end{frontmatter}
+%
+%
+\review{ GAs are repeatedly mentioned as an approach to search for locally optimal parameters, this is simply incorrect and is a very surprising statement.
+GAs are NOT a local optimizer in any shape or form.}
 
-\begin{keyword}
-evolutionary polynomial regression \sep regularization \sep feature extraction \end{keyword}
+\review{What is the reference for the GA encoding used?}
 
-\end{frontmatter}
+\review{The penalty term introduces is the main contribution of this work, this should be clearly stated in the text.
+It is basically a complexity based penalty which is really not so novel, there are even several cases in literature of penalty based GA functions.
+Please dig up further in the state of the art to back up better your claim of novelty.
+}
 
+\review{In some box plots only ten simulations are performed, in others 25.  These numbers are too low.}
+%
+%
 \linenumbers
@@
 A special case of polynomial transformation, the \emph{polynomial kernel}, $k\at{x,x'} = \left<x,x'\right>^d$ is commonly used in regression and classification tasks with \acp{SVM}.
-However general polynomial transformations do not verify the kernel trick.
+However general polynomial transformations do not verify the kernel trick.\review{do not verify does not seem like the proper term.}
 
@@
 Related research abound and spans from the 1950s seminal work of Nils Aall Barricelli \citep{Barricelli:1962aa} in the Institute for Advanced Study of Princeton to today's principal area of study for thousands of researchers, covered in hundreds of conferences, workshops and other meetings.
-Perhaps the key impulse to \acp{GA} come from John Holland's work and his book ``Adaptation in Natural and Artificial Systems'' \citep{Holland:1975aa}.
+Perhaps the key impulse to \acp{GA} came from John Holland's work and his book ``Adaptation in Natural and Artificial Systems'' \citep{Holland:1975aa}.
 
@@
 This paper describes an extension of the general \ac{EPR} method to find a regularized polynomial regression of a given dataset.
-The optimal regression results from a cost function that accounts for both the \ac{rmse} and a regularization factor to avoid overfit by penalysing polynomial complexity.
+The optimal regression results come from a cost function that accounts for both the \ac{rmse} and a regularization factor to avoid overfit by penalysing polynomial complexity.
 
@@
 
-An usual representation of polynomials is through expressions of the form
+A usual representation of polynomials is through expressions of the form
 $$
@@
 %
-\begin{eqnarray*}
+\begin{eqnarray}
 h_{\Theta,\mathcal{Q}}\left(x_1,\ldots,x_m\right) &=& \sum_{j = 1}^k \theta_j q_j|_{X_i=x_i,\forall 1 \leq i \leq m}
-\end{eqnarray*}
+\end{eqnarray}
 %
@@
 An alternative formulation would just replace $\mathcal{D}$ by $\mathcal{Q}\left( \mathcal{D} \right)$.
-It turns out that the first formulation is a special case of the second (by including the variables in the monomial set) and has better \ac{rmse} performance --- what is not surprising because it uses more features.
+It turns out that the first formulation is a special case of the second (by including the variables in the monomial set) and has better \ac{rmse} performance --- what is not surprising because it uses more features.\review{How is this known???
+Is there a reference or some experimental results backing this claim or is it simply intuition???}
 
 The second problem is treated in the \ac{GA} setting: Let $\mathcal{D}$ be a dataset as above and $\mathcal{P}$ a set of polynomials.
-For each polynomial $p\in \mathcal{P}$ let $\mathcal{Q}_p$ be the set of monomials in $p$ (without the coefficients) and define the (anti) fitness
+For each polynomial $p\in \mathcal{P}$ let $\mathcal{Q}_p$ be the set of monomials in $p$ (without the coefficients) and define the (anti) fitness\review{say what?
+Do you mean something like a minimization based fitness??
+Anti fitness seems odd.}
 %
-\begin{eqnarray*}
+\begin{eqnarray}
 \phi_p &=& \min_\Theta J_{\textrm{fit}}\left(\Theta;\mathcal{Q}_p,\mathcal{D}\right)
-\end{eqnarray*}
+\end{eqnarray}
 %
@@
 %
-The properties of \acp{GA} and linear regression entail that Algorithm \ref{alg:gapoly} converges to a polynomial that is a local minimum of the fitness function, encapsulated in the \ac{rmse} function $J_{\textrm{fit}}$.
+The properties of \acp{GA} and linear regression entail that Algorithm \ref{alg:gapoly} converges to a polynomial that is a local minimum of the fitness function\review{GAs are global optimization techniques (in spite of what wikipedia states!)}, encapsulated in the \ac{rmse} function $J_{\textrm{fit}}$.
 %
@@
 The specific encoding (representation) of a set of monomials is an important aspect in the implementation of \ac{EPR}.
-The choice described below permits active and inactive monomials for regression purposes.
+The choice described below permits active and inactive monomials for regression purposes\review{is this an original contribution????
+Otherwise it needs a reference.}.
 The active (or inactive) state of a monomial might change through mutation or crossover.
@@
 While each valid encoding represents a set of monomials the map is not bijective: each set of monomials has multiple encodings, for example by changing $d$ or the order of monomial segments.
-However, considering the \ac{EPR} task, this is a minor issue and a bijective map would add computational complexity and negative impact to the algorithm's performance.
+However, considering the \ac{EPR} task, this is a minor issue and a bijective map would add computational complexity and negative impact to the algorithm's performance\review{The encoding example should include the inactive term so that both examples are consistent (lines 113-115).
+Is there a reference some experimentation to back this up?}.
 
@@
 A simple exploration on the effect of the regularization parameter is depicted in Figure~\ref{fig:lambda.error-distribution} where it is possible to observe that the typical inflection point lies around $\lambda = 0.8$.
-This value, favoring ``larger'' polynomials is justified by the balance of the data's non-linearity and polynomial complexity: below $\lambda = 0.8$, even penalized, larger monomial sets achieve better \ac{rmse} performance than smaller ones while above that value the size of the monomial set is excessive.
+This value, favoring ``larger'' polynomials is justified by the balance of the data's non-linearity and polynomial complexity: below $\lambda = 0.8$, even penalized, larger monomial sets achieve better \ac{rmse} performance than smaller ones while above that value the size of the monomial set is excessive\review{please clarify this paragraph and justify this claim}.
 Within this tension the overall error results reduced when compared to the non-regularized \ac{EPR} version.
@@
 \subsection{Genetic Algorithm Parameterization}
+
+\review{Please indicate all these values in detail}
+
+\review{You should contrast obtained results in a table.  Results are hard to judge from Fig 3 alone.}
+
+\review{The number of 250 for iterations seems premature specially when one looks at Fig 4. Stabilize? Really?  This only seems valid in one case of lambda.}
 
@@
 
-\bibliographystyle{elsarticle-num}
+\bibliographystyle{plain}
 \bibliography{fullbib2}

### 2015-08-05 runAllDatasets.R
Including values for lambda=0.9
merging result files into single files

@@ -35,13 +35,13 @@ read_clean <- function(dataset) {
 }
 
 pop  <- 300  # GA population
-reg  <- 0.7  # amount of regularization
+reg  <- 0.9  # amount of regularization
 runs <- 75   # number of simulations
 iter <- 100  # number of interations for each simulation
 
 digits <- 5 # round errors after number of digits
 
-##################
+################## started Monday 13:43 
 
 datasets <- c("Abalone", "Auto-Mpg", "Housing", "Kinematics")
 
@@ -54,16 +54,16 @@ for (dataset in datasets) {
   report_reg <- make.report(my.data, population=pop, iterations=iter, lambda=reg, runs=runs)
   
   # Apply standard GA polynomial regression
-  report_noreg <- make.report(my.data, population=pop, iterations=iter, lambda=1, runs=runs)
+  # report_noreg <- make.report(my.data, population=pop, iterations=iter, lambda=1, runs=runs)
 
   # build and save table of error results  
   df <- data.frame(ga.eprr = report_reg$ga.error,
-                   ga.epr  = report_noreg$ga.error, 
-                   lin.reg = report_noreg$lm.error,
-                   svm     = report_noreg$svm.error,
-                   rpart   = report_noreg$rpart.error,
-                   rf      = report_noreg$rf.error,
-                   ci.tree = report_noreg$citree.error)
+                   #ga.epr  = report_noreg$ga.error, 
+                   lin.reg = report_reg$lm.error,
+                   svm     = report_reg$svm.error,
+                   rpart   = report_reg$rpart.error,
+                   rf      = report_reg$rf.error,
+                   ci.tree = report_reg$citree.error)
   
   df <- round(df, digits)
   

### 2015-08-06 classDataset.R
Added more simulation values

@@ -127,12 +127,22 @@ library(plyr)
 lambda.values <- seq(0.7,1.3,by=0.1)
 n.runs <- 50
 
-errors <- test.lambda(my.data, lambda.values, runs=n.runs, population=50, iterations=60)
+#datasets <- c("Abalone", "Auto-Mpg", "Housing", "Kinematics")
+datasets <- c("Housing", "Kinematics")
 
-df <- data.frame(error  = as.vector(matrix(errors,nrow=1)),
-                 lambda = as.factor(rep(round(lambda.values,3), each=n.runs)))
+for (dataset in datasets) {
+  # print(paste("### Starting", dataset, "dataset..."))
+  
+  my.data <- read_clean(dataset)
+  
+  errors <- test.lambda(my.data, lambda.values, runs=n.runs, population=50, iterations=60)
+  
+  df <- data.frame(error  = as.vector(matrix(errors,nrow=1)),
+                   lambda = as.factor(rep(round(lambda.values,3), each=n.runs)))
+  
+  write.table(df, paste0(dataset,"_lambda.errors.txt"))
+}
 
-write.table(df, paste0(dataset,"_lambda.errors.txt"))
 
 # to read the previous table from file:
 # df <- read.table(paste0("results/", dataset,"_lambda.errors.txt"))

### 2015-08-06 runAllDatasets.R

@@ -35,7 +35,7 @@ read_clean <- function(dataset) {
 }
 
 pop  <- 300  # GA population
-reg  <- 0.9  # amount of regularization
+reg  <- 0.7  # amount of regularization
 runs <- 75   # number of simulations
 iter <- 100  # number of interations for each simulation
 
@@ -43,7 +43,8 @@ digits <- 5 # round errors after number of digits
 
 ################## started Monday 13:43 
 
-datasets <- c("Abalone", "Auto-Mpg", "Housing", "Kinematics")
+#datasets <- c("Abalone", "Auto-Mpg", "Housing", "Kinematics")
+datasets <- c("Auto-Mpg", "Abalone")
 
 for (dataset in datasets) {
   # print(paste("### Starting", dataset, "dataset..."))

### 2015-08-06 classDataset.R
update on classDataset.R

@@ -91,6 +91,9 @@ df <- data.frame(ga.poly = report_noreg$ga.error,
 
 write.table(df, paste0(dataset, pop, "_noReg.txt"))
 
+# df <- read.table(paste0("results/", dataset, pop, "_results.txt"), header = TRUE)
+# apply(df, 2, summary)
+
 # show results
 
 size <- length(report$ga.error)

### 2015-08-07 texts/submission_03_asc/eprr.tex
Mais uma revisão

@@
 \newcommand{\review}[1]{\textcolor{blue}{~---~#1}}
+\newcommand{\revised}[2]{\review{#1}~\textcolor{red}{#2}}
 %
@@
 %
-\review{ GAs are repeatedly mentioned as an approach to search for locally optimal parameters, this is simply incorrect and is a very surprising statement.
-GAs are NOT a local optimizer in any shape or form.}
+\revised{ GAs are repeatedly mentioned as an approach to search for locally optimal parameters, this is simply incorrect and is a very surprising statement.
+GAs are NOT a local optimizer in any shape or form.}{Our intention was merely to note that GAs give no guarantee of finding a global optimum}
 
-\review{What is the reference for the GA encoding used?}
+\revised{What is the reference for the GA encoding used?}{It is ours.}
 
-\review{The penalty term introduces is the main contribution of this work, this should be clearly stated in the text.
+\revised{The penalty term introduces is the main contribution of this work, this should be clearly stated in the text.
 It is basically a complexity based penalty which is really not so novel, there are even several cases in literature of penalty based GA functions.
-Please dig up further in the state of the art to back up better your claim of novelty.
-}
+Please dig up further in the state of the art to back up better your claim of novelty.}{We included references to previous regularization in GA and outlined the role and originality of our contribution.}
 
-\review{In some box plots only ten simulations are performed, in others 25.  These numbers are too low.}
+\revised{In some box plots only ten simulations are performed, in others 25.  These numbers are too low.}{The number of samples was increased.}
 %
@@
 A special case of polynomial transformation, the \emph{polynomial kernel}, $k\at{x,x'} = \left<x,x'\right>^d$ is commonly used in regression and classification tasks with \acp{SVM}.
-However general polynomial transformations do not verify the kernel trick.\review{do not verify does not seem like the proper term.}
+However the kernel trick doesn't apply to general polynomial transformations.
 
 Polynomials, one of the most studied subjects in mathematics, generalize li\-ne\-ar functions and define, perhaps, the simplest and most used nonlinear models.
+For example, Polynomial Neural Networks \cite{dehuri2011condensed} generalize the linear component of neural networks with a polynomial function.
 Applications include colorimetric calibration \citep{Mendes:2005aa}, explicit formul\ae\ for turbulent pipe flows
@@
 Evolutionary algorithms, including \ac{GA}, were, arguably, one of the hottest topics of research in the recent decades and with good reason since they outline an optimization scheme easy to conceptualize and with very broad application.
-If a nonlinear (or otherwise) model requires parameterization, \acp{GA} provide a simple and often effective approach to search for locally optimal parameters.
+If a nonlinear (or otherwise) model requires parameterization, \acp{GA} provide a simple and often effective approach to search for global optimal parameters (although no guarantee of global optimality can be given).
 Related research abound and spans from the 1950s seminal work of Nils Aall Barricelli \citep{Barricelli:1962aa} in the Institute for Advanced Study of Princeton to today's principal area of study for thousands of researchers, covered in hundreds of conferences, workshops and other meetings.
@@
 
+ Use of regularization/penalty functions is common practice in machine learning in general and has some applications in \ac{GA}\cite{Gupta2009275}.
 This paper describes an extension of the general \ac{EPR} method to find a regularized polynomial regression of a given dataset.
-The optimal regression results come from a cost function that accounts for both the \ac{rmse} and a regularization factor to avoid overfit by penalysing polynomial complexity.
+Herein optimal regression results from a cost function that accounts for both the \ac{rmse} together with a novel regularization factor that penalizes over-fitting by polynomial complexity.
 
@@
 An alternative formulation would just replace $\mathcal{D}$ by $\mathcal{Q}\left( \mathcal{D} \right)$.
-It turns out that the first formulation is a special case of the second (by including the variables in the monomial set) and has better \ac{rmse} performance --- what is not surprising because it uses more features.\review{How is this known???
-Is there a reference or some experimental results backing this claim or is it simply intuition???}
+It turns out that the first formulation is a special case of the second (by including the variables in the monomial set) and has the potential for better \ac{rmse} performance because it uses more features.\revised{How is this known???
+Is there a reference or some experimental results backing this claim or is it simply intuition???}{We reformulated the previous sentence.}
 
 The second problem is treated in the \ac{GA} setting: Let $\mathcal{D}$ be a dataset as above and $\mathcal{P}$ a set of polynomials.
-For each polynomial $p\in \mathcal{P}$ let $\mathcal{Q}_p$ be the set of monomials in $p$ (without the coefficients) and define the (anti) fitness\review{say what?
+For each polynomial $p\in \mathcal{P}$ let $\mathcal{Q}_p$ be the set of monomials in $p$ (without the coefficients) and define the minimization based fitness\revised{say what?
 Do you mean something like a minimization based fitness??
-Anti fitness seems odd.}
+Anti fitness seems odd.}{Adopted the reviewer suggestion.}
 %
@@
 by solving the first problem.
-With a fitness of every instance, the \ac{GA} genetic operators (usually mutation and crossover) evolve the population $\mathcal{P}$ until a reasonable approximation of a local minimum is found.
+With a fitness of every instance, the \ac{GA} genetic operators (usually mutation and crossover) evolve the population $\mathcal{P}$ until a reasonable approximation of a minimum is found.
 %
@@
 %
-The properties of \acp{GA} and linear regression entail that Algorithm \ref{alg:gapoly} converges to a polynomial that is a local minimum of the fitness function\review{GAs are global optimization techniques (in spite of what wikipedia states!)}, encapsulated in the \ac{rmse} function $J_{\textrm{fit}}$.
+The properties of \acp{GA} and linear regression entail that Algorithm \ref{alg:gapoly} converges to a polynomial that is a minimum of the fitness function, encapsulated in the \ac{rmse} function $J_{\textrm{fit}}$.
 %
@@
 The specific encoding (representation) of a set of monomials is an important aspect in the implementation of \ac{EPR}.
-The choice described below permits active and inactive monomials for regression purposes\review{is this an original contribution????
-Otherwise it needs a reference.}.
+The choice described below, developed by the authors for this algorithm, permits active and inactive monomials for regression purposes\revised{is this an original contribution????
+Otherwise it needs a reference.}{it is ours}.
 The active (or inactive) state of a monomial might change through mutation or crossover.
@@
 0 & 0 & 7 \\
-1 & 1 & 0
+1 & 1 & 0 \\
+6 & 2 & 5
 \end{array}
@@
 000 & 000 & 111 \\
-001 & 001 & 000
+001 & 001 & 000 \\
+110 & 010 & 101
 \end{array}
@@
 $$
-where the right matrix is in binary form using $d = 3$ bits. An encoding of this set of monomials with the extra monomial $x_1^6x_2^2x_3^5$ inactive, setting $k = 4$, would be
+where the right matrix is in binary form using $d = 3$ bits. An example of this set of monomials with the forth monomial, $x_1^6x_2^2x_3^5$, inactive would be
 \begin{center}
@@
 While each valid encoding represents a set of monomials the map is not bijective: each set of monomials has multiple encodings, for example by changing $d$ or the order of monomial segments.
-However, considering the \ac{EPR} task, this is a minor issue and a bijective map would add computational complexity and negative impact to the algorithm's performance\review{The encoding example should include the inactive term so that both examples are consistent (lines 113-115).
-Is there a reference some experimentation to back this up?}.
+However, considering the \ac{EPR} task, this is a minor issue and a bijective map would add computational complexity and negative impact to the algorithm's performance\revised{The encoding example should include the inactive term so that both examples are consistent (lines 113-115).
+Is there a reference some experimentation to back this up?}{We changed the text according to the suggestion.}.
 
@@
 \begin{tabular}{cc}
-\includegraphics[width=0.49\textwidth]{figure_1a.pdf}
-%
-% RIGHT PANE
+\includegraphics[width=0.48\textwidth]{Fig1a.pdf}
 &
-%
-%
-\includegraphics[width=0.49\textwidth]{figure_1b.pdf}
+\includegraphics[width=0.48\textwidth]{Fig1b.pdf}
+\\
+\includegraphics[width=0.48\textwidth]{Fig1c.pdf}
+&
+\includegraphics[width=0.48\textwidth]{Fig1d.pdf}
 \end{tabular}
 
-\caption{Error distribution by regularization exponent for two common datasets.
-The box plots summarize error values of ten simulations for each value of $\lambda$.
-The smallest overall error, in grey, is achieved in both datasets when $\lambda=0.8$.
-Performance of the non-regularized \ac{EPR} is plotted in the line $\lambda = 1$.}
+\caption{Error distribution by regularization exponent for common datasets.
+The box plots summarize error values of 75 simulations (60 iterations on a population size of 50) for each value of $\lambda$.
+Performance of the non-regularized \ac{EPR} is plotted in the line $\lambda = 1$.
+Performance for $\lambda < 1$ is better than $\lambda > 1$ in all datasets.}
 \label{fig:lambda.error-distribution}\label{Abalone_dataset_lambdas}\label{Auto-MPG_dataset_lambdas}
@@
 
-A simple exploration on the effect of the regularization parameter is depicted in Figure~\ref{fig:lambda.error-distribution} where it is possible to observe that the typical inflection point lies around $\lambda = 0.8$.
-This value, favoring ``larger'' polynomials is justified by the balance of the data's non-linearity and polynomial complexity: below $\lambda = 0.8$, even penalized, larger monomial sets achieve better \ac{rmse} performance than smaller ones while above that value the size of the monomial set is excessive\review{please clarify this paragraph and justify this claim}.
-Within this tension the overall error results reduced when compared to the non-regularized \ac{EPR} version.
+A simple exploration on the effect of the regularization parameter is depicted in Figure~\ref{fig:lambda.error-distribution} where it is possible to observe that penalizing polynomial complexity (i.e.
+$\lambda < 1$) achieves better results than the opposite ($\lambda > 1$).
+This observation motivates further inquire, done in section~\ref{sec:experimental.results}.\revised{please clarify this paragraph and justify this claim}{we rephrased the statement.}
 
@@
 
-\review{Please indicate all these values in detail}
+\revised{Please indicate all these values in detail}{We described the parameterization in more detail.}
 
@@
 
-\review{The number of 250 for iterations seems premature specially when one looks at Fig 4. Stabilize? Really?  This only seems valid in one case of lambda.}
+\revised{The number of 250 for iterations seems premature specially when one looks at Fig 4.
+Stabilize?
+Really?
+This only seems valid in one case of lambda.}{we changed the number of iterations to 2000, way behind stabilization}
 
 In general \acp{GA} offer many possibilities with respect to the choice of genetic operators and respective application rates, population evolution, \emph{etc}.
-The results found here where obtained using the package \texttt{genalg}~\citep{Willighagen:2012aa} with default parameters, standard operators (crossover and mutation) and population evolution with 20\% elitism between generations.
+The results found here where obtained using the package \texttt{genalg}~\citep{Willighagen:2012aa} with standard operators (crossover and mutation) and population evolution defined by mutation rate of 5\% and 20\% elitism between generations.
 
-\section{Experimental Results}
+\section{Experimental Results}\label{sec:experimental.results}
 
@@
 
-The \ac{EPRR} method is ranked against several well-known learning algorithms for regression, namely: non-regularized \ac{EPR}, Linear Regression, Support Vector Machines~\citep{Meyer:2012aa}, Regression Trees~\citep{Therneau:2013aa} and Conditional Inference Trees~\citep{Hothorn:2006aa,Strobl:2007aa,Strobl:2008aa}.
-To achieve better error results the \ac{SVM} and Regression Tree parameters are tuned in each dataset.
+The \ac{EPRR} method is ranked against several well-known learning algorithms for regression, namely: non-regularized \ac{EPR}, Linear Regression, Support Vector Machines~\citep{Meyer:2012aa} with linear kernel, Regression Trees~\citep{Therneau:2013aa}, Random Forest~\citep{Strobl:2008aa,Strobl:2007aa} and Conditional Inference Trees~\citep{Hothorn:2006aa}.
+%
+%	INCLUDE ?
+%
+%To achieve better error results the \ac{SVM} and Regression Tree parameters are tuned in each dataset.
 
@@
 \begin{tabular}{cc}
-\includegraphics[width=0.48\textwidth]{figure_3a.pdf}
+\includegraphics[width=0.48\textwidth]{Fig3a.pdf}
 &
-\includegraphics[width=0.48\textwidth]{figure_3b.pdf}
+\includegraphics[width=0.48\textwidth]{Fig3b.pdf}
 \\
-\includegraphics[width=0.48\textwidth]{figure_3c.pdf}
+\includegraphics[width=0.48\textwidth]{Fig3c.pdf}
 &
-\includegraphics[width=0.48\textwidth]{figure_3d.pdf}
+\includegraphics[width=0.48\textwidth]{Fig3d.pdf}
 \end{tabular}
 %
-\caption{Summary results for different regression methods on diverse datasets.
+\caption{Summary results for different regression methods on common datasets.
 Although \ac{EPRR} not always achieves the smallest expected error, performance is on-par with more sophisticated methods.
-The regression methods depicted in these figures are: 1.
-\ac{EPRR}; 2.
-\ac{EPR}; 3.
-Linear Regression; 4.
-\ac{SVM}; 5.
-Regression Trees; 6.
-Conditional Inference Trees;}
+The regression methods depicted in these figures are:
+1. \ac{EPRR}, $\lambda = 0.7$;
+2. \ac{EPRR}, $\lambda = 0.8$;
+3. \ac{EPRR}, $\lambda = 0.9$;
+4. \ac{EPR} (i.e. $\lambda = 1.0$);
+5. Linear Regression;
+6. \ac{SVM} (linear kernel);
+7. Regression Trees;
+8. Random Forest (with 100 trees);
+9. Conditional Inference Trees.}
 \label{fig:four.datasets.summary}
@@
 \begin{center}
-\includegraphics[width=0.98\textwidth]{figure_4.pdf}
-\caption{Learning curve: Error progress for the \textsc{Abalone} dataset during a single execution of the genetic algorithm.
-The figure shows the fitness evolution for two different regularization values.
-The population for both consists of $200$ polynomials.
-The error values seem to stabilize around iteration~250.}
+\includegraphics[width=0.98\textwidth]{Fig4.pdf}
+\caption{Learning curve: Error progress for the \textsc{Housing} dataset during a single execution of the genetic algorithm.
+The figure shows the fitness evolution for different regularization values.
+The population for both consists of $200$ polynomials.}
 \label{Abalone_fitnessProgress}

### 2015-08-07 runAllDatasets.R

@@ -44,7 +44,7 @@ digits <- 5 # round errors after number of digits
 ################## started Monday 13:43 
 
 #datasets <- c("Abalone", "Auto-Mpg", "Housing", "Kinematics")
-datasets <- c("Auto-Mpg", "Abalone")
+datasets <- c("Housing", "Kinematics")
 
 for (dataset in datasets) {
   # print(paste("### Starting", dataset, "dataset..."))

### 2015-08-07 texts/submission_03_asc/data/gapolyfuncs.R

@@ -1,0 +1,398 @@
+#########################################################################
+# initial values, can be overriden
+max.monoids <- 4  # poly = monoid1 + monoid2 + ...
+max.degree  <- 3  # meaning exponents up to 2^max.degree - 1
+
+#########################################################################
+# Root mean square error
+rsme <- function(v1, v2=c(0)) {
+  sqrt(mean((v1-v2)^2))
+}
+
+#########################################################################
+# convert a binary vector into an integer
+bin2dec <- function(x) 
+  sum(2^(which(rev(unlist(strsplit(as.character(x), "")) == 1))-1))
+
+#########################################################################
+# The chromosome coding will as follow:
+# + an initial segment detailing which monoids are active (the 1st is always active)
+# + the chromosome is split into `max.monoids` sets of bits of equal size 
+#   (a monoid)
+# + each monoid is split into `n.vars` sets of `max.degree` size each 
+#   (a variable)
+# + for each variable, the bits give the binary description of the variable's degree 
+# 
+# Eg: consider polynomial $x_1^3 \times x_3 + x_3^7 + x_1 \times x_2$
+# 
+# One possible coding -- for the values in the previous code snippet -- would be:
+# 
+# $$ 1111 011,000,001;000,000,111;001,001,000;000,000,000 $$
+# 
+# (the semicolons separate monoids, the commas separate variables)
+# 
+# The next two functions transform the binary vector into a polynomial formula 
+# that `lm()` understands:
+
+make.monoid <- function(bits, n.vars, vars) {
+  elements <- split(bits, ceiling(seq_along(bits)/(length(bits)/n.vars)))
+  monoid <- ""
+  for(i in 1:length(elements)) {
+    power <- bin2dec(elements[[i]])
+    if (power == 1)
+      if (monoid=="")
+        monoid = vars[i]
+      else  
+        monoid = paste0(monoid, "*", vars[i])
+    if (power > 1)
+      if (monoid=="")
+        monoid = paste0(vars[i], "^", power)
+      else  
+        monoid = paste0(monoid, "*", vars[i], "^", power)
+  }
+  monoid
+}
+
+make.formula <- function(bits, n.vars, max.monoids) {
+  vars          <- paste0("x",1:n.vars)    # c("x1","x2"...)
+  active.bits   <- 1:(max.monoids-1)       # the initial activation bits
+  active.monoid <- c(1,bits[active.bits])  # get those bits
[…]

### 2015-08-07 texts/submission_03_asc/data/summary.py

@@ -1,0 +1,62 @@
+#!/usr/bin/env python3
+
+import pandas as pd
+
+def boxp(x,m = ""):
+	return "\\boxp{%5.4f}{%5.4f}{%5.4f}{%5.4f}{%5.4f}{white}; %% %s"%(
+		x.min(),
+		x.quantile(0.25),
+		x.quantile(0.50),
+		x.quantile(0.75),
+		x.max(), m)
+
+def figs_1():
+	datafiles = [
+		'Abalone_75_lambda.errors.txt',
+		'Auto-Mpg_75_lambda.errors.txt',
+		'Housing_75_lambda.errors.txt',
+		'Kinematics_75_lambda.errors.txt'
+	]
+
+	for df in datafiles:
+		print(df)
+		d = pd.read_csv( df, sep = ' ' )
+		g = d.groupby('lambda')
+
+		for key,group in g:
+			gk = group['error']
+			print(boxp(gk))
+
+		print()
+
+def figs_3():
+	datasets = [
+		'Abalone',
+		'Auto-Mpg',
+		'Housing',
+		'Kinematics'
+	]
+	colnames = [
+		'ga.eprr.0.7',
+		'ga.eprr.0.8',
+		'ga.eprr.0.9',
+		'ga.epr',
+		'lin.reg',
+		'svm',
+		'rpart',
+		'rf',
+		'ci.tree',
+	]
+	for ds_name in datasets:
+		f_name = "%s300_results.txt"%ds_name
+		ds = pd.read_csv( f_name, sep = ' ')
+		print(f_name)
+		for col in colnames:
+			if not col == 'id':
+				print(boxp(ds[col], col))
+		print('')
+
+
+
+if __name__ == '__main__':
+	figs_3()

### 2015-08-07 classDataset.R
Figure 4 with all values of lambda

@@ -63,7 +63,7 @@ head(my.data,8)
 
 pop  <- 300  # GA population
 reg  <- 0.8  # amount of regularization
-runs <- 25   # number of simulations
+runs <- 75   # number of simulations
 iter <- 100  # number of interations for each simulation
 
 # Apply regularization
@@ -167,35 +167,55 @@ qplot(lambda,error,data=df,geom=c("jitter","boxplot")) +
 ### cf. article's figure "Learning curve"
 ###############################
 
-library(ggplot2)
+### make train & test set
+set.seed(101)
+train.p.size <- 0.7 # percentage of training set
+inTrain   <- sample(1:nrow(my.data), train.p.size * nrow(my.data))
+train.set <- my.data[inTrain,]
+test.set  <- my.data[-inTrain,]
+###
+
+fitness.progress <- c() # needed to reset fitness progress
+lambda0.7 <- 
+  follow.fitness(train.set, test.set, population=200,iterations=1500,mutation.rate=0.15,lambda=0.7)
 
 fitness.progress <- c() # needed to reset fitness progress
 lambda0.8 <- 
-  follow.fitness(my.data,population=200,iterations=20,mutation.rate=0.15,lambda=0.8)
+  follow.fitness(train.set, test.set, population=200,iterations=1500,mutation.rate=0.15,lambda=0.8)
 
 fitness.progress <- c()
-lambda0.975 <- 
-  follow.fitness(my.data,population=200,iterations=20,mutation.rate=0.15,lambda=0.975)
+lambda0.9 <- 
+  follow.fitness(train.set, test.set, population=200,iterations=1500,mutation.rate=0.15,lambda=0.9)
 
 fitness.progress <- c()
-lambda1 <- 
-  follow.fitness(my.data,population=200,iterations=20,mutation.rate=0.15,lambda=1)
+lambda1.0 <- 
+  follow.fitness(train.set, test.set, population=200,iterations=1500,mutation.rate=0.15,lambda=1)
 
-lambda_Values = list(lambda0.8=lambda0.8, lambda0.975=lambda0.975, lambda1=lambda1)
+lambda_Values = list(lambda0.7=lambda0.7, 
+                     lambda0.8=lambda0.8, 
+                     lambda0.9=lambda0.9, 
+                     lambda1.0=lambda1.0)
+
[…]

### 2015-08-07 gapolyfuncs.R

@@ -261,7 +261,7 @@ test.lambda <- function(my.data, lambda.values,
 
 fitness.progress <- c()
 
-follow.fitness <- function(my.data,
+follow.fitness <- function(train.set, test.set,
                            population=100, 
                            iterations=20, 
                            mutation.rate=0.05,
@@ -269,14 +269,7 @@ follow.fitness <- function(my.data,
                            verbose=TRUE) {
   
   library(genalg)
-  
-  train.p.size <- 0.7 # percentage of training set
-  n.vars       <- ncol(my.data)-1
-  
-  # make train & test set
-  inTrain   <- sample(1:nrow(my.data), train.p.size * nrow(my.data))
-  train.set <- my.data[inTrain,]
-  test.set  <- my.data[-inTrain,]
+  n.vars <- ncol(train.set)-1
   
   GAmodel <- rbga.bin(size = max.monoids + max.degree*n.vars*max.monoids, 
                       popSize = population, 
@@ -291,7 +284,7 @@ follow.fitness <- function(my.data,
 
 # the monitor factory
 
-monitorEvalFactory <- function(train.set, n.vars, max.monoids) {
+monitorEvalFactory <- function(dataset, n.vars, max.monoids) {
   
   function(obj) {
     minEval <- min(obj$evaluations)       # get best fitness value
@@ -304,7 +297,7 @@ monitorEvalFactory <- function(train.set, n.vars, max.monoids) {
       bestSolution = obj$population[filter,];
     }
     # we have the best solution (chromosome), let's eval it and keep it
-    eval <- evalFuncFactory(train.set, n.vars, max.monoids) # get eval function
+    eval <- evalFuncFactory(dataset, n.vars, max.monoids) # get eval function
         
     fitness.progress <<- c(fitness.progress, eval(bestSolution))
     # print(make.formula(bestSolution, n.vars, max.monoids))

### 2015-08-08 texts/submission_03_asc/eprr.tex
Added table requested by reviewer
That’s a nasty thing, that :-(

@@
 
-\review{You should contrast obtained results in a table.  Results are hard to judge from Fig 3 alone.}
+\revised{You should contrast obtained results in a table.  Results are hard to judge from Fig 3 alone.}{Done that.}
 
@@
 
+\begin{table}[t]
+	\footnotesize
+	\renewcommand{\arraystretch}{0.5}
+	\begin{tabular}{ll|ccc}
+		&&& \textbf{error} \\
+		\textbf{dataset} & \textbf{method} & \textbf{quantile 25\%} & \textbf{mean} & \textbf{quantile 75\%} \\
+		\hline
+		Abalone	 & EPRR $\lambda = 0.7$ 	& 0.6392 & 0.6555 & 0.6677 \\
+					 & EPRR $\lambda = 0.8$ 	& 0.6408 & 0.6543 & 0.6636 \\
+					 & EPRR $\lambda = 0.9$ 	& 0.6481 & 0.6581 & 0.6707 \\
+					 & EPRR $\lambda = 1.0$ 	& 0.6542 & 0.6715 & 0.6816 \\
+					 & Linear Regression 	& 0.6803 & 0.6927 & 0.7078 \\
+					 & SVM (linear kernel) 	& 0.6916 & 0.7044 & 0.7205 \\
+					 & Regression Trees 	& 0.7423 & 0.7520 & 0.7621 \\
+					 & Random Forest 	& 0.6585 & 0.6695 & 0.6814 \\
+					 & Cond. Inference Trees 	& 0.7031 & 0.7126 & 0.7264 \\
+		Auto-Mpg	 & EPRR $\lambda = 0.7$ 	& 0.3635 & 0.3916 & 0.4147 \\
+					 & EPRR $\lambda = 0.8$ 	& 0.3629 & 0.3956 & 0.4228 \\
+					 & EPRR $\lambda = 0.9$ 	& 0.3646 & 0.4130 & 0.4215 \\
+					 & EPRR $\lambda = 1.0$ 	& 0.3691 & 0.3999 & 0.4057 \\
+					 & Linear Regression 	& 0.4071 & 0.4284 & 0.4473 \\
+					 & SVM (linear kernel) 	& 0.4116 & 0.4358 & 0.4613 \\
+					 & Regression Trees 	& 0.4216 & 0.4501 & 0.4785 \\
+					 & Random Forest 	& 0.3318 & 0.3624 & 0.3892 \\
+					 & Cond. Inference Trees 	& 0.4063 & 0.4372 & 0.4663 \\
+		Housing	 & EPRR $\lambda = 0.7$ 	& 0.4412 & 0.6650 & 0.5739 \\
+					 & EPRR $\lambda = 0.8$ 	& 0.4241 & 0.5274 & 0.6016 \\
+					 & EPRR $\lambda = 0.9$ 	& 0.4354 & 0.5717 & 0.6462 \\
+					 & EPRR $\lambda = 1.0$ 	& 0.4477 & 0.5417 & 0.5995 \\
+					 & Linear Regression 	& 0.4898 & 0.5313 & 0.5649 \\
+					 & SVM (linear kernel) 	& 0.4831 & 0.5469 & 0.6017 \\
+					 & Regression Trees 	& 0.4845 & 0.5232 & 0.5720 \\
+					 & Random Forest 	& 0.3283 & 0.3679 & 0.4023 \\
+					 & Cond. Inference Trees 	& 0.4676 & 0.5080 & 0.5413 \\
+		Kinematics	 & EPRR $\lambda = 0.7$ 	& 0.6600 & 0.6660 & 0.6720 \\
+					 & EPRR $\lambda = 0.8$ 	& 0.6617 & 0.6694 & 0.6751 \\
+					 & EPRR $\lambda = 0.9$ 	& 0.6636 & 0.6714 & 0.6761 \\
+					 & EPRR $\lambda = 1.0$ 	& 0.7568 & 0.7558 & 0.7739 \\
+					 & Linear Regression 	& 0.7672 & 0.7759 & 0.7849 \\
+					 & SVM (linear kernel) 	& 0.8074 & 0.8136 & 0.8247 \\
+					 & Regression Trees 	& 0.5673 & 0.6021 & 0.5803 \\
+					 & Random Forest 	& 0.7386 & 0.7344 & 0.7645 \\
+					 & Cond. Inference Trees 	& 0.7558 & 0.7620 & 0.7686
+	\end{tabular}
+	\caption{Tabular summary results for different regression methods on common datasets.
+Although \ac{EPRR} not always achieves the smallest expected error, performance is on-par with more sophisticated methods.}
+\end{table}
 Here is described the experiment setup used to gather and summarize the empirical evidence that supports this comparative study of \ac{EPR} and \ac{EPRR}.
@@
 %
-\caption{Summary results for different regression methods on common datasets.
+\caption{Graphical summary results for different regression methods on common datasets.
 Although \ac{EPRR} not always achieves the smallest expected error, performance is on-par with more sophisticated methods.
@@
 	\begin{center}
-\includegraphics[width=0.98\textwidth]{Fig4.pdf}
+		\includegraphics[width=0.98\textwidth]{Fig4xx.pdf}
 		\caption{Learning curve: Error progress for the \textsc{Housing} dataset during a single execution of the genetic algorithm.
 The figure shows the fitness evolution for different regularization values.
-The population for both consists of $200$ polynomials.}
+The population of each run consists of $200$ polynomials.}
 		\label{Abalone_fitnessProgress}

### 2015-08-08 texts/submission_03_asc/data/summary.py

@@ -56,7 +56,39 @@ def figs_3():
 				print(boxp(ds[col], col))
 		print('')
 
+def tab_3():
+	datasets = [
+		'Abalone',
+		'Auto-Mpg',
+		'Housing',
+		'Kinematics'
+	]
+	colnames = [
+		('ga.eprr.0.7', 'EPRR $\lambda = 0.7$'),
+		('ga.eprr.0.8', 'EPRR $\lambda = 0.8$'),
+		('ga.eprr.0.9', 'EPRR $\lambda = 0.9$'),
+		('ga.epr',	'EPRR $\lambda = 1.0$'),
+		('lin.reg',	'Linear Regression'),
+		('svm',	'SVM (linear kernel)'),
+		('rpart',	'Regression Trees'),
+		('rf',	'Random Forest'),
+		('ci.tree',	'Cond. Inference Trees'),
+	]
+
+	for ds_name in datasets:
+		f_name = "%s300_results.txt"%ds_name
+		ds = pd.read_csv( f_name, sep = ' ')
+		for (col,x) in colnames:
+			if not col == 'id':
+				q25 = ds[col].quantile(0.25)
+				q50 = ds[col].mean()
+				q75 = ds[col].quantile(0.75)
+				if col == 'ga.eprr.0.7':
+					print('%s\t & %s \t& %6.4f & %6.4f & %6.4f \\\\'%(ds_name,x,q25,q50, q75))
+				else:
+					print('\t\t\t & %s \t& %6.4f & %6.4f & %6.4f \\\\'%(x,q25,q50, q75))
+
 
 
 if __name__ == '__main__':
-	figs_3()
+	tab_3()

### 2015-08-08 texts/submission_03_asc/eprr.tex
Corrected wrong plot

@@
 	\footnotesize
-	\renewcommand{\arraystretch}{0.5}
+	\renewcommand{\arraystretch}{0.55}
 	\begin{tabular}{ll|ccc}
@@
 					 & Cond. Inference Trees 	& 0.7031 & 0.7126 & 0.7264 \\
+		\hline
 		Auto-Mpg	 & EPRR $\lambda = 0.7$ 	& 0.3635 & 0.3916 & 0.4147 \\
@@
 					 & Cond. Inference Trees 	& 0.4063 & 0.4372 & 0.4663 \\
+		\hline
 		Housing	 & EPRR $\lambda = 0.7$ 	& 0.4412 & 0.6650 & 0.5739 \\
@@
 					 & Cond. Inference Trees 	& 0.4676 & 0.5080 & 0.5413 \\
+		\hline
 		Kinematics	 & EPRR $\lambda = 0.7$ 	& 0.6600 & 0.6660 & 0.6720 \\

### 2015-08-10 texts/submission_03_asc/eprr.tex
minor corrections

@@
 \acrodef{SVM}[SVM]{Support Vector Machines}
+\acrodef{RF}[RF]{Random Forests}
+
 %
@@
 \ead{jpn@di.fc.ul.pt}
-\address{Departamento de Inform\'{a}tica, Faculdade de Ci\^{e}ncias da Universidade de Lisboa, Campo Grande, 1749-016 Lisboa, Portugal}
+\address{University of Lisboa, Faculty of Sciences, BioISI– Biosystems \& Integrative Sciences Institute, Campo Grande, 1749-016 Lisboa, Portugal}
 %
@@
 %
-This paper introduces Evolutionary Polynomial Regression with Regularization, an algorithm that extends the EPR method and describes a set of experiences on common datasets that compare both flavors of EPR and other methods including Linear Regression, Regression Trees and Support Vector Regression.
+This paper introduces Evolutionary Polynomial Regression with Regularization, an algorithm that extends the EPR method with a regularization term to control polynomial complexity.
+The article also describes a set of experiences on common datasets that compare both flavors of EPR and other methods including Linear Regression, Regression Trees and Support Vector Regression.
 
-The empiric conclusion of those experiments is that EPR with regularization is able to achieve better fitting than other non-ensemble methods and it has shorter computation time than plain EPR.
+The empiric conclusion of those experiments is that EPR with regularization is able to achieve better fitting than other non-ensemble methods and it needs less computation time than plain EPR.
 \end{abstract}
@@
 \revised{ GAs are repeatedly mentioned as an approach to search for locally optimal parameters, this is simply incorrect and is a very surprising statement.
-GAs are NOT a local optimizer in any shape or form.}{Our intention was merely to note that GAs give no guarantee of finding a global optimum}
+GAs are NOT a local optimizer in any shape or form.}{Our intention was merely to note that GAs give no guarantee of finding a global optimum.
+We rephrase the mentioned sentences.}
 
-\revised{What is the reference for the GA encoding used?}{It is ours.}
+\revised{What is the reference for the GA encoding used?}{It is ours. We stated this contribution more explicitly.}
 
@@
 It is basically a complexity based penalty which is really not so novel, there are even several cases in literature of penalty based GA functions.
-Please dig up further in the state of the art to back up better your claim of novelty.}{We included references to previous regularization in GA and outlined the role and originality of our contribution.}
+Please dig up further in the state of the art to back up better your claim of novelty.}{We included references to previous regularization in GA and better outlined the role and originality of our contribution.}
 
@@
 $\lambda < 1$) achieves better results than the opposite ($\lambda > 1$).
-This observation motivates further inquire, done in section~\ref{sec:experimental.results}.\revised{please clarify this paragraph and justify this claim}{we rephrased the statement.}
+This observation motivates further inquire, done in section~\ref{sec:experimental.results} where we compare the regularized versions with $\lambda < 1$ and the non-regularized algorithm against several other regression methods.\revised{please clarify this paragraph and justify this claim}{we rephrased the statement.}
 
@@
 From each dataset 70\% of the observations are reserved for training purposes and the remaining observations used to estimate the \ac{rmse}.
-To enhance the robustness of results this process is repeated $25$~times, each time with a different shuffling of the samples in the train and test sets.
-Some datasets with attribute values of different magnitudes have a pre-processing scaling transformation.
+To enhance the robustness of results this process is repeated $75$~times, each time with a different shuffling of the samples in the train and test sets.
+The datasets have been scaled to prevent problems with different magnitude attributes.
 The box plots in figures \ref{artificial_dataset1_lambda1.0} and \ref{fig:four.datasets.summary} resume the test set error distributions over these different runs.
@@
 \end{description}
+
 %
@@
 %
-Since this work is oriented to the \ac{rmse} of the \ac{EPRR} model it is necessary to assess how this depends on the number of generations of the \ac{GA}.
-As illustrated in Figure~\ref{Abalone_fitnessProgress}, the \ac{rmse} quickly drops during the initial $50$ to $100$ generations.
+Since this work is oriented to the \ac{EPRR} model \ac{rmse} it is necessary to assess how this depends on the number of generations of the \ac{GA}.
+As illustrated in Figure~\ref{Abalone_fitnessProgress}, the \ac{rmse} quickly drops during the initial $100$ generations.
 Then, it proceeds slower achieving better solutions only with marginal error reduction.
+The \ac{EPR} takes more than $300$ generations to achieve similar error results.
 
@@
 
-Of the regression methods considered \ac{SVM} achieves the best results in three out of four datasets.
-However \ac{SVM} and Conditional Inference Trees are pre-trained, having parameters tuned for each particular dataset unlike \ac{EPRR}, that runs with the same parameterization on all datasets.
-Even so it is the best estimator for the \textsc{Abalone} dataset and in the remaining datasets it outperforms most of the other estimators.
+Of the regression methods considered \ac{RF} achieves the best results in two out of four datasets.
+If we compare against non-ensemble methods, \ac{EPR} and \ac{EPRR} outperform the other estimators, with the exception of Regression Trees in the Kinematics dataset.
 
-Comparing \ac{EPR} and \ac{EPRR} --- the main article's topic --- the regularized version achieves much better results at \textsc{Abalone} and especially \textsc{Kinematics}.
-On the \textsc{Housing} dataset errors are improved  wrt \ac{EPR} in a difference in means, resulted in a 95\% HDI (Highest Density Interval) equal to $\left\lbrack 0.001, 0.119 \right\rbrack$ which, while borderline, achieves statistical significance.
-Only in the \textsc{Auto MPG} dataset \ac{EPR} achieves better results, even if not that different from \ac{EPRR}.
+Comparing \ac{EPR} and \ac{EPRR} --- the main article's topic --- the regularized version achieves better results at \textsc{Abalone} and especially \textsc{Kinematics}.
+A Bayesian Estimation was computed, which estimates the difference in means between two groups and yields a probability distribution over the difference.
+On the Abalone dataset errors are improved wrt EPR in a difference in means, for all tested $\lambda$ values less than $1$, resulted in a 95\% HDI (Highest Density Interval) equal to $\left\lbrack 0.01, 0.02 \right\rbrack$ which achieves statistical significance.
+On the Auto-MPG dataset the 95\% HDI is $\left\lbrack -0.01, 0.008 \right\rbrack$ which includes zero.
+On the Housing dataset the 95\% HDI is $\left\lbrack- 0.02, 0.05 \right\rbrack$, for $\lambda = 0.7$, which includes zero (the other values of $\lambda$ had similar statistics).
 
@@
 
-A cross-validation procedure can be implemented to refine the appropriate parameter values to achieve better errors.
-Namely, the regularization parameter, $\lambda$, can be tested with several values, instead of being fixed at $0.8$.
-Other parameters like mutation chance or the amount of elitism can also be tested.
+Parameters like mutation chance or the amount of elitism can be tested for tuning purposes.
 However, these type of tests need a low-level, fast implementation of \ac{EPR} and are postponed to future investigation.
@@
 The authors are grateful to the Fundação para a Ciência e Tecnologia (FCT) and the  R\&D laboratory LabMAg for the financial support given to this work, under the strategic project \textsc{PEst-OE/EEI/UI0434/2011}.
+
+Jo\~ao Pedro Neto's work is supported by centre grant (to BioISI, Centre Reference: UID/MULTI/04046/2013), from FCT/MCTES/PIDDAC, Portugal.
 

### 2015-08-11 texts/00-Previous/comparations/CoverinLetter.tex
Versão pré-final

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}

\title{Covering Letter}
\author{Francisco Coelho}
%\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle
There is no conflict of interest with any of the suggested reviewers.

\hspace{7cm}\raggedright{\includegraphics{fcsig.pdf}}

\end{document}

### 2015-08-11 texts/submission_03_asc/eprr.tex
Versão pré-final

@@
 \newcommand{\review}[1]{\textcolor{blue}{~---~#1}}
-\newcommand{\revised}[2]{\review{#1}~\textcolor{red}{#2}}
+%\newcommand{\revised}[2]{\review{#1}~\textcolor{red}{#2}}
+\newcommand{\revised}[2]{}
 %
@@
 \begin{tabular}{cc}
-\includegraphics[width=0.48\textwidth]{Fig1a.pdf}
+\includegraphics[width=0.475\textwidth]{Fig1a.pdf}
 &
-\includegraphics[width=0.48\textwidth]{Fig1b.pdf}
+\includegraphics[width=0.475\textwidth]{Fig1b.pdf}
 \\
-\includegraphics[width=0.48\textwidth]{Fig1c.pdf}
+\includegraphics[width=0.475\textwidth]{Fig1c.pdf}
 &
-\includegraphics[width=0.48\textwidth]{Fig1d.pdf}
+\includegraphics[width=0.475\textwidth]{Fig1d.pdf}
 \end{tabular}
@@
 					 & Regression Trees 	& 0.7423 & 0.7520 & 0.7621 \\
-					 & Random Forest 	& 0.6585 & 0.6695 & 0.6814 \\
+%					 & Random Forest 	& 0.6585 & 0.6695 & 0.6814 \\
 					 & Cond. Inference Trees 	& 0.7031 & 0.7126 & 0.7264 \\
@@
 					 & Regression Trees 	& 0.4216 & 0.4501 & 0.4785 \\
-					 & Random Forest 	& 0.3318 & 0.3624 & 0.3892 \\
+%					 & Random Forest 	& 0.3318 & 0.3624 & 0.3892 \\
 					 & Cond. Inference Trees 	& 0.4063 & 0.4372 & 0.4663 \\
@@
 					 & Regression Trees 	& 0.4845 & 0.5232 & 0.5720 \\
-					 & Random Forest 	& 0.3283 & 0.3679 & 0.4023 \\
+%					 & Random Forest 	& 0.3283 & 0.3679 & 0.4023 \\
 					 & Cond. Inference Trees 	& 0.4676 & 0.5080 & 0.5413 \\
@@
 					 & Regression Trees 	& 0.5673 & 0.6021 & 0.5803 \\
-					 & Random Forest 	& 0.7386 & 0.7344 & 0.7645 \\
+%					 & Random Forest 	& 0.7386 & 0.7344 & 0.7645 \\
 					 & Cond. Inference Trees 	& 0.7558 & 0.7620 & 0.7686
 	\end{tabular}
-	\caption{Tabular summary results for different regression methods on common datasets.
-Although \ac{EPRR} not always achieves the smallest expected error, performance is on-par with more sophisticated methods.}
+	\caption{Tabular summary results for different regression methods on common datasets.}
 \end{table}
 Here is described the experiment setup used to gather and summarize the empirical evidence that supports this comparative study of \ac{EPR} and \ac{EPRR}.
-Evaluation is focused in \ac{rmse} distribution and, besides \ac{EPR} and \ac{EPRR}, also uses several common regression methods and datasets easily accessible in \texttt{R}, the free software environment for statistical computing and graphics~\citep{R-Core-Team:2013aa}\footnote{The datasets and \texttt{R} code used to produce the results and plots in this paper are available online at \url{https://github.com/jpneto/GenAlgPoly}.}.
+Evaluation is focused in \ac{rmse} distribution and, besides \ac{EPR} and \ac{EPRR}, also uses several common regression methods and datasets easily accessible in \texttt{R}, the free software environment for statistical computing and graphics~\citep{R-Core-Team:2013aa}.
+Datasets and code used to produce the results and plots in this paper are available at \url{https://github.com/jpneto/GenAlgPoly}.
 A small consideration on the convergence speed concludes this section.
@@
 
-The \ac{EPRR} method is ranked against several well-known learning algorithms for regression, namely: non-regularized \ac{EPR}, Linear Regression, Support Vector Machines~\citep{Meyer:2012aa} with linear kernel, Regression Trees~\citep{Therneau:2013aa}, Random Forest~\citep{Strobl:2008aa,Strobl:2007aa} and Conditional Inference Trees~\citep{Hothorn:2006aa}.
+The \ac{EPRR} method is ranked against several well-known learning algorithms for regression, namely: non-regularized \ac{EPR}, Linear Regression, Support Vector Machines~\citep{Meyer:2012aa} with linear kernel, Regression Trees~\citep{Therneau:2013aa} and Conditional Inference Trees~\citep{Hothorn:2006aa,Strobl:2008aa,Strobl:2007aa}.
+No ensemble regression method was used in this comparison since these methods average between several models while \ac{EPRR} and the selected methods produce a single model.
 %
@@
 \caption{Graphical summary results for different regression methods on common datasets. 
-Although \ac{EPRR} not always achieves the smallest expected error, performance is on-par with more sophisticated methods.
 The regression methods depicted in these figures are:
@@
 7. Regression Trees;
-8. Random Forest (with 100 trees);
-9. Conditional Inference Trees.}
+%8. Random Forest (with 100 trees);
+8. Conditional Inference Trees.}
 \label{fig:four.datasets.summary}
@@
 	\begin{center}
-		\includegraphics[width=0.98\textwidth]{Fig4xx.pdf}
+		\includegraphics[width=0.98\textwidth]{Fig4.pdf}
 		\caption{Learning curve: Error progress for the \textsc{Housing} dataset during a single execution of the genetic algorithm.
@@
 
-Of the regression methods considered \ac{RF} achieves the best results in two out of four datasets.
-If we compare against non-ensemble methods, \ac{EPR} and \ac{EPRR} outperform the other estimators, with the exception of Regression Trees in the Kinematics dataset.
+Of the regression methods considered \ac{EPRR} achieves the best results in three out of four datasets, the exception being Regression Trees in the Kinematics dataset.
 

### 2015-08-11 texts/data/lambdas.py

@@ -1,507 +1,0 @@
-lambda095=[2.473536,
-2.414038,
-2.414038,
-2.414038,
-2.406747,
-2.406747,
-2.316709,
-2.350614,
-2.330452,
-2.313667,
-2.251034,
-2.251034,
-2.251034,
-2.275441,
-2.275441,
-2.275441,
-2.275307,
-2.275307,
-2.275307,
-2.268287,
-2.268287,
-2.268287,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.225473,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
-2.200263,
[…]

### 2015-08-11 texts/data/proc-lambdas.py

@@ -1,7 +1,0 @@
-f = file("lambdas.dat")
-
-for line in f.readlines():
-	if len(line) > 0:
-		eval(line)
-
-print dir()

### 2015-08-11 texts/data/retabulate.py

@@ -1,48 +1,0 @@
-import numpy as np
-
-
-def by_rows(d):
-	s = ""
-	ks = sorted(d.keys())
-	for k in ks:
-		s += "%05.3f "%eval(k)
-		for x in d[k]:
-			s+= "%05.3f "%x
-
-		s += "\n"
-	s += "\n"
-	return s
-
-def by_cols(d):
-	s = ""
-	ks = sorted(d.keys())
-	for k in ks:
-		 s += "%04.3f "%eval(k) 
-	s += "\n"
-	n = len( d[ ks[0] ] )
-
-	for i in range(n):
-		for k in ks:
-			s += "%04.3f "%d[k][i]
-		s += "\n"
-	s += "\n"
-	return s
-
-filename = "lambda.errors.Auto-Mpg.txt"
-
-f = file(filename,"r")
-
-# skip head
-f.readline()
-
-data = dict()
-
-for line in f.readlines():
-	line = line.replace("\n","").split(" ")
-	k = line[2]
-	if k in data.keys():
-		data[k].append( eval(line[1]) )
-	else:
-		data[k] = [ eval(line[1]) ]
-
-print by_cols(data)

### 2017-03-02 results/boxplot.stats/compute.boxplot.stats.Rmd
new review process
Changed project struture and updated files according to the 3rd review
process

```{r}
compute.box.plot  <- function(dataset) {
  df <- read.table(paste0("results/", dataset, "300_results_v2.txt"))

  res <- matrix(NA, ncol=5, nrow=ncol(df))

  for (i in 1:ncol(df))
    res[i,] <- boxplot.stats(df[,i])$stats

  res <- as.data.frame(res)
  names(res) <- c("min", "25%", "mean", "75%", "max")
  row.names(res) <- c("eprr.0.7", "eprr.0.8", "eprr.0.9", "epr", "lm", "svr", "rtree", "citree")
  res
}
```

```{r, fig.height=9}
datasets <- c("Abalone", "Auto-Mpg", "Housing", "Kinematics")

for(dataset in datasets) {
  print(dataset)
  df <- compute.box.plot(dataset)
  print(df)
  boxplot(t(df), main=dataset)
}
```

### 2017-03-02 results/svm_kernels/report.Rmd
new review process
Changed project struture and updated files according to the 3rd review
process

```{r}
library(readr)
```

```{r, message=FALSE}
dataset <- "Abalone"

df <- read_csv2(paste0(dataset, ".csv"))

df$`polynomial deg=4` <- pmin(df$`polynomial deg=4`, 1.0)

boxplot(df$`polynomial deg=4`, df$linear, df$radial, names=c("poly", "linear", "radial"))
```

```{r, message=FALSE}
dataset <- "Auto-Mpg"

df <- read_csv2(paste0(dataset, ".csv"))

df$`polynomial deg=4` <- pmin(df$`polynomial deg=4`, 1.0)

boxplot(df$`polynomial deg=4`, df$linear, df$radial, names=c("poly", "linear", "radial"))
```

```{r, message=FALSE}
dataset <- "Housing"

df <- read_csv2(paste0(dataset, ".csv"))

df$`polynomial deg=4` <- pmin(df$`polynomial deg=4`, 1.0)

boxplot(df$`polynomial deg=4`, df$linear, df$radial, names=c("poly", "linear", "radial"))
```

```{r, message=FALSE}
dataset <- "Kinematics"

df <- read_csv2(paste0(dataset, ".csv"))

df$`polynomial deg=4` <- pmin(df$`polynomial deg=4`, 1.0)

boxplot(df$`polynomial deg=4`, df$linear, df$radial, names=c("poly", "linear", "radial"))

```

### 2017-03-02 code/gapolyfuncs.R

@@ -1,6 +1,6 @@
 #########################################################################
 # initial values, can be overriden
-max.monoids <- 4  # poly = monoid1 + monoid2 + ...
+max.monoids <- 6  # poly = monoid1 + monoid2 + ...
 max.degree  <- 3  # meaning exponents up to 2^max.degree - 1
 
 #########################################################################
@@ -117,17 +117,16 @@ evalFuncFactory <- function(df, n.vars, max.monoids, lambda=1) {
 
 # pre: my.data must have columns named x1...xn and the last one is y (the output)
 make.report <- function(my.data, 
-                        population=100,     # GA population
-                        iterations=25,      # GA iterations
-                        runs=10,            # number of runs
-                        mutation.rate=0.05, # GA mutation rate
-                        lambda=1.0,         # regularization variable
+                        population=100,      # GA population
+                        iterations=25,       # GA iterations
+                        runs=10,             # number of runs
+                        mutation.rate=0.075, # GA mutation rate
+                        lambda=1.0,          # regularization variable
                         verbose=TRUE) {
   
   library(genalg)
   library(e1071)
   library(rpart)   
-  library(randomForest)
   library(party)
   
   train.p.size <- 0.7 # percentage of training set
@@ -138,7 +137,6 @@ make.report <- function(my.data,
   lm.error     <- rep(0,runs)
   svm.error    <- rep(0,runs)
   rpart.error  <- rep(0,runs)
-  rf.error     <- rep(0,runs)
   citree.error <- rep(0,runs)
   
   best.ga.model <- NULL  # the best GA model 
@@ -173,7 +171,9 @@ make.report <- function(my.data,
     lm.pred  <- predict(lm.model, test.set[,-ncol(test.set)]) 
     lm.error[i] <- rsme(lm.pred,test.set[,ncol(test.set)])
         
-    svm.model <- svm(y ~ ., data=train.set, kernel='linear')
+    #svm.model <- svm(y ~ ., data=train.set, kernel='polynomial', degree=4) # poorer results
+    #svm.model <- svm(y ~ ., data=train.set, kernel='linear')
+    svm.model <- svm(y ~ ., data=train.set, kernel='radial')  # best results
     svm.pred  <- predict(svm.model, test.set[,-ncol(test.set)]) 
[…]

### 2017-03-02 code/runAllDatasets.R

@@ -21,7 +21,9 @@ read_clean <- function(dataset) {
   names(my.data) <- c(paste0("x",1:(ncol(my.data)-1)),"y") # columns are named as x_i & y
   
   if (dataset=="Abalone") {
-    my.data$x1 <- as.numeric(my.data$x1) # x1 was a factor, make it a numeric value
+    f <- as.factor(my.data$x1)
+    levels(f) <- 1:3
+    my.data$x1 <- as.numeric(levels(f))[f]  # x1 was a factor, make it a numeric value
   }
   
   if (dataset=="Auto-Mpg") {
@@ -35,19 +37,22 @@ read_clean <- function(dataset) {
 }
 
 pop  <- 300  # GA population
-reg  <- 0.7  # amount of regularization
+reg  <- 0.9  # amount of regularization
 runs <- 75   # number of simulations
-iter <- 100  # number of interations for each simulation
+iter <- 150  # number of GA generations per simulation
 
 digits <- 5 # round errors after number of digits
 
 ################## started Monday 13:43 
 
-#datasets <- c("Abalone", "Auto-Mpg", "Housing", "Kinematics")
-datasets <- c("Housing", "Kinematics")
+datasets <- c("Abalone", "Auto-Mpg", "Housing", "Kinematics")
+# datasets <- c("Abalone")
+# datasets <- c("Auto-Mpg")
+# datasets <- c("Housing")
+# datasets <- c("Kinematics")
 
 for (dataset in datasets) {
-  # print(paste("### Starting", dataset, "dataset..."))
+  print(paste("\\n### Starting", dataset, "dataset..."))
         
   my.data <- read_clean(dataset)
   
@@ -63,7 +68,6 @@ for (dataset in datasets) {
                    lin.reg = report_reg$lm.error,
                    svm     = report_reg$svm.error,
                    rpart   = report_reg$rpart.error,
-                   rf      = report_reg$rf.error,
                    ci.tree = report_reg$citree.error)
   
   df <- round(df, digits)
